---
title: "Repeatability in foraging effort results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---

# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_MasterBayes(v.2.59)_, _kinship2(v.1.9.6.2)_, _quadprog(v.1.5-8)_, _genetics(v.1.3.8.1.3)_, _mvtnorm(v.1.3-3)_, _MASS(v.7.3-60)_, _gtools(v.3.9.5)_, _gdata(v.3.0.1)_, _combinat(v.0.0-8)_, _igraph(v.2.1.4)_, _MCMCglmm(v.2.36)_, _ape(v.5.8-1)_, _coda(v.0.19-4.1)_, _brms(v.2.23.0)_, _Rcpp(v.1.1.1)_, _ggridges(v.0.5.7)_, _glmmTMB(v.1.1.11)_, _stringr(v.1.6.0)_, _kableExtra(v.1.4.0)_, _tidyr(v.1.3.2)_, _gridExtra(v.2.3)_, _lme4(v.2.0-1)_, _Matrix(v.1.5-4.1)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_

**loaded via a namespace (and not attached):** 
_tidyselect(v.1.2.1)_, _viridisLite(v.0.4.3)_, _farver(v.2.1.2)_, _loo(v.2.9.0)_, _S7(v.0.2.1)_, _fastmap(v.1.2.0)_, _tensorA(v.0.36.2.1)_, _digest(v.0.6.39)_, _estimability(v.1.5.1)_, _lifecycle(v.1.0.5)_, _posterior(v.1.7.0)_, _compiler(v.4.3.1)_, _rlang(v.1.1.7)_, _tools(v.4.3.1)_, _knitr(v.1.51)_, _bridgesampling(v.1.2-1)_, _xml2(v.1.5.2)_, _RColorBrewer(v.1.1-3)_, _abind(v.1.4-8)_, _withr(v.3.0.2)_, _purrr(v.1.2.1)_, _numDeriv(v.2016.8-1.1)_, _grid(v.4.3.1)_, _xtable(v.1.8-8)_, _emmeans(v.2.0.3)_, _scales(v.1.4.0)_, _cli(v.3.6.5)_, _rmarkdown(v.2.31)_, _reformulas(v.0.4.4)_, _generics(v.0.1.4)_, _otel(v.0.2.0)_, _RcppParallel(v.5.1.11-2)_, _rstudioapi(v.0.18.0)_, _minqa(v.1.2.8)_, _pander(v.0.6.6)_, _splines(v.4.3.1)_, _bayesplot(v.1.15.0)_, _parallel(v.4.3.1)_, _matrixStats(v.1.5.0)_, _vctrs(v.0.7.2)_, _boot(v.1.3-28.1)_, _systemfonts(v.1.3.2)_, _glue(v.1.8.0)_, _nloptr(v.2.2.1)_, _distributional(v.0.7.0)_, _stringi(v.1.8.7)_, _cubature(v.2.1.4-1)_, _gtable(v.0.3.6)_, _tibble(v.3.3.1)_, _pillar(v.1.11.1)_, _htmltools(v.0.5.9)_, _Brobdingnag(v.1.2-9)_, _R6(v.2.6.1)_, _TMB(v.1.9.21)_, _textshaping(v.1.0.5)_, _Rdpack(v.2.6.3)_, _evaluate(v.1.0.5)_, _lattice(v.0.21-8)_, _rbibutils(v.2.3)_, _backports(v.1.5.0)_, _corpcor(v.1.6.10)_, _rstantools(v.2.6.0)_, _svglite(v.2.2.2)_, _nlme(v.3.1-162)_, _checkmate(v.2.3.4)_, _mgcv(v.1.8-42)_, _xfun(v.0.57)_, _forcats(v.1.0.1)_ and _pkgconfig(v.2.0.3)_

## Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4",  "gridExtra", "tidyr", "kableExtra", "stringr", "glmmTMB", "ggridges", "brms", "MCMCglmm", "igraph", "MasterBayes")

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"

```

```{r process_data}

ind_files <- list.files("Data_inputs/", pattern = "individual_trips")
ind_list <- vector(mode = "list", length = 4)

for (i in 1:length(ind_files)) {
  
  my_data <- loadRData(paste0("Data_inputs/", ind_files[i]))
  
  ind_list[[i]] <- my_data
  
}

trip_data <- do.call("rbind", ind_list)

# Load metadata ====

load("Data_inputs/ind_meta.RData")
all_ind.meta %<>% mutate(species = ifelse(species == "BBA", "bbal", "waal")) %>% select(-birthYr)

## Bind it in
trip_data <- merge(trip_data, all_ind.meta, by = c("ring", "species"), all.x = T)

# Filter data ====

# Remove any trips where landings == 0 (not real foraging)
rep_data <- trip_data %>% 
  group_by(ring) %>% 
  mutate(n_seasons = n_distinct(season),
         ring_season = paste(ring, season, sep = "_")) %>%
  filter(landing > 0) %>% 
  ungroup() %>%
  relocate(ring_season, .before = colony)

# Calculate landings per day
rep_data %<>% mutate(landings_per_day = landing/duration.days)

save(rep_data, file = "Data_inputs/behavioural_repeatability_data.RData")

```

### Sample sizes

```{r sample_sizes}

trip_sum <- rep_data %>%
  group_by(colony, species, ring) %>%
  summarise(n_trips = n(),
            n_seasons = n_distinct(season),
            .groups = "drop") %>%
  group_by(colony, species) %>%
  summarise(N = n_distinct(ring),
            'Trips per bird' = mean(n_trips),
            'Seasons per bird' = mean(n_seasons)) %>%
  mutate(species = ifelse(species == "bbal", "BBAL", "WAAL"),
         colony = ifelse(colony == "ker", "Kerguelen",
                         ifelse(colony == "cro", "Crozet", "Bird Island")))

ind_per_seasons <- rep_data %>%
  group_by(colony, species, ring) %>%
  summarise(n_seasons = n_distinct(season)) %>%
  ungroup() %>%
  mutate(rpt_category = case_when(
    n_seasons == 1 ~ "1 seasons",
    n_seasons == 2 ~ "2 seasons",
    n_seasons == 3 ~ "3 seasons",
    n_seasons > 3 ~ "4+ seasons"
  )) %>%
  group_by(colony, species, rpt_category) %>%
  summarise(num_birds = n()) %>%
  arrange(species, colony, rpt_category) %>%
  pivot_wider(
    names_from = rpt_category,       
    values_from = num_birds, 
    values_fill = 0) %>%
  mutate(species = ifelse(species == "bbal", "BBAL", "WAAL"),
         colony = ifelse(colony == "ker", "Kerguelen",
                         ifelse(colony == "cro", "Crozet", "Bird Island"))) %>%
  relocate(species, .before = colony) 

summary.table <- merge(ind_per_seasons, trip_sum, by = c("species", "colony")) %>%
  rename(Species = species, Colony = colony)


kable(summary.table, format = "html", 
      caption = 'Table 1. Sample sizes for each species and colony, showing the number of individuals for which we have 2, 3, or 4+ seasons of data, collected using GLS and/or GPS devices.') %>%
  collapse_rows(columns = c(1)) %>%
  row_spec(2, extra_css = "border-bottom: 1px solid;") %>%
  kable_classic(font_size = 12)




```

# Estimate repeatablity

```{r estimate_repeatability, eval = FALSE}

load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., landing, transit_time.pct, transit_to_rest.ratio) == 0) == 0)

# Split data
rep_bbal <- rep_data %>% filter(species == "bbal")
rep_waal <- rep_data %>% filter(species == "waal")

# Set weakly informative priors
priors <- c(
  prior(normal(0, 10), class = "Intercept"), # Intercept
  prior(normal(0, 2), class = "b"), # Fixed effects
  prior(student_t(3, 0, 5), class = "sd"), # Group-level SDs
  prior(student_t(3, 0, 10), class = "sigma"))

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "landings_per_day", "transit_time.pct", "rest_time.pct")

rpt_output.list <- vector(mode = "list", length = length(cov))

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  if(grepl("landings", cov[i])) { cov_value <- cov[i] } else {
  cov_value <- paste0("log(", cov[i], ")") }
  formula <- paste0(cov_value, " ~ phase + (1 | season) + (1 | ring ) + (1 | ring_season)") 
  
  # Fit models
  
  ## WAAL ## ----
  print("Processing wanderers...")

  ### Crozet ### ----
  waal_cro.brm_rpt <- brm(
      formula = formula,
      data = rep_waal %>% filter(colony == "cro"),
      family = gaussian(),  
      prior = priors,
      cores = 4, chains = 4, iter = 4000,
      control = list(adapt_delta = 0.95))
    
  # Calc repeatability
  posterior <- as_draws_df(waal_cro.brm_rpt)
  
  waal_cro.post_df <- data.frame(cbind(calc_rpt_brms_slope(waal_cro.brm_rpt, rep_waal %>% filter(colony == "cro"))),
                             behaviour = cov[i],
                             species = "WAAL",
                             colony = "Crozet")
  
  mod_name <- paste0("waal_cro.brm_rpt_", cov[i])
  assign(mod_name, waal_cro.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/waal_cro_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  ### Bird Island ### ----
   waal_birdis.brm_rpt <- brm(
    formula = formula,
    data = rep_waal %>% filter(colony == "birdis"),
    family = gaussian(),  
    prior = priors,
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Calc repeatability
  posterior <- as_draws_df(waal_birdis.brm_rpt)
 
  waal_birdis.post_df <- data.frame(cbind(calc_rpt_brms_slope(waal_birdis.brm_rpt, rep_waal %>% filter(colony == "birdis"))),
                              behaviour = cov[i],
                              species = "WAAL",
                              colony = "Bird Island")
    
  mod_name <- paste0("waal_birdis.brm_rpt", cov[i])
  assign(mod_name, waal_birdis.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/waal_birdis_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  
  ## BBAL ## ----
  print("Processing black-browed...")

  ### Kerguelen ### ----
  bbal_ker.brm_rpt <- brm(
      formula = formula,
      data = gls_trips,# rep_bbal %>% filter(colony == "ker"),
      family = gaussian(),  
      prior = priors,
      cores = 4, chains = 4, iter = 4000,
      control = list(adapt_delta = 0.95))
    
  # Calc repeatability
  posterior <- as_draws_df(bbal_ker.brm_rpt)

  bbal_ker.post_df <- data.frame(cbind(calc_rpt_brms_slope(bbal_ker.brm_rpt, rep_bbal %>% filter(colony == "ker"))),
                             behaviour = cov[i],
                             species = "BBAL",
                             colony = "Kerguelen")
  
  mod_name <- paste0("bbal_ker.brm_rpt_", cov[i])
  assign(mod_name, bbal_ker.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/bbal_ker_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  ### Bird Island ### ----
  bbal_birdis.brm_rpt <- brm(
    formula = formula,
    data = rep_bbal %>% filter(colony == "birdis"),
    family = gaussian(),  
    prior = priors,
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Calc repeatability
  posterior <- as_draws_df(bbal_birdis.brm_rpt)
  
  bbal_birdis.post_df <- data.frame(cbind(calc_rpt_brms_slope(bbal_birdis.brm_rpt, rep_bbal %>% filter(colony == "birdis"))),
                              behaviour = cov[i],
                              species = "BBAL",
                              colony = "Bird Island")
    
  mod_name <- paste0("bbal_birdis.brm_rpt", cov[i])
  assign(mod_name, bbal_birdis.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/bbal_birdis_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  # Output summary
  rpt_output.list[[i]] <- rbind(waal_cro.post_df, waal_birdis.post_df,
                                bbal_ker.post_df, bbal_birdis.post_df)
   
}

rpt_output.full <- do.call("rbind", rpt_output.list)
rpt_output.full %<>% arrange(species)

# Get sample sizes and add them in
rep_samples <- rep_data %>% 
  group_by(colony, species) %>% 
  summarise(n_trips = n_distinct(start_date), n_birds = n_distinct(ring)) %>% 
  mutate(species = ifelse(species == "bbal", "BBAL", "WAAL"),
         colony = case_when(colony == "birdis" ~ "Bird Island",
                            colony == "cro" ~ "Crozet",
                            colony == "ker" ~ "Kerguelen"))

rpt_output.full %<>% left_join(rep_samples, by = c("species", "colony"))

save(rpt_output.full, file = "Data_outputs/behavioural_repeatability_results.RData")

# Model checking 
# plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
# pp_check(bbal_birdis.brm_rpt)
# 
# pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

```

# Estimate heritability

```{r effort_heritability}

# Set priors
effort_priors <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(normal(0, 1), class = "b"),
  prior(student_t(3, 0, 1), class = "sd", group = "ring_pe"),
  prior(exponential(1), class = "sd", group = "animal"),
  prior(student_t(3, 0, 1), class = "sigma"))

# Load the data
load("Data_inputs/behavioural_repeatability_data.RData")

## Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, landings_per_day, transit_time.pct) == 0) == 0)

## New variable names; scale covariates
rep_data %<>% mutate(animal = ring, ring_pe = ring) 

# Loop through species and colonies
spec_col <- c("bbal_birdis", "waal_birdis", "waal_cro")
posteriors = list()

cov <- c("duration.hrs", "landings_per_day", "transit_time.pct", "rest_time.pct")

for (i in 1:length(spec_col)) {

  my_species <- strsplit(spec_col[i], "_")[[1]][1]
  my_colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  mod_data <- rep_data %>% filter(species == my_species & colony == my_colony)

  # Load the pedigree info
  pedigree.df <- read.csv(paste0("Data_inputs/", my_species, "_", my_colony, "_pedigree.csv"))[,1:3]
  pedigree.df[,c(1:3)] <- lapply(pedigree.df[,c(1:3)], as.character)
  
  ## Create the genetic matrix
  Amat <- make_Amat(mod_data, pedigree.df)
  
  # Fit the models
  posteriors_list <- list()
  
  for (c in 1:length(cov)) {
    
    print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", c, " of ", length(cov), ": ", cov[c]))
    
    if(grepl("landings", cov[c])) { cov_value <- paste0(cov[c], "_scaled") } else {
         cov_value <- paste0(cov[c], "_log_scaled") }
    
    formula <- paste0(cov_value, " ~ 1 + phase + (1 | season) + (1 | ring_pe) + (1 | ring_season) + (1 | gr(animal, cov = Amat))")

    # Scale and log the covariate
    
    if(grepl("landings", cov[c])) { 
      
      scaled_col <- paste0(cov[c], "_scaled")
      mod_data[[scaled_col]] <- scale(mod_data[[cov[c]]])
      
    } else {
         
      log_col <- paste0(cov[c], "_log")
      scaled_col <- paste0(cov[c], "_log_scaled")

      mod_data[[log_col]] <- log(mod_data[[cov[c]]])
      mod_data[[scaled_col]] <- scale(mod_data[[log_col]])[,1]
    }
    
    # Fit the model
    mod_brm <- brm(
      formula = formula,
      prior = effort_priors,
      data = mod_data,
      family = gaussian(),
      data2 = list(Amat = Amat),
      chains = 4, cores = 4, iter = 8000, warmup = 4000,
      control = list(max_treedepth = 15, adapt_delta = 0.99))
    
    ## Get posteriors 
    posterior <- as_draws_df(mod_brm)
    var_season  <- posterior$sd_season__Intercept^2
    var_ring_pe <- posterior$sd_ring_pe__Intercept^2
    var_ring_season <- posterior$sd_ring_season__Intercept^2
    var_animal  <- posterior$sd_animal__Intercept^2
    var_resid   <- posterior$sigma^2
    
    h2_post <- var_animal / (var_animal + var_season + var_ring_pe + var_ring_season + var_resid)
    
    posterior.df <- data.frame(cbind(var_animal, var_season, var_resid, h2_post),
                               behaviour = cov[c])
    
    posteriors_list[[c]] <- posterior.df
   
    # Output model
    save(mod_brm, file = paste0("Data_outputs/",
                                my_species, "_", my_colony,
                                "_brms_behaviour_heritability_", cov[c], ".RData"))

    }
     
  posteriors[[spec_col[i]]] <- bind_rows(posteriors_list)
  
  }


posteriors_df <- bind_rows(posteriors, .id = "spec_col")
save(posteriors_df, file = "Data_outputs/behaviour_heritability_posteriors.RData")

```

### Process heritability estimates

```{r plot_effort_heritability}

load("Data_outputs/behaviour_heritability_posteriors.RData")

posteriors_df %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "landings_per_day" = "Landings per day",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

posteriors_df %<>%
  group_by(spec_col, behaviour) %>%
  mutate(mean_h2 = mean(h2_post, na.rm = TRUE)) %>%
  arrange(behaviour, mean_h2) %>%
  ungroup()

## Add facet code for plotting
posteriors_df %<>%
   mutate(facet_code = case_when(
    spec_col == "bbal_birdis" ~ "A",
    spec_col == "waal_birdis" ~ "C",
    spec_col == "waal_cro" ~ "D"))


# Make the plot
design = "
A#
CD
"

facet_labels <- c(
  A = "BBAL Bird Island",
  C = "WAAL Bird Island",
  D = "WAAL Crozet"
)

heritability_plot <- ggplot(posteriors_df, aes(x = h2_post, y = behaviour)) +
  # stat_density_ridges(fill = "#1b9e77",  alpha = 0.6, scale = 1, quantile_lines = TRUE, quantiles = 2, vline_colour = "#08519c", vline_size = 1, vline_type = "dashed") +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.6, scale = 1) +
  theme_bw() +
   ggh4x::facet_manual(~facet_code, design = design, scales = "free_x",
                       labeller = as_labeller(facet_labels)) +
  labs(x = "Posterior distribution for heritability (h²)", y = "") 


```

## Supplementary analyses

### How does repeatability change with the number of individual repeats?

I have a lot of variation in the number of seasons' worth of data are
available per individual. To test whether this might be affecting my
estimates, here I have rerun the analyses using data that have been
sampled to individuals with more than or equal 2, 3, or 4 trip repeats
(within a season).

To speed things up, I'm calculating repeatability using simple linear
model versions of the brms models, as I'm not concerned about
uncertainty here.

Repeatability values stabilise at \>=3 repeats per individual-year so
ensuring this filter is included in the main analysis.

```{r SUPPLEMENTARY_repeatability_by_repeats_models}

load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct) == 0) == 0)

# Get time since first trip for each bird/season
rep_data %<>%
  group_by(ring, season) %>%
  mutate(time_since_first = as.numeric(difftime(start_date, min(start_date), units = "days"))) %>%
  ungroup()

# Split data
rep_bbal <- rep_data %>% filter(species == "bba")
rep_waal <- rep_data %>% filter(species == "waal")

cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct")

n_reps <- c(2,3,4,5)

rep_testing.output <- list()
index <- 1 

for (i in 1:length(cov)) {
  
  for (n in 1:length(n_reps)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + phase + age")
  
  # Subset to n_reps
  rep_bbal.sub <- rep_bbal %>% 
    group_by(ring_season) %>%
    mutate(n_trips = n_distinct(start_date)) %>%
    filter(n_trips >= n_reps[n])
  
  rep_waal.sub <- rep_waal %>% 
    group_by(ring_season) %>%
    mutate(n_trips = n_distinct(start_date)) %>%
    filter(n_trips >= n_reps[n])
  
  ## BBAL ##
  bbal_birdis.lmm <- lmer(formula, rep_bbal.sub %>% filter(colony == "birdis"))
  bbal_ker.lmm <- lmer(formula, rep_bbal.sub %>% filter(colony == "kerguelen"))
  
  bbal_birdis.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(bbal_birdis.lmm),
                                species = "bbal",
                                colony = "birdis")
  
  bbal_ker.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(bbal_ker.lmm),
                                species = "bbal",
                                colony = "kerguelen")
  
  ## WAAL ##
  waal_birdis.lmm <- lmer(formula, rep_waal.sub %>% filter(colony == "birdis"))
  waal_cro.lmm <- lmer(formula, rep_waal.sub %>% filter(colony == "crozet"))
  
  waal_birdis.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(waal_birdis.lmm),
                                species = "waal",
                                colony = "birdis")
  
  waal_cro.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(waal_cro.lmm),
                                species = "waal",
                                colony = "crozet")
 
  rep_all <- rbind(bbal_birdis.rpt, bbal_ker.rpt, waal_birdis.rpt, waal_cro.rpt) %>%
                mutate(behaviour = cov[i],
                       n_trips = n_reps[n]) %>%
                relocate(species, colony, behaviour, n_trips, level, rpt)
  
  rep_testing.output[[index]] <- rep_all
  index <- index + 1

  }
   
}

rep_testing.all <- bind_rows(rep_testing.output)
rep_testing.all %<>% arrange(species)

save(rep_testing.all, file = "Data_outputs/SUPP_behaviour_rpt_test_Ntrips.RData")

```
