---
title: "Spatial repeatability results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

Inputs:

-   *species_colony*\_gps_labelled.RData

Outputs:

-   Figure 2, Figure 3, Figure Sx, Figure Sx

-   all_gps.RData

-   bearings_data.Rdata

-   *species_colony*\_bearing_differences.RData

-   bearing_means.RData

-   bearing_models_summary.RData

-   *species_colony*\_ba_comparison.RData

-   ba_means.RData

-   ba_model_summary.RData

This document contains the results of the spatial repeatability section of the repeatability project. It uses outputs from the main R script, so make sure those outputs are up-to-date before running the markdown code.

## Set-up

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load functions, packages, & data ---------------------------------------------

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "ggpubr", "tidyr", "sf", "ecmwfr", "ggridges", "rnaturalearth", "purrr", "geosphere", "sp", "kableExtra", "emmeans", "data.table", "adehabitatHR", "betareg", "MasterBayes", "pheatmap", "nadiv", "brms", "circular") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
# if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"

```

## Foraging movements

#### Figure S2: Foraging tracks for all populations

```{r prep_GPS_data, eval = FALSE, include = FALSE}

# Load all the GPS data ---------------

bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

spec_col <- c("bbal_ker", "bbal_birdis", "waal_cro", "waal_birdis")
all_gps = vector(mode = "list", length = 4)

for (i in 1:length(spec_col)) {
  
  my_gps <- loadRData(paste0("Data_inputs/", spec_col[i], "_gps_labelled.RData"))
  
  # gps_file <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  # my_gps <- loadRData(paste0("Data_inputs/", gps_file)) 
  
  my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))
  
  ## Filter out individuals with only one trip
  my_gps %<>% 
    filter(loc == "trip") %>%
    group_by(ring) %>%
    mutate(n_trips = n_distinct(ID)) %>%
    filter(n_trips > 1) %>%
    select(-n_trips)
  
  ## Add species/colony info in
  my_gps %<>%
    mutate(spec_col = spec_col[i]) %>%
    separate(spec_col, into = c("species", "colony"), sep = "_")

  my_gps$col_lat = ifelse(my_gps$colony == "birdis", bi_coords$latitude, cro_coords$latitude)
  my_gps$col_lon = ifelse(my_gps$colony == "birdis", bi_coords$longitude, cro_coords$longitude)
  
  all_gps[[i]] <- my_gps
  
}

all_gps.df <- do.call("rbind", all_gps)
save(all_gps.df, file = "Data_outputs/all_gps_data.RData")

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# Loop through data
for (i in 1:length(spec_col)) {
  
  # Set projections (use same for Ker + Cro)
   
  if(grepl("birdis", spec_col[i])) {
      proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
  } else {
      proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
    }

  world2 <- sf::st_transform(world, crs = proj.utm)
  
  # Project GPS data
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  my_gps.proj <- all_gps.df %>% filter(species == my_species & colony == my_colony) %>% arrange(ringYr, datetime)
  coordinates(my_gps.proj) <- ~ longitude+latitude
  proj4string(my_gps.proj) <- proj.dec
  
  my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
  my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

  # Project the colony
  col_coords <- data.frame(longitude = my_gps.proj$col_lon[1], latitude = my_gps.proj$col_lat[1])
  
  if(grepl("ker", spec_col[i])) { col_coords <- data.frame(longitude = 70.2333, latitude = -49.6833) }
    
  coordinates(col_coords) <- ~longitude+latitude
  proj4string(col_coords) <- proj.dec
  colony.sp <- spTransform(col_coords, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)

  
  # Build the plot
  
  ## Set min/max limits
  y_min = min(my_gps.proj.df$coords.x2 - 100000)
  y_max = max(my_gps.proj.df$coords.x2 + 100000)
  x_min = min(my_gps.proj.df$coords.x1 - 100000)
  x_max = max(my_gps.proj.df$coords.x1 + 100000)
   
  gps_tracks.plot <- ggplot(data = world2) + 
    geom_sf(fill = "cadetblue", colour = "cadetblue") +
    coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
             label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
    geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
              dat = my_gps.proj.df, group = "identity") +
    scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
    scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
    scale_colour_viridis_d() +
    annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
    labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bbal", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "cro", "Crozet", "Kerguelen"))))) +
    theme(panel.background = element_rect(fill = "white"),
          panel.grid.major = element_line(colour = "grey80"),
          panel.border = element_rect(colour = "black", fill = NA),
          axis.text.x.bottom = element_blank(),
          axis.title.x.bottom = element_blank(),
          axis.text.y.right = element_blank(),
          axis.title.y.right = element_blank(),
          axis.title.y.left = element_blank(),
          axis.text.y.left = element_text(size = 10),
          axis.text.x.top = element_text(size = 10),
          legend.position = "none")

  
  assign(paste0(spec_col[i], "_gps.plot"), gps_tracks.plot)
  plot_name <- paste0(spec_col[i], "_gps.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_gps_plot.RData"))
  
}


```

```{r gps_tracks.plot, fig.height=8, fig.width=10}

plot_files <- list.files(path = "Data_outputs/R objects/", pattern = "gps_plot")
plot_paths <- paste0("Data_outputs/R objects/", plot_files)

for (file in plot_paths) {
  load(file)
}

#png("Figures/MS Figures/FIGS2_all_gps_tracks.png", width = 10, height = 10, units = "in", res = 700)
ggarrange(bba_kerguelen_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
#dev.off()

```

Here I've plotted all the GPS tracks for the different species and colonies. Note that birds from Kerguelen tend to take a NE or SW trajectory, while WAAL at Bird Island tend to go NW.

## Example tracks from each species/colony

Here I've plotted a couple of examples of within- versus between-individual tracks for each species and colony. These have been cherry-picked to show individuals with similar tracks to illustrate the greatest level of within-track similarity; note that even the most similar wandering albatross tend to show a lot of variation in their trips.

#### Figure S3: Example GPS trtacks

```{r example_gps_plots, fig.height=8, fig.width=10}

load("Data_outputs/all_gps_data.RData")

# Common parameters
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
world <- ne_countries(scale = "medium", returnclass = "sf")
spec_col <- c("bbal_ker", "bbal_birdis", "waal_cro", "waal_birdis")

targets <- c("bbal_birdis_1456598", "bbal_birdis_1441460", "bbal_birdis_1320517",
             "bbal_ker_CF27498", "bbal_ker_CF39444", "bbal_ker_CF31429",
             "waal_birdis_5145048", "waal_birdis_5144589", "waal_birdis_5132544",
             "waal_cro_BS6563", "waal_cro_BS18853", "waal_cro_BS21258")

all_gps.df %<>% mutate(plotID = paste(species, colony, ring, sep = "_"))

for (i in 1:length(spec_col)) {
  
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  # Get colony location
  if(my_colony == "birdis") { colony.df = data.frame(Lon = -38.05, Lat = -54) } else {
     colony.df = data.frame(Lon = 51.706972, Lat = -46.358639) }
  
  # Set projections
  proj.utm <- paste0("+proj=laea +lat_0=", colony.df$Lat[1], " +lon_0=", colony.df$Lon[1],
                     " +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
  
  ## Colony projection (for plotting)
  coordinates(colony.df) <- ~Lon+Lat
  proj4string(colony.df) <- proj.dec
  colony.sp <- spTransform(colony.df, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)
  
  ## Get the base world plot
  world2 <- sf::st_transform(world, crs = proj.utm)
  
  ## Subset GPS and project
  targets_speccol <- targets[grepl(spec_col[i], targets)]
  
  track.proj <- all_gps.df %>% filter(plotID %in% targets_speccol)
  coordinates(track.proj) <- ~ longitude+latitude
  proj4string(track.proj) <- proj.dec
  
  track.proj.sp <- spTransform(track.proj, CRS(proj.utm))
  track.proj.df <- as.data.frame(track.proj.sp)

  ## Build the plot
  my_xmin <- min(track.proj.df$coords.x1) - 10000
  my_xmax <- max(track.proj.df$coords.x1) + 10000
  my_ymin <- min(track.proj.df$coords.x2) - 10000
  my_ymax <- max(track.proj.df$coords.x2) + 10000
    

  example_gps.plot <- ggplot(data = world2) + 
      geom_sf(fill = "cadetblue", colour = "grey") +
      coord_sf(crs = proj.utm, xlim = c(my_xmin, my_xmax), ylim = c(my_ymin, my_ymax), 
               label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
      geom_path(aes(x = coords.x1, y = coords.x2, colour = ring), size = 1, dat = track.proj.df) +
      scale_colour_viridis_d(name = NULL) +
      labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bbal", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "cro", "Crozet", "Kerguelen"))))) +
      theme(panel.background = element_rect(fill = "white"), 
            panel.grid.major = element_line(colour = "grey80"),
            panel.border = element_rect(colour = "black", fill = NA),
            axis.text.x.top = element_blank(), 
            axis.title.x.top = element_blank(),
            axis.ticks.x.top = element_blank(),
            axis.title.x.bottom = element_blank(),
            axis.text.y.right = element_blank(), 
            axis.title.y.right = element_blank(),
            axis.title.y.left = element_blank(),
            legend.position = "none", 
            plot.caption = element_text(face = "italic")) 
  
  assign(paste0(spec_col[i], "_example_gps.plot"), example_gps.plot)
  plot_name <- paste0(spec_col[i], "_gps.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_example_gps_plot.RData"))
  
}

#png("Figures/MS Figures/FIGS3_spatial_repeatability_example_tracks.png", width = 10, height = 10, units = "in", res = 700)
ggarrange(bba_kerguelen_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
#dev.off()


```

## Bearings

### Prep data

```{r prep_bearings_data, eval = FALSE}

# Set parameters
spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

# Make list to collect bearings data
trip_bearings.list <- list()
gps_with_bearings.list <- list()

for (i in 1:length(spec_col)) {
  
  if (grepl("ker", spec_col[i])) {
    col_coords <- ker_coords
  } else if (grepl("birdis", spec_col[i])) {
    col_coords <- bi_coords
  } else if (grepl("cro", spec_col[i])) {
    col_coords <- cro_coords }
  
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
 
  my_gps %<>%
    mutate(ID = paste0(ring, "_", season, "_", boutID),
           season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                           as.numeric(format(as.Date(datetime), "%Y")),
                           as.numeric(format(as.Date(datetime), "%Y")) + 1 )) %>%
    filter(loc == "trip")
  
  ## Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>% mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))

  ## Remove colony periods & any very short trips
  my_gps %<>% filter(loc == "trip")
  my_gps %<>% group_by(tripID) %>% mutate(n_locs = n()) %>% filter(n_locs > 6) %>% select(-n_locs) %>% ungroup()
  
  ## Remove birds with only one trip
  my_gps %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% select(-n_trips) %>% ungroup()
  
  # Calculate average bearing from colony ------------
   
  trip_bearings <- my_gps %>% 
     group_by(ID) %>%
     arrange(datetime) %>%
    ## Bearing from colony to each fix
     mutate(bearing_col = bearing(
                p1 = matrix(c(col_coords$longitude, col_coords$latitude), ncol = 2, byrow = TRUE),
                p2 = matrix(c(longitude, latitude), ncol = 2)),
    ## Mean bearing across the whole trip
            bearing_avg = median.circular(circular(bearing_col, units = "degrees", template = "geographics"), na.rm = T),
            species = strsplit(spec_col[i], "_")[[1]][1],
            colony = strsplit(spec_col[i], "_")[[1]][2]) %>%
    ungroup()
  
  ## Extract complete info for plotting
  gps_with_bearings.list[[i]] <- trip_bearings

  # Isolate bearing per trip & calculate inter-trip distance
  trip_bearings %<>%
    mutate(boutID = paste0(season, "_", boutID)) %>%
    select(c(ID, ring, season, phase, boutID, bearing_avg, datetime)) %>%
    group_by(ring,season, boutID) %>%
    slice(1) %>%
    group_by(ring, season) %>%
    mutate(time_since_first = as.numeric(difftime(datetime, min(datetime), units = "days"))) %>%
    ungroup() %>%
    select(c(ID, ring, season, phase, boutID, bearing_avg, time_since_first)) %>%
    distinct()
  
  trip_bearings$species <- strsplit(spec_col[i], "_")[[1]][1]
  trip_bearings$colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  trip_bearings.list[[i]] <- trip_bearings

}

bearings.df <- do.call("rbind", trip_bearings.list)
save(bearings.df, file = "Data_outputs/bearings_data.RData")

gps_with_bearings <- do.call("rbind", gps_with_bearings.list)
save(gps_with_bearings, file = "Data_outputs/all_gps_with_bearings.RData")

```

### Plot bearings from the colony

```{r plot_bearings, fig.dim=c(16, 4.5)}

load("Data_outputs/bearings_data.RData")

bearings.df %<>% mutate(bearing_360 = ifelse(bearing_avg < 0, bearing_avg + 360, bearing_avg),
                        spec_col = case_when(
                            species == "bbal" & colony == "birdis" ~ "Black-browed albatrosses at Bird Island",
                            species == "bbal" & colony == "ker" ~ "Black-browed albatrosses at Kerguelen",
                            species == "waal" & colony == "birdis" ~ "Wandering albatrosses at Bird Island",
                            species == "waal" & colony == "cro" ~ "Wandering albatrosses at Crozet"))
breaks = seq(0, 360, 60)

#png(filename = "Figures/MS Figures/FIGSX_bearings_plot.png", width = 16, height = 4.5, units = "in", res = 900)
ggplot(bearings.df, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks, #c(0, 90, 180, 270), # breaks,
                     labels = paste0(breaks, "\u00B0")) +#c("N", "E", "S", "W")) + # paste0(breaks, "\u00B0")) +
  labs(y = "Frequency") +
  #facet_grid(spec_col ~ ., scales="free") +
  ggh4x::facet_grid2(. ~ spec_col, scales = "free_y", independent = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))
#dev.off()

  
```

### Estimate repeatability

```{r rpt_bearing_differences}

load("Data_outputs/bearings_data.RData")

bearings.df %<>% mutate(bearing_360 = ifelse(bearing_avg < 0, bearing_avg + 360, bearing_avg),
                        spec_col = case_when(
                            species == "bbal" & colony == "birdis" ~ "bbal_birdis",
                            species == "bbal" & colony == "ker" ~ "bbal_ker",
                            species == "waal" & colony == "birdis" ~ "waal_birdis",
                            species == "waal" & colony == "cro" ~ "waal_cro"))

# Convert degrees to radians
bearings.df %<>%
  mutate(bearing_rad = bearing_360 * pi / 180,
         bearing_rad = ifelse(bearing_rad >= pi, bearing_rad - 2*pi, bearing_rad),
         bearing_rad = ifelse(bearing_rad <= -pi, bearing_rad + 2*pi, bearing_rad))
  
## Add small value to avoid hitting pi exactly
eps <- 1e-6
bearings.df$bearing_rad <- pmin(pmax(bearings.df$bearing_rad, -pi + eps), pi - eps)

# Add parameters for model - ring season, time since first
bearings.df %<>% mutate(ring_season = paste0(ring, "_", season))

## Define some weakly informative priors
bearing_priors <- c(
  prior(normal(0, 1), class = "Intercept", dpar = "mu"),      # mu intercept
  prior(normal(0, 0.5), class = "b", dpar = "mu"),            # fixed effects
  prior(exponential(1), class = "sd", dpar = "mu"),           # group-level SDs
  prior(gamma(2, 0.01), class = "kappa")  # kappa intercept
)

spec_cols <- unique(bearings.df$spec_col)

bearing_posteriors.list <- list()
options(brms.backend = "cmdstanr")   

for (i in 1:length(spec_cols)) {
  
  print(paste0("Processing ", i, "..."))
  
  my_df <- subset(bearings.df, spec_col == spec_cols[i])
  
  # Set formula
  if(spec_col[i] == "waal_birdis") {
  bearing_formula <- bf("bearing_rad ~ phase + (1 | season) + (1 | ring_season)") } else {
    bearing_formula <- bf("bearing_rad ~ phase + (1 | season) + (1 | ring ) + (1 | ring_season)")  }
  
  # Fit model 
  brm_vm <- brm(
    formula = bearing_formula,
    family  = von_mises(),  
    data    = my_df,
    prior   = bearing_priors,
    chains = 4, cores = 4, iter = 8000, warmup = 4000,
    control = list(max_treedepth = 15, adapt_delta = 0.99) )

  # Calc repeatability
  if(spec_col[i] == "waal_birdis") {
    post_df <- calc_rpt_brms_VM_waBI(brm_vm) } else {
      post_df <- calc_rpt_brms_VM(brm_vm)
    }
  
  post_df %<>% mutate(spec_col = spec_col[i],
                      species = my_df$species[1],
                      colony = my_df$colony[1])
  
  mod_name <- paste0(spec_col[i], ".brm_rpt_bearings")
  assign(mod_name, brm_vm)
  
  save(list = mod_name, file = paste0("Data_outputs/", spec_col[i], "_brms_rpt_bearings.RData"))
  
  bearing_posteriors.list[[i]] <- post_df
  
}

bearing_posteriors.df <- do.call("rbind", bearing_posteriors.list)
save(bearing_posteriors.df, file = "Data_outputs/bearing_repeatability_results.RData")

```

#### Plot bearing repeatability

```{r plot_bearings_diffs, fig.dim=c(6,6)}

load("Data_outputs/bearing_repeatability_results.RData")

# Recode spec_col
bearing_posteriors.df %<>% mutate(spec_col = recode(spec_col,
   "bbal_birdis" = "BBAL Bird Island",
    "bbal_ker" = "BBAL Kerguelen",
    "waal_birdis" = "WAAL Bird Island",
    "waal_cro" = "WAAL Crozet"))

# Pivot data longer for faceted plotting
bearings_rpt.long <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") 

# Get median and 95% credible interval 
bearings_rpt.summary <- bearings_rpt.long %>%
  group_by(repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE)) %>%
  arrange(spec_col) %>%
  mutate(spec_col = factor(spec_col, levels = rev(unique(spec_col))))


bearings_rpt.plot <- ggplot(bearings_rpt.summary, aes(x = median_val, y = spec_col, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  scale_colour_manual( values = c("rpt_within" = col_within, "rpt_across" = col_between),
                       labels = c("Within-season", "Between-season") ) +
  labs(x = "Repeatability in foraging bearing (median, 95% CI)", y = "") +
  theme_bw() +
  theme(legend.position = "none")


# Repeatability estimates

# BBAL - Within
bearings_rpt.summary %>% filter(repeatability_type == "rpt_within" & spec_col == "BBAL Bird Island") %>% ungroup
bearings_rpt.summary %>% filter(repeatability_type == "rpt_within" & spec_col == "BBAL Kerguelen") %>% ungroup 

# BBAL - Within
bearings_rpt.summary %>% filter(repeatability_type == "rpt_across" & spec_col == "BBAL Bird Island") %>% ungroup
bearings_rpt.summary %>% filter(repeatability_type == "rpt_across" & spec_col == "BBAL Kerguelen") %>% ungroup 

# WAAL
bearings_rpt.summary %>% filter(spec_col == "WAAL Bird Island") %>% ungroup
bearings_rpt.summary %>% filter(spec_col == "WAAL Crozet") %>% ungroup 

```



### Heritability of bearings

```{r bearings_heritability}

load("Data_outputs/bearings_data.RData")

## Get trip ID
bearings.df %<>%
  separate(boutID, into = c("season", "trip1", "trip2"), sep = "_") %>%
  mutate(tripID = paste0(ring, "_", season, "_", trip1, trip2)) %>%
  ungroup() %>%
  select(-c(ID, trip1, trip2))

## Set weakly informative priors
bearing_priors <- c(
  prior(student_t(3, 0, 1), class = "Intercept", resp = "cosb"),
  prior(student_t(3, -0.5, 1), class = "Intercept", resp = "sinb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "ring", resp = "cosb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "animal", resp = "cosb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "ring", resp = "sinb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "animal", resp = "sinb"),
  prior(student_t(3, 0, 0.8), class = "sigma", resp = "cosb"),
  prior(student_t(3, 0, 0.8), class = "sigma", resp = "sinb")
)

# Loop through species and colonies
spec_col <- c("bbal_birdis", "waal_birdis", "waal_cro")
posteriors = list()

for (i in 1:length(spec_col)) {
  
  my_species <- strsplit(spec_col[i], "_")[[1]][1]
  my_colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  focal.bearings <- bearings.df %>% filter(species == my_species & colony == my_colony)
  focal.bearings %<>% mutate(animal = ring, ring_pe = ring, ringSeason = paste0(ring, "_", season))

  ## Convert to bearings to radians and create cos/sin
  focal.bearings$bearing_rad <- (focal.bearings$bearing_avg %% 360) * pi / 180
  focal.bearings$cos_b <- cos(focal.bearings$bearing_rad)
  focal.bearings$sin_b <- sin(focal.bearings$bearing_rad)
  
  ## Restrict to birds with > 1 trip
  focal.bearings %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% select(-n_trips)
  
  # Build the pedigree ====
  
  pedigree.df <- read.csv(paste0("Data_inputs/", spec_col[i], "_pedigree.csv"))[,1:3]
  pedigree.df[,c(1:3)] <- lapply(pedigree.df[,c(1:3)], as.character)
  
  Amat <- make_Amat(focal.bearings, pedigree.df)
  
  # Build the model ====
  
  ## Set formula
  if(spec_col[i] == "waal_birdis") {
  mv_formula <- bf(mvbind(cos_b, sin_b) ~ 1 +  (1 | gr(animal, cov = Amat)) + (1 | ring) + (1 | season)) } else {
    mv_formula <- bf(mvbind(cos_b, sin_b) ~ 1 + phase + (1 | gr(animal, cov = Amat)) + (1 | ring) + (1 | ringSeason) + (1 | season))  }

  brm_bearing_heritability <- brm(
    formula = mv_formula,
    prior = bearing_priors,
    data = focal.bearings,
    data2 = list(Amat = Amat),
    family = gaussian(),
    chains = 4, cores = 4, iter = 8000, warmup = 4000,
    control = list(max_treedepth = 15, adapt_delta = 0.99))

  #plot(brm_bearing_heritability)  
  save(brm_bearing_heritability, file = paste0("Data_outputs/", spec_col[i], "_bearing_heritability.RData"))
 
  # Calculate h2 ====
  draws <- as_draws_df(brm_bearing_heritability)
  
  ## Get variance components
  VA <- extractVar(draws, "animal")
  Vring <- extractVar(draws, "ring")
  if(spec_col[i] != "waal_birdis") { VringSeason <- extractVar(draws, "ringSeason") }
  Vseason <- extractVar(draws, "season")
  VR <- draws$sigma_cosb^2 + draws$sigma_sinb^2

  ## Calculate total phenotypic variance 
  if(spec_col[i] == "waal_birdis") { VP <- VA + Vring + Vseason + VR } else {
  VP <- VA + Vring + VR }

  ## Calculate h2
  h2 <- VA / VP

  posteriors[[i]] <- data.frame(species = my_species,
                           colony = my_colony,
                           h2 = h2)
  
}

posteriors_all <- do_call("rbind", posteriors)

posteriors_all %<>%
  mutate(spec_col = case_when(
      species == "bbal" ~ "BBAL Bird Island",
      species == "waal" & colony == "birdis" ~ "WAAL Bird Island",
      TRUE ~ "WAAL Crozet"),
      spec_col = factor(spec_col, levels = rev(unique(spec_col))))

save(posteriors_all, file = "Data_outputs/all_posteriors_spatial_heritability.RData")

```

#### Plot heritability of bearings

```{r spatial_heritability}

load("Data_outputs/all_posteriors_spatial_heritability.RData")

png(filename = "Figures/spatial_heritability.png", width = 6, height = 6, units = "in", res = 900)
ggplot(posteriors_all, aes(x = h2, y = spec_col)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw() 
dev.off()

```


## Bhattacharyya's Affinity

```{r prep_BA_data, eval = FALSE}

# Set parameters
spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

for (i in 1:length(spec_col)) {
  
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
 
  # Remove colony periods & any very short trips
  my_gps %<>% filter(loc == "trip")
  my_gps %<>% group_by(tripID) %>% mutate(n_locs = n(), max_dist = max(dist_col)/1000) %>% 
    filter(n_locs > 20 & max_dist > 65) %>% select(-c(n_locs, max_dist)) %>% ungroup()
  
  ## Remove birds with only one trip
  my_gps %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% select(-n_trips) %>% ungroup()
  
  ## Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>% mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))
  
  # Get start times look-up
  trip_starts <- my_gps %>%
    mutate(tripID = paste(season, "_", boutID)) %>%
    group_by(ring, tripID) %>%
    summarise(start_time = min(datetime), .groups = "drop")
  
  ### Meta for phase comparisons later
  my_meta <- my_gps %>%
    group_by(ID) %>%
    select(c(ring, season, boutID, ID, phase)) %>%
    distinct()

  # Generate all possible trip combinations 
  trip_combinations <- combn(unique(my_gps$ID), 2, simplify = FALSE)
  trip_combinations.df <- data.frame(do.call("rbind", trip_combinations))
  colnames(trip_combinations.df) <- c("tripID_1", "tripID_2")

  # Plotting out some UDs ----
  
  # birds <- unique(my_gps$ring)
  # 
  # for (b in 1:length(birds)) {
  # 
  #   # Get focal bird
  #   ud.df_A <- subset(my_gps, ring == birds[b]) %>% select(tripID, longitude, latitude)
  #   ud.sp_A <- SpatialPointsDataFrame(coords = ud.df_A[, c("longitude", "latitude")],
  #                                data = data.frame(id = ud.df_A$tripID))
  #   ud.est_A <- kernelUD(ud.sp_A[, 1], h = "href", extent = 3, grid = 500)
  #   ud.95_A <- getverticeshr(ud.est_A, percent = 95)
  # 
  #   ud_sf_A <- st_as_sf(ud.95_A)
  #   ud_sf_A %<>% separate(id, sep = "_", into = c("ring", "season", "dummy", "tripNo")) %>% select(-dummy)
  # 
  #   # Random bird
  #   ud.df_B <- subset(my_gps, ring == birds[sample(length(birds), 1)]) %>% select(tripID, longitude, latitude)
  #   ud.sp_B <- SpatialPointsDataFrame(coords = ud.df_B[, c("longitude", "latitude")],
  #                                data = data.frame(id = ud.df_B$tripID))
  #   ud.est_B <- kernelUD(ud.sp_B[, 1], h = "href", extent = 3, grid = 500)
  #   ud.95_B <- getverticeshr(ud.est_B, percent = 95)
  # 
  #   ud_sf_B <- st_as_sf(ud.95_B)
  #   ud_sf_B %<>% separate(id, sep = "_", into = c("ring", "season", "dummy", "tripNo")) %>% select(-dummy)
  # 
  #   ud_plot <- ggplot() +
  #     geom_sf(data = ud_sf_A, fill = "blue", alpha = 0.4) +
  #     geom_sf(data = ud_sf_B, fill = "red", alpha = 0.4) +
  #     theme_bw()
  # 
  #   png(file = paste0("Figures/Workshop/", spec_col[i], "_", birds[b], "_", ud_sf_B$ring[1], ".png"))
  #   print(ud_plot)
  #   dev.off()
  # }
  
  # ===# WITHIN INDIVIUDAL ====
  
  # Find the combinations by filtering to same rings and same phase
  ba_within.df <- trip_combinations.df %>%
     mutate(ring_1 = sub("_.*", "", tripID_1),
           ring_2 = sub("_.*", "", tripID_2)) %>%
    filter(ring_1 == ring_2) %>%
    select(-c(ring_1, ring_2)) %>%
    # Ensure comparisons are within same phase
    left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_1 = ID),
            by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID),
              by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2))

  ## Calculate the comparisons (slow!)
  ba_within.list <- vector(mode = "list", length = nrow(ba_within.df))
  pb <- txtProgressBar(min = 0, max = nrow(ba_within.df), style = 3)
  
  for (j in 1:nrow(ba_within.df)) {
    
    setTxtProgressBar(pb, j)
    
    # Subset data for both trips
    trip1 <- subset(my_gps, ID == ba_within.df$tripID_1[j])
    trip1.stt <- min(trip1$datetime)
    trip2 <- subset(my_gps, ID == ba_within.df$tripID_2[j])
    trip2.stt <- min(trip2$datetime)
    
    # Combine trips and factorize ID
    both_trips <- rbind(trip1, trip2) 
    both_trips %<>% 
      mutate(id = as.factor(both_trips$tripID)) %>%
      select(c(longitude, latitude, id))
    
    # Convert to SpatialPointsDataFrame
    coordinates(both_trips) <- ~longitude + latitude
    proj4string(both_trips) <- CRS(proj.dec)
    
    # Get bounding box
    bb <- bbox(both_trips)

    ## Define a raster grid extent slightly larger than the bounding box
    xrange <- bb[1, 2] - bb[1, 1]
    yrange <- bb[2, 2] - bb[2, 1]

    ## Expand the bounding box slightly (e.g. 20% buffer)
    bb_expanded <- bb
    bb_expanded[1, ] <- bb[1, ] + c(-0.1 * xrange, 0.1 * xrange)
    bb_expanded[2, ] <- bb[2, ] + c(-0.1 * yrange, 0.1 * yrange)

    # Get the UDs
    ud <- kernelUD(both_trips, h = "href", grid = 100, extent = 0, kern = "bivnorm", same4all = TRUE)

    # Calculate UD overlap — Bhattacharyya’s affinity
    overlap_val <- kerneloverlaphr(ud, method = "BA", percent = 95)[1,2]

    # Output
    output <- data.frame(trip1 = ba_within.df$tripID_1[j],
                         trip2 = ba_within.df$tripID_2[j],
                         ba_val = overlap_val,
                         inter_obs_dist = abs(as.numeric(difftime(trip1.stt, trip2.stt, units = "days"))))
  
    ba_within.list[[j]] <- output 
    
  }
  
  close(pb)

  ## Bind up the output
  ba_within.output <- do.call("rbind", ba_within.list)
  
  ba_within.output %<>% 
    separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
    separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
    select(-loc) %>%
    mutate(tripID_1 = paste0("trip_", tripID_1),
           tripID_2 = paste0("trip_", tripID_2),
           comp_ind = "within_ring",
           comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))

  rm(ba_within.df); rm(ba_within.list)
  gc()
  
  # ===# BETWEEN INDIVIUDAL ====
  
  set.seed(817)
  
  # Downsample the between dataset to match within 
  n_within_between = nrow(ba_within.output %>% filter(comp_season == "between_season"))
  n_within_within = nrow(ba_within.output %>% filter(comp_season == "within_season"))
  
  ba_between_between.df <- trip_combinations.df %>%
    separate(tripID_1, into = c("ring1", "season1", "dummy1", "trip1"), remove = FALSE) %>%
    separate(tripID_2, into = c("ring2", "season2", "dummy2", "trip2"), remove = FALSE) %>%
    filter(ring1 != ring2) %>%
    filter(season1 != season2) %>%
    select(c(tripID_1, tripID_2)) %>%
    # Get phase and filter for same breeding phase
    left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_1 = ID), by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID), by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2)) %>%
    sample_n(size = n_within_between, replace = FALSE)
  
  
  ba_between_within.df <- trip_combinations.df %>%
    separate(tripID_1, into = c("ring1", "season1", "dummy1", "trip1"), remove = FALSE) %>%
    separate(tripID_2, into = c("ring2", "season2", "dummy2", "trip2"), remove = FALSE) %>%
    filter(ring1 != ring2) %>%
    filter(season1 == season2) %>%
    select(c(tripID_1, tripID_2)) %>%
    # Get phase and filter for same breeding phase
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_1 = ID), by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID), by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2)) %>%
    sample_n(size = n_within_within, replace = FALSE)

  ba_between <- rbind(ba_between_between.df, ba_between_within.df)
  
  ## Calculate the comparisons
  ba_between.list <- vector(mode = "list", length = nrow(ba_between))
  pb <- txtProgressBar(min = 0, max = nrow(ba_between), style = 3)
  
  for (k in 1:nrow(ba_between)) {
    
    setTxtProgressBar(pb, k)
    
    # Subset data for both trips
    trip1 <- subset(my_gps, ID == ba_between$tripID_1[k])
    trip1.stt <- min(trip1$datetime)
    trip2 <- subset(my_gps, ID == ba_between$tripID_2[k])
    trip2.stt <- min(trip2$datetime) 
    
    # Combine trips and factorize ID
    both_trips <- rbind(trip1, trip2) 
    both_trips %<>% 
      mutate(id = as.factor(both_trips$tripID)) %>%
      select(c(longitude, latitude, id))
    
    # Convert to SpatialPointsDataFrame
    coordinates(both_trips) <- ~longitude + latitude
    proj4string(both_trips) <- CRS(proj.dec)
    
    # Get bounding box
    bb <- bbox(both_trips)

    ## Define a raster grid extent slightly larger than the bounding box
    xrange <- bb[1, 2] - bb[1, 1]
    yrange <- bb[2, 2] - bb[2, 1]

    ## Expand the bounding box slightly (e.g. 20% buffer)
    bb_expanded <- bb
    bb_expanded[1, ] <- bb[1, ] + c(-0.1 * xrange, 0.1 * xrange)
    bb_expanded[2, ] <- bb[2, ] + c(-0.1 * yrange, 0.1 * yrange)

    # Get the UDs
    ud <- kernelUD(both_trips, h = "href", grid = 100, extent = 0, kern = "bivnorm", same4all = TRUE)

    # Calculate UD overlap — Bhattacharyya’s affinity
    overlap_val <- kerneloverlaphr(ud, method = "BA", percent = 95)[1,2]

    # Output
    output <- data.frame(trip1 = ba_between$tripID_1[k],
                         trip2 = ba_between$tripID_2[k],
                         ba_val = overlap_val,
                         inter_obs_dist = abs(as.numeric(difftime(trip1.stt, trip2.stt, units = "days"))))
  
    ba_between.list[[k]] <- output 
    
  }
  
  close(pb)

  ba_between.output <- do.call("rbind", ba_between.list)
  
  ba_between.output %<>% 
    separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
    separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
    select(-loc) %>%
    mutate(tripID_1 = paste0("trip_", tripID_1),
           tripID_2 = paste0("trip_", tripID_2),
           comp_ind = "between_ring",
           comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))
  
  # Bind the two dataframes together
  ba.df <- rbind(ba_within.output, ba_between.output)
  
  save(ba.df, file = paste0("Data_outputs/", species, "_", colony, "_ba_comparison.RData"))
  
  rm(ba.df); rm(ba_within.output); rm(ba_between.list); rm(ba_between.output)
  gc()

}
  
```

### Modelling BA differences

```{r model_ba_differences, eval = FALSE}

# Species-colony combinations
spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")

# Output containers
ba_summary.list <- list()
ba_data.list <- list()
ba_means.list <- list()

for (i in seq_along(spec_col)) {
  
  # Load data and prep species/colony labels
  load(paste0("Data_outputs/", spec_col[i], "_ba_comparison.RData"))
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  meta_info <- tibble(species, colony)
  
  # Make dyad ID and subsample to ensure each comparison appears only once (circumvents random effects issues)
  set.seed(817) 
  
  ba.df %<>%
    mutate(ring_season_1 = paste(ring_1, season_1, sep = "_"),
           ring_season_2 = paste(ring_2, season_2, sep = "_"),
           dyad = if_else(ring_season_1 < ring_season_2,
                          paste(ring_season_1, ring_season_2, sep = "_"),
                          paste(ring_season_2, ring_season_1, sep = "_"))) %>%
    group_by(dyad) %>%
    slice_sample(n = 1) %>%
    ungroup() %>%
    select(-c(ring_season_1, ring_season_2))
  
  ba_data.list[[i]] <- ba.df %>%
    mutate(species = species, colony = colony)
  
  ## Relevel factors for comparison
  ba.df %<>% mutate(comp_ind = as.factor(comp_ind), comp_season = as.factor(comp_season))
  
  ba.df$comp_ind <- relevel(ba.df$comp_ind, ref = "within_ring")
  ba.df$comp_season <- relevel(ba.df$comp_season, ref = "within_season")

  ## Add small constant to handle 0 values
  n <- nrow(ba.df)
  ba.df$ba_val <- (ba.df$ba_val * (n - 1) + 0.5) / n
  
  ## Remove within ind, within season component for WAAL Bird IS
  if(i == 3) { 
    ba.df %<>% filter(!(comp_ind == "within_ring" & comp_season == "between_season"))
    ba.df <- droplevels(ba.df)
    
    m1 <- betareg(ba_val ~ comp_ind + comp_season + inter_obs_dist, data = ba.df) } else {
        m1 <- betareg(ba_val ~ comp_ind * comp_season + inter_obs_dist, data = ba.df)
    }
 
  ## Diagnostics
  # plot(m1, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
  # qqnorm(resid(m1))
  # qqline(resid(m1))
  # hist(resid(m1), breaks = 50, main = "Histogram of residuals")

  ## Extract outputs
  m1_em <- data.frame(emmeans(m1, ~comp_ind*comp_season, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring"))) %>%
    relocate(c(species, colony), .before = comp_ind) %>%
    arrange(comp_ind)
  
  m1_em.comp_ind <- data.frame(emmeans(m1, ~comp_ind, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring")),
           comp_season = NA) %>%
    relocate(c(species, colony), .before = comp_ind) %>%
    arrange(comp_ind)
  
  m1_em.comp_season <- data.frame(emmeans(m1, ~comp_season, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = NA) %>%
    relocate(c(species, colony), .before = comp_season) %>%
    arrange(comp_season)

  m1_em <- rbind(m1_em, m1_em.comp_ind, m1_em.comp_season)
  m1_out <- extract_output.lm(m1) %>% mutate(species = species, colony = colony) %>% relocate(c(species, colony), .before = Term)
  
  # Store results
  ba_summary.list[[i]] <- m1_out
  ba_means.list[[i]] <- m1_em
  
}

# Combine and save
ba_summary <- bind_rows(ba_summary.list)
ba_data    <- bind_rows(ba_data.list)
ba_means   <- bind_rows(ba_means.list)

save(ba_summary, file = "Data_outputs/ba_model_summary.RData")
save(ba_data,    file = "Data_outputs/ba_data_combined.RData")
save(ba_means,   file = "Data_outputs/ba_means.RData")

```

### Plot BA differences

```{r plot_ba_diffs, fig.dim=c(6,6)}

load("Data_outputs/ba_means.RData")
ba_means %<>% filter(!is.na(comp_season)) %>% filter(!is.na(comp_ind))

## Drop within ind, between seasons for WAAL Bird Island
ba_means %<>% filter(!(species == "waal" & colony == "birdis" & comp_ind == "within_ring" & comp_season == "between_season"))

# Data wrangling for figure
ba_plot <- ba_means %>%
  mutate(species = case_when(
          species == "bbal" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "ker" ~ "Kerguelen",
      colony == "cro" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = " ") ) %>%
  rename(lower.CL = asymp.LCL,
         upper.CL = asymp.UCL)

png(filename = "Figures/spatial_repeatability_ba_comparison.png", width = 6, height = 6, units = "in", res = 900)
ggplot(ba_plot, aes(x = comp_ind, y = emmean, colour = comp_season)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1) +
  geom_line(aes(group = comp_season)) +
  labs(x = "", colour = "", y = "Bhattacharyya’s affinity") +
  facet_wrap(~spec_col) +
  scale_colour_manual(
    values = c(col_within, col_between),
    labels = c("Within season", "Across seasons") ) +
  scale_x_discrete(labels = c("Within individual", "Among individual")) +
  theme_bw() +
  theme(legend.position = c(0.36, 0.97),
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_rect(fill = "transparent", colour = NA))
dev.off()


```

```{r BA_similarity_table, echo = FALSE}

load("Data_outputs/ba_model_summary.RData")

# Reformat the table
ba_summary %<>%
 mutate(Species = toupper(species),
        Colony = ifelse(colony == "birdis", "Bird Island",
                         ifelse(colony == "cro", "Crozet",
                                "Kerguelen")), 
        Term = case_when(
            Term == "(Intercept)"  ~ "Intercept",
            Term == "comp_seasonbetween_season" ~ "Across seasons",
            Term == "comp_indbetween_ring" ~ "Among individuals",
            Term == "comp_indbetween_ring:comp_seasonbetween_season" ~ "Among individuals * Across seasons",
            TRUE ~ "Inter-observation distance"),
         Estimate = paste0(signif(Estimate, 4), " \u00B1 ", signif(Std_Error, 4)),
         t_value = signif(t_value, 4),
         p_value = ifelse(p_value < 0.0001, "0", format(round(p_value, 4), nsmall = 4, scientific = FALSE))) %>%
  mutate(Term = factor(Term, levels = c(
    "Intercept",
    "Among individuals",
    "Across seasons",
    "Among individuals * Across seasons",
    "Inter-observation distance"))) %>%
  rename('t value' = t_value, 'p value' = p_value) %>%
  select(-c(species, colony, Std_Error)) %>%
  relocate(c(Species, Colony), .before = Term) %>%
  arrange(Species, Colony, Term) %>%
  mutate(CI = paste0("[", signif(CI_lower, 4), ", ", signif(CI_upper, 4), "]")) %>%
  relocate(CI, .before = "t value") %>%
  select(-c(CI_lower, CI_upper))
         
knitr::kable(ba_summary, format = "html", 
             caption = '<span style="font-size: 12px;">Table 3. Model outputs for BA comparisons. Bhattacharyyas Affinity measures overlap in two utilisation distributions.".</span>', digits = 3) %>%
  row_spec(which(ba_summary$`p value` < 0.05 & ba_summary$Term != "Intercept"), bold = TRUE) %>%
  collapse_rows(columns = c(1, 2)) %>%
  kable_styling(font_size = 9)


```

#### Figure 2: Bearing and space use differences

This plot shows the repeatability results from the bearings analysis, and the BA comparisons for the spatial overlap analysis. 

```{r fig2_spatial_results, fig.dim=c(6,6)}

# Bearing repeatabilities ====

load("Data_outputs/bearing_repeatability_results.RData")

# Pivot data longer for faceted plotting
bearings_rpt.long <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") 

# Get median and 95% credible interval 
bearings_rpt.summary <- bearings_rpt.long %>%
  group_by(repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) )

## Relevel factors
bearings_rpt.summary %<>%
  mutate(spec_col = recode(spec_col,
      bbal_birdis = "BBAL Bird Island",
      bbal_ker = "BBAL Kerguelen",
      waal_birdis = "WAAL Bird Island",
      waal_cro = "WAAL Crozet")) %>%
  arrange(spec_col) %>%
  mutate(spec_col = factor(spec_col, levels = rev(unique(spec_col))))

bearings_rpt.plot <- ggplot(bearings_rpt.summary, aes(x = median_val, y = spec_col, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  labs(x = "Repeatability in foraging bearing (median, 95% CI)", y = "") +
  theme_bw() +
  theme(legend.position = "none")
  
# BA comparisons ====

load("Data_outputs/ba_means.RData")
ba_means %<>% filter(!is.na(comp_season)) %>% filter(!is.na(comp_ind))

## Drop within ind, between seasons for WAAL Bird Island
ba_means %<>% filter(!(species == "waal" & colony == "birdis" & comp_ind == "within_ring" & comp_season == "between_season"))

## Data wrangling
ba_plot.df <- ba_means %>%
  mutate(species = case_when(
          species == "bbal" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "ker" ~ "Kerguelen",
      colony == "cro" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = " ") ) %>%
  rename(lower.CL = asymp.LCL,
         upper.CL = asymp.UCL) %>%
  mutate(behaviour = "overlap",
         behav_comp = paste(behaviour, comp_ind, sep = "_"),
         behav_comp = factor(behav_comp,
                             levels = c(
                               "overlap_between_ring",
                               "overlap_within_ring",
                               "bearings_between_ring",
                               "bearings_within_ring"  )))

ba_plot <- ggplot(ba_plot.df, aes(x = emmean, y = comp_ind, colour = comp_season)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = lower.CL, xmax = upper.CL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Bhattacharyya's Affinity") +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(
    values = c(col_within, col_between)) +
  scale_y_discrete(labels = c("Within individual", "Among individuals")) +
  theme_bw() +
  theme(legend.position = "none")

png(filename = "Figures/MS Figures/FIG2_spatial_results.png", width = 15, height = 7.5, units = "in", res = 900)
ggarrange(bearings_rpt.plot, ba_plot, ncol = 2)
dev.off()

```


## Supplementary analyses

#### GPS tracks by sex

Here, I've plotted the GPS tracks of all individuals across the colonies, coloured by sex (female = purple, male = yellow).

```{r process_gps_by_sex, eval = FALSE}

load("Data_outputs/all_gps_data.RData")
load("Data_inputs/ind_meta.RData")

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# Loop through data
spec_col <- c("bba_kerguelen", "bba_birdis", "waal_crozet", "waal_birdis")

for (i in 1:length(spec_col)) {
  
  # Set projections (use same for Ker + Cro)
  if(grepl("birdis", spec_col[i])) {
      proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
  } else {
      proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
    }

  world2 <- sf::st_transform(world, crs = proj.utm)
  
  # Project GPS data
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  my_gps.proj <- all_gps.df %>% filter(species == my_species & colony == my_colony) %>% arrange(ringYr, datetime)
  coordinates(my_gps.proj) <- ~ longitude+latitude
  proj4string(my_gps.proj) <- proj.dec
  
  my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
  my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

  # Project the colony
  col_coords <- data.frame(longitude = my_gps.proj$col_lon[1], latitude = my_gps.proj$col_lat[1])
  
  if(grepl("ker", spec_col[i])) { col_coords <- data.frame(longitude = 70.2333, latitude = -49.6833) }
    
  coordinates(col_coords) <- ~longitude+latitude
  proj4string(col_coords) <- proj.dec
  colony.sp <- spTransform(col_coords, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)

  
  # Merge in sex info
  my_gps.proj.df <- merge(my_gps.proj.df, all_ind.meta %>% select(-birthYr) %>% mutate(species = tolower(species)), by = c("species", "ring"), all.x = T) %>% filter(!is.na(sex))
  
  # Build the plot
  
  ## Set min/max limits
  y_min = min(my_gps.proj.df$coords.x2 - 100000)
  y_max = max(my_gps.proj.df$coords.x2 + 100000)
  x_min = min(my_gps.proj.df$coords.x1 - 100000)
  x_max = max(my_gps.proj.df$coords.x1 + 100000)
   
  gps_tracks.plot <- ggplot(data = world2) + 
    geom_sf(fill = "cadetblue", colour = "cadetblue") +
    coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
             label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
    geom_path(aes(x = coords.x1 , y = coords.x2, col = sex), size = 0.8,
              dat = my_gps.proj.df, group = "identity") +
    scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
    scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
    scale_colour_viridis_d() +
    annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
    labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bba", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "crozet", "Crozet", "Kerguelen"))))) +
    theme(panel.background = element_rect(fill = "white"),
          panel.grid.major = element_line(colour = "grey80"),
          panel.border = element_rect(colour = "black", fill = NA),
          axis.text.x.bottom = element_blank(),
          axis.title.x.bottom = element_blank(),
          axis.text.y.right = element_blank(),
          axis.title.y.right = element_blank(),
          axis.title.y.left = element_blank(),
          axis.text.y.left = element_text(size = 10),
          axis.text.x.top = element_text(size = 10),
          legend.position = "none")

  
  assign(paste0(spec_col[i], "_gps_bySex.plot"), gps_tracks.plot)
  plot_name <- paste0(spec_col[i], "_gps_bySex.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_gps_bySex_plot.RData"))
  
}
```

```{r plot_gps_tracks_by_sex}

spec_col <- c("bba_kerguelen", "bba_birdis", "waal_crozet", "waal_birdis")
plot_files <- list.files(path = "Data_outputs/R objects/", pattern = "_gps_bySex_plot")
plot_paths <- paste0("Data_outputs/R objects/", plot_files)

for (file in plot_paths) {
  load(file)
}

#png(filename = "Figures/tracks_by_sex2.png", width = 16, height = 6, units = "in", res = 900)
ggarrange(bba_kerguelen_gps_bySex.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_gps_bySex.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_gps_bySex.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_gps_bySex.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
#dev.off()

```

#### Bearings by sex

To determine whether this could have important effects on my results, I am re-running the analyses but restricting all comparisons to within-sex. If there are major differences in my model outputs, this suggests some of my results are being driven by sex differences.

```{r prep_and_model_bearings_by_sex, eval = FALSE}

# * PREP BEARING DIFFERNECES BY SEX * ----

# Load individual data
load("Data_inputs/ind_meta.RData")
all_ind.meta %<>% mutate(species = tolower(species)) %>% select(-birthYr)

# Set parameters
spec_col <- c("bba_birdis", "bba_kerguelen", "waal_birdis", "waal_crozet")

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

for (i in 1:length(spec_col)) {
  
  if (grepl("ker", spec_col[i])) {
    col_coords <- ker_coords
  } else if (grepl("birdis", spec_col[i])) {
    col_coords <- bi_coords
  } else if (grepl("cro", spec_col[i])) {
    col_coords <- cro_coords }
  
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
 
  my_gps %<>%
    mutate(ID = paste0(ring, "_", season, "_", boutID),
           season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                           as.numeric(format(as.Date(datetime), "%Y")),
                           as.numeric(format(as.Date(datetime), "%Y")) + 1 )) %>%
    filter(loc == "trip")
  
  ## Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>% mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))

  ## Integrate sex and filter
  my_ind <- subset(all_ind.meta, species == strsplit(spec_col[i], "_")[[1]][1])
  my_gps <- merge(my_gps, my_ind, by = "ring", all.x = T) %>% select(-species) %>% filter(!is.na(sex))
  
  ### Remove birds with two sexes
  my_gps %<>% group_by(ring) %>% mutate(n_sex = n_distinct(sex)) %>% filter(n_sex < 2) %>% select(-n_sex)
  
  # Calculate average bearing from colony ------------
   
  # Calculate average bearings
  trip_bearings <- my_gps %>% 
     mutate(bearing_calc = bearing(
                p1 = matrix(c(col_coords$longitude, col_coords$latitude), ncol = 2, byrow = TRUE),
                p2 = matrix(c(longitude, latitude), ncol = 2) )) %>%
    group_by(ID) %>%
    mutate(bearing_col = mean(bearing_calc, na.rm = T),
           species = strsplit(spec_col[i], "_")[[1]][1],
           colony = strsplit(spec_col[i], "_")[[1]][2])
  
  # Isolate bearing per trip
  trip_bearings %<>%
    mutate(boutID = paste0(season, "_", boutID)) %>%
    select(c(ID, ring, sex, season, phase, boutID, bearing_col)) %>%
    distinct()
  
  # Compute the bearing differences ----------
  
  # Within-individual bearings #
  bearings_within <- calculate_bearing_diffs_sex(trip_bearings, "bearing_col", "bearing_diff", comparison_type = "within")
  
  # Among-individual differences #
  bearings_between <- calculate_bearing_diffs_sex(trip_bearings, "bearing_col", "bearing_diff", comparison_type = "between")
  
  # Combine results #
  bearing_differences <- rbind(bearings_within, bearings_between)
  bearing_differences %<>% rename(sex = sex_1) %>% select(-sex_2) %>% relocate(sex, .after = comp_season)
  save(bearing_differences, file = paste0("Data_outputs/", spec_col[i], "_bearing_differences_bySex.RData"))

}


# * MODEL BEARING DIFFERNECES BY SEX * ----

# Output containers
bearings_summary.list <- list()
bearings_means.list <- list()

for (i in seq_along(spec_col)) {
  
  # Load data and prep species/colony labels
  load(paste0("Data_outputs/", spec_col[i], "_bearing_differences_bySex.RData"))
  
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  meta_info <- tibble(species, colony)
  
  # Make dyad ID and subsample to ensure each comparison appears only once (circumvents random effects issues)
  set.seed(817) 
  
  bearing_differences %<>%
    mutate(ring_season_1 = paste(ring_1, season_1, sep = "_"),
           ring_season_2 = paste(ring_2, season_2, sep = "_"),
           dyad = if_else(ring_season_1 < ring_season_2,
                          paste(ring_season_1, ring_season_2, sep = "_"),
                          paste(ring_season_2, ring_season_1, sep = "_"))) %>%
    group_by(dyad) %>%
    slice_sample(n = 1) %>%
    ungroup() %>%
    select(-c(ring_season_1, ring_season_2))
  
  ## Relevel factors for comparison
  bearing_differences %<>% mutate(comp_ind = as.factor(comp_ind), comp_season = as.factor(comp_season))
  
  bearing_differences$comp_ind <- relevel(bearing_differences$comp_ind, ref = "within_ring")
  bearing_differences$comp_season <- relevel(bearing_differences$comp_season, ref = "within_season")

  # Build the model
  m1 <- lm(bearing_diff ~ comp_season * comp_ind, data = bearing_differences)
  # m2 <- lm(bearing_diff ~ comp_season * comp_ind, data = bearing_differences)

  ## Diagnostics
  # plot(m1, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
  # qqnorm(resid(m1))
  # qqline(resid(m1))
  # hist(resid(m1), breaks = 50, main = "Histogram of residuals")

  ## Extract outputs
  m1_em <- data.frame(emmeans(m1, ~comp_ind*comp_season, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring"))) %>%
    relocate(c(species, colony), .before = comp_ind) %>%
    arrange(comp_ind)

  m1_out <- extract_output.lm(m1) %>% mutate(species = species, colony = colony) %>% relocate(c(species, colony), .before = Term)
  
  # Store results
  bearings_summary.list[[i]] <- m1_out
  bearings_means.list[[i]] <- m1_em
  
}

# Combine and save
bearings_summary <- bind_rows(bearings_summary.list)
bearings_means   <- bind_rows(bearings_means.list)

save(bearings_summary, file = "Data_outputs/trajectory_models_summary_bySex.RData")
save(bearings_means, file = "Data_outputs/trajectory_means_bySex.RData")


```

```{r plot_bearing_diffs_bySex, fig.dim=c(6,6)}

load("Data_outputs/trajectory_means_bySex.RData")

# Data wrangling for figure
bearings_plot <- bearings_means %>%
  mutate(species = case_when(
          species == "bba" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "kerguelen" ~ "Kerguelen",
      colony == "crozet" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = ", "),
    comp = case_when(
      comp_season == "within_season" & comp_ind == "within_ring" ~ "Within-individual, within season",
      comp_season == "between_season" & comp_ind == "within_ring" ~ "Within-individual, across seasons",
      comp_season == "within_season" & comp_ind == "between_ring" ~ "Among-individual, within seasons",
      comp_season == "between_season" & comp_ind == "between_ring" ~ "Among-individual, across season",
      TRUE ~ paste(comp_season, comp_ind, sep = ", "))
  )

# Define the order of Comparison levels
comparison_levels <- c(
  "Within-individual, within season",
  "Within-individual, across seasons",
  "Among-individual, within seasons",
  "Among-individual, across season"
)

bearings_plot$comp <- factor(bearings_plot$comp, levels = comparison_levels)

#png(filename = "Figures/bearings_comparison_bySex.png", width = 6, height = 6, units = "in", res = 900)
ggplot(bearings_plot, aes(x = comp_ind, y = emmean, colour = comp_season)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1) +
  geom_line(aes(group = comp_season)) +
  labs(x = "", colour = "", y = "Angular bearing differences",
       caption = "Point and line range indicate mean and 95% confidence intervals.") +
  facet_wrap(~spec_col, scales = "free") +
  scale_colour_viridis_d(labels = c("Within season", "Across seasons")) +
  scale_x_discrete(labels = c("Within individual", "Among individual")) +
  theme_bw() +
  theme(legend.position = c(0.35, 0.62),
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_rect(fill = "transparent", colour = NA))
#dev.off()


```

#### DTW by sex

```{r prep_and_model_dtw_by_sex, eval = FALSE}

# * PREP DTW BY SEX * ----

# Load individual data
load("Data_inputs/ind_meta.RData")
all_ind.meta %<>% mutate(species = tolower(species)) %>% select(-birthYr)

# Set parameters
spec_col <- c("bba_birdis", "bba_kerguelen", "waal_birdis", "waal_crozet")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

for (i in 1:length(spec_col)) {
  
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
  
  # Remove colony periods
  my_gps %<>% filter(loc == "trip")
  
  ## Integrate sex and filter
  my_ind <- subset(all_ind.meta, species == strsplit(spec_col[i], "_")[[1]][1])
  my_gps <- merge(my_gps, my_ind, by = "ring", all.x = T) %>% select(-species) %>% filter(!is.na(sex))
  
  ### Remove birds with two sexes
  my_gps %<>% group_by(ring) %>% mutate(n_sex = n_distinct(sex)) %>% filter(n_sex < 2) %>% select(-n_sex)
  
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  print(paste0("Processing ", species, " at ", colony, "."))
  
  # Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>% mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))
  
  my_meta <- my_gps %>%
    group_by(ID) %>%
    select(c(ring, sex, season, boutID, ID, phase)) %>%
    distinct()

  # Generate all possible trip combinations 
  trip_combinations <- combn(unique(my_gps$ID), 2, simplify = FALSE)
  trip_combinations.df <- data.frame(do.call("rbind", trip_combinations))
  colnames(trip_combinations.df) <- c("tripID_1", "tripID_2")

  ## Merge in sex
  trip_combinations.df <- trip_combinations.df %>%
    left_join(my_meta %>% select(ID, sex), by = c("tripID_1" = "ID")) %>%
    rename(sex_1 = sex) %>%
    left_join(my_meta %>% select(ID, sex), by = c("tripID_2" = "ID")) %>%
    rename(sex_2 = sex)
  
  # Within-individual DTW -----
  
  # Find the combinations by filtering to same rings and same phase
  dtw_within.df <- trip_combinations.df %>%
    mutate(ring_1 = sub("_.*", "", tripID_1),
           ring_2 = sub("_.*", "", tripID_2)) %>%
    filter(ring_1 == ring_2) %>%
    select(-c(ring_1, ring_2)) %>%
    # Ensure comparisons are within same phase
    left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_1 = ID),
            by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID),
              by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2))

  ## Calculate the comparisons (slow!)
  dtw_within.list <- vector(mode = "list", length = nrow(dtw_within.df))
  
  print("Processing within-individual: ")
  pb <- txtProgressBar(min = 0, max = nrow(dtw_within.df), style = 3)
  
  for (j in 1:nrow(dtw_within.df)) {
   
    setTxtProgressBar(pb, j)
    
    trip1 <- convert_to_spdf(subset(my_gps, ID == dtw_within.df$tripID_1[j]), proj.dec)
    trip2 <- convert_to_spdf(subset(my_gps, ID == dtw_within.df$tripID_2[j]), proj.dec)
    
    trip1_mat <- as.matrix(trip1@coords)
    trip2_mat <- as.matrix(trip2@coords)
   
    dtw_out <- dtw(trip1_mat, trip2_mat, dist.method = "Euclidean")$distance
    
    output <- data.frame(trip1 = dtw_within.df$tripID_1[j],
                         trip2 = dtw_within.df$tripID_2[j],
                         dtw = dtw_out)
  
    dtw_within.list[[j]] <- output 
    
  }

  close(pb) 
  
  ## Bind up the output
  dtw_within.output <- do.call("rbind", dtw_within.list)
  
  dtw_within.output %<>% 
    separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
    separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
    select(-loc) %>%
    mutate(tripID_1 = paste0("trip_", tripID_1),
           tripID_2 = paste0("trip_", tripID_2),
           comp_ind = "within_ring",
           comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))

  nrow_within <- nrow(dtw_within.df)
  rm(dtw_within.df); rm(dtw_within.list)
  gc()
  
  # Between-individual DTW -----
  
  dtw_between.df <- trip_combinations.df %>%
    filter(sex_1 == sex_2) %>%
    mutate(ring_1 = sub("_.*", "", tripID_1),
           ring_2 = sub("_.*", "", tripID_2)) %>%
    filter(ring_1 != ring_2) %>%
    select(-c(ring_1, ring_2)) %>%
    # Get phase and filter for same breeding phase
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_1 = ID),
              by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID),
              by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2))


  ## Downsample the between-individual data
  
  set.seed(817)
  n_sub = 3 * nrow_within

  dtw_between <- dtw_between.df %>% 
    sample_n(size = n_sub, replace = FALSE)
  
  ## Calculate the comparisons
  dtw_between.list <- vector(mode = "list", length = nrow(dtw_between))
  
  print("Processing between-individual: ")
  pb <- txtProgressBar(min = 0, max = nrow(dtw_between), style = 3)
  
  for (k in 1:nrow(dtw_between)) {
    
    setTxtProgressBar(pb, k)
    
    trip1 <- convert_to_spdf(subset(my_gps, ID == dtw_between$tripID_1[k]), proj.dec)
    trip2 <- convert_to_spdf(subset(my_gps, ID == dtw_between$tripID_2[k]), proj.dec)
    
    trip1_mat <- as.matrix(trip1@coords)
    trip2_mat <- as.matrix(trip2@coords)
    
    dtw_out <- dtw(trip1_mat, trip2_mat, dist.method = "Euclidean")$distance
    
    output <- data.frame(trip1 = dtw_between$tripID_1[k],
                         trip2 = dtw_between$tripID_2[k],
                         dtw = dtw_out)
    
    dtw_between.list[[k]] <- output 
    
  }

  close(pb) 
  
  dtw_between.output <- do.call("rbind", dtw_between.list)
  
  dtw_between.output %<>% 
    separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
    separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
    select(-loc) %>%
    mutate(tripID_1 = paste0("trip_", tripID_1),
           tripID_2 = paste0("trip_", tripID_2),
           comp_ind = "between_ring",
           comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))
  
  # Bind the two dataframes together
  dtw.df <- rbind(dtw_within.output, dtw_between.output)
  
  save(dtw.df, file = paste0("Data_outputs/", species, "_", colony, "_dtw_comparison_bySex.RData"))
  
  gc()

}
  
# * MODEL DTW DIFFERENCES BY SEX * ----

# Output containers
dtw_summary.list <- list()
dtw_means.list <- list()

for (i in seq_along(spec_col)) {
  
  # Load data and prep species/colony labels
  load(paste0("Data_outputs/", spec_col[i], "_dtw_comparison_bySex.RData"))
  
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  meta_info <- tibble(species, colony)
  
  # Make dyad ID and subsample to ensure each comparison appears only once (circumvents random effects issues)
  set.seed(817) 
  
  dtw.df %<>%
    mutate(ring_season_1 = paste(ring_1, season_1, sep = "_"),
           ring_season_2 = paste(ring_2, season_2, sep = "_"),
           dyad = if_else(ring_season_1 < ring_season_2,
                          paste(ring_season_1, ring_season_2, sep = "_"),
                          paste(ring_season_2, ring_season_1, sep = "_"))) %>%
    group_by(dyad) %>%
    slice_sample(n = 1) %>%
    ungroup() %>%
    select(-c(ring_season_1, ring_season_2))
  
  ## Relevel factors for comparison
  dtw.df %<>% mutate(comp_ind = as.factor(comp_ind), comp_season = as.factor(comp_season))
  
  dtw.df$comp_ind <- relevel(dtw.df$comp_ind, ref = "within_ring")
  dtw.df$comp_season <- relevel(dtw.df$comp_season, ref = "within_season")

  # Build the model
  m1 <- lm(log(dtw) ~ comp_season * comp_ind, data = dtw.df)
  
  ## Extract outputs
  m1_em <- data.frame(emmeans(m1, ~comp_ind*comp_season, pbkrtest.limit = 7000)) %>% 
    mutate(species = species, colony = colony,
           comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring"))) %>% 
    relocate(c(species, colony), .before = comp_ind) %>% 
    arrange(comp_ind)
  
  m1_out <- extract_output.lm(m1) %>% mutate(species = species, colony = colony) %>% relocate(c(species, colony), .before = Term)
  
  # Store results
  dtw_summary.list[[i]] <- m1_out
  dtw_means.list[[i]] <- m1_em
  
}

# Combine and save
dtw_summary <- bind_rows(dtw_summary.list)
dtw_means   <- bind_rows(dtw_means.list)

save(dtw_summary, file = "Data_outputs/dtw_model_summary_bySex.RData")
save(dtw_means,   file = "Data_outputs/dtw_means_bySex.RData")

```

```{r plot_dtw_diffs_by_sex, fig.dim=c(6,6)}

load("Data_outputs/dtw_means_bySex.RData")

# Data wrangling for figure
dtw_plot <- dtw_means %>%
  mutate(species = case_when(
          species == "bba" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "kerguelen" ~ "Kerguelen",
      colony == "crozet" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = ", "),
    comp = case_when(
      comp_season == "within_season" & comp_ind == "within_ring" ~ "Within-individual, within season",
      comp_season == "between_season" & comp_ind == "within_ring" ~ "Within-individual, across seasons",
      comp_season == "within_season" & comp_ind == "between_ring" ~ "Among-individual, within seasons",
      comp_season == "between_season" & comp_ind == "between_ring" ~ "Among-individual, across season",
      TRUE ~ paste(comp_season, comp_ind, sep = ", "))
  )


dtw_plot$comp <- factor(dtw_plot$comp, levels = comparison_levels)

dtw.plot_bySex <- ggplot(dtw_plot, aes(x = comp_ind, y = emmean, colour = comp_season)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1) +
  geom_line(aes(group = comp_season)) +
  labs(x = "", colour = "", y = "Logged DTW value",
       caption = "Point and line range indicate mean and 95% confidence intervals.") +
  facet_wrap(~spec_col, scales = "free") +
  scale_colour_viridis_d(labels = c("Within season", "Across seasons")) +
  scale_x_discrete(labels = c("Within individual", "Among individual")) +
  theme_bw() +
  theme(legend.position = c(0.35, 0.62),
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_rect(fill = "transparent", colour = NA))

save_plot(dtw.plot_bySex, "Figures/dtw_comparison_bySex.png", 6, 6)
dtw.plot_bySex

```

#### Overall

Looking at the outputs, there is hardly any difference between the analyses where individuals are grouped by sex and when they are not, at least for the bearing differences, so I think I'm safe to continue. I'm not interested in the effect of sex *per se*, so there's no reason to go deeper into this.




## Appendix

### Sample sizes for colony & species

```{r sample_sizes}
### Sample sizes

load("Data_outputs/all_gps_data.RData")

trips.table <- all_gps.df %>% group_by(species, colony, ring, season) %>% summarise(n_trips = n_distinct(boutID)) %>%
  filter(n_trips > 1) %>% group_by(species, colony,season) %>% summarise(n_birds = n_distinct(ring)) %>% 
  mutate(species = ifelse(species == "bba", "BBAL", "WAAL"),
         colony = ifelse(colony == "kerguelen", "Kerguelen",
                          ifelse(colony == "crozet", "Crozet", "Bird Island"))) %>%
  relocate(species, .before = colony) %>%
  rename(Species = species, Colony = colony, Season = season, `N birds` = n_birds)

trips_wide <- trips.table %>%
  arrange(Season, Species, Colony) %>%
  pivot_wider(names_from = Season, values_from = `N birds`, values_fill = 0)

kable(trips_wide, format = "html", 
             caption = 'Table 1. Sample sizes for each species, colony, and year, showing the number of individuals for which we have more than 1 foraging trip.') %>%
  collapse_rows(columns = c(1)) %>%
  row_spec(2, extra_css = "border-bottom: 1px solid;") %>%
  kable_classic(font_size = 9)


# How many birds have > 1 year of data?

seasons_table <- all_gps.df %>%
  group_by(colony, species, ring) %>%
  summarise(n_seasons = n_distinct(season)) %>%
  filter(n_seasons > 1) %>%
  group_by(colony, species) %>%
  summarise(n_birds = n_distinct(ring)) %>%
  relocate(species, colony, n_birds) %>%
  mutate(species = ifelse(species == "bba", "BBAL", "WAAL"),
         colony = case_when(
           colony == "birdis" ~ "Bird Island",
           colony == "crozet" ~ "Crozet",
           colony == "kerguelen" ~ "Kerguelen" )) %>%
  rename(Species = species, Colony = colony, Individuals = n_birds) %>%
  arrange(Species)

kable(seasons_table, format = "html", 
             caption = 'Table 2. Number of birds for each colony and species where trips are recorded in > 1 season.') %>%
  collapse_rows(columns = c(1, 2)) %>%
  kable_classic(font_size = 12)

```

### Plot bearings with tracks

```{r plot_tracks_with_bearings}

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

load("Data_outputs/all_gps_with_bearings.RData")

all_rings <- unique(gps_with_bearings$ring)

for (i in 1:length(all_rings)) {
  
  my_df <- gps_with_bearings %>% filter(ring == all_rings[[i]])
  my_df$tripID <- paste0(my_df$species[1], "_", my_df$colony[1], "_", my_df$tripID)
  
  # Get colony coords
  if (grepl("ker", my_df$colony[1])) {
    col_coords <- ker_coords
  } else if (grepl("birdis", my_df$colony[1])) {
    col_coords <- bi_coords
  } else if (grepl("cro", my_df$colony[1])) {
    col_coords <- cro_coords }
  
  # Make the bearing dataframe
  bearing_df <- my_df %>%
      group_by(tripID) %>%
      summarise(
        cx = first(longitude),
        cy = first(latitude),
        bearing_avg= as.numeric(first(bearing_avg)),
        bearing_initial = as.numeric(first(bearing_initial)),
        bearing_distal  = as.numeric(first(bearing_distal)),
        lon_range = max(longitude) - min(longitude),
        lat_range = max(latitude) - min(latitude),
        .groups = "drop" ) %>%
      pivot_longer(
        cols = starts_with("bearing_"),
        names_to = "bearing_type",
        values_to = "bearing") %>%
      mutate(
        # scale factor = how long you want the arrow to be
        bearing360 = (bearing + 360) %% 360,
        scale_factor = 0.5 * pmax(lon_range, lat_range),
        dx = scale_factor * sin(bearing360 * pi/180),  # x = sin
        dy = scale_factor * cos(bearing360 * pi/180),  # y = cos
        xend = cx + dx,
        yend = cy + dy,
        bearing_type = recode(
            bearing_type,
            "bearing_avg"     = "Average bearing",
            "bearing_initial" = "Initial bearing",
            "bearing_distal"  = "Distal bearing" ))
 
  gps_plot <- ggplot(aes(x = longitude, y = latitude), data = my_df) +
    geom_path(col = "grey50") +
    geom_segment(
      data = bearing_df,
      aes(x = cx, y = cy, xend = xend, yend = yend, col = bearing_type),
      arrow = arrow(length = unit(0.3, "cm")),
      inherit.aes = FALSE ) +
    scale_colour_viridis_d(name = "") +
    facet_wrap(~tripID, scales = "free", nrow = 1) +
    theme_bw() +
    theme(axis.title.x = element_blank(),
          axis.title.y = element_blank(),
          legend.position = "bottom",
          legend.direction = "horizontal")
  
  # Output plot
  n_facets <- n_distinct(my_df$tripID)
  width_per_facet  <- 400   
  height_per_facet <- 400   
  
  png_width  <- n_facets * width_per_facet
  png_height <- height_per_facet
  
  png(file = paste0("Figures/Workshop/", my_df$species[1], "_", my_df$colony[1], "_", my_df$ring[1], ".png"),
      width = png_width, height = png_height, res = 100)
  print(gps_plot)
  dev.off()
 
  
}


```

