---
title: "Spatial repeatability results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_circular(v.0.5-1)_, _nadiv(v.2.18.0)_, _betareg(v.3.2-0)_, _adehabitatHR(v.0.4.22)_, _adehabitatLT(v.0.3.28)_, _CircStats(v.0.2-7)_, _boot(v.1.3-28.1)_, _adehabitatMA(v.0.3.17)_, _ade4(v.1.7-24)_, _data.table(v.1.18.2.1)_, _emmeans(v.2.0.3)_, _sp(v.2.2-1)_, _geosphere(v.1.5-20)_, _purrr(v.1.2.1)_, _rnaturalearth(v.1.2.0)_, _sf(v.1.0-20)_, _ggpubr(v.0.6.3)_, _MasterBayes(v.2.59)_, _kinship2(v.1.9.6.2)_, _quadprog(v.1.5-8)_, _genetics(v.1.3.8.1.3)_, _mvtnorm(v.1.3-3)_, _MASS(v.7.3-60)_, _gtools(v.3.9.5)_, _gdata(v.3.0.1)_, _combinat(v.0.0-8)_, _igraph(v.2.1.4)_, _MCMCglmm(v.2.36)_, _ape(v.5.8-1)_, _coda(v.0.19-4.1)_, _brms(v.2.23.0)_, _Rcpp(v.1.1.1)_, _ggridges(v.0.5.7)_, _glmmTMB(v.1.1.11)_, _stringr(v.1.6.0)_, _kableExtra(v.1.4.0)_, _tidyr(v.1.3.2)_, _gridExtra(v.2.3)_, _lme4(v.2.0-1)_, _Matrix(v.1.5-4.1)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_

**loaded via a namespace (and not attached):** 
_RColorBrewer(v.1.1-3)_, _tensorA(v.0.36.2.1)_, _rstudioapi(v.0.18.0)_, _modeltools(v.0.2-24)_, _estimability(v.1.5.1)_, _farver(v.2.1.2)_, _nloptr(v.2.2.1)_, _rmarkdown(v.2.31)_, _vctrs(v.0.7.2)_, _minqa(v.1.2.8)_, _rstatix(v.0.7.3)_, _htmltools(v.0.5.9)_, _forcats(v.1.0.1)_, _distributional(v.0.7.0)_, _broom(v.1.0.12)_, _Formula(v.1.2-5)_, _StanHeaders(v.2.32.10)_, _KernSmooth(v.2.23-21)_, _sandwich(v.3.1-1)_, _plyr(v.1.8.9)_, _zoo(v.1.8-13)_, _lubridate(v.1.9.4)_, _TMB(v.1.9.21)_, _lifecycle(v.1.0.5)_, _pkgconfig(v.2.0.3)_, _R6(v.2.6.1)_, _fastmap(v.1.2.0)_, _rbibutils(v.2.3)_, _digest(v.0.6.39)_, _numDeriv(v.2016.8-1.1)_, _colorspace(v.2.1-1)_, _textshaping(v.1.0.5)_, _labeling(v.0.4.3)_, _timechange(v.0.4.0)_, _abind(v.1.4-8)_, _mgcv(v.1.8-42)_, _compiler(v.4.3.1)_, _proxy(v.0.4-27)_, _pander(v.0.6.6)_, _withr(v.3.0.2)_, _S7(v.0.2.1)_, _backports(v.1.5.0)_, _inline(v.0.3.21)_, _carData(v.3.0-6)_, _DBI(v.1.3.0)_, _QuickJSR(v.1.9.0)_, _pkgbuild(v.1.4.8)_, _ggsignif(v.0.6.4)_, _classInt(v.0.4-11)_, _corpcor(v.1.6.10)_, _loo(v.2.9.0)_, _tools(v.4.3.1)_, _units(v.0.8-7)_, _lmtest(v.0.9-40)_, _otel(v.0.2.0)_, _nnet(v.7.3-19)_, _glue(v.1.8.0)_, _callr(v.3.7.6)_, _nlme(v.3.1-162)_, _grid(v.4.3.1)_, _checkmate(v.2.3.4)_, _reshape2(v.1.4.5)_, _generics(v.0.1.4)_, _gtable(v.0.3.6)_, _class(v.7.3-22)_, _flexmix(v.2.3-20)_, _xml2(v.1.5.2)_, _car(v.3.1-5)_, _utf8(v.1.2.6)_, _pillar(v.1.11.1)_, _posterior(v.1.7.0)_, _splines(v.4.3.1)_, _lattice(v.0.21-8)_, _tidyselect(v.1.2.1)_, _knitr(v.1.51)_, _reformulas(v.0.4.4)_, _svglite(v.2.2.2)_, _stats4(v.4.3.1)_, _xfun(v.0.57)_, _bridgesampling(v.1.2-1)_, _matrixStats(v.1.5.0)_, _rstan(v.2.32.7)_, _stringi(v.1.8.7)_, _yaml(v.2.3.12)_, _evaluate(v.1.0.5)_, _codetools(v.0.2-19)_, _tibble(v.3.3.1)_, _cli(v.3.6.5)_, _RcppParallel(v.5.1.11-2)_, _xtable(v.1.8-8)_, _systemfonts(v.1.3.2)_, _Rdpack(v.2.6.3)_, _processx(v.3.8.7)_, _parallel(v.4.3.1)_, _rstantools(v.2.6.0)_, _ggh4x(v.0.3.1)_, _cubature(v.2.1.4-1)_, _bayesplot(v.1.15.0)_, _Brobdingnag(v.1.2-9)_, _viridisLite(v.0.4.3)_, _scales(v.1.4.0)_, _e1071(v.1.7-16)_ and _rlang(v.1.1.7)_

## Set-up

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load functions, packages, & data ---------------------------------------------

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "ggpubr", "tidyr", "sf", "ggridges", "rnaturalearth", "purrr", "geosphere", "sp", "kableExtra", "emmeans", "data.table", "adehabitatHR", "betareg", "MasterBayes", "nadiv", "brms", "circular") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
# if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"

```

## Bearings

### Prep data

```{r prep_bearings_data, eval = FALSE}

# Set parameters
spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

# Make list to collect bearings data
trip_bearings.list <- list()
gps_with_bearings.list <- list()

for (i in 1:length(spec_col)) {
  
  if (grepl("ker", spec_col[i])) {
    col_coords <- ker_coords
  } else if (grepl("birdis", spec_col[i])) {
    col_coords <- bi_coords
  } else if (grepl("cro", spec_col[i])) {
    col_coords <- cro_coords }
  
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
 
  my_gps %<>%
    mutate(ID = paste0(ring, "_", season, "_", boutID),
           season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                           as.numeric(format(as.Date(datetime), "%Y")),
                           as.numeric(format(as.Date(datetime), "%Y")) + 1 )) %>%
    filter(loc == "trip")
  
  ## Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>% mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))

  ## Remove colony periods & any very short trips
  my_gps %<>% filter(loc == "trip")
  my_gps %<>% group_by(tripID) %>% mutate(n_locs = n()) %>% filter(n_locs > 6) %>% select(-n_locs) %>% ungroup()
  
  ## Remove birds with only one trip
  # my_gps %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% select(-n_trips) %>% ungroup()
  
  # Calculate average bearing from colony ------------
   
  trip_bearings <- my_gps %>% 
     group_by(ID) %>%
     arrange(datetime) %>%
    ## Bearing from colony to each fix
     mutate(bearing_col = bearing(
                p1 = matrix(c(col_coords$longitude, col_coords$latitude), ncol = 2, byrow = TRUE),
                p2 = matrix(c(longitude, latitude), ncol = 2)),
    ## Mean bearing across the whole trip
            bearing_avg = median.circular(circular(bearing_col, units = "degrees", template = "geographics"), na.rm = T),
            species = strsplit(spec_col[i], "_")[[1]][1],
            colony = strsplit(spec_col[i], "_")[[1]][2]) %>%
    ungroup()
  
  ## Extract complete info for plotting
  gps_with_bearings.list[[i]] <- trip_bearings

  # Isolate bearing per trip & calculate inter-trip distance
  trip_bearings %<>%
    mutate(boutID = paste0(season, "_", boutID)) %>%
    select(c(ID, ring, season, phase, boutID, bearing_avg, datetime)) %>%
    group_by(ring,season, boutID) %>%
    slice(1) %>%
    group_by(ring, season) %>%
    mutate(time_since_first = as.numeric(difftime(datetime, min(datetime), units = "days"))) %>%
    ungroup() %>%
    select(c(ID, ring, season, phase, boutID, bearing_avg, time_since_first)) %>%
    distinct()
  
  trip_bearings$species <- strsplit(spec_col[i], "_")[[1]][1]
  trip_bearings$colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  trip_bearings.list[[i]] <- trip_bearings

}

bearings.df <- do.call("rbind", trip_bearings.list)
save(bearings.df, file = "Data_outputs/bearings_data.RData")

gps_with_bearings <- do.call("rbind", gps_with_bearings.list)
save(gps_with_bearings, file = "Data_outputs/all_gps_with_bearings.RData")

```

### Plot bearings from the colony

```{r plot_bearings, fig.dim=c(16, 4.5)}

load("Data_outputs/bearings_data.RData")

bearings.df %<>% mutate(bearing_360 = ifelse(bearing_avg < 0, bearing_avg + 360, bearing_avg),
                        spec_col = case_when(
                            species == "bbal" & colony == "birdis" ~ "Black-browed albatrosses at Bird Island",
                            species == "bbal" & colony == "ker" ~ "Black-browed albatrosses at Kerguelen",
                            species == "waal" & colony == "birdis" ~ "Wandering albatrosses at Bird Island",
                            species == "waal" & colony == "cro" ~ "Wandering albatrosses at Crozet"))
breaks = seq(0, 360, 60)

#png(filename = "Figures/MS Figures/FIGSX_bearings_plot.png", width = 16, height = 4.5, units = "in", res = 900)
ggplot(bearings.df, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks, #c(0, 90, 180, 270), # breaks,
                     labels = paste0(breaks, "\u00B0")) +#c("N", "E", "S", "W")) + # paste0(breaks, "\u00B0")) +
  labs(y = "Frequency") +
  #facet_grid(spec_col ~ ., scales="free") +
  ggh4x::facet_grid2(. ~ spec_col, scales = "free_y", independent = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))
#dev.off()

  
```

### Estimate repeatability

```{r rpt_bearings}

load("Data_outputs/bearings_data.RData")

bearings.df %<>% mutate(bearing_360 = ifelse(bearing_avg < 0, bearing_avg + 360, bearing_avg),
                        spec_col = case_when(
                            species == "bbal" & colony == "birdis" ~ "bbal_birdis",
                            species == "bbal" & colony == "ker" ~ "bbal_ker",
                            species == "waal" & colony == "birdis" ~ "waal_birdis",
                            species == "waal" & colony == "cro" ~ "waal_cro"))

# Convert degrees to radians
bearings.df %<>%
  mutate(bearing_rad = bearing_360 * pi / 180,
         bearing_rad = ifelse(bearing_rad >= pi, bearing_rad - 2*pi, bearing_rad),
         bearing_rad = ifelse(bearing_rad <= -pi, bearing_rad + 2*pi, bearing_rad))
  
## Add small value to avoid hitting pi exactly
eps <- 1e-6
bearings.df$bearing_rad <- pmin(pmax(bearings.df$bearing_rad, -pi + eps), pi - eps)

# Add parameters for model - ring season, time since first
bearings.df %<>% mutate(ring_season = paste0(ring, "_", season))

## Define some weakly informative priors
bearing_priors <- c(
  prior(normal(0, 1), class = "Intercept", dpar = "mu"),      # mu intercept
  prior(normal(0, 0.5), class = "b", dpar = "mu"),            # fixed effects
  prior(exponential(1), class = "sd", dpar = "mu"),           # group-level SDs
  prior(gamma(2, 0.01), class = "kappa")  # kappa intercept
)

spec_cols <- unique(bearings.df$spec_col)

bearing_posteriors.list <- list()
options(brms.backend = "cmdstanr")   

for (i in 1:length(spec_cols)) {
  
  print(paste0("Processing ", i, "..."))
  
  my_df <- subset(bearings.df, spec_col == spec_cols[i])
  
  # Set formula
  if(spec_col[i] == "waal_birdis") {
  bearing_formula <- bf("bearing_rad ~ phase + (1 | season) + (1 | ring_season)") } else {
    bearing_formula <- bf("bearing_rad ~ phase + (1 | season) + (1 | ring ) + (1 | ring_season)")  }
  
  # Fit model 
  brm_vm <- brm(
    formula = bearing_formula,
    family  = von_mises(),  
    data    = my_df,
    prior   = bearing_priors,
    chains = 4, cores = 4, iter = 8000, warmup = 4000,
    control = list(max_treedepth = 15, adapt_delta = 0.99) )

  # Calc repeatability
  if(spec_col[i] == "waal_birdis") {
    post_df <- calc_rpt_brms_VM_waBI(brm_vm) } else {
      post_df <- calc_rpt_brms_VM(brm_vm)
    }
  
  post_df %<>% mutate(spec_col = spec_col[i],
                      species = my_df$species[1],
                      colony = my_df$colony[1])
  
  mod_name <- paste0(spec_col[i], ".brm_rpt_bearings")
  assign(mod_name, brm_vm)
  
  save(list = mod_name, file = paste0("Data_outputs/", spec_col[i], "_brms_rpt_bearings.RData"))
  
  bearing_posteriors.list[[i]] <- post_df
  
}

bearing_posteriors.df <- do.call("rbind", bearing_posteriors.list)
save(bearing_posteriors.df, file = "Data_outputs/bearing_repeatability_results.RData")

```

#### Plot bearing repeatability

```{r plot_bearings_diffs, fig.dim=c(6,6)}

load("Data_outputs/bearing_repeatability_results.RData")

# Recode spec_col
bearing_posteriors.df %<>% mutate(spec_col = recode(spec_col,
   "bbal_birdis" = "BBAL Bird Island",
    "bbal_ker" = "BBAL Kerguelen",
    "waal_birdis" = "WAAL Bird Island",
    "waal_cro" = "WAAL Crozet"))

# Pivot data longer for faceted plotting
bearings_rpt.long <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") 

# Get median and 95% credible interval 
bearings_rpt.summary <- bearings_rpt.long %>%
  group_by(repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE)) %>%
  arrange(spec_col) %>%
  mutate(spec_col = factor(spec_col, levels = rev(unique(spec_col))))


bearings_rpt.plot <- ggplot(bearings_rpt.summary, aes(x = median_val, y = spec_col, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  scale_colour_manual( values = c("rpt_within" = col_within, "rpt_across" = col_between),
                       labels = c("Within-season", "Between-season") ) +
  labs(x = "Repeatability in foraging bearing (median, 95% CI)", y = "") +
  theme_bw() +
  theme(legend.position = "none")


# Repeatability estimates

# BBAL - Within
bearings_rpt.summary %>% filter(repeatability_type == "rpt_within" & spec_col == "BBAL Bird Island") %>% ungroup
bearings_rpt.summary %>% filter(repeatability_type == "rpt_within" & spec_col == "BBAL Kerguelen") %>% ungroup 

# BBAL - Within
bearings_rpt.summary %>% filter(repeatability_type == "rpt_across" & spec_col == "BBAL Bird Island") %>% ungroup
bearings_rpt.summary %>% filter(repeatability_type == "rpt_across" & spec_col == "BBAL Kerguelen") %>% ungroup 

# WAAL
bearings_rpt.summary %>% filter(spec_col == "WAAL Bird Island") %>% ungroup
bearings_rpt.summary %>% filter(spec_col == "WAAL Crozet") %>% ungroup 

```



### Heritability of bearings

```{r bearings_heritability}

load("Data_outputs/bearings_data.RData")

## Get trip ID
bearings.df %<>%
  separate(boutID, into = c("season", "trip1", "trip2"), sep = "_") %>%
  mutate(tripID = paste0(ring, "_", season, "_", trip1, trip2)) %>%
  ungroup() %>%
  select(-c(ID, trip1, trip2))

## Set weakly informative priors
bearing_priors <- c(
  prior(student_t(3, 0, 1), class = "Intercept", resp = "cosb"),
  prior(student_t(3, -0.5, 1), class = "Intercept", resp = "sinb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "ring", resp = "cosb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "animal", resp = "cosb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "ring", resp = "sinb"),
  prior(student_t(3, 0, 0.8), class = "sd", group = "animal", resp = "sinb"),
  prior(student_t(3, 0, 0.8), class = "sigma", resp = "cosb"),
  prior(student_t(3, 0, 0.8), class = "sigma", resp = "sinb")
)

# Loop through species and colonies
spec_col <- c("bbal_birdis", "waal_birdis", "waal_cro")
posteriors = list()

for (i in 1:length(spec_col)) {
  
  my_species <- strsplit(spec_col[i], "_")[[1]][1]
  my_colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  focal.bearings <- bearings.df %>% filter(species == my_species & colony == my_colony)
  focal.bearings %<>% mutate(animal = ring, ring_pe = ring, ringSeason = paste0(ring, "_", season))

  ## Convert to bearings to radians and create cos/sin
  focal.bearings$bearing_rad <- (focal.bearings$bearing_avg %% 360) * pi / 180
  focal.bearings$cos_b <- cos(focal.bearings$bearing_rad)
  focal.bearings$sin_b <- sin(focal.bearings$bearing_rad)
  
  ## Restrict to birds with > 1 trip
  focal.bearings %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% select(-n_trips)
  
  # Build the pedigree ====
  
  pedigree.df <- read.csv(paste0("Data_inputs/", spec_col[i], "_pedigree.csv"))[,1:3]
  pedigree.df[,c(1:3)] <- lapply(pedigree.df[,c(1:3)], as.character)
  
  Amat <- make_Amat(focal.bearings, pedigree.df)
  
  # Build the model ====
  
  ## Set formula
  if(spec_col[i] == "waal_birdis") {
  mv_formula <- bf(mvbind(cos_b, sin_b) ~ 1 +  (1 | gr(animal, cov = Amat)) + (1 | ring) + (1 | season)) } else {
    mv_formula <- bf(mvbind(cos_b, sin_b) ~ 1 + phase + (1 | gr(animal, cov = Amat)) + (1 | ring) + (1 | ringSeason) + (1 | season))  }

  brm_bearing_heritability <- brm(
    formula = mv_formula,
    prior = bearing_priors,
    data = focal.bearings,
    data2 = list(Amat = Amat),
    family = gaussian(),
    chains = 4, cores = 4, iter = 8000, warmup = 4000,
    control = list(max_treedepth = 15, adapt_delta = 0.99))

  #plot(brm_bearing_heritability)  
  save(brm_bearing_heritability, file = paste0("Data_outputs/", spec_col[i], "_bearing_heritability.RData"))
 
  # Calculate h2 ====
  draws <- as_draws_df(brm_bearing_heritability)
  
  ## Get variance components
  VA <- extractVar(draws, "animal")
  Vring <- extractVar(draws, "ring")
  if(spec_col[i] != "waal_birdis") { VringSeason <- extractVar(draws, "ringSeason") }
  Vseason <- extractVar(draws, "season")
  VR <- draws$sigma_cosb^2 + draws$sigma_sinb^2

  ## Calculate total phenotypic variance 
  if(spec_col[i] == "waal_birdis") { VP <- VA + Vring + Vseason + VR } else {
  VP <- VA + Vring + VR }

  ## Calculate h2
  h2 <- VA / VP

  posteriors[[i]] <- data.frame(species = my_species,
                           colony = my_colony,
                           h2 = h2)
  
}

posteriors_all <- do_call("rbind", posteriors)

posteriors_all %<>%
  mutate(spec_col = case_when(
      species == "bbal" ~ "BBAL Bird Island",
      species == "waal" & colony == "birdis" ~ "WAAL Bird Island",
      TRUE ~ "WAAL Crozet"),
      spec_col = factor(spec_col, levels = rev(unique(spec_col))))

save(posteriors_all, file = "Data_outputs/all_posteriors_spatial_heritability.RData")

```

#### Plot heritability of bearings

```{r spatial_heritability}

load("Data_outputs/all_posteriors_spatial_heritability.RData")

png(filename = "Figures/spatial_heritability.png", width = 6, height = 6, units = "in", res = 900)
ggplot(posteriors_all, aes(x = h2, y = spec_col)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw() 
dev.off()

```


## Bhattacharyya's Affinity

```{r prep_BA_data, eval = FALSE}

# Set parameters
spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

for (i in 1:length(spec_col)) {
  
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
 
  # Remove colony periods & any very short trips
  my_gps %<>% filter(loc == "trip")
  my_gps %<>% group_by(tripID) %>% mutate(n_locs = n(), max_dist = max(dist_col)/1000) %>% 
    filter(n_locs > 20 & max_dist > 65) %>% select(-c(n_locs, max_dist)) %>% ungroup()
  
  ## Remove birds with only one trip
  # my_gps %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% select(-n_trips) %>% ungroup()
  
  ## Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>% mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))
  
  # Get start times look-up
  trip_starts <- my_gps %>%
    mutate(tripID = paste(season, "_", boutID)) %>%
    group_by(ring, tripID) %>%
    summarise(start_time = min(datetime), .groups = "drop")
  
  ### Meta for phase comparisons later
  my_meta <- my_gps %>%
    group_by(ID) %>%
    select(c(ring, season, boutID, ID, phase)) %>%
    distinct()

  # Generate all possible trip combinations 
  trip_combinations <- combn(unique(my_gps$ID), 2, simplify = FALSE)
  trip_combinations.df <- data.frame(do.call("rbind", trip_combinations))
  colnames(trip_combinations.df) <- c("tripID_1", "tripID_2")

  # Plotting out some UDs ----
  
  # birds <- unique(my_gps$ring)
  # 
  # for (b in 1:length(birds)) {
  # 
  #   # Get focal bird
  #   ud.df_A <- subset(my_gps, ring == birds[b]) %>% select(tripID, longitude, latitude)
  #   ud.sp_A <- SpatialPointsDataFrame(coords = ud.df_A[, c("longitude", "latitude")],
  #                                data = data.frame(id = ud.df_A$tripID))
  #   ud.est_A <- kernelUD(ud.sp_A[, 1], h = "href", extent = 3, grid = 500)
  #   ud.95_A <- getverticeshr(ud.est_A, percent = 95)
  # 
  #   ud_sf_A <- st_as_sf(ud.95_A)
  #   ud_sf_A %<>% separate(id, sep = "_", into = c("ring", "season", "dummy", "tripNo")) %>% select(-dummy)
  # 
  #   # Random bird
  #   ud.df_B <- subset(my_gps, ring == birds[sample(length(birds), 1)]) %>% select(tripID, longitude, latitude)
  #   ud.sp_B <- SpatialPointsDataFrame(coords = ud.df_B[, c("longitude", "latitude")],
  #                                data = data.frame(id = ud.df_B$tripID))
  #   ud.est_B <- kernelUD(ud.sp_B[, 1], h = "href", extent = 3, grid = 500)
  #   ud.95_B <- getverticeshr(ud.est_B, percent = 95)
  # 
  #   ud_sf_B <- st_as_sf(ud.95_B)
  #   ud_sf_B %<>% separate(id, sep = "_", into = c("ring", "season", "dummy", "tripNo")) %>% select(-dummy)
  # 
  #   ud_plot <- ggplot() +
  #     geom_sf(data = ud_sf_A, fill = "blue", alpha = 0.4) +
  #     geom_sf(data = ud_sf_B, fill = "red", alpha = 0.4) +
  #     theme_bw()
  # 
  #   png(file = paste0("Figures/Workshop/", spec_col[i], "_", birds[b], "_", ud_sf_B$ring[1], ".png"))
  #   print(ud_plot)
  #   dev.off()
  # }
  
  # ===# WITHIN INDIVIUDAL ====
  
  # Find the combinations by filtering to same rings and same phase
  ba_within.df <- trip_combinations.df %>%
     mutate(ring_1 = sub("_.*", "", tripID_1),
           ring_2 = sub("_.*", "", tripID_2)) %>%
    filter(ring_1 == ring_2) %>%
    select(-c(ring_1, ring_2)) %>%
    # Ensure comparisons are within same phase
    left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_1 = ID),
            by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID),
              by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2))

  ## Calculate the comparisons (slow!)
  ba_within.list <- vector(mode = "list", length = nrow(ba_within.df))
  pb <- txtProgressBar(min = 0, max = nrow(ba_within.df), style = 3)
  
  for (j in 1:nrow(ba_within.df)) {
    
    setTxtProgressBar(pb, j)
    
    # Subset data for both trips
    trip1 <- subset(my_gps, ID == ba_within.df$tripID_1[j])
    trip1.stt <- min(trip1$datetime)
    trip2 <- subset(my_gps, ID == ba_within.df$tripID_2[j])
    trip2.stt <- min(trip2$datetime)
    
    # Combine trips and factorize ID
    both_trips <- rbind(trip1, trip2) 
    both_trips %<>% 
      mutate(id = as.factor(both_trips$tripID)) %>%
      select(c(longitude, latitude, id))
    
    # Convert to SpatialPointsDataFrame
    coordinates(both_trips) <- ~longitude + latitude
    proj4string(both_trips) <- CRS(proj.dec)
    
    # Get bounding box
    bb <- bbox(both_trips)

    ## Define a raster grid extent slightly larger than the bounding box
    xrange <- bb[1, 2] - bb[1, 1]
    yrange <- bb[2, 2] - bb[2, 1]

    ## Expand the bounding box slightly (e.g. 20% buffer)
    bb_expanded <- bb
    bb_expanded[1, ] <- bb[1, ] + c(-0.1 * xrange, 0.1 * xrange)
    bb_expanded[2, ] <- bb[2, ] + c(-0.1 * yrange, 0.1 * yrange)

    # Get the UDs
    ud <- kernelUD(both_trips, h = "href", grid = 100, extent = 0, kern = "bivnorm", same4all = TRUE)

    # Calculate UD overlap — Bhattacharyya’s affinity
    overlap_val <- kerneloverlaphr(ud, method = "BA", percent = 95)[1,2]

    # Output
    output <- data.frame(trip1 = ba_within.df$tripID_1[j],
                         trip2 = ba_within.df$tripID_2[j],
                         ba_val = overlap_val,
                         inter_obs_dist = abs(as.numeric(difftime(trip1.stt, trip2.stt, units = "days"))))
  
    ba_within.list[[j]] <- output 
    
  }
  
  close(pb)

  ## Bind up the output
  ba_within.output <- do.call("rbind", ba_within.list)
  
  ba_within.output %<>% 
    separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
    separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
    select(-loc) %>%
    mutate(tripID_1 = paste0("trip_", tripID_1),
           tripID_2 = paste0("trip_", tripID_2),
           comp_ind = "within_ring",
           comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))

  rm(ba_within.df); rm(ba_within.list)
  gc()
  
  # ===# BETWEEN INDIVIUDAL ====
  
  set.seed(817)
  
  # Downsample the between dataset to match within 
  n_within_between = nrow(ba_within.output %>% filter(comp_season == "between_season"))
  n_within_within = nrow(ba_within.output %>% filter(comp_season == "within_season"))
  
  ba_between_between.df <- trip_combinations.df %>%
    separate(tripID_1, into = c("ring1", "season1", "dummy1", "trip1"), remove = FALSE) %>%
    separate(tripID_2, into = c("ring2", "season2", "dummy2", "trip2"), remove = FALSE) %>%
    filter(ring1 != ring2) %>%
    filter(season1 != season2) %>%
    select(c(tripID_1, tripID_2)) %>%
    # Get phase and filter for same breeding phase
    left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_1 = ID), by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID), by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2)) %>%
    sample_n(size = n_within_between, replace = FALSE)
  
  
  ba_between_within.df <- trip_combinations.df %>%
    separate(tripID_1, into = c("ring1", "season1", "dummy1", "trip1"), remove = FALSE) %>%
    separate(tripID_2, into = c("ring2", "season2", "dummy2", "trip2"), remove = FALSE) %>%
    filter(ring1 != ring2) %>%
    filter(season1 == season2) %>%
    select(c(tripID_1, tripID_2)) %>%
    # Get phase and filter for same breeding phase
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_1 = ID), by = "tripID_1") %>%
    rename(phase_1 = phase) %>%
    left_join(my_meta %>% select(ID, phase) %>% 
                rename(tripID_2 = ID), by = "tripID_2") %>%
    rename(phase_2 = phase) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2)) %>%
    sample_n(size = n_within_within, replace = FALSE)

  ba_between <- rbind(ba_between_between.df, ba_between_within.df)
  
  ## Calculate the comparisons
  ba_between.list <- vector(mode = "list", length = nrow(ba_between))
  pb <- txtProgressBar(min = 0, max = nrow(ba_between), style = 3)
  
  for (k in 1:nrow(ba_between)) {
    
    setTxtProgressBar(pb, k)
    
    # Subset data for both trips
    trip1 <- subset(my_gps, ID == ba_between$tripID_1[k])
    trip1.stt <- min(trip1$datetime)
    trip2 <- subset(my_gps, ID == ba_between$tripID_2[k])
    trip2.stt <- min(trip2$datetime) 
    
    # Combine trips and factorize ID
    both_trips <- rbind(trip1, trip2) 
    both_trips %<>% 
      mutate(id = as.factor(both_trips$tripID)) %>%
      select(c(longitude, latitude, id))
    
    # Convert to SpatialPointsDataFrame
    coordinates(both_trips) <- ~longitude + latitude
    proj4string(both_trips) <- CRS(proj.dec)
    
    # Get bounding box
    bb <- bbox(both_trips)

    ## Define a raster grid extent slightly larger than the bounding box
    xrange <- bb[1, 2] - bb[1, 1]
    yrange <- bb[2, 2] - bb[2, 1]

    ## Expand the bounding box slightly (e.g. 20% buffer)
    bb_expanded <- bb
    bb_expanded[1, ] <- bb[1, ] + c(-0.1 * xrange, 0.1 * xrange)
    bb_expanded[2, ] <- bb[2, ] + c(-0.1 * yrange, 0.1 * yrange)

    # Get the UDs
    ud <- kernelUD(both_trips, h = "href", grid = 100, extent = 0, kern = "bivnorm", same4all = TRUE)

    # Calculate UD overlap — Bhattacharyya’s affinity
    overlap_val <- kerneloverlaphr(ud, method = "BA", percent = 95)[1,2]

    # Output
    output <- data.frame(trip1 = ba_between$tripID_1[k],
                         trip2 = ba_between$tripID_2[k],
                         ba_val = overlap_val,
                         inter_obs_dist = abs(as.numeric(difftime(trip1.stt, trip2.stt, units = "days"))))
  
    ba_between.list[[k]] <- output 
    
  }
  
  close(pb)

  ba_between.output <- do.call("rbind", ba_between.list)
  
  ba_between.output %<>% 
    separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
    separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
    select(-loc) %>%
    mutate(tripID_1 = paste0("trip_", tripID_1),
           tripID_2 = paste0("trip_", tripID_2),
           comp_ind = "between_ring",
           comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))
  
  # Bind the two dataframes together
  ba.df <- rbind(ba_within.output, ba_between.output)
  
  save(ba.df, file = paste0("Data_outputs/", species, "_", colony, "_ba_comparison.RData"))
  
  rm(ba.df); rm(ba_within.output); rm(ba_between.list); rm(ba_between.output)
  gc()

}
  
```

### Modelling BA differences

```{r model_ba_differences, eval = FALSE}

# Species-colony combinations
spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")

# Output containers
ba_summary.list <- list()
ba_data.list <- list()
ba_means.list <- list()

for (i in seq_along(spec_col)) {
  
  # Load data and prep species/colony labels
  load(paste0("Data_outputs/", spec_col[i], "_ba_comparison.RData"))
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  meta_info <- tibble(species, colony)
  
  # Make dyad ID and subsample to ensure each comparison appears only once (circumvents random effects issues)
  set.seed(817) 
  
  ba.df %<>%
    mutate(ring_season_1 = paste(ring_1, season_1, sep = "_"),
           ring_season_2 = paste(ring_2, season_2, sep = "_"),
           dyad = if_else(ring_season_1 < ring_season_2,
                          paste(ring_season_1, ring_season_2, sep = "_"),
                          paste(ring_season_2, ring_season_1, sep = "_"))) %>%
    group_by(dyad) %>%
    slice_sample(n = 1) %>%
    ungroup() %>%
    select(-c(ring_season_1, ring_season_2))
  
  ba_data.list[[i]] <- ba.df %>%
    mutate(species = species, colony = colony)
  
  ## Relevel factors for comparison
  ba.df %<>% mutate(comp_ind = as.factor(comp_ind), comp_season = as.factor(comp_season))
  
  ba.df$comp_ind <- relevel(ba.df$comp_ind, ref = "within_ring")
  ba.df$comp_season <- relevel(ba.df$comp_season, ref = "within_season")

  ## Add small constant to handle 0 values
  n <- nrow(ba.df)
  ba.df$ba_val <- (ba.df$ba_val * (n - 1) + 0.5) / n
  
  ## Remove within ind, within season component for WAAL Bird IS
  if(i == 3) { 
    ba.df %<>% filter(!(comp_ind == "within_ring" & comp_season == "between_season"))
    ba.df <- droplevels(ba.df)
    
    m1 <- betareg(ba_val ~ comp_ind + comp_season + inter_obs_dist, data = ba.df) } else {
        m1 <- betareg(ba_val ~ comp_ind * comp_season + inter_obs_dist, data = ba.df)
    }
 
  ## Diagnostics
  # plot(m1, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
  # qqnorm(resid(m1))
  # qqline(resid(m1))
  # hist(resid(m1), breaks = 50, main = "Histogram of residuals")

  ## Extract outputs
  m1_em <- data.frame(emmeans(m1, ~comp_ind*comp_season, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring"))) %>%
    relocate(c(species, colony), .before = comp_ind) %>%
    arrange(comp_ind)
  
  m1_em.comp_ind <- data.frame(emmeans(m1, ~comp_ind, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring")),
           comp_season = NA) %>%
    relocate(c(species, colony), .before = comp_ind) %>%
    arrange(comp_ind)
  
  m1_em.comp_season <- data.frame(emmeans(m1, ~comp_season, pbkrtest.limit = 7000)) %>%
    mutate(species = species, colony = colony,
           comp_ind = NA) %>%
    relocate(c(species, colony), .before = comp_season) %>%
    arrange(comp_season)

  m1_em <- rbind(m1_em, m1_em.comp_ind, m1_em.comp_season)
  m1_out <- extract_output.lm(m1) %>% mutate(species = species, colony = colony) %>% relocate(c(species, colony), .before = Term)
  
  # Store results
  ba_summary.list[[i]] <- m1_out
  ba_means.list[[i]] <- m1_em
  
}

# Combine and save
ba_summary <- bind_rows(ba_summary.list)
ba_data    <- bind_rows(ba_data.list)
ba_means   <- bind_rows(ba_means.list)

save(ba_summary, file = "Data_outputs/ba_model_summary.RData")
save(ba_data,    file = "Data_outputs/ba_data_combined.RData")
save(ba_means,   file = "Data_outputs/ba_means.RData")

```

### Plot BA differences

```{r plot_ba_diffs, fig.dim=c(6,6)}

load("Data_outputs/ba_means.RData")
ba_means %<>% filter(!is.na(comp_season)) %>% filter(!is.na(comp_ind))

## Drop within ind, between seasons for WAAL Bird Island
ba_means %<>% filter(!(species == "waal" & colony == "birdis" & comp_ind == "within_ring" & comp_season == "between_season"))

# Data wrangling for figure
ba_plot <- ba_means %>%
  mutate(species = case_when(
          species == "bbal" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "ker" ~ "Kerguelen",
      colony == "cro" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = " ") ) %>%
  rename(lower.CL = asymp.LCL,
         upper.CL = asymp.UCL)

png(filename = "Figures/spatial_repeatability_ba_comparison.png", width = 6, height = 6, units = "in", res = 900)
ggplot(ba_plot, aes(x = comp_ind, y = emmean, colour = comp_season)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL),
                width = 0.1) +
  geom_line(aes(group = comp_season)) +
  labs(x = "", colour = "", y = "Bhattacharyya’s affinity") +
  facet_wrap(~spec_col) +
  scale_colour_manual(
    values = c(col_within, col_between),
    labels = c("Within season", "Across seasons") ) +
  scale_x_discrete(labels = c("Within individual", "Among individual")) +
  theme_bw() +
  theme(legend.position = c(0.36, 0.97),
        legend.background = element_rect(fill = "transparent", colour = NA),
        legend.key = element_rect(fill = "transparent", colour = NA))
dev.off()


```

```{r BA_similarity_table, echo = FALSE}

load("Data_outputs/ba_model_summary.RData")

# Reformat the table
ba_summary %<>%
 mutate(Species = toupper(species),
        Colony = ifelse(colony == "birdis", "Bird Island",
                         ifelse(colony == "cro", "Crozet",
                                "Kerguelen")), 
        Term = case_when(
            Term == "(Intercept)"  ~ "Intercept",
            Term == "comp_seasonbetween_season" ~ "Across seasons",
            Term == "comp_indbetween_ring" ~ "Among individuals",
            Term == "comp_indbetween_ring:comp_seasonbetween_season" ~ "Among individuals * Across seasons",
            TRUE ~ "Inter-observation distance"),
         Estimate = paste0(signif(Estimate, 4), " \u00B1 ", signif(Std_Error, 4)),
         t_value = signif(t_value, 4),
         p_value = ifelse(p_value < 0.0001, "0", format(round(p_value, 4), nsmall = 4, scientific = FALSE))) %>%
  mutate(Term = factor(Term, levels = c(
    "Intercept",
    "Among individuals",
    "Across seasons",
    "Among individuals * Across seasons",
    "Inter-observation distance"))) %>%
  rename('t value' = t_value, 'p value' = p_value) %>%
  select(-c(species, colony, Std_Error)) %>%
  relocate(c(Species, Colony), .before = Term) %>%
  arrange(Species, Colony, Term) %>%
  mutate(CI = paste0("[", signif(CI_lower, 4), ", ", signif(CI_upper, 4), "]")) %>%
  relocate(CI, .before = "t value") %>%
  select(-c(CI_lower, CI_upper))
         
knitr::kable(ba_summary, format = "html", 
             caption = '<span style="font-size: 12px;">Table 3. Model outputs for BA comparisons. Bhattacharyyas Affinity measures overlap in two utilisation distributions.".</span>', digits = 3) %>%
  row_spec(which(ba_summary$`p value` < 0.05 & ba_summary$Term != "Intercept"), bold = TRUE) %>%
  collapse_rows(columns = c(1, 2)) %>%
  kable_styling(font_size = 9)


```

