---
title: "Habitat selection analyses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---

**Note:** Naming conventions for files:

-   Bird Island = birdis

-   Crozet= cro

-   Kerguelen = ker

-   Chlorophyll = chl

Inputs:

-   *species_colony*\_sst-bathy-chla_data_population_level.RData

Outputs:

-   *species_colony*\_habitat_data.RData

-   Figure 4, Figure 5, Table 4, Table 5

-   *species_colony*\_habitat_used_variances.RData

-   *species_colony*\_habitat_used_ranef.Rdata

-   *species_colony*\_glmm\_*[between/within]*\_*environment*.RData

-   *species_colony*\_habitat_slopes.RData

-   *species_colony*\_pedigree.csv

# Set-up

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Functions for GPS processing and plotting
source("RPT_functions.R")
source("condR.R")

## MasterBayes installation
# install.packages(c('genetics', 'gtools', 'kinship2'))
# install.packages("https://cran.r-project.org/src/contrib/Archive/MasterBayes/MasterBayes_2.58.tar.gz", repos = NULL, type = "source", dependencies = T)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "tidyr", "lubridate", "DHARMa", "data.table", "ggpubr",  "kableExtra", "ggridges", "corrplot", "pROC", "ggeffects", "glmmTMB", "brms", "MASS", "MCMCglmm", "igraph", "MasterBayes", "pheatmap", "nadiv") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Scale function
scale_2SD <- function(variable) { (variable - mean(variable, na.rm = TRUE)) / sd(variable, na.rm = TRUE) * 2 }

# Set colours
bathy_col = "#0072B2" #"#1E88E5"
chl_col = "#009E73" #"#004D40"
sst_col = "#F0E442" #"#D81B60"

col_between <- "#91c5f2"
col_within <- "#0b3860"

# Heritability priors

# Set priors
priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(exponential(1), class = "sd", group = "ring"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(exponential(1), class = "sigma") )

```



# Wandering albatrosses

## Bird Island

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )


# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/waal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r waal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  print(paste0("Processing ", env_vars[i]))
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season)")), data = pnts.used)
  
  # Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = 500)
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

waal_birdis_var_used <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(waal_birdis_var_used, file = "Data_outputs/waal_birdis_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/waal_birdis_habitat_used_ranef.RData")


```

#### Plot 'used' repeatability

```{r waal_birdis_used_rpt}

load("Data_outputs/waal_birdis_habitat_used_variances.RData")

# Wrangle the data into long form
waal_birdis_var_used.long <- waal_birdis_var_used %>%
      pivot_longer(cols = matches("_(within|across)$"),
                   names_to = c(".value", "group"),
                   names_pattern = "(.*)_(within|across)") %>%
  select(-c(Effect, Variance, `Std. dev`)) %>%
  distinct()

# Change factors
waal_birdis_var_used.long %<>%
 mutate(Environment = recode(Environment,
      "bathy" = "Bathymetry",
      "chlA" = "Chlorophyll",
      "SST" = "SST")) %>%
  arrange(Environment) %>%
  mutate(Environment = factor(Environment, levels = rev(unique(Environment))))

ggplot(waal_birdis_var_used.long, aes(x = rpt, y = Environment, colour = group)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = LCL, xmax = UCL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Repeatability (95% CI)") +
  scale_colour_manual(
    values = c(col_within, col_between),
    labels = c("Within season", "Across seasons") ) +
  theme_bw() +
  theme(legend.position = "none")

```

#### Plot random intercepts

```{r waal_birdis_habitat_used_ranef, fig.dim = c(10,5)}

load("Data_outputs/waal_birdis_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
waal_birdis_used_plot <- annotate_figure(waal_birdis_used_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_used_plot, file = "Data_outputs/waal_birdis_used_plot.RData")

png(filename = "Figures//waal_birdis_habitat_used_ranef_plot.png", width = 14, height = 6, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


load("Data_outputs/waal_birdis_habitat_used_variances.RData")

head(waal_birdis_var_used)

```

### Habitat selection

#### Fit the models

```{r waal_birdis_fit_models, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

# Within season #

waal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_bathy, file = "Data_outputs/waal_birdis_glmm_within_bathy.RData")
rm(waal_birdis_glmm.within_bathy); gc()


waal_birdis_glmm.within_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_chl, file = "Data_outputs/waal_birdis_glmm_within_chl.RData")
rm(waal_birdis_glmm.within_chl); gc()


waal_birdis_glmm.within_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),      
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_sst, file = "Data_outputs/waal_birdis_glmm_within_sst.RData")
rm(waal_birdis_glmm.within_sst); gc()


# Between season #

waal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0 + bathy_scaled | ID ) + (1 | ID ) + (0 + bathy_scaled | ID ),
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.between_bathy, file = "Data_outputs/waal_birdis_glmm_between_bathy.RData")
rm(waal_birdis_glmm.between_bathy); gc()


waal_birdis_glmm.between_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0 + chl_scaled_log | ID ) + (1 | ID ) + (0 + chl_scaled_log | ID ),
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.between_chl, file = "Data_outputs/waal_birdis_glmm_between_chl.RData")
rm(waal_birdis_glmm.between_chl); gc()


waal_birdis_glmm.between_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0 + sst_scaled | ID ) + (1 | ID ) + (0 + sst_scaled | ID ),
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.between_sst, file = "Data_outputs/waal_birdis_glmm_between_sst.RData")
rm(waal_birdis_glmm.between_sst); gc()

```


#### Conditional repeatability

```{r waal_birdis_conditional_repeatability}

load("Data_inputs/waal_birdis_habitat_data.RData")

env_vars = c("sst", "bathy", "chl")

# [ 1 ] Calculate conditional repeatability ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_IDseason <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
sst_scaled <- unlist(pnts_all$sst_scaled)
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled_log)

var_map <- list(
  sst = sst_scaled,
  bathy = bathy_scaled,
  chl = chl_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_birdis_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Bootstrap conditional repeatability uncertainty ====
  
  ## ~ Between ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)) %>% 
    ungroup() %>%
    mutate(level = "between")
  
  
  ## ~ Within ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])   
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_IDseason) / df_IDseason
    var_slope_sim <- var_slope * rchisq(1, df = df_IDseason) / df_IDseason
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  
  # Estimate conditional repeatability ====
  
  ## ~ Within ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var_between = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Bird Island", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
   
  ## ~ Between ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Bird Island", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("waal_birdis_glmm.within_", env_vars[i]))
  rm(list = paste0("waal_birdis_glmm.between_", env_vars[i]))
  gc()
  
}

waal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(waal_birdis_condR_results, file = "Data_outputs/waal_birdis_condR_results.RData")

```

```{r waal_birdis_plot_slopes}

load("Data_inputs/waal_birdis_habitat_data.RData")
env_vars = c("sst", "bathy", "chl")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_birdis_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between)
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID, colour = line_colour), alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
   
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_birdis_rpt_slopes_plot}

# [ 3 ] Final plot ----

waal_birdis_pref_plot <- ggarrange(bathy_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
          bathy_condR_plot + labs(x = "Scaled bathymetry (m below sea level)") + ylim(0, 0.25),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 0.25) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 0.25) + theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)

waal_birdis_pref_plot <- annotate_figure(waal_birdis_pref_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_pref_plot, file = "Data_outputs/waal_birdis_selection_plot.RData")


png("Figures/waal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
waal_birdis_pref_plot
dev.off()



```

### Heritability

#### Used habitat

Change to individual intercept and try again

```{r waal_birdis_heritability_used }

# (1) Load the data ----
load("Data_inputs/waal_birdis_habitat_data.RData")
pnts.used <- pnts_all %>% filter(used == 1)

pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)


# (2) Check the pedigree -----

## Are the ID in pnts.used the same as id in pedigree.df?
pedigree.df$id <- as.character(pedigree.df$id)
pnts.used$ID <- as.character(pnts.used$ID)

model_ids <- unique(pnts.used$ID)
pedigree_ids <- pedigree.df$id

### IDs in model AND pedigree
ids_in_both <- intersect(model_ids, pedigree_ids)

### IDs in model NOT in pedigree
ids_missing_from_pedigree <- setdiff(model_ids, pedigree_ids)


# (3) Subset the pedigree to include all ancestors ----

## Check formatting
pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

ped_ids_needed <- get_ancestors(model_ids, pedigree.df)
length(ped_ids_needed)

## Subset the pedigree
pedigree.pruned <- pedigree.df[pedigree.df$id %in% ped_ids_needed, ]

## Check for missing IDs - NONE, so end here
setdiff(model_ids, pedigree.pruned$id) 


# (4) Make the genetic matrix A ----

ped <- pedigree.pruned
names(ped) <- c("animal", "dam", "sire")

# Deal with dams coming before offspring
ped <- orderPed(ped)
Amat <- as.matrix(nadiv::makeA(ped))


# (5) Fit multivariate model covering responses to each environment ----

pnts.used %<>% mutate(ring_pe = ID, animal = ID)

# Build models
waal_birdis_brm.heritability_used_bathy <- brm(
  bathy ~ 1 + phase + (1 | ID_season ) + (1 | tripID) + (1 | gr(animal, cov = Amat)) ,
  prior = priors,
  data = pnts.used,
  data2 = list(Amat = Amat),
  cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95) ) 

save(waal_birdis_brm.heritability_used_bathy, file = "Data_outputs/waal_birdis_heritability_used_bathy.RData")

waal_birdis_brm.heritability_used_chl <- brm(
  chlA ~ 1 + phase + (1 | season) + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors,
  data = pnts.used,
  data2 = list(Amat = Amat),
  chains = 2, cores = 2, iter = 4000 ) 

save(waal_birdis_brm.heritability_used_bathy, file = "Data_outputs/waal_birdis_heritability_used_chl.RData")

waal_birdis_brm.heritability_used_sst <- brm(
  SST ~ 1 + phase + (1 | season) + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors,
  data = pnts.used,
  data2 = list(Amat = Amat),
  chains = 2, cores = 2, iter = 4000 ) 

save(waal_birdis_brm.heritability_used_bathy, file = "Data_outputs/waal_birdis_heritability_used_sst.RData")


# Estimate heritability

get_h2 <- function(model) {
  
  v_animal <- (VarCorr(model, summary = FALSE)$animal$sd)^2
  v_r <- (VarCorr(model, summary = FALSE)$residual$sd)^2
  
  h2_est <- as.mcmc(v_animal[, 1] / (v_animal[, 1] + v_r[, 1]))
  
  return(h2_est)
  
}

v_animal <- (VarCorr(waal_birdis_brm.heritability_used, summary = FALSE)$animal$sd)^2
v_r <- (VarCorr(waal_birdis_brm.heritability_used, summary = FALSE)$residual$sd)^2

h2_bathy <- as.mcmc(v_animal[, 1] / (v_animal[, 1] + v_r[, 1]))
h2_chl <- as.mcmc(v_animal[, 2] / (v_animal[, 2] + v_r[, 2]))
h2_sst <- as.mcmc(v_animal[, 3] / (v_animal[, 3] + v_r[, 3]))

summary(h2_bathy)
plot(h2_bathy)

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/waal_birdis_habitat_used_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()


```

#### Extract individual slopes

```{r waal_birdis_extract_slopes}

load("Data_inputs/waal_birdis_habitat_data.RData")

## Store envronmental variables
env_vars = c("sst", "bathy", "chl")

## Get meta to recover later
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

slopes.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars[i], ".RData"))
  my_mod = get(paste0("waal_birdis_glmm.within_", env_vars[i]))
  
  # Extract slopes - mean and SE
  slopes <- ranef(my_mod)$cond$tripID
  
  slopes_df <- data.frame(
        ID = rownames(slopes),
        slope = slopes[,2])

  slopes_df %<>% 
    left_join(pnts_meta %>% rename(ID = tripID), by = "ID") %>%
    separate(ID, into = c("ring", "season", "tripID"), remove = FALSE) %>%
    mutate(env = env_vars[i]) %>% 
    relocate(c(ring, season, tripID))

  slopes.list[[i]] <- slopes_df
  
}

waal_birdis_habitat_slopes <- do.call("rbind", slopes.list)
save(waal_birdis_habitat_slopes, file = "Data_outputs/waal_birdis_habitat_slopes.RData")


```

#### Estimate heritability

Try and run each model separately

```{r waal_birdis_heritability_pref }

# (1) Load the data ----
load("Data_outputs/waal_birdis_habitat_slopes.RData")

pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(waal_birdis_habitat_slopes, pedigree.df)

## Optional - visualise the pruned peigree (go into function to get pruned pedigree)

# edges <- pedigree.pruned %>%
#   tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
#   dplyr::select(from = value, to = id)
# 
# ped_graph <- graph_from_data_frame(edges, directed = TRUE)
# 
# plot(ped_graph,
#      vertex.size = 3,
#      vertex.label = NA,
#      edge.arrow.size = 0.2,
#      layout = layout_with_sugiyama(ped_graph)$layout)


## Optional - visualise sparse matrix

# A_sparse <- as(Amat, "dgCMatrix")
# summary(A_sparse)
#
### Check how many non-zero off-diagonal elements
# nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
# density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
# cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")
# 
# pheatmap(as.matrix(Amat), show_rownames = FALSE, show_colnames = FALSE)


# (3) Fit multivariate model covering responses to each environment ----
waal_birdis_habitat_slopes %<>%
  mutate(animal = ring, 
         tripID = ID, 
         ring_pe = ring)

## Make dataframe wide (i.e. slope per environment)
waal_birdis_slopes.wide <- waal_birdis_habitat_slopes %>%
  pivot_wider(names_from = env, 
              values_from = slope,
              names_glue = "{env}_slope")

## Only retain birds with > 1 trip
waal_birdis_slopes.wide %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% ungroup() %>% select(-n_trips)

# Build univariate models and extract h2
waal_birdis_brm.heritability_bathy <- brm(
  bathy_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_bathy <- (VarCorr(waal_birdis_brm.heritability_bathy, summary = FALSE)$animal$sd)^2
v_r_bathy <- (VarCorr(waal_birdis_brm.heritability_bathy, summary = FALSE)$residual$sd)^2
h2_bathy <- as.mcmc(v_animal_bathy / (v_animal_bathy + v_r_bathy) )

save(waal_birdis_brm.heritability_bathy, file = "Data_outputs/waal_birdis_habitat_pref_heritability_bathy.RData")


waal_birdis_brm.heritability_chl <- brm(
  chl_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_chl <- (VarCorr(waal_birdis_brm.heritability_chl, summary = FALSE)$animal$sd)^2
v_r_chl <- (VarCorr(waal_birdis_brm.heritability_chl, summary = FALSE)$residual$sd)^2
h2_chl <- as.mcmc(v_animal_chl / (v_animal_chl + v_r_chl) )

save(waal_birdis_brm.heritability_chl, file = "Data_outputs/waal_birdis_habitat_pref_heritability_chl.RData")


waal_birdis_brm.heritability_sst <- brm(
  sst_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_sst <- (VarCorr(waal_birdis_brm.heritability_sst, summary = FALSE)$animal$sd)^2
v_r_sst <- (VarCorr(waal_birdis_brm.heritability_sst, summary = FALSE)$residual$sd)^2
h2_sst <- as.mcmc(v_animal_sst / (v_animal_sst + v_r_sst) )

save(waal_birdis_brm.heritability_sst, file = "Data_outputs/waal_birdis_habitat_pref_heritability_sst.RData")


posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
colnames(posteriors) <- c("h2_bathy", "h2_chl", "h2_sst")
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/waal_birdis_habitat_pref_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```

## Crozet

Chlorophyll values are heavily skewed, with extremely high
concentrations occurring near coastlines in the ‘available’ dataset. To
prevent these large outliers from unduly influencing the model results,
I apply a log-transformation to chlorophyll.

```{r waal_cro_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_cro_sst-bathy-chla_data_population_level.RData")

## Fix some rings
pnts_all %<>%
  mutate(ring = recode(ring,
                       "11376" = "BS11376",
                       "11378" = "BS11378",
                       "25076" = "BS25076",
                       "8196" = "BS28916",
                       "8337" = "BS28337",
                       "BS8337" = "BS28337",
                       "BB21651" = "BS81651",
                       "BS7625" = "BS27625",
                       "BS8402" = "BS28402",
                       "BS8404" = "BS28404",
                       "BS9084" = "BS19084",
                       "BS9444" = "BS19444"))

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )

save(pnts_all, file = "Data_inputs/waal_cro_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r waal_cro_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.used)
  
  ## Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = 500)
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  
  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

waal_cro_var_used <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(waal_cro_var_used, file = "Data_outputs/waal_cro_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/waal_cro_habitat_used_ranef.RData")

```

#### Plot random intercepts

```{r waal_cro_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/waal_cro_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_cro_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
waal_cro_used_plot <- annotate_figure(waal_cro_used_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_used_plot, file = "Data_outputs/waal_cro_used_plot.RData")

png(filename = "Figures/waal_cro_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r waal_cro_table_variance_habitat_used}

load("Data_outputs/waal_cro_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

# Add whitespace to avoid collapsing bathymetry & chlorophyll together
var.table$`Across season rpt`[6:10] <- paste0(var.table$`Across season rpt`[6:10], ".00")

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

The random slope for chlorophyll-a was retained to account for potential
individual variation in response to environmental conditions. However,
the estimated correlation between the random intercept and slope for
chl_scaled was extremely high (0.99), suggesting that these components
are statistically confounded in the current model. This may reflect
either collinearity in the data or insufficient within-individual
variation in chlorophyll exposure to reliably separate baseline use from
individual responses.

As a result, I have droppped the polynomial terms as I think the models
are unable to cope with this degree of complexity.

```{r waal_cro_fit_models, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

# try logging chl-A

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

# Within season #

waal_cro_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),              
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_bathy, file = "Data_outputs/waal_cro_glmm_within_bathy.RData")
rm(waal_cro_glmm.within_bathy); gc()


waal_cro_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),         
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_chl, file = "Data_outputs/waal_cro_glmm_within_chl.RData")
rm(waal_cro_glmm.within_chl); gc()


waal_cro_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),    
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_sst, file = "Data_outputs/waal_cro_glmm_within_sst.RData")
rm(waal_cro_glmm.within_sst); gc()

# Between season #

waal_cro_glmm.between_bathy <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + bathy_scaled | ID) + (1 | ID) + (0  + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_bathy, file = "Data_outputs/waal_cro_glmm_between_bathy.RData")
rm(waal_cro_glmm.between_bathy); gc()


waal_cro_glmm.between_chl <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + chl_scaled_log | ID) + (1 | ID) + (0  + chl_scaled_log | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_chl, file = "Data_outputs/waal_cro_glmm_between_chl.RData")
rm(waal_cro_glmm.between_chl); gc()

waal_cro_glmm.between_sst <-  glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + sst_scaled | ID) + (1 | ID) + (0  + sst_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_sst, file = "Data_outputs/waal_cro_glmm_between_sst.RData")
rm(waal_cro_glmm.between_sst); gc()


```

#### Diagnostics

```{r waal_cro_diagnostics, eval = FALSE}

env_vars = c("sst", "bathy", "chl")

auc_results <- data.frame(
  model = character(),
  auc = numeric()  )

for (i in 1:length(env_vars)) {

  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))

  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))

  # print("Covariance structure: ")
  # print(VarCorr(mod_between))
  # print(VarCorr(mod_within))
  # 
  # print("Collinearity: ")
  # print(performance::check_collinearity(mod_between))
  # print(performance::check_collinearity(mod_within))
  # 
  # print(paste0("DONE. ---------------------------"))
  
  preds.between <- predict(mod_between, type = "response")
  auc_value.between <- as.numeric(auc(pnts_all$used, preds.between))
  
  preds.within <- predict(mod_within, type = "response")
  auc_value.within <- as.numeric(auc(pnts_all$used, preds.within))
  
  auc_between <- cbind(auc_value.between, "between")
  auc_within <- cbind(auc_value.within, "within")
  
  waal_cro_auc_results <- data.frame(rbind(auc_between, auc_within)) %>% 
                              mutate(env = env_vars[i]) %>%
                              rename(level = V2, auc = auc_value.between) %>%
                              relocate(auc, .after = env)
  
}

waal_cro_auc_results %<>% mutate(species = "WAAL", colony = "crozet") %>% relocate(c(species, colony), .before = level)
save(waal_cro_auc_results, file = "Data_outputs/waal_cro_auc_habitat_selection.RData")

```

#### Fixed effects

```{r waal_cro_fixed_effects_predictions }

load("Data_outputs/waal_cro_glmm_between_sst.RData")
load("Data_outputs/waal_cro_glmm_between_bathy.RData")
load("Data_outputs/waal_cro_glmm_between_chl.RData")


# Plot effects
pred_plot.sst <-   plot(ggeffects::ggpredict(waal_cro_glmm.between_sst, terms = "sst_scaled [all]", type = "fixed"))
pred_plot.bathy <- plot(ggeffects::ggpredict(waal_cro_glmm.between_bathy, terms = "bathy_scaled [all]", type = "fixed"))
pred_plot.chl <-  plot(ggeffects::ggpredict(waal_cro_glmm.between_chl, terms = "chl_scaled [all]", type = "fixed"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)", y = "Used")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)", y = "Used") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")", y = "Used")) + theme(plot.title = element_blank()),
          ncol = 2, nrow = 2)

png("Figures/waal_cro_habitat_fixed_effects.png")
all_plots
dev.off()

rm(pred_plot.sst); rm(pred_plot.bathy); rm(pred_plot.chl); rm(all_plots)
gc()

```

#### Conditional repeatability

```{r waal_cro_conditional_repeatability}

load("Data_inputs/waal_cro_habitat_data.RData")

env_vars = c("sst", "bathy", "chl")

# [ 1 ] Calculate conditional repeatability ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_IDseason <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
sst_scaled <- unlist(pnts_all$sst_scaled)
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled_log)

var_map <- list(
  sst = sst_scaled,
  bathy = bathy_scaled,
  chl = chl_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics ====
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Bootstrap conditional repeatability uncertainty ====
  
  ## ~ Between ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 

  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "between")
  
  
  ## ~ Within ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])   
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_IDseason) / df_IDseason
    var_slope_sim <- var_slope * rchisq(1, df = df_IDseason) / df_IDseason
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID

    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  
  # Estimate conditional repeatability ====
  
  ## ~ Within ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var_between = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Crozet", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
   
  ## ~ Between ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Crozet", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("waal_cro_glmm.within_", env_vars[i]))
  rm(list = paste0("waal_cro_glmm.between_", env_vars[i]))
  gc()
  
}

waal_cro_condR_results <- do.call("rbind", cond_res.list)
save(waal_cro_condR_results, file = "Data_outputs/waal_cro_condR_results.RData")

```

```{r waal_cro_plot_slopes}

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between)
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_cro_rpt_slopes_plot}

# [ 3 ] Final plot ----

waal_cro_pref_plot <- ggarrange(bathy_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
          bathy_condR_plot + labs(x = "Scaled bathymetry (m below sea level)") + ylim(0, 0.6),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 0.6) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 0.6) + theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)

waal_cro_pref_plot <- annotate_figure(waal_cro_pref_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_pref_plot, file = "Data_outputs/waal_cro_selection_plot.RData")


png("Figures/waal_cro_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
waal_cro_pref_plot
dev.off()



```

### Heritability

#### Extract individual slopes

```{r waal_cro_extract_slopes}

load("Data_inputs/waal_cro_habitat_data.RData")

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

## Store envronmental variables
env_vars = c("sst", "bathy", "chl")

slopes.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  my_mod = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  
  # Extract slopes - mean and SE
  slopes <- ranef(my_mod)$cond$ID_season
  
  slopes_df <- data.frame(
        ID = rownames(slopes),
        slope = slopes[,2])

  slopes_df %<>% 
    separate(ID, into = c("ring", "season"), remove = FALSE) %>%
    mutate(env = env_vars[i]) %>% 
    relocate(c(ring, season))

  slopes.list[[i]] <- slopes_df
  
}

waal_cro_habitat_slopes <- do.call("rbind", slopes.list)
save(waal_cro_habitat_slopes, file = "Data_outputs/waal_cro_habitat_slopes.RData")

```

#### Estimate heritability

```{r waal_cro_heritability }

# (1) Load the data ----
load("Data_outputs/waal_cro_habitat_slopes.RData")

pedigree.df <- read.csv("Data_inputs/waal_cro_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(waal_cro_habitat_slopes, pedigree.df)

## Optional - visualise the pruned peigree (go into function to get pruned pedigree)

# edges <- pedigree.pruned %>%
#   tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
#   dplyr::select(from = value, to = id)
# 
# ped_graph <- graph_from_data_frame(edges, directed = TRUE)
# 
# plot(ped_graph,
#      vertex.size = 3,
#      vertex.label = NA,
#      edge.arrow.size = 0.2,
#      layout = layout_with_sugiyama(ped_graph)$layout)


## Optional - visualise sparse matrix

A_sparse <- as(Amat, "dgCMatrix")
summary(A_sparse)

### Check how many non-zero off-diagonal elements
# nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
# density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
# cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")
# 
# pheatmap(as.matrix(Amat), show_rownames = FALSE, show_colnames = FALSE)


# (3) Fit heritability models ----
waal_cro_habitat_slopes %<>%
  mutate(animal = ring, 
         ring_pe = ring)

## Make dataframe wide (i.e. slope per environment)
waal_cro_slopes.wide <- waal_cro_habitat_slopes %>%
  pivot_wider(names_from = env, 
              values_from = slope,
              names_glue = "{env}_slope")

## Only retain birds with > 1 trip
waal_cro_slopes.wide %<>% group_by(ring) %>% mutate(n_trips = n_distinct(season)) %>% filter(n_trips > 1) %>% ungroup() %>% select(-n_trips)

# Build univariate models and extract h2
waal_cro_brm.heritability_bathy <- brm(
  bathy_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_cro_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_bathy <- (VarCorr(waal_cro_brm.heritability_bathy, summary = FALSE)$animal$sd)^2
v_r_bathy <- (VarCorr(waal_cro_brm.heritability_bathy, summary = FALSE)$residual$sd)^2
h2_bathy <- as.mcmc(v_animal_bathy / (v_animal_bathy + v_r_bathy) )

save(waal_cro_brm.heritability_bathy, file = "Data_outputs/waal_cro_habitat_pref_heritability_bathy.RData")


waal_cro_brm.heritability_chl <- brm(
  chl_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_cro_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_chl <- (VarCorr(waal_cro_brm.heritability_chl, summary = FALSE)$animal$sd)^2
v_r_chl <- (VarCorr(waal_cro_brm.heritability_chl, summary = FALSE)$residual$sd)^2
h2_chl <- as.mcmc(v_animal_chl / (v_animal_chl + v_r_chl) )

save(waal_cro_brm.heritability_chl, file = "Data_outputs/waal_cro_habitat_pref_heritability_chl.RData")


waal_cro_brm.heritability_sst <- brm(
  sst_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_cro_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_sst <- (VarCorr(waal_cro_brm.heritability_sst, summary = FALSE)$animal$sd)^2
v_r_sst <- (VarCorr(waal_cro_brm.heritability_sst, summary = FALSE)$residual$sd)^2
h2_sst <- as.mcmc(v_animal_sst / (v_animal_sst + v_r_sst) )

save(waal_cro_brm.heritability_sst, file = "Data_outputs/waal_cro_habitat_pref_heritability_sst.RData")


# Extract posteriors 

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
colnames(posteriors) <- c("h2_bathy", "h2_chl", "h2_sst")
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/waal_cro_habitat_pref_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```

# Black-browed albatrosses

## Bird Island

```{r bbal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/bbal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2 )

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r bbal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.used)
  
  ## Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = 500)
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

bbal_birdis_var_used <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(bbal_birdis_var_used, file = "Data_outputs/bbal_birdis_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/bbal_birdis_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_birdis_used_plot <- annotate_figure(bbal_birdis_used_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_used_plot, file = "Data_outputs/bbal_birdis_used_plot.RData")

png(filename = "Figures/bbal_birdis_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r bbal_birdis_table_variance_habitat_used}

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

```{r bbal_birdis_fit_models, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

# Within season #

bbal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | season ) + (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_bathy, file = "Data_outputs/bbal_birdis_glmm_within_bathy.RData")
rm(bbal_birdis_glmm.within_bathy); gc()


bbal_birdis_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_chl, file = "Data_outputs/bbal_birdis_glmm_within_chl.RData")
rm(bbal_birdis_glmm.within_chl); gc()


bbal_birdis_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_sst, file = "Data_outputs/bbal_birdis_glmm_within_sst.RData")
rm(bbal_birdis_glmm.within_sst); gc()


# Between season #

bbal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID) + (0 + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_bathy, file = "Data_outputs/bbal_birdis_glmm_between_bathy.RData")
rm(bbal_birdis_glmm.between_bathy); gc()


bbal_birdis_glmm.between_sst <-  glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_birdis_glmm.between_sst, file = "Data_outputs/bbal_birdis_glmm_between_sst.RData")
rm(bbal_birdis_glmm.between_sst); gc()


bbal_birdis_glmm.between_chl <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_chl, file = "Data_outputs/bbal_birdis_glmm_between_chl.RData")
rm(bbal_birdis_glmm.between_chl); gc()

```

#### Conditional repeatability

```{r bbal_birdis_cond_repeatability}

load("Data_inputs/bbal_birdis_habitat_data.RData")

env_vars = c("bathy", "chl", "sst")

# [ 1 ] Calculate conditional repeatability ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_IDseason <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(
  bathy = bathy_scaled,
  chl = chl_scaled,
  sst = sst_scaled )

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics ====
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Bootstrap conditional repeatability uncertainty ====
  
  ## ~ Between ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 

  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "between")
  
  
  ## ~ Within ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])   
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_IDseason) / df_IDseason
    var_slope_sim <- var_slope * rchisq(1, df = df_IDseason) / df_IDseason
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID

    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  
  # Estimate conditional repeatability ====
  
  ## ~ Within ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var_between = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Bird Island", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
   
  ## ~ Between ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Bird Island", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_birdis_glmm.within_", env_vars[i]))
  rm(list = paste0("bbal_birdis_glmm.between_", env_vars[i]))
  gc()
  
}

bbal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(bbal_birdis_condR_results, file = "Data_outputs/bbal_birdis_condR_results.RData")

```

```{r bbal_birdis_plot_slopes}

# [ 2 ] Plot individual slopes ----

rnd_effs.list <- list()

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)
save(rnd_effs.out, file = "Data_outputs/bbal_birdis_habitat_selection.RData")

```

```{r bbal_birdis_rpt_slopes_plot}

# [ 3 ] Final plot ----

bbal_birdis_pref_plot <- ggarrange(bathy_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
          bathy_condR_plot + labs(x = "Scaled bathymetry (m below sea level)") + ylim(0, 1.1) ,
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 1.1) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 1.1) + theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)

bbal_birdis_pref_plot <- annotate_figure(bbal_birdis_pref_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_pref_plot, file = "Data_outputs/bbal_birdis_selection_plot.RData")


png("Figures/bbal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
bbal_birdis_pref_plot
dev.off()


```

### Heritability

#### Extract individual slopes

```{r bbal_birdis_extract_slopes}

load("Data_inputs/bbal_birdis_habitat_data.RData")

## Store envronmental variables
env_vars = c("bathy", "chl", "sst")

## Get meta to recover later
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

slopes.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  my_mod = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  
  # TESTING ----
  
  fit <- my_mod
  
  # get conditional modes & (if available) conditional variances
  re <- ranef(fit, condVar = TRUE)  
  
  # re is a list with elements for each grouping factor, e.g. re$cond$tripID
  # For trip slopes:
  trip_ranef <- re$cond$tripID   # data.frame with columns like "(Intercept)" and "bathy_scaled"
  head(trip_ranef)
  
  # extract the postVar attribute (3D array: npar x npar x nlevels)
  pv <- attr(trip_ranef, "postVar")  
  
  # sanity check
  if (is.null(pv)) stop("No 'postVar' found. Your glmmTMB version may not have condVar support; see fallback below.")
  
  # find which column index corresponds to the bathy slope
  colnames(trip_ranef)
  # find index
  slope_col <- which(colnames(trip_ranef) == "bathy_scaled")
  if (length(slope_col) != 1) stop("Couldn't find 'bathy_scaled' in trip random effects columns; adjust name.")
  
  # extract per-trip variances: postVar is array [npar, npar, nlevels]
  trip_var <- sapply(seq_len(dim(pv)[3]), function(i) pv[slope_col, slope_col, i])
  trip_se  <- sqrt(trip_var)
  
  # assemble dataframe
  trip_slopes_df <- tibble(
    tripID = rownames(trip_ranef),
    slope_hat = trip_ranef$bathy_scaled,
    slope_se  = trip_se,
    slope_var = trip_var
  )
  
  # map trip -> ID (from your data)
  trip_meta <- pnts_all %>% distinct(tripID, ID)
  slopes_per_trip <- trip_slopes_df %>% left_join(trip_meta, by = "tripID")
  
  # check for NAs or absurd SEs
  summary(slopes_per_trip$slope_se)
  sum(is.na(slopes_per_trip$slope_se))
  
  #### -----
  
  
  # Extract slopes - mean and SE
  slopes <- ranef(my_mod)$cond$tripID
  
  slopes_df <- data.frame(
        ID = rownames(slopes),
        slope = slopes[,2])

  slopes_df %<>% 
    left_join(pnts_meta %>% rename(ID = tripID), by = "ID") %>%
    separate(ID, into = c("ring", "season", "tripID"), remove = FALSE) %>%
    mutate(env = env_vars[i]) %>%
    relocate(c(ring, season, tripID))

  slopes.list[[i]] <- slopes_df
  
}

bbal_birdis_habitat_slopes <- do.call("rbind", slopes.list)
save(bbal_birdis_habitat_slopes, file = "Data_outputs/bbal_birdis_habitat_slopes.RData")

```

#### Estimate heritability

```{r bbal_birdis_heritability }

# (1) Load the data ----
load("Data_outputs/bbal_birdis_habitat_slopes.RData")

pedigree.df <- read.csv("Data_inputs/bbal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
pedigree.df$sire[pedigree.df$id == "1271009"] <- "1318880"
pedigree.df$dam[pedigree.df$id == "1271009"] <- "1318939"

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(bbal_birdis_habitat_slopes, pedigree.df)

## Optional - visualise the pruned peigree (go into function to get pruned pedigree)

# edges <- pedigree.pruned %>%
#   tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
#   dplyr::select(from = value, to = id)
# 
# ped_graph <- graph_from_data_frame(edges, directed = TRUE)
# 
# plot(ped_graph,
#      vertex.size = 3,
#      vertex.label = NA,
#      edge.arrow.size = 0.2,
#      layout = layout_with_sugiyama(ped_graph)$layout)


## Optional - visualise sparse matrix

# A_sparse <- as(Amat, "dgCMatrix")
# summary(A_sparse)

### Check how many non-zero off-diagonal elements
# nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
# density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
# cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")
# 
# pheatmap(as.matrix(Amat), show_rownames = FALSE, show_colnames = FALSE)


# (3) Fit multivariate model covering responses to each environment ----
bbal_birdis_habitat_slopes %<>%
  mutate(animal = ring, 
         tripID = ID, 
         ring_pe = ring)

## Make dataframe wide (i.e. slope per environment)
bbal_birdis_slopes.wide <- bbal_birdis_habitat_slopes %>%
  pivot_wider(names_from = env, 
              values_from = slope,
              names_glue = "{env}_slope")

## Only retain birds with > 1 trip
bbal_birdis_slopes.wide %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% ungroup() %>% select(-n_trips)

# Build univariate models and extract h2

bbal_birdis_brm.heritability_bathy <- brm(
  bathy_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = bbal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_bathy <- (VarCorr(bbal_birdis_brm.heritability_bathy, summary = FALSE)$animal$sd)^2
v_r_bathy <- (VarCorr(bbal_birdis_brm.heritability_bathy, summary = FALSE)$residual$sd)^2
h2_bathy <- as.mcmc(v_animal_bathy / (v_animal_bathy + v_r_bathy) )

save(bbal_birdis_brm.heritability_bathy, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_bathy.RData")


bbal_birdis_brm.heritability_chl <- brm(
  chl_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = bbal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_chl <- (VarCorr(bbal_birdis_brm.heritability_chl, summary = FALSE)$animal$sd)^2
v_r_chl <- (VarCorr(bbal_birdis_brm.heritability_chl, summary = FALSE)$residual$sd)^2
h2_chl <- as.mcmc(v_animal_chl / (v_animal_chl + v_r_chl) )

save(bbal_birdis_brm.heritability_chl, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_chl.RData")


bbal_birdis_brm.heritability_sst <- brm(
  sst_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = bbal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_sst <- (VarCorr(bbal_birdis_brm.heritability_sst, summary = FALSE)$animal$sd)^2
v_r_sst <- (VarCorr(bbal_birdis_brm.heritability_sst, summary = FALSE)$residual$sd)^2
h2_sst <- as.mcmc(v_animal_sst / (v_animal_sst + v_r_sst) )

save(bbal_birdis_brm.heritability_sst, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_sst.RData")


# Extract posteriors of h2

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
colnames(posteriors) <- c("h2_bathy", "h2_chl", "h2_sst")
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()


```

- Posterior distributions for h2 are biomdal, indicating stong confounding between additive genetic variance and permanent environmental (repeatability) effects
- Collapsing to one slope per bird gives a model that fails to converge with highly uncertain posteriors - i.e. when repeatability information is removed, there is insufficient signal to estimate heritability
- Overall, reasonable evidence that while there may be individual consistency, preference is unlikely to be heritable

e.g. In multivariate animal models, posterior distributions of heritability estimates for bathymetry and chlorophyll were bimodal, reflecting poor separation of additive genetic variance from permanent environmental effects. A diagnostic single-slope model, in which one slope per bird was analysed, failed to converge and yielded highly uncertain variance component estimates. Together these results suggest that while individual foraging preferences are partially repeatable, there is no strong evidence that this variation is heritable.

## Kerguelen

For Kerguelen especially, birds' preference for shelf zones makes the
bathymetry model fall apart. To tackle this, I have calculated distance
from the shelf, and split preferences into on shelf, shelf break, and
off shelf. This seems to resolve my problems. The shelf break is
delimited at 1000m following Hindell et al. 2011.

```{r bbal_ker_process_data, eval = FALSE}

load("Data_inputs/bbal_ker_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )

## Add binomial bathymetry variable
pnts_all %<>%
  mutate(dist_shelf_bin = case_when(
            dist_shelf_break_signed > 20  ~ "on_shelf",       # >20 km inside shelf
            dist_shelf_break_signed >= -20 & dist_shelf_break_signed <= 20 ~ "near_break", # ±20 km of break
            dist_shelf_break_signed < -20  ~ "off_shelf"       # >20 km offshore (deeper)
    ) )

pnts_all$dist_shelf_bin <- factor(pnts_all$dist_shelf_bin, levels = c("on_shelf", "near_break", "off_shelf"))

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_ker_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r bbal_ker_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.used)
  
  ## Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = 500)
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}


bbal_ker_var_used <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(bbal_ker_var_used, file = "Data_outputs/bbal_ker_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/bbal_ker_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_ker_plot_habitat_used }

load("Data_outputs/bbal_ker_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)

split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_ker_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_ker_used_plot <- annotate_figure(bbal_ker_used_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_used_plot, file = "Data_outputs/bbal_ker_used_plot.RData")

png(filename = "Figures/bbal_ker_habitat_used_ranef_plot.png", width = 10, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r bbal_ker_table_variance_habitat_used}

load("Data_outputs/bbal_ker_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

```{r bbal_ker_fit_model, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

## Try dropping individuals with less than a day of data
pnts_all %<>% group_by(ID) %>% mutate(n_recs = n_distinct(date_hourly)) %>% filter(n_recs > 24) %>% select(-n_recs) %>% ungroup()

# Within season #

bbal_ker_glmm.within_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID_season) + (0 + dist_shelf_bin | ID_season),              
  data = pnts_all,
  family = binomial() )

save(bbal_ker_glmm.within_bathy, file = "Data_outputs/bbal_ker_glmm_within_bathy.RData")
rm(bbal_ker_glmm.within_bathy); gc()


bbal_ker_glmm.within_chl <-  glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),           
  data = pnts_all,
  family = binomial() )

save(bbal_ker_glmm.within_chl, file = "Data_outputs/bbal_ker_glmm_within_chl.RData")
rm(bbal_ker_glmm.within_chl); gc()


bbal_ker_glmm.within_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),       
  family = binomial() )

save(bbal_ker_glmm.within_sst, file = "Data_outputs/bbal_ker_glmm_within_sst.RData")
rm(bbal_ker_glmm.within_sst); gc()


# Between season #

bbal_ker_glmm.between_bathy <- glmmTMB(
  used ~ sst_scaled + dist_shelf_bin + chl_scaled + dist_col_scaled + phase + 
         (1 | tripID) + (0 + dist_shelf_bin | ID) + (1 | ID) + (0 + dist_shelf_bin | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_ker_glmm.between_bathy, file = "Data_outputs/bbal_ker_glmm_between_bathy.RData")
rm(bbal_ker_glmm.between_bathy); gc()


bbal_ker_glmm.between_chl <- glmmTMB(
  used ~ sst_scaled + dist_shelf_bin + chl_scaled + dist_col_scaled + phase + 
         (1 | tripID) + (0 + chl_scaled | ID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_ker_glmm.between_chl, file = "Data_outputs/bbal_ker_glmm_between_chl.RData")
rm(bbal_ker_glmm.between_chl); gc()


bbal_ker_glmm.between_sst <- glmmTMB(
  used ~ sst_scaled + dist_shelf_bin + chl_scaled + dist_col_scaled + phase + 
         (1 | tripID) + (0 + sst_scaled | ID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_sst, file = "Data_outputs/bbal_ker_glmm_between_sst.RData")
rm(bbal_ker_glmm.between_sst); gc()


```

#### Diagnostics

```{r bbal_ker_diagnostics, eval = FALSE}

env_vars = c("sst", "bathy", "chl")

auc_results <- data.frame(
  model = character(),
  auc = numeric()  )

for (i in 1:length(env_vars)) {

  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))

  # print("Covariance structure: ")
  # print(VarCorr(mod_between))
  # 
  # print("Collinearity: ")
  # print(performance::check_collinearity(mod_between))
  # 
  # print(paste0("DONE. ---------------------------"))
  
  preds <- predict(mod_between, type = "response")
  auc_value <- as.numeric(auc(pnts_all$used, preds))
  
  bbal_ker_auc_results <- rbind(auc_results, data.frame(level = "between", env = env_vars[i], auc = auc_value))
  
}

bbal_ker_auc_results %<>% mutate(species = "BBAL", colony = "kerguelen") %>% relocate(c(species, colony), .before = level)
save(bbal_ker_auc_results, file = "Data_outputs/bbal_ker_auc_habitat_selection.RData")

```

#### Fixed effects

```{r bbal_ker_fixed_effects_predictions }

load("Data_outputs/bbal_ker_glmm_between_sst.RData")
load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
load("Data_outputs/bbal_ker_glmm_between_chl.RData")

# Plot effects
pred_plot.sst <-   plot(ggeffects::ggpredict(bbal_ker_glmm.between_sst, terms = "sst_scaled [all]", type = "fixed"))
pred_plot.bathy <- plot(ggeffects::ggpredict(bbal_ker_glmm.between_bathy, terms = "bathy_scaled [all]", type = "fixed"))
pred_plot.chl <-   plot(ggeffects::ggpredict(bbal_ker_glmm.between_chl, terms = "chl_scaled [all]", type = "fixed"))

all_plots <- ggarrange(pred_plot.sst + 
                         labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)", y = "Used")) +
                         theme(plot.title = element_blank()), 
                       pred_plot.bathy + 
                         labs(x = "Scaled bathymetry (m below sea level)", y = "Used") + 
                         theme(plot.title = element_blank(),
                               axis.title.y = element_blank()),
                      pred_plot.chl + 
                        labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")", y = "Used")) +
                        theme(plot.title = element_blank(),
                              axis.title.y = element_blank()),
          ncol = 3, nrow = 1)

png("Figures/bbal_ker_habitat_fixed_effects.png", width = 9, height = 4, units = "in", res = 600)
all_plots
dev.off()

rm(pred_plot.sst); rm(pred_plot.bathy); rm(pred_plot.chl); rm(all_plots)
gc()

```

#### Conditional repeatability


```{r bbal_ker_cond_repeatability}

load("Data_inputs/bbal_ker_habitat_data.RData")

# [ 1A ] Calculate conditional repeatability - Chl & SST ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_IDseason <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
env_vars = c("sst", "chl", "bathy")

sst_scaled <- unlist(pnts_all$sst_scaled)
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)

var_map <- list(
  sst = sst_scaled,
  chl = chl_scaled )

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:2) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics ====
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Extract uncertainty in repeatability estimates ====
  
  ## ~ Between ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  
  # Bootstrap conditional repeatability uncertainty ====
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "between")
  
  
  ## ~ Within ~ ----
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])
  var_tripID <- as.numeric(vc$tripID[1,1]) 

  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_IDseason) / df_IDseason
    var_slope_sim <- var_slope * rchisq(1, df = df_IDseason) / df_IDseason
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID

    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  
  # Estimate conditional repeatability ====
  
  ## ~ Within ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$tripID), 0,
    0, as.numeric(VarCorr(mod_within)$cond$tripID.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var_slope = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Kerguelen", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
   
  ## ~ Between ~ ----
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var_slope = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Kerguelen", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var_slope)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.between %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty.within %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_ker_glmm.within_", env_vars[i]))
  rm(list = paste0("bbal_ker_glmm.between_", env_vars[i]))
  gc()
  
}

condR_results <- do.call("rbind", cond_res.list)

# [ 1B ] Calculate conditional repeatability - Bathymetry ------

# Between ====
  
## Calculate repeatability -----
load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
mod_between = get("bbal_ker_glmm.between_bathy")

# Extract variance-covariance matrices of random effects
vc <- VarCorr(mod_between)$cond
V_ID_intercept <- as.numeric(vc$ID[1,1])
V_tripID <- as.numeric(vc$tripID[1,1])
V_slopes_matrix <- vc$ID.1 

# Extract variances (diagonal) for each bathymetry level slope
V_slope <- diag(V_slopes_matrix)
names(V_slope) <- colnames(V_slopes_matrix)
  
# Residual variance for binomial logistic link
V_residual <- (pi^2) / 3

# Total random variance per level
V_random <- V_ID_intercept + V_slope
V_total <- V_ID_intercept + V_slope + V_tripID + V_residual

# Conditional repeatability per level
R2 <- V_random / V_total

R_cond.df_between <- data.frame(R2) 
R_cond.df_between <- cbind(rownames(R_cond.df_between), data.frame(R_cond.df_between, row.names = NULL))
R_cond.df_between %<>%
  mutate(species = "BBAL",
         colony = "Kerguelen",
         level = "Between", 
         var_slope = V_slope,
         R2 = R2,
         marginal_R2 = NA,
         med_rpt = NA) %>%
  rename(env = `rownames(R_cond.df_between)`) %>%
  relocate(c(species, colony, level, env, var_slope, R2, marginal_R2, med_rpt))

R_cond.df_between$env <- factor(R_cond.df_between$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

# Boostrap uncertainty ----

# Storage matrix for simulated repeatabilities (rows = simulations, cols = slope levels)
rpt_sim <- matrix(NA, nrow = n_sim, ncol = length(V_slope))
colnames(rpt_sim) <- names(V_slope)

for(i in seq_len(n_sim)){
  # Simulate variance components from scaled chi-square
  var_int_sim <- V_ID_intercept * rchisq(1, df_ID) / df_ID
  var_tripID_sim <- V_tripID * rchisq(1, df_tripID) / df_tripID
  
  var_slope_sim <- sapply(V_slope, function(v) v * rchisq(1, df_ID) / df_ID)

  # Calculate conditional repeatability for each slope level
  for(j in seq_along(var_slope_sim)){
    V_random_sim <- var_int_sim + var_slope_sim[j] 
    V_total_sim <- V_random_sim + var_tripID_sim + V_residual
    rpt_sim[i, j] <- V_random_sim / V_total_sim
  }
}

# Summarize simulation results
rpt_summary <- apply(rpt_sim, 2, function(x) {
  c(
    median = median(x, na.rm=TRUE),
    lower = quantile(x, 0.025, na.rm=TRUE),
    upper = quantile(x, 0.975, na.rm=TRUE)
  )
}) %>% t() %>% as.data.frame()

rpt_summary$env <- rownames(rpt_summary)
rpt_summary %<>%
  rename(lower = `lower.2.5%`, upper = `upper.97.5%`) %>%
  select(env, median, lower, upper) %>%
  mutate(env = as.factor(env))

## Merge them with the main dataframe for plotting 
R_cond.df_between %<>% left_join(rpt_summary, by = "env")



# Within ####

# Calculate repeatability ----

load("Data_outputs/bbal_ker_glmm_within_bathy.RData")
mod_within = get("bbal_ker_glmm.within_bathy")

# Extract variance-covariance matrices of random effects
vc <- VarCorr(mod_within)$cond
V_IDseason_intercept <- as.numeric(vc$ID_season[1,1])
V_slopes_matrix <- vc$ID_season.1
V_tripID <- as.numeric(vc$tripID[1,1])
 
# Extract variances (diagonal) for each bathymetry level slope
V_slope <- diag(V_slopes_matrix)
names(V_slope) <- colnames(V_slopes_matrix)
  
# Residual variance for binomial logistic link
V_residual <- (pi^2) / 3

# Total random variance per level
V_random <- V_ID_intercept + V_slope 
V_total <- V_ID_intercept + V_slope + V_tripID + V_residual

# Conditional repeatability per level
R2 <- V_random / V_total

R_cond.df_within <- data.frame(R2) 
R_cond.df_within <- cbind(rownames(R_cond.df_within), data.frame(R_cond.df_within, row.names = NULL))
R_cond.df_within %<>%
  mutate(species = "BBAL",
         colony = "Kerguelen",
         level = "within", 
         var_slope = V_slope,
         R2 = R2,
         marginal_R2 = NA,
         med_rpt = NA) %>%
  rename(env = `rownames(R_cond.df_within)`) %>%
  relocate(c(species, colony, level, env, var_slope, R2, marginal_R2, med_rpt))

R_cond.df_within$env <- factor(R_cond.df_within$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))


# Boostrap uncertainty ----

# Storage matrix for simulated repeatabilities (rows = simulations, cols = slope levels)
rpt_sim <- matrix(NA, nrow = n_sim, ncol = length(V_slope))
colnames(rpt_sim) <- names(V_slope)

for(i in seq_len(n_sim)){
  # Simulate variance components from scaled chi-square
  var_int_sim <- V_ID_intercept * rchisq(1, df_IDseason) / df_ID
  var_slope_sim <- sapply(V_slope, function(v) v * rchisq(1, df_IDseason) / df_IDseason)
  var_tripID_sim <- V_tripID * rchisq(1, df_tripID) / df_tripID

  # Calculate conditional repeatability for each slope level
  for(j in seq_along(var_slope_sim)){
    V_random_sim <- var_int_sim + var_slope_sim[j] 
    V_total_sim <- V_random_sim + var_tripID_sim + V_residual
    rpt_sim[i, j] <- V_random_sim / V_total_sim
  }
}

# Summarize simulation results
rpt_summary <- apply(rpt_sim, 2, function(x) {
  c(
    median = median(x, na.rm=TRUE),
    lower = quantile(x, 0.025, na.rm=TRUE),
    upper = quantile(x, 0.975, na.rm=TRUE)
  )
}) %>% t() %>% as.data.frame()

rpt_summary$env <- rownames(rpt_summary)
rpt_summary %<>%
  rename(lower = `lower.2.5%`, upper = `upper.97.5%`) %>%
  select(env, median, lower, upper) %>%
  mutate(env = as.factor(env))

## Merge them with the main dataframe for plotting 
R_cond.df_within %<>% left_join(rpt_summary, by = "env")



# Bind it all together ====

bbal_ker_condR_results <- rbind(condR_results, R_cond.df_between[,1:8], R_cond.df_within[,1:8])

save(bbal_ker_condR_results, file = "Data_outputs/bbal_ker_condR_results.RData")




# Make the plot ====
bathy_labels <- c("On shelf", "Near shelf break", "Off shelf")

condR_plot <- ggplot() +
    geom_point(data = R_cond.df_between, aes(x= env, y = median), col = col_between, size = 3) +
    geom_errorbar(data = R_cond.df_between, aes(x = env, ymin = lower, ymax = upper), width = 0.1, colour = col_between) +
    geom_point(data = R_cond.df_within, aes(x= env, y = median), col = col_within, size = 3) +
    geom_errorbar(data = R_cond.df_within, aes(x = env, ymin = lower, ymax = upper), width = 0.1, colour = col_within) +
    labs(y = "Conditional repeatability") +
    scale_x_discrete(labels = bathy_labels) +
    expand_limits(y = c(0, 1)) +
    theme_bw() 

assign("bathy_condR_plot", condR_plot)




# [ 2 ] Plot individual slopes ---- 

rnd_effs.list <- list()

for (i in 1:2) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring")
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)


## Bathymetry only 

levels_bathy <- levels(pnts_all$dist_shelf_bin)

fixef_vals.between <- fixef(bbal_ker_glmm.between_bathy)$cond
fixef_vals.within <- fixef(bbal_ker_glmm.within_bathy)$cond

# Prepare a data frame to hold predictions
ranef_slopes.between <- ranef(bbal_ker_glmm.between_bathy)$cond$ID
ranef_slopes.within  <- ranef(bbal_ker_glmm.within_bathy)$cond$tripID

pred_df.between <- expand.grid(ID = rownames(ranef_slopes.between),
                       dist_shelf_bin = levels_bathy)
pred_df.within <- expand.grid(ID = rownames(ranef_slopes.within),
                       dist_shelf_bin = levels_bathy)

# Calculate linear predictor per bird & category: - STOPPED HERE
pred_df.between %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.between[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.between[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.between[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.between[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.between["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.between["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.between["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()


pred_df.within %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.within[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.within[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.within[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.within[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.within["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.within["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.within["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()

## Bind together
pred_df.bathy <- rbind(pred_df.between %>% mutate(season_type = "Between seasons"),
                       pred_df.within %>% mutate(season_type = "Within season"))

# Plot lines connecting points per bird
pred_plot <- ggplot() +
  geom_line(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  geom_point(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  facet_grid(~season_type) +
  scale_x_discrete(labels = c("On shelf", "Near shelf break", "Off shelf")) +
  labs(x = "Bathymetry category",
       y = "Probability of use") +
  theme_bw()

assign("bathy_pred_plot", pred_plot)

save(rnd_effs.out, file = "Data_outputs/bbal_birdis_habitat_selection.RData")

# Make the plot ====

bbal_ker_pref_plot <- ggarrange(bathy_pred_plot + ylim(0, 1) + theme(axis.title.x = element_blank()), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
          bathy_condR_plot + labs(x = "Bathymetry category"),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0,1) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0,1) + theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)

bbal_ker_pref_plot <- annotate_figure(bbal_ker_pref_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_pref_plot, file = "Data_outputs/bbal_ker_selection_plot.RData")


png("Figures/bbal_ker_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
bbal_ker_pref_plot
dev.off()


```

# Final results

For the manuscript, we want to show:

-   Differences in average habitat use by species
-   Individual curves for each habitat type by species
-   Repeatability summaries for each species

### Combine AUC values

```{r auc_values}

load("Data_outputs/bbal_birdis_auc_habitat_selection.RData")
load("Data_outputs/bbal_ker_auc_habitat_selection.RData")
load("Data_outputs/waal_birdis_auc_habitat_selection.RData")
load("Data_outputs/waal_cro_auc_habitat_selection.RData")

auc_results.all <- rbind(bbal_birdis_auc_results, bbal_ker_auc_results,
                         waal_birdis_auc_results, waal_cro_auc_results)

```

#### Figure 3 - Plot used repeatability

```{r fig3_plot_rpt_used } 

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

used_variances <- rbind(bbal_birdis_var_used %>% mutate(spec_col = "BBAL Bird Island"),
                        bbal_ker_var_used %>% mutate(spec_col = "BBAL Kerguelen"),
                        waal_birdis_var_used %>% mutate(spec_col = "WAAL Bird Island"),
                        waal_cro_var_used %>% mutate(spec_col = "WAAL Crozet"))

# Wrangle the data into long form
used_variances %<>%
  select(-c(Effect, Variance, `Std. dev`)) %>%
  distinct() %>%
  group_by(spec_col) %>%
  pivot_longer(cols = matches("_(within|across)$"),
                   names_to = c(".value", "group"),
                   names_pattern = "(.*)_(within|across)")
# Change factors
used_variances %<>%
 mutate(Environment = recode(Environment,
      "bathy" = "Bathymetry",
      "chlA" = "Chlorophyll",
      "SST" = "SST")) %>%
  arrange(Environment) %>%
  mutate(Environment = factor(Environment, levels = rev(unique(Environment))))

rpt_used.plot <- ggplot(used_variances, aes(x = rpt, y = Environment, colour = group)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = LCL, xmax = UCL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Repeatability (95% CI)") +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(
    values = c(col_between, col_within)) +
  theme_bw() +
  theme(legend.position = "none")



png(file = "Figures/MS Figures/FIG3_habitat_used_repeatability.png", width = 10, height = 10, units = "in", res = 300)
rpt_used.plot
dev.off()


```

#### Figure S5 - Plot random intercepts

```{r figs5_plot_used}

load("Data_outputs/bbal_birdis_used_plot.RData")
load("Data_outputs/bbal_ker_used_plot.RData")
load("Data_outputs/waal_birdis_used_plot.RData")
load("Data_outputs/waal_cro_used_plot.RData")

png(file = "Figures/MS Figures/FIGS5_habitat_used_intercepts.png", width = 25, height = 17, units = "in", res = 300)
ggarrange(bbal_birdis_used_plot, bbal_ker_used_plot,
          waal_cro_used_plot, waal_birdis_used_plot,
          ncol = 2, nrow = 2,
          widths = c(1, 0.8))
dev.off()

```

#### Table 4 - Random intercept variances

```{r habitat_used_variance}

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")


var.table <- rbind(bbal_birdis_var_used %>% mutate(Species = "BBAL", Colony = "Bird Island"), 
                   bbal_ker_var_used %>% mutate(Species = "BBAL", Colony = "Kerguelen"), 
                   waal_birdis_var_used %>% mutate(Species = "WAAL", Colony = "Bird Island"), 
                   waal_birdis_var_used %>% mutate(Species = "WAAL", Colony = "Crozet")) %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorophyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(where(is.numeric), 
                ~ ifelse(.x < 1e-4 & .x > 0, 
                         "< 0.0001", 
                         formatC(.x, digits = 2, format = "fg")))) %>%
  filter(Effect == "ID_season" | Effect == "ID") %>%
  rename(Level = Effect) %>%
  relocate(c(Species, Colony), .before = Environment)

kable(var.table, format = "html", 
  caption = 'Table 1A. Variance associated with random intercepts from the habitat used model.') %>%
 # row_spec(c(6,11), extra_css = "border-top: 1px solid;") %>%
  collapse_rows(columns = c(1, 2, 3, 7, 8)) %>%
  kable_classic(font_size = 12)

```

#### Figure 4 - Plot preferred habitat

```{r plot_used}

load("Data_outputs/bbal_birdis_selection_plot.RData")
load("Data_outputs/bbal_ker_selection_plot.RData")
load("Data_outputs/waal_birdis_selection_plot.RData")
load("Data_outputs/waal_cro_selection_plot.RData")

png(file = "Figures/MS Figures/FIG4_habitat_preference.png", width = 25, height = 15, units = "in", res = 300)
ggarrange(bbal_birdis_pref_plot, 
          bbal_ker_pref_plot,
          waal_cro_pref_plot, 
          waal_birdis_pref_plot,
          ncol = 2, nrow = 2,
          widths = c(1, 1))
dev.off()

```

#### Table 5 - Repeatability of preference

```{r table_rpt_preference}

load("Data_outputs/bbal_birdis_condR_results.RData")
load("Data_outputs/bbal_ker_condR_results.RData")
load("Data_outputs/waal_birdis_condR_results.RData")
load("Data_outputs/waal_cro_condR_results.RData")

rpt.table <- rbind(bbal_birdis_condR_results, 
                   bbal_ker_condR_results %>% rename(var_between = var_slope) %>% arrange(env), 
                   waal_birdis_condR_results, 
                   waal_cro_condR_results) %>%
  mutate(med_rpt = med_rpt / 100) %>%
  rename(Environment = env, 'Slope Variance' = var_between, "Median cond R" = med_rpt,
         Species = species, Colony = colony, Level = level) %>%
  relocate(Environment, .before = Level) %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chl" ~ "Chlorophyll-a",
    Environment == "sst" ~ "SST",
    Environment == "dist_shelf_binon_shelf" ~ "On shelf",
    Environment == "dist_shelf_binnear_break" ~ "Near shelf break",
    Environment == "dist_shelf_binoff_shelf" ~ "Off shelf"),
    Level = ifelse(Level == "Between", "Between", "Within")) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  select(-marginal_R2) %>%
  arrange(Species, Colony, Environment, Level)

kable(rpt.table, format = "html") %>% 
  collapse_rows(columns = c(1, 2, 3)) %>%
  kable_classic(font_size = 12)

```

#### Plot heritability

```{r plot_used}

bbal_birdis.heritability <- loadRData("Data_outputs/bbal_birdis_habitat_pref_heritability_posteriors.RData") %>% mutate(spec_col = "BBAL Bird Island")
waal_birdis.heritability <- loadRData("Data_outputs/waal_birdis_habitat_pref_heritability_posteriors.RData") %>% mutate(spec_col = "WAAL Bird Island")
waal_cro.heritability <- loadRData("Data_outputs/waal_cro_habitat_pref_heritability_posteriors.RData") %>% mutate(spec_col = "WAAL Crozet")

posteriors_all <- rbind(bbal_birdis.heritability, waal_birdis.heritability, waal_cro.heritability)

## Add facet code for plotting
posteriors_all %<>%
   mutate(facet_code = case_when(
    spec_col == "BBAL Bird Island" ~ "A",
    spec_col == "WAAL Bird Island" ~ "C",
    spec_col == "WAAL Crozet" ~ "D"))


# Make the plot
design = "
A#
CD
"

facet_labels <- c(
  A = "BBAL Bird Island",
  C = "WAAL Bird Island",
  D = "WAAL Crozet"
)

heritability_plot <- ggplot(posteriors_all, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.6, scale = 1) +
  theme_bw() +
  ggh4x::facet_manual(~facet_code, design = design,
                        labeller = as_labeller(facet_labels)) +
  labs(x = "Posterior distribution for heritability (h²)", y = "") +
  theme_bw()


png(file = "Figures/MS Figures/FigureX_habitat_heritability.png", width = 10, height = 10, units = "in", res = 300)
heritability_plot
dev.off()

```

<!-- #### Habitat preference -->

<!-- After discussion with Sam & Al it was suggested that I could also look at mean habitat use across individuals; essentially I want to look at whether birds are specialists or generalists. Then, the habitat modelling could tell me whether individuals vary in plasticity i.e. are they flexible or specific.  -->

<!-- If I just want to see whether individuals have different preferences, I can do a simple LMM, e.g.:  -->

<!-- $$y = \beta_0 + \beta_1 \cdot \text{phase} + \beta_2 \cdot \text{age} + U_{\text{ring}} + U_{\text{season}} + U_{\text{ring:season}} + \varepsilon$$ -->

<!-- where: -->

<!-- -   $y$ represents each of the 3 environmental variables, -->

<!-- -   $\beta_0$ is the intercept, -->

<!-- -   $\beta_1$ - $\beta_{2}$ are fixed effect coefficients for $phase$ and $age$, -->

<!-- -   $U_{\text{ring}}$, $U_{\text{season}}$, and $U_{\text{ring:season}}$ are random intercepts for ring, season, and ring within season respectively, -->

<!--     $$U_{\text{ring}} \sim \mathcal{N}(0, \sigma^2_{\text{ring}})$$ -->

<!--     $$U_{\text{season}} \sim \mathcal{N}(0, \sigma^2_{\text{season}})$$ -->

<!--     $$U_{\text{ring_season}} \sim \mathcal{N}(0, \sigma^2_{\text{ring_season}})$$ -->

<!-- -   $ε$ is the residual error. -->

<!--     $$ε \sim \mathcal{N}(0, \sigma^2_{\text{res}})$$ -->

<!-- But if I want to know whether some individuals are more selective than others, I would need to use DHGLMMs. This would allow me to model the variance in selection as a function of individual identity. However, this requires some thought - I tried running a simple DHGLMM, and after 3 days it had only completed 30% of chains. Come back to this. -->

```{r bba_birdis_habitat_preferences, eval = FALSE}

# Integrate meta
load("Data_inputs/ind_meta.RData")
all_ind.meta %<>% filter(species == "BBA") %>% mutate(birthYr = as.numeric(birthYr)) %>% rename(ID = ring)

pnts_all <- merge(pnts_all, all_ind.meta, by = "ID", all.x = T) %>% 
  mutate(ID_season = paste(ID, season, sep = "_"), 
         age = season - birthYr)

pnts_all %<>% filter(!is.na(age))

# Build models

## SST ##
pnts_subset <- pnts_all[sample(nrow(pnts_all), 20000),]

library(brms)

sst_dhglmm <- brm(
  bf(SST ~ phase + age + (1 | ID) + (1 | season)) +
  lf(sigma ~ 0 + ID),
  iter = 2000, warmup = 1000, thin = 10,
  data = pnts_subset)

summary(sst_dhglmm)
save(sst_dhglmm, file = "Data_outputs/test_dhglmm.RData")


sst_pref <- lmer(SST ~ (1 | ID) + (1 | season) + (1 | ID_season) + phase + age,
                 data = pnts_all)

summary(sst_pref)

## Diagnostics
  # plot(m1, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
  # qqnorm(resid(m1))
  # qqline(resid(m1))
  # hist(resid(m1), breaks = 50, main = "Histogram of residuals")

```

<!-- ## Wandering albatrosses -->

<!-- ### Bird Island -->

<!-- Again, for a subsample of 20 birds (for readability), I'm plotting the environmental variability used by each individual. -->

```{r waal_birdis_plot_env, eval = FALSE, fig.cap="**Figure 2.1.** *Individual density histograms for use of SST, bathymetry, chlorophyll, and sea surface height. Random sample of 20 individuals.*", fig.dim = c(8, 10), fig.topcaption = TRUE}

load("Data_inputs/waal_birdis_sst-bathy-chl-ssh_data.RData")

subset_ids <- sample(unique(pnts_all$ring), 30)

#### ~# SST # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_sst = mean(SST, na.rm = T)) %>%
  ungroup() %>%
  mutate(ring = factor(ring, levels = unique(ring[order(mean_sst)]))) 

plot.sst <- ggplot(pnts_subset, 
       aes(x = SST, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#0D0887FF", "#CC4678FF", "#F0F921FF"),
    name = "SST (°C)" ) +
  labs(x = "Sea surface temperature (°C)", y= "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())

#### ~# Bathymetry # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
   filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_bathy = mean(bathy, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_bathy)]))) 

plot.bathy <- ggplot(pnts_subset, 
       aes(x = bathy, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_viridis_c( name = "Sea floor depth (m below sea level)" ) +
  labs(x = "Sea floor depth (m below sea level)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


#### ~# Chlorophyll # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
   filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_chl = mean(chl, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_chl)]))) 

plot.chl <- ggplot(pnts_subset, 
       aes(x = chl, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1A6D2A", "#6CDC82", "#D3F5D9"),
    name = expression("Chlorophyll (" * mg ~ Chl ~ m^-3 * ")") ) +
  labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"), y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


#### ~# Sea surface height # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_ssh = mean(SSH, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_ssh)]))) 

plot.ssh <- ggplot(pnts_subset, 
       aes(x = SSH, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1a2290", "#B2B6F1", "#D8DCE0"),
    name = "SSH (m)") +
  labs(x = "Sea surface height (m)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


ggarrange(plot.sst, plot.bathy + theme(axis.title.y = element_blank()),
          plot.chl, plot.ssh + theme(axis.title.y = element_blank()), 
          ncol = 2, nrow = 2)

```

<!-- #### Modelling -->

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_birdis_sst-bathy-chl-ssh_data.RData")

## Scale the variables
pnts_all %<>%
  mutate(sst_scaled = scale(SST),
         bathy_scaled = scale(bathy),
         ssh_scaled = scale(SSH),
         chl_scaled = scale(chl),
         dist_col_scaled = scale(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 )) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID)) 

## Remove birds with only a couple of points
pnts_all %<>% 
  group_by(ID, year) %>%
  mutate(n_recs = n(),
         flag = ifelse(n_recs < 20, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-flag)

## Remove NAs and set distance to km
pnts_all %<>% filter(!is.na(SST) & !is.na(used) & !is.na(bathy) & !is.na(SSH) &!is.na(chl)) %>%
  mutate(dist_col.km = dist_col/1000)


```

```{r waal_birdis_models, eval = FALSE}

# Within-season ----

mod_within.waal_bi.sst <- glmmTMB(
  used ~ poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial()
)

save(mod_within.waal_bi.sst, file = "Data_outputs/waal_birdis_habitat_model_within_sst.RData")

mod_within.waal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.waal_bi.bathy, file = "Data_outputs/waal_birdis_habitat_model_within_bathy.RData")

mod_within.waal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.waal_bi.chl, file = "Data_outputs/waal_birdis_habitat_model_within_chl.RData")


# Between-season ----

mod_between.waal_bi.sst <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.waal_bi.sst, file = "Data_outputs/waal_birdis_habitat_model_between_sst.RData")

mod_between.waal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 +bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.waal_bi.bathy, file = "Data_outputs/waal_birdis_habitat_model_between_bathy.RData")

mod_between.waal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.waal_bi.chl, file = "Data_outputs/waal_birdis_habitat_model_between_chl.RData")

```

```{r waal_birdis_diagnostics, eval = FALSE}

load("Data_outputs/waal_birdis_habitat_model_between.RData")

### Diagnostics ----

# AUC around 0.5 = essential random; closer to 1 = better fit
preds <- predict(mod_between.waal_bi, type = "response")  
auc.mod <- as.numeric(auc(pnts_all$used, preds))

# Check for correlations
vif.mod <- car::vif(mod_between.waal_bi)

# Check for patterns and that most points are within 95% confidence intervals
plot_binnedplot(mod_between.waal_bi)

# Check ratio is close to 1
overdisp.mod <- overdisp_fun(mod_between.waal_bi)

# Check ratio: >1 may indicate zero inflation (more 0s than expected); < 1 zero deflation
# Small P = evidence to suggest zero-inflation/deflation
zeroinf.mod <- testZeroInflation(mod_between.waal_bi, plot = FALSE)

```

<!-- #### Diagnostics -->

<!-- -   AUC = `r signif(as.numeric(auc.mod), 2)`, which suggests the model separates 'used' vs. 'not used' locations well. -->

<!-- -   The VIF values are all \< 2, so there is no concerning multicollinearity. -->

<!-- -   The overdispersion function returns `r as.numeric(overdisp.mod[2])`, which is slightly high. This could indicate moderate overdispersion (i.e. variance is higher than expected in a binomial model), but probably not a major concern. -->

<!-- -   There is some evidence for zero-inflation - p = `r zeroinf.mod$p.value`. However, this might be to expected, as zero counts are an inherent part of the process I am trying to model, and so a zero-inflated approach is likely unncessary. -->

<!-- #### Predictions -->

```{r waal_birdis_model_outputs, eval = FALSE, fig.cap = "**Figure 2.3.** *Predicted probabilities of use for each fixed effect.*", fig.topcaption = TRUE}

# Plot effects
pred_plot.sst <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "sst_scaled [all]"))
pred_plot.bathy <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "bathy_scaled [all]"))
pred_plot.chl <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "chl_scaled [all]"))
pred_plot.dist <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "dist_col_scaled [all]"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(plot.title = element_blank()), 
          pred_plot.dist + labs(x = "Scaled distance from colony (km)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          ncol = 2, nrow = 2)

all_plots

```

<!-- The results show that in general, birds select for median SST, deeper waters, higher chlorophyll concentrations, and locations closer to the colony. There is a high degree of individual variation, especially in the use of SST and bathymetry. -->

<!-- #### Individual effects -->

```{r waal_birdis_individual, eval = FALSE, fig.cap = "**Figure 2.4.** *Individual level responses to changes in the environmental variables. Individual slopes are plotted for the extent of each individual's range. Black line and grey interval shows population-level response and associated confidence interval.*", fig.topcaption = TRUE}

# Build plots

env_vars = c("sst_scaled", "bathy_scaled", "chl_scaled")

for (i in 1:length(env_vars)) {

  habitat_var = env_vars[i]
  rnd_eff <- get_rnd_effects(mod_between.bbal_bi, habitat_var)
  
  # Get range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_sst = min(!!sym(habitat_var), na.rm = TRUE),
              max_sst = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(habitat_val)) 
  
  predictions.ind <- merge(predictions.ind, rnd_eff, by = "ID", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Get global predictions
  predictions.global <- data.frame(ggpredict(mod_between.bbal_bi, terms = paste0(habitat_var, " [all]")))
  
  ## Plot
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, colour = abs(slope), group = ID)) +
      scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      geom_line(data = predictions.global, aes(x = x, y = predicted), col = "black",
                 linetype = "dashed", linewidth = 1) +
      geom_ribbon(data = predictions.global, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "grey90", alpha = 0.5, col = "grey10") +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

ggarrange(sst_scaled_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) , 
          bathy_scaled_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.y = element_blank()), 
          chl_scaled_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")),
          ncol = 2, nrow = 2)


```

```{r waal_birdis_calc_repeatability, eval = FALSE}

# Within season -----
summary(mod_within.waal_bi)

## manual test
vc <- VarCorr(mod_within.waal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.within <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.within <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.within <- V_ID_chl / (V_ID_chl + V_residual)



# Between season -----
summary(mod_between.waal_bi)

## manual test
vc <- VarCorr(mod_between.waal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.between <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.between <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.between <- V_ID_chl / (V_ID_chl + V_residual)


```

# APPENDICES

At this point I have (1) randomly sampled available habitat on a ratio
of 6 available to 1 used points; and (2) matched up all of these points
with environmental data - specifically, SST, bathymetry, SSH, and
chlorophyll.

To ensure appropriate model fit, I'll be working on each species/colony
separately. I can think about integrating them later, if the residual
distributions etc are similar.

## EXPLORATION: Black-browed albatrosses

### Bird Island

First, I'm exploring the environmental variation experienced by
individuals. Here, for a subsample of 20 birds (for readability), I'm
plotting the environmental variability used by each individual.

```{r bbal_birdis_plot_env, fig.cap="**Figure 1.1** *Individual density histograms for use of SST, bathymetry, chlorophyll, and sea surface height. Random sample of 20 individuals.*", fig.dim = c(8, 10), fig.topcaption = TRUE}

load("Data_inputs/bba_birdis_sst-bathy-chl-ssh_data.RData")

subset_ids <- sample(unique(pnts_all$ring), 30)

#### ~# SST # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_sst = mean(SST, na.rm = T)) %>%
  ungroup() %>%
  mutate(ring = factor(ring, levels = unique(ring[order(mean_sst)]))) 

plot.sst <- ggplot(pnts_subset, 
       aes(x = SST, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#0D0887FF", "#CC4678FF", "#F0F921FF"),
    name = "SST (°C)" ) +
  labs(x = "Sea surface temperature (°C)", y= "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())

png(filename = "Figures/sst_use.png", width = 4.5, height = 6, units = "in", res = 600)
plot.sst
dev.off()

#### ~# Bathymetry # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_bathy = mean(bathy, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_bathy)]))) 

plot.bathy <- ggplot(pnts_subset, 
       aes(x = bathy, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_viridis_c( name = "Sea floor depth (m below sea level)" ) +
  labs(x = "Sea floor depth (m below sea level)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())

png(filename = "Figures/bathy_use.png", width = 4.5, height = 6, units = "in", res = 600)
plot.bathy
dev.off()


#### ~# Chlorophyll # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_chl = mean(chl, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_chl)]))) 

plot.chl <- ggplot(pnts_subset, 
       aes(x = chl, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1A6D2A", "#6CDC82", "#D3F5D9"),
    name = expression("Chlorophyll (" * mg ~ Chl ~ m^-3 * ")") ) +
  labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"), y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


png(filename = "Figures/chl_use.png", width = 4.5, height = 6, units = "in", res = 600)
plot.chl
dev.off()

#### ~# Sea surface height # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_ssh = mean(SSH, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_ssh)]))) 

plot.ssh <- ggplot(pnts_subset, 
       aes(x = SSH, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1a2290", "#B2B6F1", "#D8DCE0"),
    name = "SSH (m)") +
  labs(x = "Sea surface height (m)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


ggarrange(plot.sst, plot.bathy + theme(axis.title.y = element_blank()),
          plot.chl, plot.ssh + theme(axis.title.y = element_blank()), 
          ncol = 2, nrow = 2)

```

Next, I looked at potential covariation between the different variables.

```{r bbal_birdis_correlations, fig.cap = "**Figure 1.2.** *Correlations between chosen environmental variables.*", fig.dim = c(5, 5), fig.topcaption = TRUE}

corr_matrix <- cor(na.omit(pnts_all[, c(8, 12, 13, 15, 16)]))
corrplot.mixed(corr_matrix, order = "AOE")

```

Surprisingly, SST and SSH are extremely correlated - it turns out that
they do indeed have a strong spatial correlation, particularly due to
e.g. El Nino events, e.g. see:

Leuliette, E. W., & Wahr, J. M. (1999). Coupled pattern analysis of sea
surface temperature and TOPEX/Poseidon sea surface height. *Journal of
Physical Oceanography, 29*(4), 599–611.
[https://doi.org/10.1175/1520-0485(1999)029\<0599:CPAOSS\>2.0.CO;2](https://doi.org/10.1175/1520-0485(1999)029%3C0599:CPAOSS%3E2.0.CO;2)

Apart from this, the correlations don't seem very worrying. In response
to these correlations, I dropped SSH.

```{r bbal_birdis_process_data, eval = TRUE}

# Do some data cleaning
load("Data_inputs/bba_birdis_sst-bathy-chl-ssh_data.RData")

scale_2SD <- function(variable) { (variable - mean(variable, na.rm = TRUE)) / sd(variable, na.rm = TRUE) * 2 }

## Scale the variables
pnts_all %<>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         #ssh_scaled = scale_2SD(chl),
         chl_scaled = scale_2SD(chl),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
                     bathy_scaled2 = bathy_scaled^2) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID)) 

## Remove birds with only a couple of points
pnts_all %<>% 
  group_by(ID, year) %>%
  mutate(n_recs = n(),
         flag = ifelse(n_recs < 20, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-flag)

## Remove NAs and set distance to km
pnts_all %<>% filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chl)) %>%
  mutate(dist_col.km = dist_col/1000)

## Calculate within-individual centered SST
# pnts_all %<>%
#   group_by(ID) %>%
#   mutate(sst_within = SST - mean(SST, na.rm = TRUE),
#          sst_within2 = sst_within^2) 


```

#### Average preference

Here I'm looking at differences in average habitat choice amongst
individuals. Here, I fit simple linear models where the average
environmental variable experienced is predicted by distance from the
colony, breeding phase, and - critically - the random effect of
individual.

From this I can extract random intercepts as the variation in
environment use attributable to individual - e.g. average value of SST
for each individual - and then the repeatability of individual choice
for the environmental variable.

The model I fitted was:

$$y = \beta_0 + \beta_1 \cdot \text{dist_col} + \beta_2 \cdot \text{phase} + U_{\text{ID}} + U_{\text{tripID}}$$

where:

-   $\beta_0$ is the intercept,
-   $\beta_1$ is the fixed effect of the distance from the colony (to
    control for environmental variation associated with this),
-   $\beta_2$ is the fixed effects of breeding phase,
-   $U_{\text{ID}}$ is the individual-level random intercept,
-   $U_{\text{ID}}$ is the random intercept for trip.

```{r average_habitat}

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chl")
rpt_df <- data.frame(env_var = character(), rpt_R = numeric())

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1 | ID) + (1 | tripID)")), data = pnts.used)
  
  # Extract random effects for ID
  random_effects <- ranef(mod)$ID
  ind_intercepts <- random_effects[, 1]
  
  # Extract fixed effects for the intercept
  ind_avg <- ind_intercepts + fixef(mod)[1]
  assign(paste0("ind_avg.", env_vars[i]), ind_avg)
  
  # Calculate variance components
  var_components <- as.data.frame(VarCorr(mod))
  var_id <- var_components[var_components$grp == "ID", "vcov"]
  var_residual <- attr(VarCorr(mod), "sc")^2
  
  # Total variance = variance from ID + residual variance
  total_var <- var_id + var_residual
  
  # Calculate repeatability
  repeatability <- var_id / total_var
  
  # Output
  rpt_df <- rbind(rpt_df, data.frame(variable = env_vars[i], repeatability = repeatability))
 
}

## Plot outputs
ind_avg.plot <- data.frame(SST = ind_avg.sst,
                           bathy = ind_avg.bathy,
                           chl = ind_avg.chl)

ind_avg_long <- ind_avg.plot %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  arrange(variable) %>%
  mutate(dummy_x = 1)

# Add y position for plot: SST - bathy - chl
rpt_df$y_position = c(8, 1200, 4)

ind_avg_long_with_rpt <- ind_avg_long %>%
  left_join(rpt_df, by = "variable")

# Plot
label_map <- c(
  bathy = "Bathymetry (m below sea level)",
  chl = "Chlorophyll (mg Chl m^-3)",
  SST = "SST (°C)")

png(filename = "Figures/avg_environment.png", width = 10, height = 5, units = "in", res = 600)
ggplot(ind_avg_long_with_rpt, aes(x = dummy_x, y = value, fill = variable)) +
  geom_violin(trim = FALSE, colour = "black") +
  geom_jitter(width = 0.1, alpha = 0.5, size = 1) +
   # Add the repeatability text
  geom_text(aes(x = 1, y = y_position, label = paste("Repeatability: ", round(repeatability, 3))),
            fontface = "plain", colour = "black", size = 3) +
  facet_wrap(~ variable, scales = "free_y", 
             labeller = labeller(variable = label_map)) + 
  scale_fill_manual(values = c("#37D5AC", "#339F48", "#A0307E")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank())
dev.off()



```

#### Habitat selection models

To understand individual variation in habitat preference, I am fitting
habitat selection models. These are binomial models that look at
differences in environmental conditions between 'used' and 'available'
habitats. I have fitted a random slope to allow indiviudal selection to
vary.

After having issues fitting each environmental variable in turn
(specifically, the AUC was very low, suggesting poor predictive
performance), I realised that a maximal model including all the
predictors might be better. Indeed, I get better AUC and fit (see
below).

This current model is only looking at between season individual
variation, I still need to look at within season.

The generalized linear mixed model (GLMM) used in this analysis is:

$$\text{logit}(\Pr(\text{used} = 1)) = \beta_0 + \beta_1 \cdot \text{sst} + \beta_2 \cdot \text{sst}^2 + \beta_3 \cdot \text{bathy} + \beta_4 \cdot \text{bathy}^2 + \beta_5 \cdot \text{chl} + \beta_6 \cdot \text{dist col} + \beta_7 \cdot \text{phase} + U_{\text{ID}} + U_{\text{ID}}^{\text{sst}} \cdot \text{sst} + U_{\text{ID}}^{\text{bathy}} \cdot \text{bathy} + U_{\text{ID}}^{\text{chl}} \cdot \text{chl}$$

where:

-   $\beta_0$ is the intercept,
-   $\beta_1$–$\beta_{7}$ are the fixed effects of each environmental
    variable ($sst$, $bathy$, $chl$), breeding phase ($phase$), and
    distance from the colony in km ($dist col$)
-   $U_{\text{ID}}$ is the individual-level random intercept,
-   $U_{\text{ID}}^{\text{sst}}$, $U_{\text{ID}}^{\text{bathy}}$, and
-   $U_{\text{ID}}^{\text{chl}}$ are **individual-level random slopes**
    for **SST, bathymetry, and chlorophyll**, allowing individuals to
    have **different responses** to these variables.

Both SST and bathymetry were fitted with quadratic effects based on
exploratory modelling. For SST, it looks like birds select medium-low
temperatures; for bathymetry it looks like birds prefer shallow and deep
areas (possibly confounded by colony distance, but we'll get on to
that). Models fitted with quadratic SST and bathymetry are a
significiantly better fit based on AIC and an ANOVA comparison. The
binned residual plot also looks considerably better, and the AUC
improves (marginally, as it was reasonably good to begin with). Most
importantly, I have no convergence issues with this model structure.

I have not included quadratic random slopes as this gets very
complicated. So for now, we are assuming that all birds have the same
quadratic preference (same slope shape) but this can 'shift'.

```{r bbal_birdis_models, eval = FALSE}

pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

## currently working here 13/05/2025
mod_linear <- glmmTMB(
  used ~ sst_scaled + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

mod_quadratic <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

AIC(mod_linear, mod_quadratic)
anova(mod_linear, mod_quadratic)

tok <- Sys.time()
mod_brms.quad <- brm(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase + 
        (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_all,
  family = bernoulli(),
  chains = 4,       
  cores = 4,        
  iter = 2000,      
  warmup = 1000     
)
tik <- Sys.time()

save(mod_brms.quad, file = "Data_outputs/mod_brms_quad.RData")




mod_between.bbal_bi.sst <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

summary(mod_between.bbal_bi.sst)
save(mod_between.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_between_sst.RData")

# Within-season ----

mod_within.bbal_bi.sst <- glmmTMB(
  used ~ poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial()
)

save(mod_within.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_within_sst.RData")

mod_within.bbal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.bbal_bi.bathy, file = "Data_outputs/bba_birdis_habitat_model_within_bathy.RData")

mod_within.bbal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.bbal_bi.chl, file = "Data_outputs/bba_birdis_habitat_model_within_chl.RData")


# Between-season ----

mod_between.bbal_bi.sst <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

summary(mod_between.bbal_bi.sst)
save(mod_between.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_between_sst.RData")


mod_between.bbal_bi.sst <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_between_sst.RData")

mod_between.bbal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 +bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.bbal_bi.bathy, file = "Data_outputs/bba_birdis_habitat_model_between_bathy.RData")

mod_between.bbal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.bbal_bi.chl, file = "Data_outputs/bba_birdis_habitat_model_between_chl.RData")

```

```{r plot_individual_slopes}

env_variables <- c("sst", "bathy", "chl")
env_colours <- c(
  "sst" = "#A0307E",
  "bathy" = "#37D5AC",
  "chl" = "#339F48")

env_labels <- c(
  "sst" = "Scaled SST",
  "bathy" = "Scaled bathymetry",
  "chl" = "Scaled chlorophyll conc.")

rpt_list <- list() 
within = TRUE

for (i in seq_along(env_variables)) {
  
  if (within == TRUE) {
    mod_loaded <-loadRData(paste0("Data_outputs/mod_brms_quad_", env_variables[i], "_within.RData")) } else {
      mod_loaded <-loadRData(paste0("Data_outputs/mod_brms_quad_", env_variables[i], ".RData"))
    }
  
  env_scaled <- paste0(env_variables[i], "_scaled")
  
  # Get variances
  posterior <- as_draws_df(mod_loaded)
  var_col <- ifelse(within == TRUE, paste0("sd_ID_season__", env_scaled), paste0("sd_ID__", env_scaled))
  var_among <- posterior[[var_col]]^2 
  var_resid <- pi^2 / 3
  
  # Calculate repeatability per posterior sample
  repeatability <- var_among / (var_among + var_resid)
  repeatability_summary <- data.frame(
    var = env_variables[i],
    mean = mean(repeatability),
    median = median(repeatability),
    CI_lower = quantile(repeatability, 0.025),
    CI_upper = quantile(repeatability, 0.975))
  
  rpt_list[[i]] <- repeatability_summary
  
  # Extract random & fixed effects
  re <- ranef(mod_loaded)[[1]]
  fixed <- fixef(mod_loaded)
  
  # Build data frame of individual slopes
  slopes <- data.frame(
    ID = rownames(re),
    intercept = fixed["Intercept", "Estimate"] + re[, "Estimate", "Intercept"],
    slope = fixed[env_scaled, "Estimate"] + re[, "Estimate", env_scaled])
  
  if(env_variables[i] == "chl") {
    slopes$quad = NA
  } else {
    slopes$quad = fixed[paste0(env_scaled, 2), "Estimate"] + re[, "Estimate", paste0(env_scaled, 2)] 
  }
  
   
  png(file = ifelse(within == TRUE, 
                    paste0("Figures/individual_", env_variables[i], "_within_slopes.png"), 
                    paste0("Figures/individual_", env_variables[i], "_between_slopes.png")),
                    width = 5, height = 4, units = "in", res = 600)
  ggplot(slopes, aes(x = reorder(ID, slope), y = slope)) +
    geom_point() +
    geom_hline(yintercept = fixed[env_scaled, "Estimate"], linetype = "dashed", colour = "red") +
    coord_flip() +
    labs(y = "Individual slope", x = "Individual ID") +
    theme_bw() +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank()) 
  dev.off()
  
  png(file =ifelse(within == TRUE, 
                    paste0("Figures/histogram_", env_variables[i], "_within_slopes.png"), 
                    paste0("Figures/histogram_", env_variables[i], "_between_slopes.png")),
      width = 5, height = 4, units = "in", res = 600)
  ggplot(slopes, aes(x = slope)) +
    geom_histogram() +
    labs(x = "Individual slope", y = "Count") +
    theme_bw() 
  dev.off()
  
  # Plot individual curves
  ## Create a sequence of environmental variables
  env_seq <- seq(-3, 3, length.out = 100)
  
  # Expand to a grid: one row per individual per sst_scaled value
  curve_data <- slopes %>%
    mutate(ID = as.character(ID)) %>%
    slice(rep(1:n(), each = length(env_seq))) %>%
    mutate(env_scaled = rep(env_seq, times = nrow(slopes)),
           env_scaled2 = env_scaled^2)
  
  if(env_variables[i] == "chl") {
    curve_data %<>%
      mutate(logit_p = intercept + slope * env_scaled,
             prob = plogis(logit_p))
  } else {
    curve_data %<>%
      mutate(logit_p = intercept + slope * env_scaled + quad * env_scaled2,
             prob = plogis(logit_p))
  }

  colour_choice <- env_colours[env_variables[i]]
  x_label <- env_labels[env_variables[i]]
  
  png(file = ifelse(within == TRUE, 
                    paste0("Figures/individual_", env_variables[i], "_within_curves.png"), 
                    paste0("Figures/individual_", env_variables[i], "_between_curves.png")), 
      width = 6.5, height = 6, units = "in", res = 600)
  ggplot(curve_data, aes(x = env_scaled, y = prob, group = ID)) +
    geom_line(alpha = 0.3, colour = colour_choice) +
    labs(x = x_label, y = "P|used") +
    theme_bw()
  dev.off()

}


```

```{r code from Bertille}

load("Data_outputs/mod_brms.quad.Rdata")
summary(mod_brms.quad)

set.seed(817)
pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

# Onset of senescence
df_rawAge_randomSlopesAndInterceptsIDlevel$OnsetSenescence = - df_rawAge_randomSlopesAndInterceptsIDlevel$Slope / (2*df_rawAge_randomSlopesAndInterceptsIDlevel$Slope2 )


# extract rate of senescence
# Function that calculates the derivative of 
# exp(quad*xval^2+lin*xval+const)/(1+exp(quad*xval^2+lin*xval+const))
deriv_logit_quad<-function(xval,quad,lin,const)
{
  output<-(2*quad*xval+lin)*exp(quad*xval^2+lin*xval+const)/((1+exp(quad*xval^2+lin*xval+const))^2)
  return(output)
}
 
# Find minimum of the derivative 
xinfl<-c()
gradmins<-c()
for(index in 1:length(df_rawAge_randomSlopesAndInterceptsIDlevel$Slope2))
{
  minval<-optimize(deriv_logit_quad,c(0,100),df_rawAge_randomSlopesAndInterceptsIDlevel$Slope2[index],df_rawAge_randomSlopesAndInterceptsIDlevel$Slope[index],df_rawAge_randomSlopesAndInterceptsIDlevel$Intercept[index]) 
  xinfl<-c(xinfl,minval$minimum)
  gradmins<-c(gradmins,minval$objective)
}
 
# Add to df_rawAge_randomSlopesAndInterceptsIDlevelframe and put out to CSV
df_rawAge_randomSlopesAndInterceptsIDlevel$SenesceRate<-gradmins
df_rawAge_randomSlopesAndInterceptsIDlevel$ageInfl<-xinfl

```

```{r bbal_birdis_diagnostics, eval = FALSE}

# Need to expand to all models #

# load("Data_outputs/bba_birdis_habitat_model_within_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_chl.RData")

load("Data_outputs/bba_birdis_habitat_model_between_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_chl.RData")


### Diagnostics ----

model_info <- list(
  within_sst = list(model = mod_within.bbal_bi.sst, data = pnts_subset),
  within_bathy = list(model = mod_within.bbal_bi.bathy, data = pnts_subset),
  within_chl = list(model = mod_within.bbal_bi.chl, data = pnts_subset)#,
  # between_sst = list(model = mod_between.bbal_bi.sst, data = pnts_all),
  # between_bathy = list(model = mod_between.bbal_bi.bathy, data = pnts_all),
  # between_chl = list(model = mod_between.bbal_bi.chl, data = pnts_all)
)

auc_results <- data.frame(
  model = character(),
  auc = numeric(),
  stringsAsFactors = FALSE
)

# Loop through models and compute AUC
for (model_name in names(model_info)) {
  this_model <- model_info[[model_name]]$model
  this_data <- model_info[[model_name]]$data
  
  preds <- predict(this_model, type = "response")
  auc_value <- as.numeric(auc(this_data$used, preds))
  
  auc_results <- rbind(auc_results, data.frame(model = model_name, auc = auc_value))
}



# AUC around 0.5 = essential random; closer to 1 = better fit
preds <- predict(mod_between.bbal_bi, type = "response")  
auc.mod <- as.numeric(auc(pnts_all$used, preds))

# Check for correlations
vif.mod <- car::vif(mod_between.bbal_bi)

# Check for patterns and that most points are within 95% confidence intervals
plot_binnedplot(mod_between.bbal_bi)

# Check ratio is close to 1
overdisp.mod <- overdisp_fun(mod_between.bbal_bi)

# Check ratio: >1 may indicate zero inflation (more 0s than expected); < 1 zero deflation
# Small P = evidence to suggest zero-inflation/deflation
zeroinf.mod <- testZeroInflation(mod_between.bbal_bi, plot = FALSE)

```

#### Model diagnostics

-   In the most simple model, season explained essentially no variation,
    so I dropped this.
-   The AUC value was \> 0.5, which suggests the model separates 'used'
    vs. 'not used' locations very well, or in other words, the
    predictors are strongly informative.
-   The VIF values are all \< 2, so there is no concerning
    multicollinearity.
-   The overdispersion function is slightly high. This could indicate
    moderate overdispersion (i.e. variance is higher than expected in a
    binomial model), but probably not a major concern.
-   There is some evidence for zero-inflation. However, this might be to
    expected, as zero counts are an inherent part of the process I am
    trying to model, and so a zero-inflated approach is likely
    unnecessary.

I am getting some convergence warnings, which, coupled with the high
variance in my slopes, might suggest that the random effects can't be
trusted. This could be because I have too many complex random effects,
or sparse data per group. I should probably try to fit the random
effects separately...

#### Model predictions

```{r model_outputs, fig.cap = "**Figure 1.3.** *Predicted probabilities of use for each fixed effect.*", fig.topcaption = TRUE}

# load("Data_outputs/bba_birdis_habitat_model_within_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_chl.RData")

load("Data_outputs/bba_birdis_habitat_model_between_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_chl.RData")

# Plot effects
pred_plot.sst <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "sst_scaled [all]"))
pred_plot.bathy <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "bathy_scaled [all]"))
pred_plot.chl <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "chl_scaled [all]"))
pred_plot.dist <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "dist_col_scaled [all]"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(plot.title = element_blank()), 
          pred_plot.dist + labs(x = "Scaled distance from colony (km)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          ncol = 2, nrow = 2)

all_plots

```

The results show that in general, birds select for median SST, deeper
waters, higher chlorophyll concentrations, and locations closer to the
colony. There is a high degree of individual variation, especially in
the use of SST and bathymetry.

#### Individual effects

```{r bbal_birdis_individual_within, fig.cap = "**Figure 1.4.** *Within-season, individual level responses to changes in the environmental variables. Individual slopes are plotted for the extent of each individual's range. Black line and grey interval shows population-level response and associated confidence interval.*", fig.topcaption = TRUE}

# Build plots
load("Data_outputs/bba_birdis_habitat_model_within_sst.RData")

env_vars = c("sst_scaled", "bathy_scaled", "chl_scaled")

for (i in 1:length(env_vars)) {

  habitat_var = env_vars[i]
  rnd_eff <- get_rnd_effects(mod_within.bbal_bi, habitat_var)
  
  # Get range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_sst = min(!!sym(habitat_var), na.rm = TRUE),
              max_sst = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(habitat_val)) 
  
  predictions.ind <- merge(predictions.ind, rnd_eff, by = "ID", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Get global predictions
  predictions.global <- data.frame(ggpredict(mod_within.bbal_bi, terms = paste0(habitat_var, " [all]")))
  
  ## Plot
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, colour = abs(slope), group = ID)) +
      scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      geom_line(data = predictions.global, aes(x = x, y = predicted), col = "black",
                 linetype = "dashed", linewidth = 1) +
      geom_ribbon(data = predictions.global, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "grey90", alpha = 0.5, col = "grey10") +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

ggarrange(sst_scaled_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) , 
          bathy_scaled_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.y = element_blank()), 
          chl_scaled_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")),
          ncol = 2, nrow = 2)


```

```{r bbal_birdis_individual_between, fig.cap = "**Figure 1.5.** *Between-season, individual level responses to changes in the environmental variables. Individual slopes are plotted for the extent of each individual's range. Black line and grey interval shows population-level response and associated confidence interval.*", fig.topcaption = TRUE}

# Build plots

env_vars = c("sst_scaled", "bathy_scaled", "chl_scaled")

for (i in 1:length(env_vars)) {

  habitat_var = env_vars[i]
  rnd_eff <- get_rnd_effects(mod_between.bbal_bi, habitat_var)
  
  # Get range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_sst = min(!!sym(habitat_var), na.rm = TRUE),
              max_sst = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(habitat_val)) 
  
  predictions.ind <- merge(predictions.ind, rnd_eff, by = "ID", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Get global predictions
  predictions.global <- data.frame(ggpredict(mod_between.bbal_bi, terms = paste0(habitat_var, " [all]")))
  
  ## Plot
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, colour = abs(slope), group = ID)) +
      scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      geom_line(data = predictions.global, aes(x = x, y = predicted), col = "black",
                 linetype = "dashed", linewidth = 1) +
      geom_ribbon(data = predictions.global, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "grey90", alpha = 0.5, col = "grey10") +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

ggarrange(sst_scaled_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) , 
          bathy_scaled_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.y = element_blank()), 
          chl_scaled_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")),
          ncol = 2, nrow = 2)


```

```{r bbal_birdis_calc_repeatability}

# Within season -----
summary(mod_within.bbal_bi)

## manual test
vc <- VarCorr(mod_within.bbal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.within <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.within <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.within <- V_ID_chl / (V_ID_chl + V_residual)



# Between season -----
summary(mod_between.bbal_bi)

## manual test
vc <- VarCorr(mod_between.bbal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.between <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.between <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.between <- V_ID_chl / (V_ID_chl + V_residual)


```

# Understanding within-individual centering

Currently, the model is fitted with within-population centering of the
environmental variables - i.e., each variable has been scaled to the
mean and 2SD of the entire datasets. Conversely, within-individual
centering removes between-individual variation, allowing me to focus
just on within-individual responses (i.e. plasticity).

Here I am trying a few different model structures to better understand
how within-individual centering, random slopes, and quadratic random
slopes change the predictions of my models. For simplicity, I have
removed the distance effect and the phase effect.

```{r exploring_structures, eval = FALSE}

#set.seed(817)
#pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

#pnts_subset <- pnts_all

# Population centering, linear slope ----

mod_pop_linear <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled +
         (1 + sst_scaled | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

summary(mod_pop_linear)
save(mod_pop_linear, file = "Data_outputs/bba_birdis_habitat_model_pop_linear.RData")

# Population centering, quadratic slope ----

mod_pop_quad <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (sst_scaled + sst_scaled2 | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

summary(mod_pop_quad)
save(mod_pop_quad, file = "Data_outputs/bba_birdis_habitat_model_pop_quad.RData")


# Within-individual centering, linear slope ----

mod_ind_linear <- glmmTMB(
  used ~ sst_within + sst_within2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_within | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

save(mod_ind_linear, file = "Data_outputs/bba_birdis_habitat_model_ind_linear.RData")

# Within-individual centering, quadratic slope ----

mod_ind_quad <- glmmTMB(
  used ~ sst_within + sst_within2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (sst_within + sst_within2 | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

save(mod_ind_quad, file = "Data_outputs/bba_birdis_habitat_model_ind_quad.RData")


```

```{r exploring_structures_plots, eval = TRUE}

load("Data_outputs/bba_birdis_habitat_model_pop_linear.RData")
load("Data_outputs/bba_birdis_habitat_model_pop_quad.RData")
load("Data_outputs/bba_birdis_habitat_model_ind_linear.RData")
load("Data_outputs/bba_birdis_habitat_model_ind_quad.RData")

# set.seed(817)
# pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

# Get individual range
ind_ranges <- pnts_subset %>%
    mutate(ID_season = paste(ring, season, sep = "_")) %>%
    group_by(ID_season) %>%
    summarise(min_sst = min(sst_scaled, na.rm = TRUE),
              max_sst = max(sst_scaled, na.rm = TRUE)) 

predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(sst_scaled = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(sst_scaled)) %>%
    mutate(sst_scaled2 = sst_scaled^2)

predictions.ind_within <- pnts_subset %>% group_by(ID_season) %>% select(ID_season, sst_within, sst_within2)

# Population centering, linear slope ----
rnd_eff <- ranef(mod_pop_linear)
rnd_eff <- rnd_eff$cond$ID_season
rnd_eff$ID_season <- rownames(rnd_eff)
rownames(rnd_eff) <- NULL
rnd_eff %<>% rename(slope = sst_scaled, intercept = '(Intercept)')

predictions.ind_pop.linear <- merge(predictions.ind, rnd_eff, by = "ID_season", all.x = TRUE)
predictions.ind_pop.linear$logit <- predictions.ind_pop.linear$intercept + predictions.ind_pop.linear$slope * predictions.ind_pop.linear$sst_scaled
predictions.ind_pop.linear$probability <- 1 / (1 + exp(-predictions.ind_pop.linear$logit))  
  
## Plot
predictions.ind_pop.linear %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_pop.linear <- ggplot() +
    geom_line(data = predictions.ind_pop.linear, aes(x = sst_scaled, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
  scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use", x = "SST, scaled", title = "Linear random slope, population-level environmental centering")


# Population centering, quadratic slope ----

rnd_eff_pop_quad <- ranef(mod_pop_quad)
rnd_eff_pop_quad <- rnd_eff_pop_quad$cond$ID_season
rnd_eff_pop_quad$ID_season <- rownames(rnd_eff_pop_quad)
rownames(rnd_eff_pop_quad) <- NULL
rnd_eff_pop_quad %<>% rename(slope_linear = sst_scaled, slope_quadratic = sst_scaled2, intercept = '(Intercept)')

predictions.ind_pop.quad <- merge(predictions.ind, rnd_eff_pop_quad, by = "ID_season", all.x = TRUE)
predictions.ind_pop.quad$logit <- predictions.ind_pop.quad$intercept + 
                                  predictions.ind_pop.quad$slope_linear * predictions.ind_pop.quad$sst_scaled + 
                                  predictions.ind_pop.quad$slope_quadratic * predictions.ind_pop.quad$sst_scaled2
predictions.ind_pop.quad$probability <- 1 / (1 + exp(-predictions.ind_pop.quad$logit))

## Plot
predictions.ind_pop.quad %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_pop.quad <- ggplot() +
    geom_line(data = predictions.ind_pop.quad, aes(x = sst_scaled, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
    scale_colour_gradient(low = "#f4edff", high = "#340088") +  
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Probability of use", x = "SST, scaled", title = "Quadratic random slope, population-level environmental centering")


# Within-individual centering, linear slope ----

rnd_eff_ind_linear <- ranef(mod_ind_linear)
rnd_eff_ind_linear <- rnd_eff_ind_linear$cond$ID_season
rnd_eff_ind_linear$ID_season <- rownames(rnd_eff_ind_linear)
rownames(rnd_eff_ind_linear) <- NULL
rnd_eff_ind_linear %<>% rename(slope = sst_within, intercept = '(Intercept)')

predictions.ind_within.linear <- merge(predictions.ind_within, rnd_eff_ind_linear, by = "ID_season", all.x = TRUE)
predictions.ind_within.linear$logit <- predictions.ind_within.linear$intercept + predictions.ind_within.linear$slope * predictions.ind_within.linear$sst_within
predictions.ind_within.linear$probability <- 1 / (1 + exp(-predictions.ind_within.linear$logit))

## Plot
predictions.ind_within.linear %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_ind.linear <- ggplot() +
    geom_line(data = predictions.ind_within.linear, aes(x = sst_within, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
    scale_colour_gradient(low = "#f4edff", high = "#340088") +  
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Probability of use", x = "SST, within-individual centered", title = "Linear random slope, within-individual centering")

# Within-individual centering, quadratic slope ----

rnd_eff_ind_quad <- ranef(mod_ind_quad)
rnd_eff_ind_quad <- rnd_eff_ind_quad$cond$ID_season
rnd_eff_ind_quad$ID_season <- rownames(rnd_eff_ind_quad)
rownames(rnd_eff_ind_quad) <- NULL
rnd_eff_ind_quad %<>% rename(slope_linear = sst_within, slope_quadratic = sst_within2, intercept = '(Intercept)')

predictions.ind_within.quad <- merge(predictions.ind_within, rnd_eff_ind_quad, by = "ID_season", all.x = TRUE)
predictions.ind_within.quad$logit <- predictions.ind_within.quad$intercept + 
                                     predictions.ind_within.quad$slope_linear * predictions.ind_within.quad$sst_within + 
                                     predictions.ind_within.quad$slope_quadratic * predictions.ind_within.quad$sst_within2
predictions.ind_within.quad$probability <- 1 / (1 + exp(-predictions.ind_within.quad$logit))

## Plot
predictions.ind_within.quad %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_ind.quad <- ggplot() +
    geom_line(data = predictions.ind_within.quad, aes(x = sst_within, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
    scale_colour_gradient(low = "#f4edff", high = "#340088") +  
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Probability of use", x = "SST, within-individual centered", title = "Quadratic random slope, within-individual centering")


# Arrange plots ----------

ggarrange(plot_pop.linear, plot_pop.quad,
          plot_ind.linear, plot_ind.quad,
          ncol = 2, nrow = 2)

#knitr::knit_exit()

```
