---
title: "Habitat selection analyses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---


# Set-up

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Functions for GPS processing and plotting
source("RPT_functions.R")
source("condR.R")

## MasterBayes installation
# install.packages(c('genetics', 'gtools', 'kinship2'))
# install.packages("https://cran.r-project.org/src/contrib/Archive/MasterBayes/MasterBayes_2.58.tar.gz", repos = NULL, type = "source", dependencies = T)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "tidyr", "lubridate", "DHARMa", "data.table", "ggpubr",  "kableExtra", "ggridges", "corrplot", "pROC", "ggeffects", "glmmTMB", "brms", "MASS", "MCMCglmm", "igraph", "MasterBayes", "pheatmap", "nadiv") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Scale function
scale_2SD <- function(variable) { (variable - mean(variable, na.rm = TRUE)) / sd(variable, na.rm = TRUE) * 2 }

# Set colours
bathy_col = "#0072B2" #"#1E88E5"
chl_col = "#009E73" #"#004D40"
sst_col = "#F0E442" #"#D81B60"

col_between <- "#91c5f2"
col_within <- "#0b3860"

# Heritability priors

priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(exponential(1), class = "sd", group = "ring"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(exponential(1), class = "sigma") )


# Sim number for bootstrapping CIs (INCREASE FOR PUBLICATION)
my_nsim = 500

```

# Black-browed albatrosses

## Bird Island

```{r bbal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/bbal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2 )

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r bbal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

## Remove birds with only one trip
pnts.used %<>%
  group_by(ID_season) %>%
  mutate(n_trips = n_distinct(tripID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Thin to every 3 hours to reduce autocorrelation
pnts.thinned <- pnts.used %>%
  # group_by(tripID, date_hourly) %>%
  # slice(1) %>%
  mutate(time_3h = floor_date(date_hourly, "3 hours")) %>%
  group_by(tripID, time_3h) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(tripID, date_hourly)

## Set loop parameters
env_vars <- c("bathy", "chlA", "SST")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  print(paste0("Processing ", env_vars[i], "..."))
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.thinned)

  ## Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = my_nsim, .progress = "txt")
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "tripID", ] %>% 
    separate(level, into = c("ring", "season", "trip")) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 

  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

bbal_birdis_var_used <- do.call("rbind", var.list)
bbal_birdis_ranefs_used <- do.call("rbind", ranef.list)

save(bbal_birdis_var_used, file = "Data_outputs/bbal_birdis_habitat_used_variances.RData")
save(bbal_birdis_ranefs_used, file = "Data_outputs/bbal_birdis_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_birdis_used_plot <- annotate_figure(bbal_birdis_used_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_used_plot, file = "Data_outputs/bbal_birdis_used_plot.RData")

png(filename = "Figures/bbal_birdis_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r bbal_birdis_table_variance_habitat_used}

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

```{r bbal_birdis_fit_models, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Fit the models ====

# Within season #

bbal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_bathy, file = "Data_outputs/bbal_birdis_glmm_within_bathy.RData")
rm(bbal_birdis_glmm.within_bathy); gc()


bbal_birdis_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_chl, file = "Data_outputs/bbal_birdis_glmm_within_chl.RData")
rm(bbal_birdis_glmm.within_chl); gc()


bbal_birdis_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_sst, file = "Data_outputs/bbal_birdis_glmm_within_sst.RData")
rm(bbal_birdis_glmm.within_sst); gc()


# Between season #

bbal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID) + (0 + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_bathy, file = "Data_outputs/bbal_birdis_glmm_between_bathy.RData")
rm(bbal_birdis_glmm.between_bathy); gc()


bbal_birdis_glmm.between_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_chl, file = "Data_outputs/bbal_birdis_glmm_between_chl.RData")
rm(bbal_birdis_glmm.between_chl); gc()


bbal_birdis_glmm.between_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_birdis_glmm.between_sst, file = "Data_outputs/bbal_birdis_glmm_between_sst.RData")
rm(bbal_birdis_glmm.between_sst); gc()


```

#### Conditional repeatability

```{r bbal_birdis_cond_repeatability}

load("Data_inputs/bbal_birdis_habitat_data.RData")

env_vars = c("bathy", "chl", "sst")

# [ 1 ] Calculate conditional repeatability ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(
  bathy = bathy_scaled,
  chl = chl_scaled,
  sst = sst_scaled )

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {
  
  habitat_var = paste0(env_vars[i], "_scaled")

  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars[i], ".RData"))
  
  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics ====
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Between ==== 
  
  ## Estimate conditional repeatability ####
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
    nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var = res_between$Vi,
                                    R2 = res_between$R2s, 
                                    marginal_R2 = res_between$Rmar,
                                    med_rpt = median(res_between$condR$condR),
                                    LCL_rpt = quantile(res_between$condR$condR, 0.025, na.rm = TRUE),
                                    UCL_rpt = quantile(res_between$condR$condR, 0.975, na.rm = TRUE),
                                    env = env_vars[i],
                                    species = "BBAL",
                                    colony = "Bird Island", 
                                    level = "between")
  
  rownames(res_between.summary) <- NULL
  
  ## Bootstrap conditional repeatability uncertainty ####
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "between")
  
  
  # Within ====
  
  ## Estimate conditional repeatability ####
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
    nrow = 2) 
  
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var = res_within$Vi,
                                   R2 = res_within$R2s, 
                                   marginal_R2 = res_within$Rmar,
                                   med_rpt = median(res_within$condR$condR),
                                   LCL_rpt = quantile(res_within$condR$condR, 0.025, na.rm = TRUE),
                                   UCL_rpt = quantile(res_within$condR$condR, 0.975, na.rm = TRUE),
                                   env = env_vars[i],
                                   species = "BBAL",
                                   colony = "Bird Island", 
                                   level = "within")
  
  rownames(res_within.summary) <- NULL
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])    
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID_season) / df_ID_season
    var_slope_sim <- var_slope * rchisq(1, df = df_ID_season) / df_ID_season
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("bbal_birdis_", env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_birdis_glmm.within_", env_vars[i]))
  rm(list = paste0("bbal_birdis_glmm.between_", env_vars[i]))
  gc()
  
}

bbal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(bbal_birdis_condR_results, file = "Data_outputs/bbal_birdis_condR_results.RData")

```

```{r bbal_birdis_plot_slopes}

# [ 2 ] Plot individual slopes ----

rnd_effs.list <- list()

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("bbal_birdis_", env_vars[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)
save(rnd_effs.out, file = "Data_outputs/bbal_birdis_habitat_selection.RData")

```

```{r bbal_birdis_preference_figures}

# [ 3 ] Final plot ----

bbal_birdis_pref_plot <- ggarrange(
  bbal_birdis_bathy_pred_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
  bbal_birdis_chl_pred_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  bbal_birdis_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
  bbal_birdis_bathy_condR_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + ylim(0, 0.4) ,
  bbal_birdis_chl_condR_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 0.4) + theme(axis.title.y = element_blank()),
  bbal_birdis_sst_condR_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 0.4) + theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)

bbal_birdis_pref_plot <- annotate_figure(bbal_birdis_pref_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_pref_plot, file = "Data_outputs/bbal_birdis_selection_plot.RData")


png("Figures/bbal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
bbal_birdis_pref_plot
dev.off()


```

### Heritability

#### Habitat used

```{r bbal_birdis_heritability_used }

# (1) Load the data ----
load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")

ranefs.df_wide <- bbal_birdis_ranefs_used %>%
  select(c(ring, season, trip, estimate, std.error, env)) %>%
  pivot_wider(names_from = env, 
              values_from = c(estimate, std.error),
               names_glue = "{env}_{.value}")

pedigree.df <- read.csv("Data_inputs/bbal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# Fix birds that are recorded as both dams and sires
pedigree.df$sire[pedigree.df$id == "1271009"] <- "1318880"
pedigree.df$dam[pedigree.df$id == "1271009"] <- "1318939"

# (2) Make the pedigree ----

Amat <- make_Amat(ranefs.df_wide, pedigree.df)

# (3) Fit univariate models covering responses to each environment ----

ranefs.df_wide %<>% mutate(ring_pe = ring, animal = ring)

priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(student_t(3,0,1), class = "sd", group = "ring_pe"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(student_t(3, 0, 1), class = "sigma"))

# Build univariate models
bbal_birdis_brm.heritability_used_bathy <- brm(
  bathy_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(bbal_birdis_brm.heritability_used_bathy, file = "Data_outputs/bbal_birdis_heritability_used_bathy.RData")

bbal_birdis_brm.heritability_used_chl <- brm(
  chlA_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(bbal_birdis_brm.heritability_used_chl, file = "Data_outputs/bbal_birdis_heritability_used_chl.RData")

bbal_birdis_brm.heritability_used_sst <- brm(
  SST_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(bbal_birdis_brm.heritability_used_sst, file = "Data_outputs/bbal_birdis_heritability_used_sst.RData")


# Estimate heritability
h2_bathy <- get_h2(bbal_birdis_brm.heritability_used_bathy)
h2_chl <- get_h2(bbal_birdis_brm.heritability_used_chl)
h2_sst <- get_h2(bbal_birdis_brm.heritability_used_sst)

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))


ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

save(posteriors, file = "Data_outputs/bbal_birdis_habitat_used_heritability_posteriors.RData")


```

#### Habitat preference


#### Extract individual slopes

```{r bbal_birdis_extract_slopes}

load("Data_inputs/bbal_birdis_habitat_data.RData")

## Store envronmental variables
env_vars = c("bathy", "chl", "sst")

## Get meta to recover later
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

slopes.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  my_mod = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  
  # Extract slopes - mean and SE
  slopes <- ranef(my_mod)$cond$tripID
  
  slopes_df <- data.frame(
        ID = rownames(slopes),
        slope = slopes[,2])

  slopes_df %<>% 
    left_join(pnts_meta %>% rename(ID = tripID), by = "ID") %>%
    separate(ID, into = c("ring", "season", "tripID"), remove = FALSE) %>%
    mutate(env = env_vars[i]) %>%
    relocate(c(ring, season, tripID))

  slopes.list[[i]] <- slopes_df
  
}

bbal_birdis_habitat_slopes <- do.call("rbind", slopes.list)
save(bbal_birdis_habitat_slopes, file = "Data_outputs/bbal_birdis_habitat_slopes.RData")

```

#### Estimate heritability

```{r bbal_birdis_heritability }

# (1) Load the data ----
load("Data_outputs/bbal_birdis_habitat_slopes.RData")

pedigree.df <- read.csv("Data_inputs/bbal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
pedigree.df$sire[pedigree.df$id == "1271009"] <- "1318880"
pedigree.df$dam[pedigree.df$id == "1271009"] <- "1318939"

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(bbal_birdis_habitat_slopes, pedigree.df)

## Optional - visualise the pruned peigree (go into function to get pruned pedigree)

# edges <- pedigree.pruned %>%
#   tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
#   dplyr::select(from = value, to = id)
# 
# ped_graph <- graph_from_data_frame(edges, directed = TRUE)
# 
# plot(ped_graph,
#      vertex.size = 3,
#      vertex.label = NA,
#      edge.arrow.size = 0.2,
#      layout = layout_with_sugiyama(ped_graph)$layout)


## Optional - visualise sparse matrix

# A_sparse <- as(Amat, "dgCMatrix")
# summary(A_sparse)

### Check how many non-zero off-diagonal elements
# nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
# density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
# cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")
# 
# pheatmap(as.matrix(Amat), show_rownames = FALSE, show_colnames = FALSE)


# (3) Fit multivariate model covering responses to each environment ----
bbal_birdis_habitat_slopes %<>%
  mutate(animal = ring, 
         tripID = ID, 
         ring_pe = ring)

## Make dataframe wide (i.e. slope per environment)
bbal_birdis_slopes.wide <- bbal_birdis_habitat_slopes %>%
  pivot_wider(names_from = env, 
              values_from = slope,
              names_glue = "{env}_slope")

## Only retain birds with > 1 trip
bbal_birdis_slopes.wide %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% ungroup() %>% select(-n_trips)

# Build univariate models and extract h2

bbal_birdis_brm.heritability_bathy <- brm(
  bathy_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = bbal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_bathy <- (VarCorr(bbal_birdis_brm.heritability_bathy, summary = FALSE)$animal$sd)^2
v_r_bathy <- (VarCorr(bbal_birdis_brm.heritability_bathy, summary = FALSE)$residual$sd)^2
h2_bathy <- as.mcmc(v_animal_bathy / (v_animal_bathy + v_r_bathy) )

save(bbal_birdis_brm.heritability_bathy, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_bathy.RData")


bbal_birdis_brm.heritability_chl <- brm(
  chl_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = bbal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_chl <- (VarCorr(bbal_birdis_brm.heritability_chl, summary = FALSE)$animal$sd)^2
v_r_chl <- (VarCorr(bbal_birdis_brm.heritability_chl, summary = FALSE)$residual$sd)^2
h2_chl <- as.mcmc(v_animal_chl / (v_animal_chl + v_r_chl) )

save(bbal_birdis_brm.heritability_chl, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_chl.RData")


bbal_birdis_brm.heritability_sst <- brm(
  sst_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = bbal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_sst <- (VarCorr(bbal_birdis_brm.heritability_sst, summary = FALSE)$animal$sd)^2
v_r_sst <- (VarCorr(bbal_birdis_brm.heritability_sst, summary = FALSE)$residual$sd)^2
h2_sst <- as.mcmc(v_animal_sst / (v_animal_sst + v_r_sst) )

save(bbal_birdis_brm.heritability_sst, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_sst.RData")


# Extract posteriors of h2

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
colnames(posteriors) <- c("h2_bathy", "h2_chl", "h2_sst")
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/bbal_birdis_habitat_pref_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()


```

- Posterior distributions for h2 are biomdal, indicating stong confounding between additive genetic variance and permanent environmental (repeatability) effects
- Collapsing to one slope per bird gives a model that fails to converge with highly uncertain posteriors - i.e. when repeatability information is removed, there is insufficient signal to estimate heritability
- Overall, reasonable evidence that while there may be individual consistency, preference is unlikely to be heritable

e.g. In multivariate animal models, posterior distributions of heritability estimates for bathymetry and chlorophyll were bimodal, reflecting poor separation of additive genetic variance from permanent environmental effects. A diagnostic single-slope model, in which one slope per bird was analysed, failed to converge and yielded highly uncertain variance component estimates. Together these results suggest that while individual foraging preferences are partially repeatable, there is no strong evidence that this variation is heritable.

## Kerguelen

For Kerguelen especially, birds' preference for shelf zones makes the
bathymetry model fall apart. To tackle this, I have calculated distance
from the shelf, and split preferences into on shelf, shelf break, and
off shelf. This seems to resolve my problems. The shelf break is
delimited at 1000m following Hindell et al. 2011.

```{r bbal_ker_process_data, eval = FALSE}

load("Data_inputs/bbal_ker_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )

## Add binomial bathymetry variable
pnts_all %<>%
  mutate(dist_shelf_bin = case_when(
            dist_shelf_break_signed > 20  ~ "on_shelf",       # >20 km inside shelf
            dist_shelf_break_signed >= -20 & dist_shelf_break_signed <= 20 ~ "near_break", # ±20 km of break
            dist_shelf_break_signed < -20  ~ "off_shelf"       # >20 km offshore (deeper)
    ) )

pnts_all$dist_shelf_bin <- factor(pnts_all$dist_shelf_bin, levels = c("on_shelf", "near_break", "off_shelf"))

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_ker_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r bbal_ker_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

## Remove birds with only one trip
pnts.used %<>%
  group_by(ID_season) %>%
  mutate(n_trips = n_distinct(tripID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Thin to every 3 hours to reduce autocorrelation
pnts.thinned <- pnts.used %>%
  # group_by(tripID, date_hourly) %>%
  # slice(1) %>%
  mutate(time_3h = floor_date(date_hourly, "3 hours")) %>%
  group_by(tripID, time_3h) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(tripID, date_hourly)

## Set loop parameters
env_vars <- c("bathy", "chlA", "SST")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.thinned)
  
  ## Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = my_nsim, .progress = "txt")
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "tripID", ] %>% 
    separate(level, into = c("ring", "season", "trip")) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}


bbal_ker_var_used <- do.call("rbind", var.list)
bbal_ker_ranefs_used <- do.call("rbind", ranef.list)

save(bbal_ker_var_used, file = "Data_outputs/bbal_ker_habitat_used_variances.RData")
save(bbal_ker_ranefs_used, file = "Data_outputs/bbal_ker_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_ker_plot_habitat_used }

load("Data_outputs/bbal_ker_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)

split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_ker_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_ker_used_plot <- annotate_figure(bbal_ker_used_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_used_plot, file = "Data_outputs/bbal_ker_used_plot.RData")

png(filename = "Figures/bbal_ker_habitat_used_ranef_plot.png", width = 10, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r bbal_ker_table_variance_habitat_used}

load("Data_outputs/bbal_ker_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

```{r bbal_ker_fit_model, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Fit the models ====

# Within season #

bbal_ker_glmm.within_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID_season) + (0 + dist_shelf_bin | ID_season),              
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_bathy, file = "Data_outputs/bbal_ker_glmm_within_bathy.RData")
rm(bbal_ker_glmm.within_bathy); gc()


bbal_ker_glmm.within_chl <-  glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),           
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_chl, file = "Data_outputs/bbal_ker_glmm_within_chl.RData")
rm(bbal_ker_glmm.within_chl); gc()


bbal_ker_glmm.within_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),       
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_sst, file = "Data_outputs/bbal_ker_glmm_within_sst.RData")
rm(bbal_ker_glmm.within_sst); gc()


# Between season #

bbal_ker_glmm.between_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID) + (0 + dist_shelf_bin | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_bathy, file = "Data_outputs/bbal_ker_glmm_between_bathy.RData")
rm(bbal_ker_glmm.between_bathy); gc()


bbal_ker_glmm.between_chl <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_chl, file = "Data_outputs/bbal_ker_glmm_between_chl.RData")
rm(bbal_ker_glmm.between_chl); gc()


bbal_ker_glmm.between_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_sst, file = "Data_outputs/bbal_ker_glmm_between_sst.RData")
rm(bbal_ker_glmm.between_sst); gc()


```


#### Conditional repeatability

```{r bbal_ker_cond_repeatability}

load("Data_inputs/bbal_ker_habitat_data.RData")

# [ 1A ] Calculate conditional repeatability - Chl & SST ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
env_vars = c("chl", "sst", "bathy")

bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(
  sst = sst_scaled,
  chl = chl_scaled )

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:2) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics ====
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Between ==== 
  
  ## Estimate conditional repeatability ####
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
    nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var = res_between$Vi,
                                    R2 = res_between$R2s, 
                                    marginal_R2 = res_between$Rmar,
                                    med_rpt = median(res_between$condR$condR),
                                    LCL_rpt = quantile(res_between$condR$condR, 0.025, na.rm = TRUE),
                                    UCL_rpt = quantile(res_between$condR$condR, 0.975, na.rm = TRUE),
                                    env = env_vars[i],
                                    species = "BBAL",
                                    colony = "Kerguelen", 
                                    level = "between")
  
  rownames(res_between.summary) <- NULL
  
  ## Bootstrap conditional repeatability uncertainty ####
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "between")
  
  
  # Within ====
  
  ## Estimate conditional repeatability ####
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
    nrow = 2) 
  
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var = res_within$Vi,
                                   R2 = res_within$R2s, 
                                   marginal_R2 = res_within$Rmar,
                                   med_rpt = median(res_within$condR$condR),
                                   LCL_rpt = quantile(res_within$condR$condR, 0.025, na.rm = TRUE),
                                   UCL_rpt = quantile(res_within$condR$condR, 0.975, na.rm = TRUE),
                                   env = env_vars[i],
                                   species = "BBAL",
                                   colony = "Kerguelen", 
                                   level = "within")
  
  rownames(res_within.summary) <- NULL
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])    
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID_season) / df_ID_season
    var_slope_sim <- var_slope * rchisq(1, df = df_ID_season) / df_ID_season
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.between %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty.within %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0("bbal_ker_", env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_ker_glmm.within_", env_vars[i]))
  rm(list = paste0("bbal_ker_glmm.between_", env_vars[i]))
  gc()
  
}

condR_results <- do.call("rbind", cond_res.list)

# [ 1B ] Calculate conditional repeatability - Bathymetry ------

# Between ====
  
## Calculate repeatability -----
load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
mod_between = get("bbal_ker_glmm.between_bathy")

## Extract variance-covariance matrices of random effects
vc <- VarCorr(mod_between)$cond
V_ID_intercept <- as.numeric(vc$ID[1,1])
V_tripID <- as.numeric(vc$tripID[1,1])
V_slopes_matrix <- vc$ID.1 

## Extract variances (diagonal) for each bathymetry level slope
V_slope <- diag(V_slopes_matrix)
names(V_slope) <- colnames(V_slopes_matrix)
  
## Residual variance for binomial logistic link
V_residual <- (pi^2) / 3

## Total random variance per level
V_random <- V_ID_intercept + V_slope + V_tripID
V_total <- V_ID_intercept + V_slope + V_tripID + V_residual

## Conditional repeatability per level
R2 <- V_random / V_total

R_cond.df_between <- data.frame(R2) 
R_cond.df_between <- cbind(rownames(R_cond.df_between), data.frame(R_cond.df_between, row.names = NULL))
R_cond.df_between %<>%
  mutate(species = "BBAL",
         colony = "Kerguelen",
         level = "between", 
         var_slope = V_slope,
         R2 = R2,
         marginal_R2 = NA) %>%
  rename(env = `rownames(R_cond.df_between)`) %>%
  relocate(c(species, colony, level, env, var_slope, R2, marginal_R2))

R_cond.df_between$env <- factor(R_cond.df_between$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

# Boostrap uncertainty ----

# Storage matrix for simulated repeatabilities (rows = simulations, cols = slope levels)
rpt_sim <- matrix(NA, nrow = n_sim, ncol = length(V_slope))
colnames(rpt_sim) <- names(V_slope)

for(i in seq_len(n_sim)){
  # Simulate variance components from scaled chi-square
  var_int_sim <- V_ID_intercept * rchisq(1, df_ID) / df_ID
  var_tripID_sim <- V_tripID * rchisq(1, df_tripID) / df_tripID
  
  var_slope_sim <- sapply(V_slope, function(v) v * rchisq(1, df_ID) / df_ID)

  # Calculate conditional repeatability for each slope level
  for(j in seq_along(var_slope_sim)){
    V_random_sim <- var_int_sim + var_slope_sim[j] + var_tripID_sim
    V_total_sim <- V_random_sim + V_residual
    rpt_sim[i, j] <- V_random_sim / V_total_sim
  }
}

# Summarize simulation results
rpt_summary <- apply(rpt_sim, 2, function(x) {
  c(
    median = median(x, na.rm=TRUE),
    lower = quantile(x, 0.025, na.rm=TRUE),
    upper = quantile(x, 0.975, na.rm=TRUE)
  )
}) %>% t() %>% as.data.frame()

rpt_summary$env <- rownames(rpt_summary)
rpt_summary %<>%
  rename(med_rpt = median, LCL_rpt = `lower.2.5%`, UCL_rpt = `upper.97.5%`) %>%
  select(env, med_rpt, LCL_rpt, UCL_rpt) %>%
  mutate(env = as.factor(env))

## Merge them with the main dataframe for plotting 
R_cond.df_between %<>% left_join(rpt_summary, by = "env")



# Within ####

# Calculate repeatability ----

load("Data_outputs/bbal_ker_glmm_within_bathy.RData")
mod_within = get("bbal_ker_glmm.within_bathy")

# Extract variance-covariance matrices of random effects
vc <- VarCorr(mod_within)$cond
V_IDseason_intercept <- as.numeric(vc$ID_season[1,1])
V_tripID <- as.numeric(vc$tripID[1,1])
V_slopes_matrix <- vc$ID_season.1 

# Extract variances (diagonal) for each bathymetry level slope
V_slope <- diag(V_slopes_matrix)
names(V_slope) <- colnames(V_slopes_matrix)
  
# Residual variance for binomial logistic link
V_residual <- (pi^2) / 3

# Total random variance per level
V_random <- V_IDseason_intercept + V_slope + V_tripID
V_total <- V_IDseason_intercept + V_slope + V_tripID + V_residual

# Conditional repeatability per level
R2 <- V_random / V_total

R_cond.df_within <- data.frame(R2) 
R_cond.df_within <- cbind(rownames(R_cond.df_within), data.frame(R_cond.df_within, row.names = NULL))
R_cond.df_within %<>%
  mutate(species = "BBAL",
         colony = "Kerguelen",
         level = "within", 
         var_slope = V_slope,
         R2 = R2,
         marginal_R2 = NA) %>%
  rename(env = `rownames(R_cond.df_within)`) %>%
  relocate(c(species, colony, level, env, var_slope, R2, marginal_R2))

R_cond.df_within$env <- factor(R_cond.df_within$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))


# Boostrap uncertainty ----

# Storage matrix for simulated repeatabilities (rows = simulations, cols = slope levels)
rpt_sim <- matrix(NA, nrow = n_sim, ncol = length(V_slope))
colnames(rpt_sim) <- names(V_slope)

for(i in seq_len(n_sim)){
  # Simulate variance components from scaled chi-square
  var_int_sim <- V_IDseason_intercept * rchisq(1, df_ID_season) / df_ID_season
  var_tripID_sim <- V_tripID * rchisq(1, df_tripID) / df_tripID
  
  var_slope_sim <- sapply(V_slope, function(v) v * rchisq(1, df_ID_season) / df_ID_season)

  # Calculate conditional repeatability for each slope level
  for(j in seq_along(var_slope_sim)){
    V_random_sim <- var_int_sim + var_slope_sim[j] + var_tripID_sim
    V_total_sim <- V_random_sim + V_residual
    rpt_sim[i, j] <- V_random_sim / V_total_sim
  }
}

# Summarize simulation results
rpt_summary <- apply(rpt_sim, 2, function(x) {
  c(
    median = median(x, na.rm=TRUE),
    lower = quantile(x, 0.025, na.rm=TRUE),
    upper = quantile(x, 0.975, na.rm=TRUE)
  )
}) %>% t() %>% as.data.frame()

rpt_summary$env <- rownames(rpt_summary)
rpt_summary %<>%
  rename(med_rpt = median, LCL_rpt = `lower.2.5%`, UCL_rpt = `upper.97.5%`) %>%
  select(env, med_rpt, LCL_rpt, UCL_rpt) %>%
  mutate(env = as.factor(env))

## Merge them with the main dataframe for plotting 
R_cond.df_within %<>% left_join(rpt_summary, by = "env")


# Bind it all together ====

bbal_ker_condR_results <- rbind(condR_results %>% rename(var_slope = var), R_cond.df_between, R_cond.df_within)
bbal_ker_condR_results %<>% rename(var = var_slope)

save(bbal_ker_condR_results, file = "Data_outputs/bbal_ker_condR_results.RData")




# Make the plot ====
bathy_labels <- c("On shelf", "Near shelf break", "Off shelf")

condR_plot <- ggplot() +
    geom_point(data = R_cond.df_between, aes(x= env, y = med_rpt), col = col_between, size = 3) +
    geom_errorbar(data = R_cond.df_between, aes(x = env, ymin = LCL_rpt, ymax = UCL_rpt), width = 0.1, colour = col_between) +
    geom_point(data = R_cond.df_within, aes(x= env, y = med_rpt), col = col_within, size = 3) +
    geom_errorbar(data = R_cond.df_within, aes(x = env, ymin = LCL_rpt, ymax = UCL_rpt), width = 0.1, colour = col_within) +
    labs(y = "Conditional repeatability") +
    scale_x_discrete(labels = bathy_labels) +
    expand_limits(y = c(0, 0.2)) +
    theme_bw() 

assign("bbal_ker_bathy_condR_plot", condR_plot)

```

```{r bbal_ker_plot_slopes}

# [ 2 ] Plot individual slopes ---- 

env_vars = c("chl", "sst", "bathy")

rnd_effs.list <- list()

for (i in 1:2) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring")
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("bbal_ker_", env_vars[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)


## Bathymetry only ====

levels_bathy <- levels(mod_df$dist_shelf_bin)

fixef_vals.between <- fixef(bbal_ker_glmm.between_bathy)$cond
fixef_vals.within <- fixef(bbal_ker_glmm.within_bathy)$cond

# Prepare a data frame to hold predictions
ranef_slopes.between <- ranef(bbal_ker_glmm.between_bathy)$cond$ID
ranef_slopes.within  <- ranef(bbal_ker_glmm.within_bathy)$cond$tripID

pred_df.between <- expand.grid(ID = rownames(ranef_slopes.between),
                       dist_shelf_bin = levels_bathy)
pred_df.within <- expand.grid(ID = rownames(ranef_slopes.within),
                       dist_shelf_bin = levels_bathy)

# Calculate linear predictor per bird & category: - STOPPED HERE
pred_df.between %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.between[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.between[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.between[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.between[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.between["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.between["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.between["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()


pred_df.within %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.within[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.within[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.within[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.within[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.within["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.within["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.within["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()

## Bind together
pred_df.bathy <- rbind(pred_df.between %>% mutate(season_type = "Between seasons"),
                       pred_df.within %>% mutate(season_type = "Within season"))

# Plot lines connecting points per bird
pred_plot <- ggplot() +
  geom_line(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  geom_point(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  facet_grid(~season_type) +
  scale_x_discrete(labels = c("On shelf", "Near shelf", "Off shelf")) +
  labs(x = "Bathymetry category",
       y = "Probability of use") +
  theme_bw()

assign("bbal_ker_bathy_pred_plot", pred_plot)

save(rnd_effs.out, file = "Data_outputs/bbal_ker_habitat_selection.RData")

```

```{r bbal_ker_preference_figures}

# Make the plot ====

bbal_ker_pref_plot <- ggarrange(
  bbal_ker_bathy_pred_plot +
    ylim(0, 1) + theme(axis.title.x = element_blank()), 
  bbal_ker_chl_pred_plot +
    labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  bbal_ker_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
  bbal_ker_bathy_condR_plot +
    labs(x = "Bathymetry category") + ylim(0, 0.45),
  bbal_ker_chl_condR_plot +
    labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 0.45) + theme(axis.title.y = element_blank()),
  bbal_ker_sst_condR_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 0.45) + theme(axis.title.y = element_blank()), 
                    ncol = 3, nrow = 2)

bbal_ker_pref_plot <- annotate_figure(bbal_ker_pref_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_pref_plot, file = "Data_outputs/bbal_ker_selection_plot.RData")


png("Figures/bbal_ker_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
bbal_ker_pref_plot
dev.off()


```


# Wandering albatrosses

## Bird Island

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )


# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/waal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r waal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

## Remove birds with only one trip
pnts.used %<>%
  group_by(ID_season) %>%
  mutate(n_trips = n_distinct(tripID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Thin to every 3 hours to reduce autocorrelation
pnts.thinned <- pnts.used %>%
  # group_by(tripID, date_hourly) %>%
  # slice(1) %>%
  mutate(time_3h = floor_date(date_hourly, "3 hours")) %>%
  group_by(tripID, time_3h) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(tripID, date_hourly)

## Set loop parameters
env_vars <- c("bathy", "chlA", "SST")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  print(paste0("Processing ", env_vars[i]))
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1 | tripID) + (1| season) + (1 | ID_season)")), data = pnts.thinned)

  # Bootstrap 95% CIs
  set.seed(123)

  boot <- bootMer(mod, FUN = repeat_func_waBI, nsim = my_nsim, .progress = "txt")
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "tripID", ] %>% 
    separate(level, into = c("ring", "season", "trip")) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_within <- var_components[var_components$grp == "ID_season", "vcov"]
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = NA,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = NA,
            UCL_across = NA) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

waal_birdis_var_used <- do.call("rbind", var.list)
waal_birdis_ranefs_used <- do.call("rbind", ranef.list)

save(waal_birdis_var_used, file = "Data_outputs/waal_birdis_habitat_used_variances.RData")
save(waal_birdis_ranefs_used, file = "Data_outputs/waal_birdis_habitat_used_ranef.RData")

```

#### Plot 'used' repeatability

```{r waal_birdis_used_rpt}

load("Data_outputs/waal_birdis_habitat_used_variances.RData")

# Wrangle the data into long form
waal_birdis_var_used.long <- waal_birdis_var_used %>%
      pivot_longer(cols = matches("_(within|across)$"),
                   names_to = c(".value", "group"),
                   names_pattern = "(.*)_(within|across)") %>%
  select(-c(Effect, Variance, `Std. dev`)) %>%
  distinct()

# Change factors
waal_birdis_var_used.long %<>%
 mutate(Environment = recode(Environment,
      "bathy" = "Bathymetry",
      "chlA" = "Chlorophyll",
      "SST" = "SST")) %>%
  arrange(Environment) %>%
  mutate(Environment = factor(Environment, levels = rev(unique(Environment))))

ggplot(waal_birdis_var_used.long, aes(x = rpt, y = Environment, colour = group)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = LCL, xmax = UCL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Repeatability (95% CI)") +
  scale_colour_manual(
    values = c(col_between, col_within),
    labels = c("Across season", "Within seasons") ) +
  theme_bw() +
  theme(legend.position = "none")

```

#### Plot random intercepts

```{r waal_birdis_habitat_used_ranef, fig.dim = c(10,5)}

load("Data_outputs/waal_birdis_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
waal_birdis_used_plot <- annotate_figure(waal_birdis_used_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_used_plot, file = "Data_outputs/waal_birdis_used_plot.RData")

png(filename = "Figures//waal_birdis_habitat_used_ranef_plot.png", width = 14, height = 6, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


load("Data_outputs/waal_birdis_habitat_used_variances.RData")

head(waal_birdis_var_used)

```

### Habitat selection

#### Fit the models

```{r waal_birdis_fit_models, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )


# Fit the models ====

# Within season #

waal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_bathy, file = "Data_outputs/waal_birdis_glmm_within_bathy.RData")
rm(waal_birdis_glmm.within_bathy); gc()


waal_birdis_glmm.within_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_chl, file = "Data_outputs/waal_birdis_glmm_within_chl.RData")
rm(waal_birdis_glmm.within_chl); gc()


waal_birdis_glmm.within_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),      
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_sst, file = "Data_outputs/waal_birdis_glmm_within_sst.RData")
rm(waal_birdis_glmm.within_sst); gc()

```


#### Conditional repeatability

```{r waal_birdis_conditional_repeatability}

load("Data_inputs/waal_birdis_habitat_data.RData")

env_vars = c("bathy", "chl", "sst")

# [ 1 ] Calculate conditional repeatability ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled_log)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(
  bathy = bathy_scaled,
  chl = chl_scaled,
  sst = sst_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars[i], ".RData"))
  
  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars[i]))
  
  ## Get environmental metrics
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Within ====
  
  ## Estimate conditional repeatability ====

  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,   
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            LCL_rpt = quantile(res_within$condR$condR, 0.025, na.rm = TRUE),
                            UCL_rpt = quantile(res_within$condR$condR, 0.975, na.rm = TRUE),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Bird Island", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
  
  ## Bootstrap conditional repeatability uncertainty ####
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1]) 
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID_season) / df_ID_season
    var_slope_sim <- var_slope * rchisq(1, df = df_ID_season) / df_ID_season
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  res.summary <- res_within.summary %>% 
    relocate(c(species, colony, level, env), .before = var)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0("waal_birdis_", env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within)
  rm(list = paste0("waal_birdis_glmm.within_", env_vars[i]))
  gc()
  
}

waal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(waal_birdis_condR_results, file = "Data_outputs/waal_birdis_condR_results.RData")

```

```{r waal_birdis_plot_slopes}

load("Data_inputs/waal_birdis_habitat_data.RData")
env_vars = c("bathy", "chl", "sst")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars[i], ".RData"))
  
  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  
  rnd_eff.all <- rnd_eff.within
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID), alpha = 0.2, colour = line_colour) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
   
  assign(paste0("waal_birdis_", env_vars[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_birdis_preference_figures}

# [ 3 ] Final plot ----

waal_birdis_pref_plot <- ggarrange(
  waal_birdis_bathy_pred_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
  waal_birdis_chl_pred_plot + 
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  waal_birdis_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
  waal_birdis_bathy_condR_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + ylim(0, 0.15),
  waal_birdis_chl_condR_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 0.15) + theme(axis.title.y = element_blank()),
  waal_birdis_sst_condR_plot + 
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 0.15) + theme(axis.title.y = element_blank()), 
                         ncol = 3, nrow = 2)

waal_birdis_pref_plot <- annotate_figure(waal_birdis_pref_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_pref_plot, file = "Data_outputs/waal_birdis_selection_plot.RData")


png("Figures/waal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
waal_birdis_pref_plot
dev.off()



```

### Heritability

#### Habitat used

```{r waal_birdis_heritability_used }

# (1) Load the data ----
load("Data_outputs/waal_birdis_habitat_used_ranef.RData")

ranefs.df_wide <- waal_birdis_ranefs_used %>%
  select(c(ring, season, trip, estimate, std.error, env)) %>%
  pivot_wider(names_from = env, 
              values_from = c(estimate, std.error),
               names_glue = "{env}_{.value}")

pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(ranefs.df_wide, pedigree.df)

# (3) Fit univariate models covering responses to each environment ----

ranefs.df_wide %<>% mutate(ring_pe = ring, animal = ring)

priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(student_t(3,0,1), class = "sd", group = "ring_pe"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(student_t(3, 0, 1), class = "sigma"))

# Build univariate models
waal_birdis_brm.heritability_used_bathy <- brm(
  bathy_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(waal_birdis_brm.heritability_used_bathy, file = "Data_outputs/waal_birdis_heritability_used_bathy.RData")

waal_birdis_brm.heritability_used_chl <- brm(
  chlA_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(waal_birdis_brm.heritability_used_chl, file = "Data_outputs/waal_birdis_heritability_used_chl.RData")

waal_birdis_brm.heritability_used_sst <- brm(
  SST_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(waal_birdis_brm.heritability_used_sst, file = "Data_outputs/waal_birdis_heritability_used_sst.RData")


# Estimate heritability
h2_bathy <- get_h2(waal_birdis_brm.heritability_used_bathy)
h2_chl <- get_h2(waal_birdis_brm.heritability_used_chl)
h2_sst <- get_h2(waal_birdis_brm.heritability_used_sst)

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

save(posteriors, file = "Data_outputs/waal_birdis_habitat_used_heritability_posteriors.RData")


```

#### Habitat preference

##### Extract slopes

```{r waal_birdis_extract_slopes}

load("Data_inputs/waal_birdis_habitat_data.RData")

## Store envronmental variables
env_vars = c("bathy", "chl", "sst")

## Get meta to recover later
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

slopes.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars[i], ".RData"))
  my_mod = get(paste0("waal_birdis_glmm.within_", env_vars[i]))
  
  # Extract slopes - mean and SE
  slopes <- ranef(my_mod)$cond$tripID
  
  slopes_df <- data.frame(
        ID = rownames(slopes),
        slope = slopes[,2])

  slopes_df %<>% 
    left_join(pnts_meta %>% rename(ID = tripID), by = "ID") %>%
    separate(ID, into = c("ring", "season", "tripID"), remove = FALSE) %>%
    mutate(env = env_vars[i]) %>% 
    relocate(c(ring, season, tripID))

  slopes.list[[i]] <- slopes_df
  
}

waal_birdis_habitat_slopes <- do.call("rbind", slopes.list)
save(waal_birdis_habitat_slopes, file = "Data_outputs/waal_birdis_habitat_slopes.RData")


```

##### Estimate heritability

Try and run each model separately

```{r waal_birdis_heritability_pref }

# (1) Load the data ----
load("Data_outputs/waal_birdis_habitat_slopes.RData")

pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(waal_birdis_habitat_slopes, pedigree.df)

## Optional - visualise the pruned peigree (go into function to get pruned pedigree)

# edges <- pedigree.pruned %>%
#   tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
#   dplyr::select(from = value, to = id)
# 
# ped_graph <- graph_from_data_frame(edges, directed = TRUE)
# 
# plot(ped_graph,
#      vertex.size = 3,
#      vertex.label = NA,
#      edge.arrow.size = 0.2,
#      layout = layout_with_sugiyama(ped_graph)$layout)


## Optional - visualise sparse matrix

# A_sparse <- as(Amat, "dgCMatrix")
# summary(A_sparse)
#
### Check how many non-zero off-diagonal elements
# nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
# density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
# cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")
# 
# pheatmap(as.matrix(Amat), show_rownames = FALSE, show_colnames = FALSE)


# (3) Fit multivariate model covering responses to each environment ----
waal_birdis_habitat_slopes %<>%
  mutate(animal = ring, 
         tripID = ID, 
         ring_pe = ring)

## Make dataframe wide (i.e. slope per environment)
waal_birdis_slopes.wide <- waal_birdis_habitat_slopes %>%
  pivot_wider(names_from = env, 
              values_from = slope,
              names_glue = "{env}_slope")

## Only retain birds with > 1 trip
waal_birdis_slopes.wide %<>% group_by(ring) %>% mutate(n_trips = n_distinct(tripID)) %>% filter(n_trips > 1) %>% ungroup() %>% select(-n_trips)

# Build univariate models and extract h2
waal_birdis_brm.heritability_bathy <- brm(
  bathy_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_bathy <- (VarCorr(waal_birdis_brm.heritability_bathy, summary = FALSE)$animal$sd)^2
v_r_bathy <- (VarCorr(waal_birdis_brm.heritability_bathy, summary = FALSE)$residual$sd)^2
h2_bathy <- as.mcmc(v_animal_bathy / (v_animal_bathy + v_r_bathy) )

save(waal_birdis_brm.heritability_bathy, file = "Data_outputs/waal_birdis_habitat_pref_heritability_bathy.RData")


waal_birdis_brm.heritability_chl <- brm(
  chl_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_chl <- (VarCorr(waal_birdis_brm.heritability_chl, summary = FALSE)$animal$sd)^2
v_r_chl <- (VarCorr(waal_birdis_brm.heritability_chl, summary = FALSE)$residual$sd)^2
h2_chl <- as.mcmc(v_animal_chl / (v_animal_chl + v_r_chl) )

save(waal_birdis_brm.heritability_chl, file = "Data_outputs/waal_birdis_habitat_pref_heritability_chl.RData")


waal_birdis_brm.heritability_sst <- brm(
  sst_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_birdis_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_sst <- (VarCorr(waal_birdis_brm.heritability_sst, summary = FALSE)$animal$sd)^2
v_r_sst <- (VarCorr(waal_birdis_brm.heritability_sst, summary = FALSE)$residual$sd)^2
h2_sst <- as.mcmc(v_animal_sst / (v_animal_sst + v_r_sst) )

save(waal_birdis_brm.heritability_sst, file = "Data_outputs/waal_birdis_habitat_pref_heritability_sst.RData")


posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
colnames(posteriors) <- c("h2_bathy", "h2_chl", "h2_sst")
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/waal_birdis_habitat_pref_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```

## Crozet

Chlorophyll values are heavily skewed, with extremely high
concentrations occurring near coastlines in the ‘available’ dataset. To
prevent these large outliers from unduly influencing the model results,
I apply a log-transformation to chlorophyll.

```{r waal_cro_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_cro_sst-bathy-chla_data_population_level.RData")

## Fix some rings
pnts_all %<>%
  mutate(ring = recode(ring,
                       "11376" = "BS11376",
                       "11378" = "BS11378",
                       "25076" = "BS25076",
                       "8196" = "BS28916",
                       "8337" = "BS28337",
                       "BS8337" = "BS28337",
                       "BB21651" = "BS81651",
                       "BS7625" = "BS27625",
                       "BS8402" = "BS28402",
                       "BS8404" = "BS28404",
                       "BS9084" = "BS19084",
                       "BS9444" = "BS19444"))

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )

save(pnts_all, file = "Data_inputs/waal_cro_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r waal_cro_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

## Remove birds with only one trip
pnts.used %<>%
  group_by(ID_season) %>%
  mutate(n_trips = n_distinct(tripID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Thin to every 3 hours to reduce autocorrelation
pnts.thinned <- pnts.used %>%
  # group_by(tripID, date_hourly) %>%
  # slice(1) %>%
  mutate(time_3h = floor_date(date_hourly, "3 hours")) %>%
  group_by(tripID, time_3h) %>%
  slice(1) %>%
  ungroup() %>%
  arrange(tripID, date_hourly)

## Set loop parameters
env_vars <- c("bathy", "chlA", "SST")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  print(paste0("Processing ", env_vars[i], "..."))
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1 | season) + (1 | tripID) + (1 | ID) + (1 | ID_season)")), data = pnts.thinned)
  
  ## Bootstrap 95% CIs
  set.seed(123)
  boot <- bootMer(mod, FUN = repeat_func, nsim = my_nsim, .progress = "txt")
  rpt_CI <- apply(boot$t, 2, function(x) quantile(x, c(0.025, 0.975)))

  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "tripID", ] %>%
    separate(level, into = c("ring", "season", "trip")) %>%
    mutate(env = env_vars[i])

  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"] + var_across
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 

  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across,
            LCL_within = rpt_CI[1,1],
            UCL_within = rpt_CI[2,1],
            LCL_across = rpt_CI[1,2],
            UCL_across = rpt_CI[2,2]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

waal_cro_var_used <- do.call("rbind", var.list)
waal_cro_ranefs_used <- do.call("rbind", ranef.list)

save(waal_cro_var_used, file = "Data_outputs/waal_cro_habitat_used_variances.RData")
save(waal_cro_ranefs_used, file = "Data_outputs/waal_cro_habitat_used_ranef.RData")

```

#### Plot random intercepts

```{r waal_cro_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/waal_cro_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_cro_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
waal_cro_used_plot <- annotate_figure(waal_cro_used_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_used_plot, file = "Data_outputs/waal_cro_used_plot.RData")

png(filename = "Figures/waal_cro_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r waal_cro_table_variance_habitat_used}

load("Data_outputs/waal_cro_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

# Add whitespace to avoid collapsing bathymetry & chlorophyll together
var.table$`Across season rpt`[6:10] <- paste0(var.table$`Across season rpt`[6:10], ".00")

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

The random slope for chlorophyll-a was retained to account for potential
individual variation in response to environmental conditions. However,
the estimated correlation between the random intercept and slope for
chl_scaled was extremely high (0.99), suggesting that these components
are statistically confounded in the current model. This may reflect
either collinearity in the data or insufficient within-individual
variation in chlorophyll exposure to reliably separate baseline use from
individual responses.

As a result, I have droppped the polynomial terms as I think the models
are unable to cope with this degree of complexity.

```{r waal_cro_fit_models, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

# Add a small constant to avoid log(0) when logging chl
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

# Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Within season #

waal_cro_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),    data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_bathy, file = "Data_outputs/waal_cro_glmm_within_bathy.RData")
rm(waal_cro_glmm.within_bathy); gc()


waal_cro_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),         
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_chl, file = "Data_outputs/waal_cro_glmm_within_chl.RData")
rm(waal_cro_glmm.within_chl); gc()


waal_cro_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),  
  data = pnts_all,
  family = binomial())

save(waal_cro_glmm.within_sst, file = "Data_outputs/waal_cro_glmm_within_sst.RData")
rm(waal_cro_glmm.within_sst); gc()

# Between season #

waal_cro_glmm.between_bathy <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + bathy_scaled | tripID) + (1 | ID) + (0  + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_bathy, file = "Data_outputs/waal_cro_glmm_between_bathy.RData")
rm(waal_cro_glmm.between_bathy); gc()


waal_cro_glmm.between_chl <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + chl_scaled_log | tripID) + (1 | ID) + (0  + chl_scaled_log | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_chl, file = "Data_outputs/waal_cro_glmm_between_chl.RData")
rm(waal_cro_glmm.between_chl); gc()

waal_cro_glmm.between_sst <-  glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + sst_scaled | tripID) + (1 | ID) + (0  + sst_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_sst, file = "Data_outputs/waal_cro_glmm_between_sst.RData")
rm(waal_cro_glmm.between_sst); gc()


```


#### Conditional repeatability

```{r waal_cro_conditional_repeatability}

load("Data_inputs/waal_cro_habitat_data.RData")

env_vars = c("bathy", "chl", "sst")

# [ 1 ] Calculate conditional repeatability ------

## Decide on simulation number
n_sim <- 100

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled_log)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(
  bathy = bathy_scaled,
  chl = chl_scaled,
  sst = sst_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))
  
  ## Get environmental metrics ====
  env_current <- var_map[[env_vars[i]]]
  
  ## Summarise data 
  meanX <- mean(env_current)
  varX <- var(env_current)
  xrange <- range(env_current)
  
  # Between ==== 
  
  ## Estimate conditional repeatability ####
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            LCL_rpt = quantile(res_between$condR$condR, 0.025, na.rm = TRUE),
                            UCL_rpt = quantile(res_between$condR$condR, 0.975, na.rm = TRUE),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Crozet", 
                            level = "between")
  
  rownames(res_between.summary) <- NULL
  
  ## Bootstrap conditional repeatability uncertainty ####
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_between)$cond
  vcov_beta <- vcov(mod_between)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_between)$cond
  var_int <- as.numeric(vc$ID[1,1])         
  var_slope <- as.numeric(vc$ID.1[1,1])     
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID) / df_ID
    var_slope_sim <- var_slope * rchisq(1, df = df_ID) / df_ID
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID
    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.between <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "between")
  
  
  # Within ====
  
  ## Estimate conditional repeatability ####
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            LCL_rpt = quantile(res_within$condR$condR, 0.025, na.rm = TRUE),
                            UCL_rpt = quantile(res_within$condR$condR, 0.975, na.rm = TRUE),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Crozet", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
  
  ## Extract fixed effects and their covariance matrix
  beta_hat <- fixef(mod_within)$cond
  vcov_beta <- vcov(mod_within)
  
  ## Extract variance components (point estimates)
  vc <- VarCorr(mod_within)$cond
  var_int <- as.numeric(vc$ID_season[1,1])         
  var_slope <- as.numeric(vc$ID_season.1[1,1])    
  var_tripID <- as.numeric(vc$tripID[1,1]) 
  var_resid <- (pi^2)/3                     
  
  ## Get covariance matrix for intercept/slope random effects
  Sigma = matrix( c(
    var_int, 0,
    0, var_slope),
    nrow = 2) 
  
  set.seed(123)
  
  condR_sim <- list()
  
  for(n in 1:n_sim){
    
    # 1. Sample fixed effects from multivariate normal
    vcov_mat_numeric <- vcov_beta$cond
    beta_sim <- mvrnorm(1, mu = beta_hat, Sigma = vcov_mat_numeric)
    
    # 2. Sample variance components from scaled chi-square approximations:
    # (This is a rough approximation — bootstrapping is better but slower)
    
    var_int_sim <- var_int * rchisq(1, df = df_ID_season) / df_ID_season
    var_slope_sim <- var_slope * rchisq(1, df = df_ID_season) / df_ID_season
    var_tripID_sim <- var_tripID * rchisq(1, df = df_tripID) / df_tripID

    
    ## Make the covariance matrix
    Sigma_sim = matrix( c( var_int_sim, 0,
                           0, var_slope_sim), nrow = 2)
    
    # 3. Compute conditional repeatability
    condR_df <- condR(
      betaX = beta_sim[habitat_var], 
      meanX = meanX,
      Vx = varX,
      Vr = var_resid,
      Sigma = Sigma_sim, 
      Vo = var_tripID_sim,
      xrange = xrange)$condR
    
    condR_sim[[n]] <- condR_df %>% mutate(sim = n)
    
  }
  
  ## Summarise output
  condR_all <- bind_rows(condR_sim)
  
  condR_uncertainty.within <- condR_all %>%
    group_by(x) %>%
    summarise(
      median = median(condR, na.rm = TRUE),
      lower = quantile(condR, 0.025, na.rm = TRUE),
      upper = quantile(condR, 0.975, na.rm = TRUE)
    ) %>% ungroup() %>%
    mutate(level = "within")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower/100, ymax = upper/100), fill = col_within, alpha = 0.4) +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = col_between, size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = col_within, size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0("waal_cro_", env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("waal_cro_glmm.within_", env_vars[i]))
  rm(list = paste0("waal_cro_glmm.between_", env_vars[i]))
  gc()
  
}

waal_cro_condR_results <- do.call("rbind", cond_res.list)
save(waal_cro_condR_results, file = "Data_outputs/waal_cro_condR_results.RData")

```

```{r waal_cro_plot_slopes}

env_vars = c("bathy", "chl", "sst")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between)
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("waal_cro_", env_vars[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_cro_preference_figures}

# [ 3 ] Final plot ----

waal_cro_pref_plot <- ggarrange(
  waal_cro_bathy_pred_plot + 
    labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
  waal_cro_chl_pred_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  waal_cro_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  waal_cro_bathy_condR_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + ylim(0, 0.1),
  waal_cro_chl_condR_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + ylim(0, 0.1) + theme(axis.title.y = element_blank()),
  waal_cro_sst_condR_plot + 
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + ylim(0, 0.1) + theme(axis.title.y = element_blank()), 
                     ncol = 3, nrow = 2)

waal_cro_pref_plot <- annotate_figure(waal_cro_pref_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_pref_plot, file = "Data_outputs/waal_cro_selection_plot.RData")


png("Figures/waal_cro_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
waal_cro_pref_plot
dev.off()



```

### Heritability

#### Habitat used

```{r waal_cro_heritability_used }

# (1) Load the data ----
load("Data_outputs/waal_cro_habitat_used_ranef.RData")

ranefs.df_wide <- waal_cro_ranefs_used %>%
  select(c(ring, season, trip, estimate, std.error, env)) %>%
  pivot_wider(names_from = env, 
              values_from = c(estimate, std.error),
               names_glue = "{env}_{.value}")

pedigree.df <- read.csv("Data_inputs/waal_cro_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(ranefs.df_wide, pedigree.df)

# (3) Fit univariate models covering responses to each environment ----

ranefs.df_wide %<>% mutate(ring_pe = ring, animal = ring)

priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(student_t(3,0,1), class = "sd", group = "ring_pe"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(student_t(3, 0, 1), class = "sigma"))

# Build univariate models
waal_cro_brm.heritability_used_bathy <- brm(
  bathy_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(waal_cro_brm.heritability_used_bathy, file = "Data_outputs/waal_cro_heritability_used_bathy.RData")

waal_cro_brm.heritability_used_chl <- brm(
  chlA_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(waal_cro_brm.heritability_used_chl, file = "Data_outputs/waal_cro_heritability_used_chl.RData")

waal_cro_brm.heritability_used_sst <- brm(
  SST_estimate ~ 1 + (1 | ring_pe ) + (1 | gr(animal, cov = Amat)) ,
  prior = priors_animal,
  data = ranefs.df_wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99)) 

save(waal_cro_brm.heritability_used_sst, file = "Data_outputs/waal_cro_heritability_used_sst.RData")


# Estimate heritability
h2_bathy <- get_h2(waal_cro_brm.heritability_used_bathy)
h2_chl <- get_h2(waal_cro_brm.heritability_used_chl)
h2_sst <- get_h2(waal_cro_brm.heritability_used_sst)

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))


ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

save(posteriors, file = "Data_outputs/waal_cro_habitat_used_heritability_posteriors.RData")


```

#### Habitat preference


#### Extract individual slopes

```{r waal_cro_extract_slopes}

load("Data_inputs/waal_cro_habitat_data.RData")

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

## Store envronmental variables
env_vars = c("sst", "bathy", "chl")

slopes.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  my_mod = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  
  # Extract slopes - mean and SE
  slopes <- ranef(my_mod)$cond$ID_season
  
  slopes_df <- data.frame(
        ID = rownames(slopes),
        slope = slopes[,2])

  slopes_df %<>% 
    separate(ID, into = c("ring", "season"), remove = FALSE) %>%
    mutate(env = env_vars[i]) %>% 
    relocate(c(ring, season))

  slopes.list[[i]] <- slopes_df
  
}

waal_cro_habitat_slopes <- do.call("rbind", slopes.list)
save(waal_cro_habitat_slopes, file = "Data_outputs/waal_cro_habitat_slopes.RData")

```

#### Estimate heritability

```{r waal_cro_heritability }

# (1) Load the data ----
load("Data_outputs/waal_cro_habitat_slopes.RData")

pedigree.df <- read.csv("Data_inputs/waal_cro_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)

# (2) Make the pedigree ----

Amat <- make_Amat(waal_cro_habitat_slopes, pedigree.df)

## Optional - visualise the pruned peigree (go into function to get pruned pedigree)

# edges <- pedigree.pruned %>%
#   tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
#   dplyr::select(from = value, to = id)
# 
# ped_graph <- graph_from_data_frame(edges, directed = TRUE)
# 
# plot(ped_graph,
#      vertex.size = 3,
#      vertex.label = NA,
#      edge.arrow.size = 0.2,
#      layout = layout_with_sugiyama(ped_graph)$layout)


## Optional - visualise sparse matrix

A_sparse <- as(Amat, "dgCMatrix")
summary(A_sparse)

### Check how many non-zero off-diagonal elements
# nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
# density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
# cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")
# 
# pheatmap(as.matrix(Amat), show_rownames = FALSE, show_colnames = FALSE)


# (3) Fit heritability models ----
waal_cro_habitat_slopes %<>%
  mutate(animal = ring, 
         ring_pe = ring)

## Make dataframe wide (i.e. slope per environment)
waal_cro_slopes.wide <- waal_cro_habitat_slopes %>%
  pivot_wider(names_from = env, 
              values_from = slope,
              names_glue = "{env}_slope")

## Only retain birds with > 1 trip
waal_cro_slopes.wide %<>% group_by(ring) %>% mutate(n_trips = n_distinct(season)) %>% filter(n_trips > 1) %>% ungroup() %>% select(-n_trips)

# Build univariate models and extract h2
waal_cro_brm.heritability_bathy <- brm(
  bathy_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_cro_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_bathy <- (VarCorr(waal_cro_brm.heritability_bathy, summary = FALSE)$animal$sd)^2
v_r_bathy <- (VarCorr(waal_cro_brm.heritability_bathy, summary = FALSE)$residual$sd)^2
h2_bathy <- as.mcmc(v_animal_bathy / (v_animal_bathy + v_r_bathy) )

save(waal_cro_brm.heritability_bathy, file = "Data_outputs/waal_cro_habitat_pref_heritability_bathy.RData")


waal_cro_brm.heritability_chl <- brm(
  chl_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_cro_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_chl <- (VarCorr(waal_cro_brm.heritability_chl, summary = FALSE)$animal$sd)^2
v_r_chl <- (VarCorr(waal_cro_brm.heritability_chl, summary = FALSE)$residual$sd)^2
h2_chl <- as.mcmc(v_animal_chl / (v_animal_chl + v_r_chl) )

save(waal_cro_brm.heritability_chl, file = "Data_outputs/waal_cro_habitat_pref_heritability_chl.RData")


waal_cro_brm.heritability_sst <- brm(
  sst_slope ~ 1 + (1 | ring) + (1 | gr(animal, cov = Amat)),
  prior = priors_animal,
  data = waal_cro_slopes.wide,
  data2 = list(Amat = Amat),
  chains = 4, cores = 4, iter = 8000, warmup = 4000,
  control = list(max_treedepth = 15, adapt_delta = 0.99) )

v_animal_sst <- (VarCorr(waal_cro_brm.heritability_sst, summary = FALSE)$animal$sd)^2
v_r_sst <- (VarCorr(waal_cro_brm.heritability_sst, summary = FALSE)$residual$sd)^2
h2_sst <- as.mcmc(v_animal_sst / (v_animal_sst + v_r_sst) )

save(waal_cro_brm.heritability_sst, file = "Data_outputs/waal_cro_habitat_pref_heritability_sst.RData")


# Extract posteriors 

posteriors <- data.frame(cbind(h2_bathy, h2_chl, h2_sst))
colnames(posteriors) <- c("h2_bathy", "h2_chl", "h2_sst")
posteriors %<>% 
  pivot_longer(cols = starts_with("h2"),
               names_to = "env",
               values_to = "h2") %>%
  mutate(env = recode(env,
      "h2_bathy" = "Bathymetry",
      "h2_chl" = "Chlorophyll",
      "h2_sst" = "SST")) %>%
  arrange(env) %>%
  mutate(env = factor(env, levels = rev(unique(env))))

save(posteriors, file = "Data_outputs/waal_cro_habitat_pref_heritability_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```


