---
title: "Habitat selection analyses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---

# Session info

**R version 4.5.0 (2025-04-11 ucrt)**

**Platform:** x86_64-w64-mingw32/x64

**locale:** *LC_COLLATE=English_United Kingdom.utf8*,
*LC_CTYPE=English_United Kingdom.utf8*, *LC_MONETARY=English_United
Kingdom.utf8*, *LC_NUMERIC=C* and *LC_TIME=English_United Kingdom.utf8*

**attached base packages:** *stats*, *graphics*, *grDevices*, *utils*,
*datasets*, *methods* and *base*

**other attached packages:** *pander(v.0.6.6)*, *nadiv(v.2.18.0)*,
*pheatmap(v.1.0.13)*, *MasterBayes(v.2.58)*, *kinship2(v.1.9.6.2)*,
*quadprog(v.1.5-8)*, *genetics(v.1.3.8.1.3)*, *mvtnorm(v.1.3-3)*,
*gtools(v.3.9.5)*, *gdata(v.3.0.1)*, *combinat(v.0.0-8)*,
*igraph(v.2.2.1)*, *MCMCglmm(v.2.36)*, *ape(v.5.8-1)*,
*coda(v.0.19-4.1)*, *MASS(v.7.3-65)*, *brms(v.2.23.0)*,
*Rcpp(v.1.0.14)*, *glmmTMB(v.1.1.14)*, *ggeffects(v.2.3.2)*,
*pROC(v.1.19.0.1)*, *corrplot(v.0.95)*, *ggridges(v.0.5.7)*,
*kableExtra(v.1.4.0)*, *ggpubr(v.0.6.2)*, *data.table(v.1.18.2.1)*,
*DHARMa(v.0.4.7)*, *lubridate(v.1.9.4)*, *tidyr(v.1.3.1)*,
*lme4(v.1.1-38)*, *Matrix(v.1.7-3)*, *ggplot2(v.4.0.1)*,
*magrittr(v.2.0.3)* and *dplyr(v.1.1.4)*

**loaded via a namespace (and not attached):** *Rdpack(v.2.6.5)*,
*sandwich(v.3.1-1)*, *rlang(v.1.1.7)*, *otel(v.0.2.0)*,
*matrixStats(v.1.5.0)*, *compiler(v.4.5.0)*, *mgcv(v.1.9-4)*,
*loo(v.2.9.0)*, *systemfonts(v.1.3.1)*, *vctrs(v.0.6.5)*,
*stringr(v.1.6.0)*, *pkgconfig(v.2.0.3)*, *fastmap(v.1.2.0)*,
*backports(v.1.5.0)*, *rmarkdown(v.2.30)*, *nloptr(v.2.2.1)*,
*purrr(v.1.0.4)*, *xfun(v.0.52)*, *broom(v.1.0.12)*,
*parallel(v.4.5.0)*, *R6(v.2.6.1)*, *stringi(v.1.8.7)*,
*RColorBrewer(v.1.1-3)*, *car(v.3.1-3)*, *boot(v.1.3-31)*,
*numDeriv(v.2016.8-1.1)*, *estimability(v.1.5.1)*, *knitr(v.1.51)*,
*zoo(v.1.8-15)*, *bayesplot(v.1.15.0)*, *splines(v.4.5.0)*,
*timechange(v.0.3.0)*, *tidyselect(v.1.2.1)*, *rstudioapi(v.0.18.0)*,
*abind(v.1.4-8)*, *yaml(v.2.3.12)*, *TMB(v.1.9.20)*,
*lattice(v.0.22-6)*, *tibble(v.3.3.0)*, *withr(v.3.0.2)*,
*bridgesampling(v.1.2-1)*, *S7(v.0.2.0)*, *posterior(v.1.6.1)*,
*evaluate(v.1.0.5)*, *RcppParallel(v.5.1.10)*, *xml2(v.1.5.2)*,
*pillar(v.1.11.1)*, *carData(v.3.0-6)*, *tensorA(v.0.36.2.1)*,
*checkmate(v.2.3.2)*, *reformulas(v.0.4.4)*, *insight(v.1.4.5)*,
*distributional(v.0.6.0)*, *generics(v.0.1.4)*, *rstantools(v.2.6.0)*,
*scales(v.1.4.0)*, *minqa(v.1.2.8)*, *xtable(v.1.8-4)*, *glue(v.1.8.0)*,
*cubature(v.2.1.4-1)*, *emmeans(v.2.0.1)*, *tools(v.4.5.0)*,
*ggsignif(v.0.6.4)*, *grid(v.4.5.0)*, *rbibutils(v.2.4.1)*,
*nlme(v.3.1-168)*, *Formula(v.1.2-5)*, *cli(v.3.6.5)*,
*textshaping(v.1.0.4)*, *viridisLite(v.0.4.2)*, *svglite(v.2.2.2)*,
*Brobdingnag(v.1.2-9)*, *corpcor(v.1.6.10)*, *gtable(v.0.3.6)*,
*rstatix(v.0.7.3)*, *digest(v.0.6.39)*, *farver(v.2.1.2)*,
*htmltools(v.0.5.9)* and *lifecycle(v.1.0.5)*

# Set-up

```{r setup_packages_functions, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Functions for GPS processing and plotting
source("RPT_functions.R")
source("condR.R")

## MasterBayes installation
# install.packages(c('genetics', 'gtools', 'kinship2'))
# install.packages("https://cran.r-project.org/src/contrib/Archive/MasterBayes/MasterBayes_2.58.tar.gz", repos = NULL, type = "source", dependencies = T)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "tidyr", "lubridate", "DHARMa", "data.table", "ggpubr",  "kableExtra", "ggridges", "corrplot", "pROC", "ggeffects", "glmmTMB", "brms", "MASS", "MCMCglmm", "igraph", "MasterBayes", "pheatmap", "nadiv") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Scale function
scale_2SD <- function(variable) { (variable - mean(variable, na.rm = TRUE)) / sd(variable, na.rm = TRUE) * 2 }

```

```{r setup_parameters}

# Set colours
bathy_col = "#0072B2" 
chl_col = "#009E73" 
sst_col = "#F0E442" 

col_between <- "#91c5f2"
col_within <- "#0b3860"

# Heritability priors
priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(exponential(1), class = "sd", group = "ring_pe"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(exponential(1), class = "sigma") )

# Sim number for bootstrapping CIs 
n_sim = 500

# Environmental variable names
env_vars.used <- c("bathy", "chlA", "SST")
env_vars.pref = c("bathy", "chl", "sst")

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


```

# Black-browed albatrosses

## Bird Island

```{r bbal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/bbal_birdis_sst-bathy-chla_data_population_level.RData")

pnts_all <- process_habitat_data(pnts_all, 6, 50) # ratio used:avail; min recs per individual
### NEED TO TEST THIS ^ ####

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r bbal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

bbal_birdis_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) "

bbal_birdis_vars <- model_habitat_use(bbal_birdis_pnts_all, env_vars.used, "BBAL", "Bird Island", n_sim, my_mod)
bbal_birdis_var_used <- bbal_birdis_vars$variance
bbal_birdis_ranefs_used <- bbal_birdis_vars$ranef

save(bbal_birdis_var_used, file = "Data_outputs/bbal_birdis_habitat_used_variances.RData")
save(bbal_birdis_ranefs_used, file = "Data_outputs/bbal_birdis_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")

bbal_birdis_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" )) 

split_df <- bbal_birdis_ranefs_used  %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
bbal_birdis_used_plot <- annotate_figure(bbal_birdis_used_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_used_plot, file = "Data_outputs/bbal_birdis_used_plot.RData")

png(filename = "Figures/bbal_birdis_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()

```

### Habitat selection

#### Fit the models

```{r bbal_birdis_fit_models, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Fit the models ====

# Within season #

bbal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_bathy, file = "Data_outputs/bbal_birdis_glmm_within_bathy.RData")
rm(bbal_birdis_glmm.within_bathy); gc()


bbal_birdis_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_chl, file = "Data_outputs/bbal_birdis_glmm_within_chl.RData")
rm(bbal_birdis_glmm.within_chl); gc()


bbal_birdis_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_sst, file = "Data_outputs/bbal_birdis_glmm_within_sst.RData")
rm(bbal_birdis_glmm.within_sst); gc()


# Between season #

bbal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID) + (0 + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_bathy, file = "Data_outputs/bbal_birdis_glmm_between_bathy.RData")
rm(bbal_birdis_glmm.between_bathy); gc()


bbal_birdis_glmm.between_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_chl, file = "Data_outputs/bbal_birdis_glmm_between_chl.RData")
rm(bbal_birdis_glmm.between_chl); gc()


bbal_birdis_glmm.between_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_birdis_glmm.between_sst, file = "Data_outputs/bbal_birdis_glmm_between_sst.RData")
rm(bbal_birdis_glmm.between_sst); gc()


```

#### Conditional repeatability

We will calculate metrics of repeatability from our preference models.
We measure both raw variation and derive conditional repeatability
across the environmental covariates.

Raw variation includes: - *slope_var*: Among-individual variance in
plasticity (measured as individual preference for different
environmental levels) - *marginal_R2*: Average repeatability across the
entire environmental range - *R2*: Proportion of total variance
attributed to the random slopes

Conditional repeatability metrics include: - *med_rpt*: Median
repeatability. Since repeatability varies across the 90% range of
chlorophyll, condR calculates it at many points. med_rpt is the average
(median) repeatability across that whole environmental range

```{r bbal_birdis_cond_repeatability}

# [ 1 ] Calculate conditional repeatability ----

load("Data_inputs/bbal_birdis_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = length(env_vars.pref))

for (i in 1:length(env_vars.pref)) {
  
  env_current <- var_map[[env_vars.pref[i]]]
  
  # Load models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_between.summary <- extract_pref_metrics(mod_between, env_vars.pref[i], "between", env_current, "BBAL", "Bird Island")
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "BBAL", "Bird Island")
  
  res.summary <- rbind(res_between.summary, res_within.summary)
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.between <- bootstrap_condR(mod_between, env_vars.pref[i], "between", n_sim, env_current)
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current)

  # Make the conditional repeatability plot
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.between, aes(x = x, ymin = lower, ymax = upper), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty.within, aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_between, size = 1) +
    geom_line(data = condR_uncertainty.within, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("bbal_birdis_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_birdis_glmm.within_", env_vars.pref[i]))
  rm(list = paste0("bbal_birdis_glmm.between_", env_vars.pref[i]))
  gc()
  
}

bbal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(bbal_birdis_condR_results, file = "Data_outputs/bbal_birdis_condR_results.RData")

```

```{r bbal_birdis_plot_slopes}

# [ 2 ] Plot individual slopes ----

rnd_effs.list <- list()

for (i in 1:length(env_vars.pref)) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars.pref[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  

  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("bbal_birdis_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)
save(rnd_effs.out, file = "Data_outputs/bbal_birdis_habitat_selection.RData")

```

```{r bbal_birdis_preference_figures}

# [ 3 ] Final plot ----

bbal_birdis_pref_plot <- ggarrange(
  bbal_birdis_bathy_pred_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + 
    theme(axis.title.x = element_blank()), 
  bbal_birdis_chl_pred_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  bbal_birdis_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
  bbal_birdis_bathy_condR_plot +
    ylim(0, 1) +
    labs(x = "Scaled bathymetry (m below sea level)"),
  bbal_birdis_chl_condR_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    ylim(0, 1) +
    theme(axis.title.y = element_blank()),
  bbal_birdis_sst_condR_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) +
    theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)

bbal_birdis_pref_plot <- annotate_figure(bbal_birdis_pref_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_pref_plot, file = "Data_outputs/bbal_birdis_selection_plot.RData")


png("Figures/bbal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
bbal_birdis_pref_plot
dev.off()


```

### Heritability

#### Habitat used

```{r bbal_birdis_heritability_used }

# Load & prep data ====
load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")
bbal_birdis_pedigree.df <- read.csv("Data_inputs/bbal_birdis_pedigree.csv") %>% select(-Year)

# Fix birds that are recorded as both dams and sires
bbal_birdis_pedigree.df$sire[bbal_birdis_pedigree.df$id == "1271009"] <- "1318880"
bbal_birdis_pedigree.df$dam[bbal_birdis_pedigree.df$id == "1271009"] <- "1318939"

data_prep <- prep_used_heritability_data(bbal_birdis_ranefs_used, bbal_birdis_pedigree.df)
bbal_birdis_used_h2.df <- data_prep$data
bbal_birdis_used_h2.Amat <- data_prep$Amat


# Fit models ====
models_list <- lapply(env_vars.used, function(e) {
  fit_habitat_heritability_model("estimate", e, bbal_birdis_used_h2.df, bbal_birdis_used_h2.Amat, priors_animal, "Data_outputs/bbal_birdis_")
})
names(models_list) <- env_vars.used


# Extract h2 ==== 
bbal_birdis_h2_used.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chlA"   = "Chlorophyll",
      "SST"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) %>%
  rename(h2 = var1)

rownames(bbal_birdis_h2_used.posteriors) <- NULL

save(bbal_birdis_h2_used.posteriors, file = "Data_outputs/bbal_birdis_habitat_used_h2_posteriors.RData")

```

#### Habitat preference

```{r bbal_birdis_heritability_pref }

# Extract slopes ====

load("Data_inputs/bbal_birdis_habitat_data.RData")
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

bbal_birdis_habitat_slopes <- extract_habitat_slopes(env_vars.pref, "bbal_birdis", pnts_meta)
save(bbal_birdis_habitat_slopes, file = "Data_outputs/bbal_birdis_habitat_slopes.RData")


# Estimate heritability ===

## Prep the data
bbal_birdis_pedigree.df <- read.csv("Data_inputs/bbal_birdis_pedigree.csv") %>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
bbal_birdis_pedigree.df$sire[bbal_birdis_pedigree.df$id == "1271009"] <- "1318880"
bbal_birdis_pedigree.df$dam[bbal_birdis_pedigree.df$id == "1271009"] <- "1318939"

data_prep <- prep_slope_heritability_data(bbal_birdis_habitat_slopes, bbal_birdis_pedigree.df)
bbal_birdis_pref_h2.df <- data_prep$data
bbal_birdis_pref_h2.Amat <- data_prep$Amat


# Fit models ====

models_list <- lapply(env_vars.pref, function(e) {
  fit_habitat_heritability_model("slope", e, bbal_birdis_pref_h2.df, bbal_birdis_pref_h2.Amat, priors_animal, "Data_outputs/bbal_birdis_")
})
names(models_list) <- env_vars.pref


# Extract h2 ==== 
bbal_birdis_h2_slope.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chl"   = "Chlorophyll",
      "sst"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) %>%
  rename(h2 = var1)

rownames(bbal_birdis_h2_slope.posteriors) <- NULL

save(bbal_birdis_h2_slope.posteriors, file = "Data_outputs/bbal_birdis_habitat_pref_h2_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```

## Kerguelen

For Kerguelen especially, birds' preference for shelf zones makes the
bathymetry model fall apart. To tackle this, I have calculated distance
from the shelf, and split preferences into on shelf, shelf break, and
off shelf. This seems to resolve my problems. The shelf break is
delimited at 1000m following Hindell et al. 2011.

```{r bbal_ker_process_data, eval = FALSE}

load("Data_inputs/bbal_ker_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )

## Add binomial bathymetry variable
pnts_all %<>%
  mutate(dist_shelf_bin = case_when(
            dist_shelf_break_signed > 20  ~ "on_shelf",       # >20 km inside shelf
            dist_shelf_break_signed >= -20 & dist_shelf_break_signed <= 20 ~ "near_break", # ±20 km of break
            dist_shelf_break_signed < -20  ~ "off_shelf"       # >20 km offshore (deeper)
    ) )

pnts_all$dist_shelf_bin <- factor(pnts_all$dist_shelf_bin, levels = c("on_shelf", "near_break", "off_shelf"))

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_ker_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r bbal_ker_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

bbal_ker_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) "

bbal_ker_vars <- model_habitat_use(bbal_ker_pnts_all, env_vars.used, "BBAL", "Kerguelen", n_sim, my_mod)
bbal_ker_var_used <- bbal_ker_vars$variance
bbal_ker_ranefs_used <- bbal_ker_vars$ranef

save(bbal_ker_var_used, file = "Data_outputs/bbal_ker_habitat_used_variances.RData")
save(bbal_ker_ranefs_used, file = "Data_outputs/bbal_ker_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_ker_plot_habitat_used }

load("Data_outputs/bbal_ker_habitat_used_ranef.RData")

bbal_ker_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" ))

split_df <- bbal_ker_ranefs_used %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_ker_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_ker_used_plot <- annotate_figure(bbal_ker_used_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_used_plot, file = "Data_outputs/bbal_ker_used_plot.RData")

png(filename = "Figures/bbal_ker_habitat_used_ranef_plot.png", width = 10, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

### Habitat selection

#### Fit the models

```{r bbal_ker_fit_model, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Fit the models ====

# Within season #

bbal_ker_glmm.within_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID_season) + (0 + dist_shelf_bin | ID_season),              
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_bathy, file = "Data_outputs/bbal_ker_glmm_within_bathy.RData")
rm(bbal_ker_glmm.within_bathy); gc()


bbal_ker_glmm.within_chl <-  glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),           
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_chl, file = "Data_outputs/bbal_ker_glmm_within_chl.RData")
rm(bbal_ker_glmm.within_chl); gc()


bbal_ker_glmm.within_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),       
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_sst, file = "Data_outputs/bbal_ker_glmm_within_sst.RData")
rm(bbal_ker_glmm.within_sst); gc()


# Between season #

bbal_ker_glmm.between_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID) + (0 + dist_shelf_bin | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_bathy, file = "Data_outputs/bbal_ker_glmm_between_bathy.RData")
rm(bbal_ker_glmm.between_bathy); gc()


bbal_ker_glmm.between_chl <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_chl, file = "Data_outputs/bbal_ker_glmm_between_chl.RData")
rm(bbal_ker_glmm.between_chl); gc()


bbal_ker_glmm.between_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_sst, file = "Data_outputs/bbal_ker_glmm_between_sst.RData")
rm(bbal_ker_glmm.between_sst); gc()


```

#### Conditional repeatability

```{r bbal_ker_cond_repeatability}

load("Data_inputs/bbal_ker_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = 2)

# [ 1A ] Calculate conditional repeatability - Chl & SST ------

for (i in 2:3) { 

  env_current <- var_map[[env_vars.pref[i]]]
  print(env_vars.pref[i])
  
  # Load models
  load(paste0("Data_outputs/bbal_ker_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars.pref[i], ".RData"))
  
  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_between.summary <- extract_pref_metrics(mod_between, env_vars.pref[i], "between", env_current, "BBAL", "Kerguelen")
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "BBAL", "Kerguelen")
  
  res.summary <- rbind(res_between.summary, res_within.summary)
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.between <- bootstrap_condR(mod_between, env_vars.pref[i], "between", n_sim, env_current)
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current)
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  # Make the conditional repeatability plot
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower, ymax = upper), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_between, size = 1) +
    geom_line(data = condR_uncertainty.within, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("bbal_ker_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_ker_glmm.within_", env_vars.pref[i]))
  rm(list = paste0("bbal_ker_glmm.between_", env_vars.pref[i]))
  gc()
}

condR_results <- do.call("rbind", cond_res.list)


# [ 1B ] Calculate conditional repeatability - Bathymetry ------

# Between ====
  
## Calculate repeatability -----
load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
mod_between = get("bbal_ker_glmm.between_bathy")

load("Data_outputs/bbal_ker_glmm_within_bathy.RData")
mod_within = get("bbal_ker_glmm.within_bathy")

R_cond.df_between <- process_categorical_slopes(mod_between, "between", n_sim)
R_cond.df_between$env <- factor(R_cond.df_between$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

R_cond.df_within <- process_categorical_slopes(mod_within, "within", n_sim)
R_cond.df_within$env <- factor(R_cond.df_within$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

bbal_ker_condR_results <- rbind(condR_results, R_cond.df_between, R_cond.df_within)
save(bbal_ker_condR_results, file = "Data_outputs/bbal_ker_condR_results.RData")


# Make the plot ====
bathy_labels <- c("On shelf", "Near shelf break", "Off shelf")

condR_plot <- ggplot() +
    geom_point(data = R_cond.df_between, aes(x= env, y = med_rpt), col = col_between, size = 3) +
    geom_errorbar(data = R_cond.df_between, aes(x = env, ymin = LCL_rpt, ymax = UCL_rpt), width = 0.1, colour = col_between) +
    geom_point(data = R_cond.df_within, aes(x= env, y = med_rpt), col = col_within, size = 3) +
    geom_errorbar(data = R_cond.df_within, aes(x = env, ymin = LCL_rpt, ymax = UCL_rpt), width = 0.1, colour = col_within) +
    labs(y = "Conditional repeatability") +
    scale_x_discrete(labels = bathy_labels) +
    expand_limits(y = c(0, 0.2)) +
    theme_bw() 

assign("bbal_ker_bathy_condR_plot", condR_plot)

```

```{r bbal_ker_plot_slopes}

# [ 2 ] Plot individual slopes ---- 

env_vars.pref = c("chl", "sst", "bathy")

rnd_effs.list <- list()

for (i in 1:2) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars.pref[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring")
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("bbal_ker_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)


## Bathymetry only ====

levels_bathy <- levels(mod_df$dist_shelf_bin)

fixef_vals.between <- fixef(bbal_ker_glmm.between_bathy)$cond
fixef_vals.within <- fixef(bbal_ker_glmm.within_bathy)$cond

# Prepare a data frame to hold predictions
ranef_slopes.between <- ranef(bbal_ker_glmm.between_bathy)$cond$ID
ranef_slopes.within  <- ranef(bbal_ker_glmm.within_bathy)$cond$tripID

pred_df.between <- expand.grid(ID = rownames(ranef_slopes.between),
                       dist_shelf_bin = levels_bathy)
pred_df.within <- expand.grid(ID = rownames(ranef_slopes.within),
                       dist_shelf_bin = levels_bathy)

# Calculate linear predictor per bird & category: - STOPPED HERE
pred_df.between %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.between[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.between[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.between[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.between[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.between["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.between["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.between["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()


pred_df.within %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.within[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.within[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.within[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.within[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.within["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.within["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.within["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()

## Bind together
pred_df.bathy <- rbind(pred_df.between %>% mutate(season_type = "Between seasons"),
                       pred_df.within %>% mutate(season_type = "Within season"))

# Plot lines connecting points per bird
pred_plot <- ggplot() +
  geom_line(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  geom_point(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  facet_grid(~season_type) +
  scale_x_discrete(labels = c("On shelf", "Near shelf", "Off shelf")) +
  labs(x = "Bathymetry category",
       y = "Probability of use") +
  theme_bw()

assign("bbal_ker_bathy_pred_plot", pred_plot)

save(rnd_effs.out, file = "Data_outputs/bbal_ker_habitat_selection.RData")

```

```{r bbal_ker_preference_figures}

# Make the plot ====

bbal_ker_pref_plot <- ggarrange(
  bbal_ker_bathy_pred_plot +
    ylim(0, 1) +
    theme(axis.title.x = element_blank()), 
  bbal_ker_chl_pred_plot +
    labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  bbal_ker_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
  bbal_ker_bathy_condR_plot +
    labs(x = "Bathymetry category") + 
    ylim(0, 1),
  bbal_ker_chl_condR_plot +
    labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank()),
  bbal_ker_sst_condR_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank()), 
                    ncol = 3, nrow = 2)

bbal_ker_pref_plot <- annotate_figure(bbal_ker_pref_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_pref_plot, file = "Data_outputs/bbal_ker_selection_plot.RData")


png("Figures/bbal_ker_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
bbal_ker_pref_plot
dev.off()


```

# Wandering albatrosses

## Bird Island

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )


# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/waal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r waal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

waal_birdis_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID_season) "

waal_birdis_vars <- model_habitat_use(waal_birdis_pnts_all, env_vars.used, "WAAL", "Bird Island", n_sim, my_mod)
waal_birdis_var_used <- waal_birdis_vars$variance
waal_birdis_ranefs_used <- waal_birdis_vars$ranef

save(waal_birdis_var_used, file = "Data_outputs/waal_birdis_habitat_used_variances.RData")
save(waal_birdis_ranefs_used, file = "Data_outputs/waal_birdis_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r waal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/waal_birdis_habitat_used_ranef.RData")

waal_birdis_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" )) 

split_df <- waal_birdis_ranefs_used  %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
waal_birdis_used_plot <- annotate_figure(waal_birdis_used_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_used_plot, file = "Data_outputs/waal_birdis_used_plot.RData")

png(filename = "Figures/waal_birdis_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()

```

### Habitat selection

#### Fit the models

```{r waal_birdis_fit_models, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )


# Fit the models ====

# Within season #

waal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_bathy, file = "Data_outputs/waal_birdis_glmm_within_bathy.RData")
rm(waal_birdis_glmm.within_bathy); gc()


waal_birdis_glmm.within_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_chl, file = "Data_outputs/waal_birdis_glmm_within_chl.RData")
rm(waal_birdis_glmm.within_chl); gc()


waal_birdis_glmm.within_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),      
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_sst, file = "Data_outputs/waal_birdis_glmm_within_sst.RData")
rm(waal_birdis_glmm.within_sst); gc()

```

#### Conditional repeatability

```{r waal_birdis_conditional_repeatability}

# [ 1 ] Calculate conditional repeatability ----

load("Data_inputs/waal_birdis_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = length(env_vars.pref))

for (i in 1:length(env_vars.pref)) {
  
  env_current <- var_map[[env_vars.pref[i]]]
  
  # Load models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "WAAL", "Bird Island")
  
  res.summary <- res_within.summary
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current)

  # Make the conditional repeatability plot
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.within, aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.within, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("waal_birdis_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within)
  rm(list = paste0("waal_birdis_glmm.within_", env_vars.pref[i]))
  gc()
  
}

waal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(waal_birdis_condR_results, file = "Data_outputs/waal_birdis_condR_results.RData")

```

```{r waal_birdis_plot_slopes}

load("Data_inputs/waal_birdis_habitat_data.RData")
env_vars.pref = c("bathy", "chl", "sst")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars.pref)) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  
  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  
  rnd_eff.all <- rnd_eff.within
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID), alpha = 0.2, colour = line_colour) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
   
  assign(paste0("waal_birdis_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_birdis_preference_figures}

# [ 3 ] Final plot ----

waal_birdis_pref_plot <- ggarrange(
  waal_birdis_bathy_pred_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + 
    theme(axis.title.x = element_blank()), 
  waal_birdis_chl_pred_plot + 
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  waal_birdis_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()), 
  waal_birdis_bathy_condR_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + 
    ylim(0, 1),
  waal_birdis_chl_condR_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) +
    ylim(0, 1) + 
    theme(axis.title.y = element_blank()),
  waal_birdis_sst_condR_plot + 
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank()), 
                         ncol = 3, nrow = 2)

waal_birdis_pref_plot <- annotate_figure(waal_birdis_pref_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_pref_plot, file = "Data_outputs/waal_birdis_selection_plot.RData")


png("Figures/waal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
waal_birdis_pref_plot
dev.off()



```

### Heritability

#### Habitat used

```{r waal_birdis_heritability_used }

# Load & prep data ====
load("Data_outputs/waal_birdis_habitat_used_ranef.RData")
waal_birdis_pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv") %>% select(-Year)

data_prep <- prep_used_heritability_data(waal_birdis_ranefs_used, waal_birdis_pedigree.df)
waal_birdis_used_h2.df <- data_prep$data
waal_birdis_used_h2.Amat <- data_prep$Amat


# Fit models ====
models_list <- lapply(env_vars.used, function(e) {
  fit_habitat_heritability_model("estimate", e, waal_birdis_used_h2.df, waal_birdis_used_h2.Amat, priors_animal, "Data_outputs/waal_birdis_")
})
names(models_list) <- env_vars.used


# Extract h2 ==== 
waal_birdis_h2_used.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chlA"   = "Chlorophyll",
      "SST"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_birdis_h2_used.posteriors) <- NULL

save(waal_birdis_h2_used.posteriors, file = "Data_outputs/waal_birdis_habitat_used_h2_posteriors.RData")

```

#### Habitat preference

```{r waal_birdis_heritability_pref }

# Extract slopes ====

load("Data_inputs/waal_birdis_habitat_data.RData")
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

waal_birdis_habitat_slopes <- extract_habitat_slopes(env_vars.pref, "waal_birdis", pnts_meta)
save(waal_birdis_habitat_slopes, file = "Data_outputs/waal_birdis_habitat_slopes.RData")


# Estimate heritability ===

## Prep the data
waal_birdis_pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv") %>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
waal_birdis_pedigree.df$sire[waal_birdis_pedigree.df$id == "1271009"] <- "1318880"
waal_birdis_pedigree.df$dam[waal_birdis_pedigree.df$id == "1271009"] <- "1318939"

data_prep <- prep_slope_heritability_data(waal_birdis_habitat_slopes, waal_birdis_pedigree.df)
waal_birdis_pref_h2.df <- data_prep$data
waal_birdis_pref_h2.Amat <- data_prep$Amat


# Fit models ====

models_list <- lapply(env_vars.pref, function(e) {
  fit_habitat_heritability_model("slope", e, waal_birdis_pref_h2.df, waal_birdis_pref_h2.Amat, priors_animal, "Data_outputs/waal_birdis_")
})
names(models_list) <- env_vars.pref


# Extract h2 ==== 
waal_birdis_h2_slope.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chl"   = "Chlorophyll",
      "sst"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_birdis_h2_slope.posteriors) <- NULL

save(waal_birdis_h2_slope.posteriors, file = "Data_outputs/waal_birdis_habitat_pref_h2_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```

## Crozet

Chlorophyll values are heavily skewed, with extremely high
concentrations occurring near coastlines in the 'available' dataset. To
prevent these large outliers from unduly influencing the model results,
I apply a log-transformation to chlorophyll.

```{r waal_cro_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_cro_sst-bathy-chla_data_population_level.RData")

## Fix some rings
pnts_all %<>%
  mutate(ring = recode(ring,
                       "11376" = "BS11376",
                       "11378" = "BS11378",
                       "25076" = "BS25076",
                       "8196" = "BS28916",
                       "8337" = "BS28337",
                       "BS8337" = "BS28337",
                       "BB21651" = "BS81651",
                       "BS7625" = "BS27625",
                       "BS8402" = "BS28402",
                       "BS8404" = "BS28404",
                       "BS9084" = "BS19084",
                       "BS9444" = "BS19444"))

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(tripID = gsub("_", "", tripID),
         ID_season = paste0(ID, "_", season),
         tripID = paste(ID, season, tripID, sep = "_"))  %>%
  group_by(ID_season) %>%
  ungroup()

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Drop individuals with insufficient data - aim for 50 points per individual
pnts_all %<>%
  group_by(tripID) %>%
  mutate(n_recs = n()) %>%
  filter(n_recs > 50) %>%
  select(-c(n_recs)) %>%
  ungroup()

## Scale the variables 
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
         bathy_scaled2 = bathy_scaled^2,
         chl_scaled_log = scale(log(chlA + 0.01)) )

save(pnts_all, file = "Data_inputs/waal_cro_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r waal_cro_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

waal_cro_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) "

waal_cro_vars <- model_habitat_use(waal_cro_pnts_all, env_vars.used, "WAAL", "Crozet", n_sim, my_mod)
waal_cro_var_used <- waal_cro_vars$variance
waal_cro_ranefs_used <- waal_cro_vars$ranef

save(waal_cro_var_used, file = "Data_outputs/waal_cro_habitat_used_variances.RData")
save(waal_cro_ranefs_used, file = "Data_outputs/waal_cro_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r waal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/waal_cro_habitat_used_ranef.RData")

waal_cro_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" )) 

split_df <- waal_cro_ranefs_used  %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_cro_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
waal_cro_used_plot <- annotate_figure(waal_cro_used_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_used_plot, file = "Data_outputs/waal_cro_used_plot.RData")

png(filename = "Figures/waal_cro_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()

```

### Habitat selection

#### Fit the models

The random slope for chlorophyll-a was retained to account for potential
individual variation in response to environmental conditions. However,
the estimated correlation between the random intercept and slope for
chl_scaled was extremely high (0.99), suggesting that these components
are statistically confounded in the current model. This may reflect
either collinearity in the data or insufficient within-individual
variation in chlorophyll exposure to reliably separate baseline use from
individual responses.

As a result, I have droppped the polynomial terms as I think the models
are unable to cope with this degree of complexity.

```{r waal_cro_fit_models, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

# Add a small constant to avoid log(0) when logging chl
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

# Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Within season #

waal_cro_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),    data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_bathy, file = "Data_outputs/waal_cro_glmm_within_bathy.RData")
rm(waal_cro_glmm.within_bathy); gc()


waal_cro_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),         
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_chl, file = "Data_outputs/waal_cro_glmm_within_chl.RData")
rm(waal_cro_glmm.within_chl); gc()


waal_cro_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),  
  data = pnts_all,
  family = binomial())

save(waal_cro_glmm.within_sst, file = "Data_outputs/waal_cro_glmm_within_sst.RData")
rm(waal_cro_glmm.within_sst); gc()

# Between season #

waal_cro_glmm.between_bathy <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + bathy_scaled | tripID) + (1 | ID) + (0  + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_bathy, file = "Data_outputs/waal_cro_glmm_between_bathy.RData")
rm(waal_cro_glmm.between_bathy); gc()


waal_cro_glmm.between_chl <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + chl_scaled_log | tripID) + (1 | ID) + (0  + chl_scaled_log | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_chl, file = "Data_outputs/waal_cro_glmm_between_chl.RData")
rm(waal_cro_glmm.between_chl); gc()

waal_cro_glmm.between_sst <-  glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + sst_scaled | tripID) + (1 | ID) + (0  + sst_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_sst, file = "Data_outputs/waal_cro_glmm_between_sst.RData")
rm(waal_cro_glmm.between_sst); gc()


```

#### Conditional repeatability

```{r waal_cro_conditional_repeatability}

# [ 1 ] Calculate conditional repeatability ----

load("Data_inputs/waal_cro_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = length(env_vars.pref))

for (i in 1:length(env_vars.pref)) {
  
  env_current <- var_map[[env_vars.pref[i]]]
  
  # Load models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_between.summary <- extract_pref_metrics(mod_between, env_vars.pref[i], "between", env_current, "WAAL", "Crozet")
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "WAAL", "Crozet")
  
  res.summary <- rbind(res_between.summary, res_within.summary)
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.between <- bootstrap_condR(mod_between, env_vars.pref[i], "between", n_sim, env_current)
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current)

  # Make the conditional repeatability plot
  #xrange_90 <- quantile(env_current, probs = c(0.05, 0.95), na.rm = TRUE)
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.between, aes(x = x, ymin = lower, ymax = upper), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty.within, aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_between, size = 1) +
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("waal_cro_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("waal_cro_glmm.within_", env_vars.pref[i]))
  rm(list = paste0("waal_cro_glmm.between_", env_vars.pref[i]))
  gc()
  
}

waal_cro_condR_results <- do.call("rbind", cond_res.list)
save(waal_cro_condR_results, file = "Data_outputs/waal_cro_condR_results.RData")

```

```{r waal_cro_plot_slopes}

env_vars.pref = c("bathy", "chl", "sst")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars.pref)) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between)
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("waal_cro_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_cro_preference_figures}

# [ 3 ] Final plot ----

waal_cro_pref_plot <- ggarrange(
  waal_cro_bathy_pred_plot + 
    labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.x = element_blank()), 
  waal_cro_chl_pred_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  waal_cro_sst_pred_plot +
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), axis.title.x = element_blank()),
  waal_cro_bathy_condR_plot +
    labs(x = "Scaled bathymetry (m below sea level)") + 
    ylim(0, 1),
  waal_cro_chl_condR_plot +
    labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank()),
  waal_cro_sst_condR_plot + 
    labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank()), 
                     ncol = 3, nrow = 2)

waal_cro_pref_plot <- annotate_figure(waal_cro_pref_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_pref_plot, file = "Data_outputs/waal_cro_selection_plot.RData")


png("Figures/waal_cro_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
waal_cro_pref_plot
dev.off()



```

### Heritability

#### Habitat used

```{r waal_cro_heritability_used }

# Load & prep data ====
load("Data_outputs/waal_cro_habitat_used_ranef.RData")
waal_cro_pedigree.df <- read.csv("Data_inputs/waal_cro_pedigree.csv") %>% select(-Year)

data_prep <- prep_used_heritability_data(waal_cro_ranefs_used, waal_cro_pedigree.df)
waal_cro_used_h2.df <- data_prep$data
waal_cro_used_h2.Amat <- data_prep$Amat


# Fit models ====
models_list <- lapply(env_vars.used, function(e) {
  fit_habitat_heritability_model("estimate", e, waal_cro_used_h2.df, waal_cro_used_h2.Amat, priors_animal, "Data_outputs/waal_cro_")
})
names(models_list) <- env_vars.used


# Extract h2 ==== 
waal_cro_h2_used.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chlA"   = "Chlorophyll",
      "SST"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_cro_h2_used.posteriors) <- NULL

save(waal_cro_h2_used.posteriors, file = "Data_outputs/waal_cro_habitat_used_h2_posteriors.RData")

```

#### Habitat preference

```{r waal_cro_heritability_pref }

# Extract slopes ====

load("Data_inputs/waal_cro_habitat_data.RData")
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

waal_cro_habitat_slopes <- extract_habitat_slopes(env_vars.pref, "waal_cro", pnts_meta)
save(waal_cro_habitat_slopes, file = "Data_outputs/waal_cro_habitat_slopes.RData")


# Estimate heritability ===

## Prep the data
waal_cro_pedigree.df <- read.csv("Data_inputs/waal_cro_pedigree.csv") %>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
waal_cro_pedigree.df$sire[waal_cro_pedigree.df$id == "1271009"] <- "1318880"
waal_cro_pedigree.df$dam[waal_cro_pedigree.df$id == "1271009"] <- "1318939"

data_prep <- prep_slope_heritability_data(waal_cro_habitat_slopes, waal_cro_pedigree.df)
waal_cro_pref_h2.df <- data_prep$data
waal_cro_pref_h2.Amat <- data_prep$Amat


# Fit models ====

models_list <- lapply(env_vars.pref, function(e) {
  fit_habitat_heritability_model("slope", e, waal_cro_pref_h2.df, waal_cro_pref_h2.Amat, priors_animal, "Data_outputs/waal_cro_")
})
names(models_list) <- env_vars.pref


# Extract h2 ==== 
waal_cro_h2_slope.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chl"   = "Chlorophyll",
      "sst"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_cro_h2_slope.posteriors) <- NULL

save(waal_cro_h2_slope.posteriors, file = "Data_outputs/waal_cro_habitat_pref_h2_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```
