---
title: "Main results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "ggridges", "gridExtra", "tidyr", "kableExtra", "stringr", "ggpubr") 

source("RPT_functions.R")

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

## Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

## Make sure using dplyr select
select <- dplyr::select

## Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"

col_effort <- "#1b9e77"
col_bearing <- "#aaf0db"
col_slope <- "#0d4e3a"
col_used <- "#3cddac"

```

## Figure S2: Foraging tracks for all populations

```{r prep_GPS_tracks, eval = FALSE, include = FALSE}

# Load all the GPS data ---------------

bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

spec_col <- c("bbal_ker", "bbal_birdis", "waal_cro", "waal_birdis")
all_gps = vector(mode = "list", length = 4)

for (i in 1:length(spec_col)) {
  
  my_gps <- loadRData(paste0("Data_inputs/", spec_col[i], "_gps_labelled.RData"))
  
  # gps_file <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  # my_gps <- loadRData(paste0("Data_inputs/", gps_file)) 
  
  my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))
  
  ## Filter out individuals with only one trip
  my_gps %<>% 
    filter(loc == "trip") %>%
    group_by(ring) %>%
    mutate(n_trips = n_distinct(ID)) %>%
    filter(n_trips > 1) %>%
    select(-n_trips)
  
  ## Add species/colony info in
  my_gps %<>%
    mutate(spec_col = spec_col[i]) %>%
    separate(spec_col, into = c("species", "colony"), sep = "_")

  my_gps$col_lat = ifelse(my_gps$colony == "birdis", bi_coords$latitude, cro_coords$latitude)
  my_gps$col_lon = ifelse(my_gps$colony == "birdis", bi_coords$longitude, cro_coords$longitude)
  
  all_gps[[i]] <- my_gps
  
}

all_gps.df <- do.call("rbind", all_gps)
save(all_gps.df, file = "Data_outputs/all_gps_data.RData")

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# Loop through data
for (i in 1:length(spec_col)) {
  
  # Set projections (use same for Ker + Cro)
   
  if(grepl("birdis", spec_col[i])) {
      proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
  } else {
      proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
    }

  world2 <- sf::st_transform(world, crs = proj.utm)
  
  # Project GPS data
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  my_gps.proj <- all_gps.df %>% filter(species == my_species & colony == my_colony) %>% arrange(ringYr, datetime)
  coordinates(my_gps.proj) <- ~ longitude+latitude
  proj4string(my_gps.proj) <- proj.dec
  
  my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
  my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

  # Project the colony
  col_coords <- data.frame(longitude = my_gps.proj$col_lon[1], latitude = my_gps.proj$col_lat[1])
  
  if(grepl("ker", spec_col[i])) { col_coords <- data.frame(longitude = 70.2333, latitude = -49.6833) }
    
  coordinates(col_coords) <- ~longitude+latitude
  proj4string(col_coords) <- proj.dec
  colony.sp <- spTransform(col_coords, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)

  
  # Build the plot
  
  ## Set min/max limits
  y_min = min(my_gps.proj.df$coords.x2 - 100000)
  y_max = max(my_gps.proj.df$coords.x2 + 100000)
  x_min = min(my_gps.proj.df$coords.x1 - 100000)
  x_max = max(my_gps.proj.df$coords.x1 + 100000)
   
  gps_tracks.plot <- ggplot(data = world2) + 
    geom_sf(fill = "cadetblue", colour = "cadetblue") +
    coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
             label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
    geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
              dat = my_gps.proj.df, group = "identity") +
    scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
    scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
    scale_colour_viridis_d() +
    annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
    labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bbal", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "cro", "Crozet", "Kerguelen"))))) +
    theme(panel.background = element_rect(fill = "white"),
          panel.grid.major = element_line(colour = "grey80"),
          panel.border = element_rect(colour = "black", fill = NA),
          axis.text.x.bottom = element_blank(),
          axis.title.x.bottom = element_blank(),
          axis.text.y.right = element_blank(),
          axis.title.y.right = element_blank(),
          axis.title.y.left = element_blank(),
          axis.text.y.left = element_text(size = 10),
          axis.text.x.top = element_text(size = 10),
          legend.position = "none")

  
  assign(paste0(spec_col[i], "_gps.plot"), gps_tracks.plot)
  plot_name <- paste0(spec_col[i], "_gps.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_gps_plot.RData"))
  
}


```

```{r figS2_gps_tracks }

plot_files <- list.files(path = "Data_outputs/R objects/", pattern = "gps_plot")
plot_paths <- paste0("Data_outputs/R objects/", plot_files)

for (file in plot_paths) {
  load(file)
}

png("Figures/MS Figures/FIGS2_all_gps_tracks.png", width = 10, height = 10, units = "in", res = 700)
ggarrange(bba_kerguelen_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
dev.off()

```


# Foraging effort

```{r effort_summary_stats, echo = TRUE }

effort_rpt.full <- loadRData("Data_outputs/behavioural_repeatability_results.RData")

summaries <- effort_rpt.full %>%
  group_by(species, colony, behaviour) %>%
  summarise(
    r_within = mean(rpt_within, na.rm = TRUE),
    r_across = mean(rpt_across, na.rm = TRUE) )


summaries %>%
  group_by(species) %>%
  pivot_longer(cols = c(r_within, r_across), 
               values_to = "r_all") %>%
  summarise(r_med = median(r_all),
            r_min = min(r_all),
            r_max = max(r_all))

summaries %>%
  group_by(species) %>%
  summarise(
    # within-season
    within_min = min(r_within),
    within_max = max(r_within),
    within_med = median(r_within),
    
    # across-season
    across_min = min(r_across),
    across_max = max(r_across),
    across_med = median(r_across)  )


effort_rpt.full %>%
  group_by(species) %>%
  summarise(med_diff = median(rpt_within) - median(rpt_across))

```

## Figure S1: Repeatability by repeat trips

```{r FigS1_repeatability_by_repeats_figure}

load("Data_outputs/SUPP_behaviour_rpt_test_Ntrips.RData")

rep_testing.all %<>%
  mutate(spec_col = case_when( colony == "kerguelen" ~ "BBAL Kerguelen",
                               colony == "crozet" ~ "WAAL Crozet",
                               colony == "birdis" & species == "bbal" ~ "BBAL Birdisland",
                               TRUE ~ "WAAL Bird Island"),
         behaviour = recode(behaviour,
                "duration.hrs" = "Trip duration",
                "transitions" = "Landings",
                "transit_time.pct" = "Flight time",
                "rest_time.pct" = "Rest time",
                "transit_to_rest.ratio" = "Flight:Rest ratio"),
        behaviour = factor(behaviour),              
        behaviour = forcats::fct_rev(behaviour)) %>%
  filter(level != "annual")

rpt_test.plot <- ggplot(rep_testing.all, aes(x = n_trips, y = rpt, group = level, col = level)) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = c(col_between, col_within),
                   labels = c("Between season", "Within-season")) +
  labs(x = "Minimum number of trips per individual-year", y = "Estimated repeatability") +
  facet_grid(spec_col ~ behaviour) +
  theme_bw() +
  theme(legend.position = "none")


png(filename = "Figures/MS Figures/FIGS1_repeatability_by_sample.png", width = 12, height = 10, units = "in", res = 900)
rpt_test.plot
dev.off()


```


## Figure 2: Repeatability by species & colony

```{r fig2_effort_rpt}

## Recode behaviour
effort_rpt.full %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "landings_per_day" = "Landings per day",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

# Pivot data longer for faceted plotting
effort_rpt.long <- effort_rpt.full %>%
  mutate(rpt_within_total = rpt_within + rpt_across) %>%
  pivot_longer(cols = c(rpt_within_total, rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  filter(repeatability_type != "rpt_within") %>%
  mutate(spec_col = paste(species, colony, sep = " "))


# Get median and 95% credible interval 
effort_rpt.summary <- effort_rpt.long %>%
  group_by(behaviour, repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) )


rptplot <- ggplot(effort_rpt.summary, aes(x = median_val, y = behaviour, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season"),
                      name = NULL) +
  labs(x = "Repeatability (median, 95% CI)", y = "") +
  theme_bw() +
  theme(legend.position = "top")
  

png(filename = "Figures/MS Figures/FIG2_behaviour_repeatability.png", width = 10, height = 10, units = "in", res = 900)
rptplot
dev.off()

```


## Figure S3: Repeatability by species & colony (posteriors)

```{r figs3_effort_rpt}

rptplot_posteriors <- ggplot(effort_rpt.long, aes(x = rpt, y = behaviour, fill = repeatability_type)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  scale_fill_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  facet_wrap(~spec_col, nrow = 2, scales = "free_x") +
  labs(x = "Posterior distribution for repeatability", y = "") +
  theme_bw() +
  theme(legend.position = "none")


png(filename = "Figures/MS Figures/FIGS3_behaviour_repeatability.png", width = 10, height = 10, units = "in", res = 900)
rptplot_posteriors
dev.off()

```

## Table S3: Repeatability parameters

```{r tableS3_effort_rpt}

effort_rpt.full <- loadRData("Data_outputs/behavioural_repeatability_results.RData")

# Reorder for table
effort_rpt.full$behaviour <- factor(effort_rpt.full$behaviour, levels = rev(c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "landings_per_day",
  "duration.hrs")))

effort_rpt.full %<>% arrange(species, behaviour) 

# Summarise output
effort_rpt.table <- effort_rpt.full %>%
  group_by(species, colony, behaviour) %>%
  summarise(var_ringSeason = sprintf("%.2f [%.2f, %.2f]",
              mean(var_ringSeason),
              quantile(var_ringSeason, 0.025),
              quantile(var_ringSeason, 0.975)),
      
            var_ring = sprintf("%.2f [%.2f, %.2f]",
              mean(var_ring),
              quantile(var_ring, 0.025),
              quantile(var_ring, 0.975)),
      
            var_season = sprintf("%.2f [%.2f, %.2f]",
              mean(var_season),
              quantile(var_season, 0.025),
              quantile(var_season, 0.975)),
      
            var_resid = sprintf("%.2f [%.2f, %.2f]",
              mean(var_resid),
              quantile(var_resid, 0.025),
              quantile(var_resid, 0.975)),
      
            rpt_within = sprintf("%.2f [%.2f, %.2f]",
              mean(rpt_within),
              quantile(rpt_within, 0.025),
              quantile(rpt_within, 0.975)),
      
            rpt_across = sprintf("%.2f [%.2f, %.2f]",
              mean(rpt_across),
              quantile(rpt_across, 0.025),
              quantile(rpt_across, 0.975)),
                  
            n_trips = n_trips[1],
            n_birds = n_birds[1]) 

# Make table
kable(effort_rpt.table, format = "html") %>%
  kable_styling("striped", full_width = FALSE) %>%
  column_spec(1, background = "white") %>%  # remove stripe from column 1
  column_spec(2, background = "white") %>%  # remove stripe from column 2
  collapse_rows(columns = c(1,2,10, 11)) %>%
  kable_classic(font_size = 12)

```

# Space use

## Figure 3: Bearing and space use differences

```{r fig3_spatial_rpt, fig.dim=c(6,6)}

# Bearings ====
load("Data_outputs/bearing_repeatability_results.RData")

# Pivot data longer for faceted plotting
bearings_rpt.long <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") 

# Get median and 95% credible interval 
bearings_rpt.summary <- bearings_rpt.long %>%
  group_by(repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) )

## Relevel factors
bearings_rpt.summary %<>%
  mutate(spec_col = recode(spec_col,
      bbal_birdis = "BBAL Bird Island",
      bbal_ker = "BBAL Kerguelen",
      waal_birdis = "WAAL Bird Island",
      waal_cro = "WAAL Crozet"),
      behaviour = "Foraging bearing") %>%
  arrange(spec_col) %>%
  mutate(spec_col = factor(spec_col, levels = rev(unique(spec_col))))

bearings_rpt.plot <- ggplot(bearings_rpt.summary, aes(x = median_val, y = behaviour, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season"),
                      name = NULL) +
  labs(x = "Repeatability (median, 95% CI)", y = "") +
  theme_bw() 
  
# BA comparisons ====

load("Data_outputs/ba_means.RData")
ba_means %<>% filter(!is.na(comp_season)) %>% filter(!is.na(comp_ind))

## Drop within ind, between seasons for WAAL Bird Island
ba_means %<>% filter(!(species == "waal" & colony == "birdis" & comp_ind == "within_ring" & comp_season == "between_season"))

## Data wrangling
ba_plot.df <- ba_means %>%
  mutate(species = case_when(
          species == "bbal" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "ker" ~ "Kerguelen",
      colony == "cro" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = " ") ) %>%
  rename(lower.CL = asymp.LCL,
         upper.CL = asymp.UCL) %>%
  mutate(behaviour = "overlap",
         behav_comp = paste(behaviour, comp_ind, sep = "_"),
         behav_comp = factor(behav_comp,
                             levels = c(
                               "overlap_between_ring",
                               "overlap_within_ring",
                               "bearings_between_ring",
                               "bearings_within_ring"  )))

ba_plot <- ggplot(ba_plot.df, aes(x = emmean, y = comp_ind, colour = comp_season)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = lower.CL, xmax = upper.CL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Bhattacharyya's Affinity") +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(
    values = c(col_within, col_between)) +
  scale_y_discrete(labels = c("Within individual", "Among individuals"),
                      name = NULL) +
  theme_bw()

png(filename = "Figures/MS Figures/FIG3_spatial_results.png", width = 15, height = 7.5, units = "in", res = 900)
ggarrange(bearings_rpt.plot, ba_plot, ncol = 2,
          common.legend = TRUE, legend = "top",
          labels = c("(A)", "(B)"))
dev.off()

```


## Figure S4: Bearing repeatability (posteriors)

```{r figS4_effort_rpt}

bearings_rpt.long %<>%
  mutate(repeatability_type = ifelse(repeatability_type == "rpt_within", "Within-season", "Between-season"),
         spec_col = case_when(
    spec_col == "bbal_birdis" ~ "BBAL Bird Island",
    spec_col == "bbal_ker" ~ "BBAL Kerguelen",
    spec_col == "waal_birdis" ~ "WAAL Bird Island",
    spec_col == "waal_cro"    ~ "WAAL Crozet"))  


rptplot_posteriors <- ggplot(bearings_rpt.long, aes(x = rpt, y = repeatability_type, fill = repeatability_type)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  scale_fill_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  facet_wrap(~spec_col, nrow = 2, scales = "free_x") +
  labs(x = "Posterior distribution for repeatability of foraging bearing", y = "") +
  theme_bw() +
  theme(legend.position = "none")


png(filename = "Figures/MS Figures/FIGS4_bearings_repeatability.png", width = 10, height = 10, units = "in", res = 900)
rptplot_posteriors
dev.off()

```

## Table S4: BA comparison statistics

```{r tableS4_BA_comparison, echo = FALSE}

load("Data_outputs/ba_model_summary.RData")

# Reformat the table
ba_summary %<>%
 mutate(Species = toupper(species),
        Colony = ifelse(colony == "birdis", "Bird Island",
                         ifelse(colony == "cro", "Crozet",
                                "Kerguelen")), 
        Term = case_when(
            Term == "(Intercept)"  ~ "Intercept",
            Term == "comp_seasonbetween_season" ~ "Across seasons",
            Term == "comp_indbetween_ring" ~ "Among individuals",
            Term == "comp_indbetween_ring:comp_seasonbetween_season" ~ "Among individuals * Across seasons",
            TRUE ~ "Inter-observation distance"),
         Estimate = paste0(signif(Estimate, 4), " \u00B1 ", signif(Std_Error, 4)),
         t_value = signif(t_value, 4),
         p_value = ifelse(p_value < 0.0001, "0", format(round(p_value, 4), nsmall = 4, scientific = FALSE))) %>%
  mutate(Term = factor(Term, levels = c(
    "Intercept",
    "Among individuals",
    "Across seasons",
    "Among individuals * Across seasons",
    "Inter-observation distance"))) %>%
  rename('t value' = t_value, 'p value' = p_value) %>%
  select(-c(species, colony, Std_Error)) %>%
  relocate(c(Species, Colony), .before = Term) %>%
  arrange(Species, Colony, Term) %>%
  mutate(CI = paste0("[", signif(CI_lower, 4), ", ", signif(CI_upper, 4), "]")) %>%
  relocate(CI, .before = "t value") %>%
  select(-c(CI_lower, CI_upper))
         
knitr::kable(ba_summary, format = "html", 
             caption = '<span style="font-size: 12px;">Table S4. Model outputs for BA comparisons. Bhattacharyyas Affinity measures overlap in two utilisation distributions.".</span>', digits = 3) %>%
  row_spec(which(ba_summary$`p value` < 0.05 & ba_summary$Term != "Intercept"), bold = TRUE) %>%
  collapse_rows(columns = c(1, 2)) %>%
  kable_styling(font_size = 9)

```


## Figure S5: Example GPS tracks

```{r figs5_example_gps_plots, fig.height=8, fig.width=10}

load("Data_outputs/all_gps_data.RData")

# Common parameters
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
world <- ne_countries(scale = "medium", returnclass = "sf")
spec_col <- c("bbal_ker", "bbal_birdis", "waal_cro", "waal_birdis")

targets <- c("bbal_birdis_1456598", "bbal_birdis_1441460", "bbal_birdis_1320517",
             "bbal_ker_CF27498", "bbal_ker_CF39444", "bbal_ker_CF31429",
             "waal_birdis_5145048", "waal_birdis_5144589", "waal_birdis_5132544",
             "waal_cro_BS6563", "waal_cro_BS18853", "waal_cro_BS21258")

all_gps.df %<>% mutate(plotID = paste(species, colony, ring, sep = "_"))

for (i in 1:length(spec_col)) {
  
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  # Get colony location
  if(my_colony == "birdis") { colony.df = data.frame(Lon = -38.05, Lat = -54) } else {
     colony.df = data.frame(Lon = 51.706972, Lat = -46.358639) }
  
  # Set projections
  proj.utm <- paste0("+proj=laea +lat_0=", colony.df$Lat[1], " +lon_0=", colony.df$Lon[1],
                     " +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
  
  ## Colony projection (for plotting)
  coordinates(colony.df) <- ~Lon+Lat
  proj4string(colony.df) <- proj.dec
  colony.sp <- spTransform(colony.df, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)
  
  ## Get the base world plot
  world2 <- sf::st_transform(world, crs = proj.utm)
  
  ## Subset GPS and project
  targets_speccol <- targets[grepl(spec_col[i], targets)]
  
  track.proj <- all_gps.df %>% filter(plotID %in% targets_speccol)
  coordinates(track.proj) <- ~ longitude+latitude
  proj4string(track.proj) <- proj.dec
  
  track.proj.sp <- spTransform(track.proj, CRS(proj.utm))
  track.proj.df <- as.data.frame(track.proj.sp)

  ## Build the plot
  my_xmin <- min(track.proj.df$coords.x1) - 10000
  my_xmax <- max(track.proj.df$coords.x1) + 10000
  my_ymin <- min(track.proj.df$coords.x2) - 10000
  my_ymax <- max(track.proj.df$coords.x2) + 10000
    

  example_gps.plot <- ggplot(data = world2) + 
      geom_sf(fill = "cadetblue", colour = "grey") +
      coord_sf(crs = proj.utm, xlim = c(my_xmin, my_xmax), ylim = c(my_ymin, my_ymax), 
               label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
      geom_path(aes(x = coords.x1, y = coords.x2, colour = ring), size = 1, dat = track.proj.df) +
      scale_colour_viridis_d(name = NULL) +
      labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bbal", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "cro", "Crozet", "Kerguelen"))))) +
      theme(panel.background = element_rect(fill = "white"), 
            panel.grid.major = element_line(colour = "grey80"),
            panel.border = element_rect(colour = "black", fill = NA),
            axis.text.x.top = element_blank(), 
            axis.title.x.top = element_blank(),
            axis.ticks.x.top = element_blank(),
            axis.title.x.bottom = element_blank(),
            axis.text.y.right = element_blank(), 
            axis.title.y.right = element_blank(),
            axis.title.y.left = element_blank(),
            legend.position = "none", 
            plot.caption = element_text(face = "italic")) 
  
  assign(paste0(spec_col[i], "_example_gps.plot"), example_gps.plot)
  plot_name <- paste0(spec_col[i], "_gps.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_example_gps_plot.RData"))
  
}

png("Figures/MS Figures/FIGS5_spatial_repeatability_example_tracks.png", width = 10, height = 10, units = "in", res = 700)
ggarrange(bba_kerguelen_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
dev.off()

```



# Habitat 

## Used habitat

```{r habitat_used_summary_stats, echo = TRUE }

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

used_variances <- rbind(bbal_birdis_var_used,
                        bbal_ker_var_used,
                        waal_birdis_var_used,
                        waal_cro_var_used)

used_variances %>%
  group_by(Species, Colony) %>%
  summarise(r_within = median(rpt_within, na.rm = TRUE),
            r_across = median(rpt_across, na.rm = TRUE) )

summaries <- used_variances %>%
  group_by(Species, Colony, Environment) %>%
  summarise(r_within = median(rpt_within, na.rm = TRUE),
            r_across = median(rpt_across, na.rm = TRUE, scientific = F) )

summaries %>%
  group_by(Species) %>%
  filter(!(Species == "WAAL" & Colony == "Bird Island")) %>%
  summarise(med_across = median(r_across),
            med_within = median(r_within)) %>%
  mutate(diff = med_within - med_across)

summaries %>%
  group_by(Species, Colony) %>%
  pivot_longer(cols = c(r_within, r_across), 
               values_to = "r_all") %>%
  summarise(r_med = median(r_all, na.rm = TRUE),
            r_min = min(r_all, na.rm = TRUE),
            r_max = max(r_all, na.rm = TRUE))

summaries %>%
  group_by(species) %>%
  filter(!(species == "WAAL" & colony == "Bird Island")) %>%
  summarise(
    # within-season
    within_med = median(r_within, na.rm = TRUE),
    within_min = min(r_within, na.rm = TRUE),
    within_max = max(r_within, na.rm = TRUE),
    
    # across-season
    across_med = median(r_across, na.rm = TRUE),
    across_min = min(r_across, na.rm = TRUE),
    across_max = max(r_across, na.rm = TRUE) )

```

### Figure 4: Repeatability of used habitat

```{r fig4_rpt_used } 

# Wrangle the data into long form
used_variances %<>%
  select(-c(Effect, Variance, `Std. dev`)) %>%
  distinct() %>%
  group_by(Species, Colony) %>%
  pivot_longer(cols = matches("_(within|across)$"),
                   names_to = c(".value", "group"),
                   names_pattern = "(.*)_(within|across)") %>%
  mutate(spec_col = paste(Species, Colony, sep = " "))

# Change factors
used_variances %<>%
 mutate(Environment = recode(Environment,
      "bathy" = "Bathymetry",
      "chlA" = "Chlorophyll",
      "SST" = "SST")) %>%
  arrange(Environment) %>%
  mutate(Environment = factor(Environment, levels = rev(unique(Environment))))

rpt_used.plot <- ggplot(used_variances, aes(x = rpt, y = Environment, colour = group)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = LCL, xmax = UCL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Repeatability (95% CI)") +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(
    values = c(col_between, col_within)) +
  theme_bw() +
  theme(legend.position = "none")


png(file = "Figures/MS Figures/FIG4_habitat_used_repeatability.png", width = 10, height = 10, units = "in", res = 300)
rpt_used.plot
dev.off()


```

### Figure S6: Random intercepts of habitat use

```{r figS6_plot_used}

load("Data_outputs/bbal_birdis_used_plot.RData")
load("Data_outputs/bbal_ker_used_plot.RData")
load("Data_outputs/waal_birdis_used_plot.RData")
load("Data_outputs/waal_cro_used_plot.RData")

png(file = "Figures/MS Figures/FIGS6_habitat_used_intercepts.png", width = 25, height = 17, units = "in", res = 300)
ggarrange(bbal_birdis_used_plot, bbal_ker_used_plot,
          waal_birdis_used_plot, waal_cro_used_plot, 
          ncol = 2, nrow = 2,
          widths = c(1, 0.8))
dev.off()

```


### Table S5: Habitat used variances

```{r tableS5_habitat_used_variance}

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

var.table <- rbind(bbal_birdis_var_used, 
                   bbal_ker_var_used, 
                   waal_birdis_var_used, 
                   waal_cro_var_used) %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorophyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(where(is.numeric), 
                ~ ifelse(.x < 1e-4 & .x > 0, 
                         "0.00", 
                         formatC(.x, digits = 2, format = "fg")))) %>%
  mutate(R_across = paste0(rpt_across, " [", LCL_across, ", ", UCL_across, "]"),
         R_within = paste0(rpt_within, " [", LCL_within, ", ", UCL_within, "]")) %>%
  select(-c(rpt_across, LCL_across, UCL_across, rpt_within, LCL_within, UCL_within)) %>%
  filter(Effect == "ID_season" | Effect == "ID") %>%
  rename(Level = Effect) %>%
  relocate(c(Species, Colony), .before = Environment)

kable(var.table, format = "html", 
  caption = 'Table S5. Variance and repeatability measures associated with random intercepts from the habitat used model.') %>%
  collapse_rows(columns = c(1, 2, 3, 7, 8)) %>%
  kable_classic(font_size = 12)

```




## Habitat preference

### Table S6: Conditional repeatability of preference

```{r tableS6_rpt_preference}

load("Data_outputs/bbal_birdis_condR_results.RData")
load("Data_outputs/bbal_ker_condR_results.RData")
load("Data_outputs/waal_birdis_condR_results.RData")
load("Data_outputs/waal_cro_condR_results.RData")

all_condR_results <- rbind(bbal_birdis_condR_results, 
                           bbal_ker_condR_results, 
                           waal_birdis_condR_results, 
                           waal_cro_condR_results)

all_condR_results %>%
  select(species, colony, env, level, med_rpt) %>%
  pivot_wider(names_from = level, values_from = med_rpt) %>%
  summarise(diff = median(within - between, na.rm = TRUE))

rpt.table <- all_condR_results %>%
  select(-marginal_R2) %>%
  rename(Species = species, Colony = colony, Level = level, Environment = env) %>%
  relocate(Environment, .before = Level) %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chl" ~ "Chlorophyll-a",
    Environment == "sst" ~ "SST",
    Environment == "dist_shelf_binon_shelf" ~ "On shelf",
    Environment == "dist_shelf_binnear_break" ~ "Near shelf break",
    Environment == "dist_shelf_binoff_shelf" ~ "Off shelf"),
    Level = ifelse(Level == "between", "Between", "Within")) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  mutate(`Slope Variance` = paste0(slope_var, " [", LCL_slope_var, ", ", UCL_slope_var, "]"),
         `R_cond` = paste0(med_rpt, " [", LCL_rpt, ", ", UCL_rpt, "]")) %>%
  select(-c(slope_var, LCL_slope_var, UCL_slope_var, med_rpt, LCL_rpt, UCL_rpt)) %>%
  arrange(Species, Colony, Environment, Level)

kable(rpt.table, format = "html") %>% 
  collapse_rows(columns = c(1, 2, 3)) %>%
  kable_classic(font_size = 12)

```


### Figure 5: Preferred habitat

```{r fig5_habitat_pref}

load("Data_outputs/bbal_birdis_selection_plot.RData")
load("Data_outputs/bbal_ker_selection_plot.RData")
load("Data_outputs/waal_birdis_selection_plot.RData")
load("Data_outputs/waal_cro_selection_plot.RData")

png(file = "Figures/MS Figures/FIG5_habitat_preference.png", width = 25, height = 15, units = "in", res = 300)
ggarrange(bbal_birdis_pref_plot, 
          bbal_ker_pref_plot,
          waal_birdis_pref_plot,
          waal_cro_pref_plot, 
          ncol = 2, nrow = 2,
          widths = c(1, 1))
dev.off()

```


# Heritability

## Table S2: Summary stats

```{r tableS2_heritability_summary}

load("Data_outputs/table_s2_data.RData")

table_s2 %<>%
  relocate(population, .before = N) %>%
  mutate(population = case_when(
    population == "bbal_birdis" ~ "BBAL Bird Island",
    population == "waal_birdis" ~ "WAAL Bird Island",
    population == "waal_cro"    ~ "WAAL Crozet"))  

kable(table_s2, format = "html", 
  caption = "Pedigree descriptive statistics for the three study populations, where BBAL = black-browed albatross and WAAL = wandering albatross. $N$ = total individuals, founders = number individuals with no known parents, paternal and maternal percentage give the percentage of individuals with known paternal (sire) and maternal (dam) links respectively. Generation depth represents the traceability of lineages, where mean depth indicates the average number of ancestral generations recorded per individual and maximum depth reflects the longest lineage tracked within each colony monitoring program.") %>%
  kable_classic(font_size = 12)

```


## Figure 6: Plot heritability

```{r fig6_plot_heritability}

load("Data_outputs/heritability_posteriors_all.RData")

# Get median and 95% credible interval 
posteriors.summary <- posteriors_all %>%
  group_by(behaviour, group, spec_col) %>%
  summarise(
    median_val = median(h2, na.rm = TRUE),
    lower_ci = quantile(h2, 0.025, na.rm = TRUE),
    upper_ci = quantile(h2, 0.975, na.rm = TRUE) )

heritability_plot_medians <- ggplot(posteriors.summary, aes(x = median_val, y = behaviour, col = group)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  scale_colour_manual(values = c(
    "effort" = col_effort,
    "bearings" = col_bearing,
    "used" = col_used,
    "slopes" = col_slope)) +
  facet_wrap(~spec_col, nrow = 1, ncol = 3, scales = "free_x") +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw() +
  theme(legend.position = "none")


png(file = "Figures/MS Figures/FIG6_heritability.png", width = 15, height = 9, units = "in", res = 300)
heritability_plot_medians
dev.off()

```

## Figure S7: Heritability posteriors

```{r figS7_heritability_posteriors }

load("Data_outputs/heritability_posteriors_all.RData")

heritability_plot <- ggplot(posteriors_all, aes(x = h2, y = behaviour, fill = group)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  scale_fill_manual(values = c(
    "effort" = col_effort,
    "bearings" = col_bearing,
    "slopes" = col_slope,
    "used" = col_used)) +
  facet_wrap(~spec_col, nrow = 1, ncol = 3, scales = "free_x") +
  labs(x = "Posterior distribution for heritability (h²)", y = "") +
  theme_bw() +
  theme(legend.position = "none")


png(file = "Figures/MS Figures/FIGS7_heritability.png", width = 15, height = 9, units = "in", res = 300)
heritability_plot
dev.off()

```

## Table S7: Posteriors of heritability

```{r tableS7_heritability}

load("Data_outputs/heritability_posteriors_all.RData")

posteriors_all.summary <- posteriors_all %>%
  group_by(spec_col, behaviour) %>%
  summarise(
    median = median(h2),
    lower = quantile(h2, 0.025),
    upper = quantile(h2, 0.975))

posteriors_all.summary %>%
  group_by(spec_col) %>%
  summarise(median_val = median(median),
            min_val = min(median),
            max_val = max(median))

heritability_table <- posteriors_all %>%
  group_by(spec_col, behaviour) %>%
  summarise(
    median = median(h2),
    lower = quantile(h2, 0.025),
    upper = quantile(h2, 0.975)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  mutate("Estimated h2" = paste0(median, " [", lower, ", ", upper, "]")) %>%
  select(c(spec_col, behaviour, 'Estimated h2')) %>%
  arrange(spec_col, behaviour) %>%
  separate(spec_col, into = c("Species", "Colony"), sep = " ", extra = "merge") %>%
  rename(Behaviour = behaviour)
 
heritability_table$Behaviour <- factor(
  heritability_table$Behaviour,
  levels = rev(levels(heritability_table$Behaviour)))

heritability_table %<>% arrange(Species, Colony, Behaviour)

kable(heritability_table, format = "html") %>% 
  collapse_rows(columns = c(1, 2)) %>%
  kable_classic(font_size = 12)

```


# Overall repeatability table

## Table S8: Repeatabilities summary

```{r table_rpt_overall}

# Effort ====

effort_rpt.full <- loadRData("Data_outputs/behavioural_repeatability_results.RData")

## Recode behaviour
effort_rpt.full %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "landings_per_day" = "Landings per day",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

# Pivot data longer for faceted plotting
effort_rpt.summary <- effort_rpt.full %>%
  select(c(species, colony, behaviour, rpt_within, rpt_across)) %>%
  mutate(spec_col = paste(species, colony, sep = " ")) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  group_by(spec_col, repeatability_type, behaviour) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) ) %>%
  mutate(
    formatted = paste0(
      format_numbers(median_val), " [",
      format_numbers(lower_ci), ", ",
      format_numbers(upper_ci), "]")) %>%
  select(spec_col, behaviour, repeatability_type, formatted) %>%
  pivot_wider(
    names_from = repeatability_type,
    values_from = formatted  )


# Spatial ====

load("Data_outputs/bearing_repeatability_results.RData")

bearings_rpt.summary <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  group_by(spec_col, repeatability_type) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) ) %>%
  mutate(
    formatted = paste0(
      format_numbers(median_val), " [",
      format_numbers(lower_ci), ", ",
      format_numbers(upper_ci), "]"),
    behaviour = "Foraging bearing",
    spec_col = recode(spec_col,
    "bbal_birdis" = "BBAL Bird Island",
    "bbal_ker" = "BBAL Kerguelen",
    "waal_birdis" = "WAAL Bird Island",
    "waal_cro" = "WAAL Crozet")) %>%
  select(spec_col, behaviour, repeatability_type, formatted) %>%
  pivot_wider(
    names_from = repeatability_type,
    values_from = formatted)
  
  
# Habitat used ====

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

used_rpt.summary <- rbind(bbal_birdis_var_used, 
                   bbal_ker_var_used, 
                   waal_birdis_var_used, 
                   waal_cro_var_used) %>%
  mutate(behaviour = case_when(
    Environment == "bathy" ~ "Bathymetry used",
    Environment == "chlA" ~ "Chlorophyll-a used",
    TRUE ~ "SST used" ),
    spec_col = paste0(Species, " ", Colony)) %>%
  mutate(rpt_across =  paste0(
        format_numbers(rpt_across), " [",
        format_numbers(LCL_across), ", ",
        format_numbers(UCL_across), "]"),
      rpt_within =  paste0(
        format_numbers(rpt_within), " [",
        format_numbers(LCL_within), ", ",
        format_numbers(UCL_within), "]")) %>%
  select(c(spec_col, behaviour, rpt_across, rpt_within)) %>%
  distinct()

# Habitat preference ====

load("Data_outputs/bbal_birdis_condR_results.RData")
load("Data_outputs/bbal_ker_condR_results.RData")
load("Data_outputs/waal_birdis_condR_results.RData")
load("Data_outputs/waal_cro_condR_results.RData")

preference_rpt.summary <- rbind(bbal_birdis_condR_results, 
                   bbal_ker_condR_results %>% arrange(env), 
                   waal_birdis_condR_results, 
                   waal_cro_condR_results) %>%
  mutate(med_rpt = ifelse(env == "chl" | env == "sst" | env == "bathy", 
                          med_rpt / 100,
                          med_rpt),
         spec_col = paste0(species, " ", colony),
         med_rpt = ifelse(is.na(med_rpt), R2, med_rpt),
         level = tolower(level),
         env = recode(env,
            "bathy" = "Bathymetry preference",
            "chl" = "Chlorophyll-a preference",
            "sst" = "SST preference",
            "dist_shelf_binon_shelf" = "On shelf preference",
            "dist_shelf_binnear_break" = "Near shelf break preference",
            "dist_shelf_binoff_shelf" = "Off shelf preference"),
         level = ifelse(level == "between", "across", "within")) %>%
  mutate(med_rpt =  paste0(
        format_numbers(med_rpt), " [",
        format_numbers(LCL_rpt), ", ",
        format_numbers(UCL_rpt), "]")) %>%
  select(c(spec_col, env, level, med_rpt)) %>%
  pivot_wider(
    names_from = level,
    names_prefix = "rpt_",
    values_from = med_rpt ) %>%
  rename(behaviour = env) %>%
  relocate(spec_col, behaviour, rpt_across, rpt_within)

# Bind all repeatability values

behaviour_order <- c("Flight time", "Flight:Rest ratio", "Landings per day", "Rest time", "Trip duration",
                     "Foraging bearing",
                     "Bathymetry used", "Chlorophyll-a used", "SST used",
                     "Bathymetry preference", "Chlorophyll-a preference", "SST preference", "Near shelf break preference",
                     "On shelf preference", "Off shelf preference")

rpt_all.summary <- rbind(effort_rpt.summary,
                         bearings_rpt.summary,
                         used_rpt.summary,
                         preference_rpt.summary) %>%
  mutate(behaviour = factor(behaviour, levels = behaviour_order)) %>%
  arrange(spec_col, behaviour)

```

```{r tableS7_all_rpt}

kable(rpt_all.summary, format = "html") %>%
  collapse_rows(columns = 1) %>%
  kable_classic(font_size = 12)

```
