---
title: "Main results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

This script contains the main results from the paper, using outputs produced in the other scripts.

Inputs:

-   *species_colony*\_heritability_posteriors.RData

-   all_posteriors_spatial_heritability.RData

-   *species_colony*\_habitat_pref_heritability_posteriors.RData

Outputs:

-   Figure 6, Table 6

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "ggridges", "gridExtra", "tidyr", "kableExtra", "stringr", "ggpubr") 

source("RPT_functions.R")

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

## Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

## Make sure using dplyr select
select <- dplyr::select

## Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"

col_effort <- "#1b9e77"
col_bearing <- "#aaf0db"
col_slope <- "#0d4e3a"
col_used <- "#3cddac"

```

# Foraging effort

## Figure 1: Repeatability by species & colony

```{r fig1_effort_rpt}

effort_rpt.full <- loadRData("Data_outputs/behavioural_repeatability_results.RData")

## Recode behaviour
effort_rpt.full %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

# Pivot data longer for faceted plotting
effort_rpt.long <- rpt_output.full %>%
  mutate(rpt_within_total = rpt_within + rpt_across) %>%
  pivot_longer(cols = c(rpt_within_total, rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  filter(repeatability_type != "rpt_within") %>%
  mutate(spec_col = paste(species, colony, sep = " "))


# Get median and 95% credible interval 
effort_rpt.summary <- effort_rpt.long %>%
  group_by(behaviour, repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) )


rptplot <- ggplot(effort_rpt.summary, aes(x = median_val, y = behaviour, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  labs(x = "Repeatability (median, 95% CI)", y = "") +
  theme_bw() +
  theme(legend.position = "none")
  

png(filename = "Figures/MS Figures/FIG1_behaviour_repeatability.png", width = 10, height = 10, units = "in", res = 900)
rptplot
dev.off()

```


## Figure S1: Repeatability by repeat trips

```{r FigS1_repeatability_by_repeats_figure}

load("Data_outputs/SUPP_behaviour_rpt_test_Ntrips.RData")

rep_testing.all %<>%
  mutate(spec_col = case_when( colony == "kerguelen" ~ "BBAL Kerguelen",
                               colony == "crozet" ~ "WAAL Crozet",
                               colony == "birdis" & species == "bbal" ~ "BBAL Birdisland",
                               TRUE ~ "WAAL Bird Island"),
         behaviour = recode(behaviour,
                "duration.hrs" = "Trip duration",
                "transitions" = "Landings",
                "transit_time.pct" = "Flight time",
                "rest_time.pct" = "Rest time",
                "transit_to_rest.ratio" = "Flight:Rest ratio"),
        behaviour = factor(behaviour),              
        behaviour = forcats::fct_rev(behaviour)) %>%
  filter(level != "annual")

rpt_test.plot <- ggplot(rep_testing.all, aes(x = n_trips, y = rpt, group = level, col = level)) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = c(col_between, col_within),
                   labels = c("Between season", "Within-season")) +
  labs(x = "Minimum number of trips per individual-year", y = "Estimated repeatability") +
  facet_grid(spec_col ~ behaviour) +
  theme_bw() +
  theme(legend.position = "none")


png(filename = "Figures/MS Figures/FIGS1_repeatability_by_sample.png", width = 12, height = 10, units = "in", res = 900)
rpt_test.plot
dev.off()


```


## Table S2: Repeatability parameters

```{r tableS2_effort_rpt}

effort_rpt.full <- loadRData("Data_outputs/behavioural_repeatability_results.RData")

# Reorder for table
effort_rpt.full$behaviour <- factor(effort_rpt.full$behaviour, levels = rev(c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs")))

effort_rpt.full %<>% arrange(species, behaviour) 

# Summarise output
effort_rpt.table <- effort_rpt.full %>%
  group_by(species, colony, behaviour) %>%
  summarise(var_ringSeason_x = sprintf("%.2f [%.2f, %.2f]",
              mean(var_ringSeason_x),
              quantile(var_ringSeason_x, 0.025),
              quantile(var_ringSeason_x, 0.975)),
      
            var_ring = sprintf("%.2f [%.2f, %.2f]",
              mean(var_ring),
              quantile(var_ring, 0.025),
              quantile(var_ring, 0.975)),
      
            var_season = sprintf("%.2f [%.2f, %.2f]",
              mean(var_season),
              quantile(var_season, 0.025),
              quantile(var_season, 0.975)),
      
            var_resid = sprintf("%.2f [%.2f, %.2f]",
              mean(var_resid),
              quantile(var_resid, 0.025),
              quantile(var_resid, 0.975)),
      
            sd_slope = sprintf("%.2f [%.2f, %.2f]",
              mean(sd_slope),
              quantile(sd_slope, 0.025),
              quantile(sd_slope, 0.975)),
      
            rpt_within = sprintf("%.2f [%.2f, %.2f]",
              mean(rpt_within),
              quantile(rpt_within, 0.025),
              quantile(rpt_within, 0.975)),
      
            rpt_across = sprintf("%.2f [%.2f, %.2f]",
              mean(rpt_across),
              quantile(rpt_across, 0.025),
              quantile(rpt_across, 0.975)),
                  
            n_trips = n_trips[1],
            n_birds = n_birds[1]) %>%
  select(-sd_slope)

# Make table
kable(effort_rpt.table, format = "html") %>%
  kable_styling("striped", full_width = FALSE) %>%
  column_spec(1, background = "white") %>%  # remove stripe from column 1
  column_spec(2, background = "white") %>%  # remove stripe from column 2
  collapse_rows(columns = c(1,2,10, 11)) %>%
  kable_classic(font_size = 12)

```

# Space use

## Figure 2: Bearing and space use differences

```{r fig2_spatial_rpt, fig.dim=c(6,6)}

# Bearings ====
load("Data_outputs/bearing_repeatability_results.RData")

# Pivot data longer for faceted plotting
bearings_rpt.long <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") 

# Get median and 95% credible interval 
bearings_rpt.summary <- bearings_rpt.long %>%
  group_by(repeatability_type, spec_col) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) )

## Relevel factors
bearings_rpt.summary %<>%
  mutate(spec_col = recode(spec_col,
      bbal_birdis = "BBAL Bird Island",
      bbal_ker = "BBAL Kerguelen",
      waal_birdis = "WAAL Bird Island",
      waal_cro = "WAAL Crozet")) %>%
  arrange(spec_col) %>%
  mutate(spec_col = factor(spec_col, levels = rev(unique(spec_col))))

bearings_rpt.plot <- ggplot(bearings_rpt.summary, aes(x = median_val, y = spec_col, colour = repeatability_type)) +
  geom_point(position = position_dodge(width = 0.6), size = 3) +
  geom_errorbarh(aes(xmin = lower_ci, xmax = upper_ci),
                 position = position_dodge(width = 0.6), height = 0.3) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  labs(x = "Repeatability in foraging bearing (median, 95% CI)", y = "") +
  theme_bw() +
  theme(legend.position = "none")
  
# BA comparisons ====

load("Data_outputs/ba_means.RData")
ba_means %<>% filter(!is.na(comp_season)) %>% filter(!is.na(comp_ind))

## Drop within ind, between seasons for WAAL Bird Island
ba_means %<>% filter(!(species == "waal" & colony == "birdis" & comp_ind == "within_ring" & comp_season == "between_season"))

## Data wrangling
ba_plot.df <- ba_means %>%
  mutate(species = case_when(
          species == "bbal" ~ "BBAL",
          species == "waal" ~ "WAAL",
          TRUE ~ species),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "ker" ~ "Kerguelen",
      colony == "cro" ~ "Crozet",
      TRUE ~ colony),
    spec_col = paste(species, colony, sep = " ") ) %>%
  rename(lower.CL = asymp.LCL,
         upper.CL = asymp.UCL) %>%
  mutate(behaviour = "overlap",
         behav_comp = paste(behaviour, comp_ind, sep = "_"),
         behav_comp = factor(behav_comp,
                             levels = c(
                               "overlap_between_ring",
                               "overlap_within_ring",
                               "bearings_between_ring",
                               "bearings_within_ring"  )))

ba_plot <- ggplot(ba_plot.df, aes(x = emmean, y = comp_ind, colour = comp_season)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = lower.CL, xmax = upper.CL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Bhattacharyya's Affinity") +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(
    values = c(col_within, col_between)) +
  scale_y_discrete(labels = c("Within individual", "Among individuals")) +
  theme_bw() +
  theme(legend.position = "none")

png(filename = "Figures/MS Figures/FIG2_spatial_results.png", width = 15, height = 7.5, units = "in", res = 900)
ggarrange(bearings_rpt.plot, ba_plot, ncol = 2)
dev.off()

```

## Figure S2: Foraging tracks for all populations

```{r prep_GPS_tracks, eval = FALSE, include = FALSE}

# Load all the GPS data ---------------

bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

spec_col <- c("bbal_ker", "bbal_birdis", "waal_cro", "waal_birdis")
all_gps = vector(mode = "list", length = 4)

for (i in 1:length(spec_col)) {
  
  my_gps <- loadRData(paste0("Data_inputs/", spec_col[i], "_gps_labelled.RData"))
  
  # gps_file <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  # my_gps <- loadRData(paste0("Data_inputs/", gps_file)) 
  
  my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))
  
  ## Filter out individuals with only one trip
  my_gps %<>% 
    filter(loc == "trip") %>%
    group_by(ring) %>%
    mutate(n_trips = n_distinct(ID)) %>%
    filter(n_trips > 1) %>%
    select(-n_trips)
  
  ## Add species/colony info in
  my_gps %<>%
    mutate(spec_col = spec_col[i]) %>%
    separate(spec_col, into = c("species", "colony"), sep = "_")

  my_gps$col_lat = ifelse(my_gps$colony == "birdis", bi_coords$latitude, cro_coords$latitude)
  my_gps$col_lon = ifelse(my_gps$colony == "birdis", bi_coords$longitude, cro_coords$longitude)
  
  all_gps[[i]] <- my_gps
  
}

all_gps.df <- do.call("rbind", all_gps)
save(all_gps.df, file = "Data_outputs/all_gps_data.RData")

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

# Loop through data
for (i in 1:length(spec_col)) {
  
  # Set projections (use same for Ker + Cro)
   
  if(grepl("birdis", spec_col[i])) {
      proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
  } else {
      proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
    }

  world2 <- sf::st_transform(world, crs = proj.utm)
  
  # Project GPS data
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  my_gps.proj <- all_gps.df %>% filter(species == my_species & colony == my_colony) %>% arrange(ringYr, datetime)
  coordinates(my_gps.proj) <- ~ longitude+latitude
  proj4string(my_gps.proj) <- proj.dec
  
  my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
  my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

  # Project the colony
  col_coords <- data.frame(longitude = my_gps.proj$col_lon[1], latitude = my_gps.proj$col_lat[1])
  
  if(grepl("ker", spec_col[i])) { col_coords <- data.frame(longitude = 70.2333, latitude = -49.6833) }
    
  coordinates(col_coords) <- ~longitude+latitude
  proj4string(col_coords) <- proj.dec
  colony.sp <- spTransform(col_coords, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)

  
  # Build the plot
  
  ## Set min/max limits
  y_min = min(my_gps.proj.df$coords.x2 - 100000)
  y_max = max(my_gps.proj.df$coords.x2 + 100000)
  x_min = min(my_gps.proj.df$coords.x1 - 100000)
  x_max = max(my_gps.proj.df$coords.x1 + 100000)
   
  gps_tracks.plot <- ggplot(data = world2) + 
    geom_sf(fill = "cadetblue", colour = "cadetblue") +
    coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
             label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
    geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
              dat = my_gps.proj.df, group = "identity") +
    scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
    scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
    scale_colour_viridis_d() +
    annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
    labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bbal", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "cro", "Crozet", "Kerguelen"))))) +
    theme(panel.background = element_rect(fill = "white"),
          panel.grid.major = element_line(colour = "grey80"),
          panel.border = element_rect(colour = "black", fill = NA),
          axis.text.x.bottom = element_blank(),
          axis.title.x.bottom = element_blank(),
          axis.text.y.right = element_blank(),
          axis.title.y.right = element_blank(),
          axis.title.y.left = element_blank(),
          axis.text.y.left = element_text(size = 10),
          axis.text.x.top = element_text(size = 10),
          legend.position = "none")

  
  assign(paste0(spec_col[i], "_gps.plot"), gps_tracks.plot)
  plot_name <- paste0(spec_col[i], "_gps.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_gps_plot.RData"))
  
}


```

```{r figS2_gps_tracks }

plot_files <- list.files(path = "Data_outputs/R objects/", pattern = "gps_plot")
plot_paths <- paste0("Data_outputs/R objects/", plot_files)

for (file in plot_paths) {
  load(file)
}

png("Figures/MS Figures/FIGS2_all_gps_tracks.png", width = 10, height = 10, units = "in", res = 700)
ggarrange(bba_kerguelen_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
dev.off()

```

## Figure S3: Example GPS tracks

```{r example_gps_plots, fig.height=8, fig.width=10}

load("Data_outputs/all_gps_data.RData")

# Common parameters
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"
world <- ne_countries(scale = "medium", returnclass = "sf")
spec_col <- c("bbal_ker", "bbal_birdis", "waal_cro", "waal_birdis")

targets <- c("bbal_birdis_1456598", "bbal_birdis_1441460", "bbal_birdis_1320517",
             "bbal_ker_CF27498", "bbal_ker_CF39444", "bbal_ker_CF31429",
             "waal_birdis_5145048", "waal_birdis_5144589", "waal_birdis_5132544",
             "waal_cro_BS6563", "waal_cro_BS18853", "waal_cro_BS21258")

all_gps.df %<>% mutate(plotID = paste(species, colony, ring, sep = "_"))

for (i in 1:length(spec_col)) {
  
  my_species = strsplit(spec_col[i], "_")[[1]][1]
  my_colony = strsplit(spec_col[i], "_")[[1]][2]
  
  # Get colony location
  if(my_colony == "birdis") { colony.df = data.frame(Lon = -38.05, Lat = -54) } else {
     colony.df = data.frame(Lon = 51.706972, Lat = -46.358639) }
  
  # Set projections
  proj.utm <- paste0("+proj=laea +lat_0=", colony.df$Lat[1], " +lon_0=", colony.df$Lon[1],
                     " +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0")
  
  ## Colony projection (for plotting)
  coordinates(colony.df) <- ~Lon+Lat
  proj4string(colony.df) <- proj.dec
  colony.sp <- spTransform(colony.df, CRS(proj.utm))
  colony.df <- as.data.frame(colony.sp)
  
  ## Get the base world plot
  world2 <- sf::st_transform(world, crs = proj.utm)
  
  ## Subset GPS and project
  targets_speccol <- targets[grepl(spec_col[i], targets)]
  
  track.proj <- all_gps.df %>% filter(plotID %in% targets_speccol)
  coordinates(track.proj) <- ~ longitude+latitude
  proj4string(track.proj) <- proj.dec
  
  track.proj.sp <- spTransform(track.proj, CRS(proj.utm))
  track.proj.df <- as.data.frame(track.proj.sp)

  ## Build the plot
  my_xmin <- min(track.proj.df$coords.x1) - 10000
  my_xmax <- max(track.proj.df$coords.x1) + 10000
  my_ymin <- min(track.proj.df$coords.x2) - 10000
  my_ymax <- max(track.proj.df$coords.x2) + 10000
    

  example_gps.plot <- ggplot(data = world2) + 
      geom_sf(fill = "cadetblue", colour = "grey") +
      coord_sf(crs = proj.utm, xlim = c(my_xmin, my_xmax), ylim = c(my_ymin, my_ymax), 
               label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
      geom_path(aes(x = coords.x1, y = coords.x2, colour = ring), size = 1, dat = track.proj.df) +
      scale_colour_viridis_d(name = NULL) +
      labs(caption = stringr::str_wrap( paste0(
                          ifelse(my_species == "bbal", "Black-browed albatrosses", "Wandering albatrosses"),
                          " at ", 
                          ifelse(my_colony == "birdis", "Bird Island",
                                 ifelse(my_colony == "cro", "Crozet", "Kerguelen"))))) +
      theme(panel.background = element_rect(fill = "white"), 
            panel.grid.major = element_line(colour = "grey80"),
            panel.border = element_rect(colour = "black", fill = NA),
            axis.text.x.top = element_blank(), 
            axis.title.x.top = element_blank(),
            axis.ticks.x.top = element_blank(),
            axis.title.x.bottom = element_blank(),
            axis.text.y.right = element_blank(), 
            axis.title.y.right = element_blank(),
            axis.title.y.left = element_blank(),
            legend.position = "none", 
            plot.caption = element_text(face = "italic")) 
  
  assign(paste0(spec_col[i], "_example_gps.plot"), example_gps.plot)
  plot_name <- paste0(spec_col[i], "_gps.plot")
  save(list = plot_name, file = paste0("Data_outputs/R objects/", spec_col[i], "_example_gps_plot.RData"))
  
}

png("Figures/MS Figures/FIGS3_spatial_repeatability_example_tracks.png", width = 10, height = 10, units = "in", res = 700)
ggarrange(bba_kerguelen_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')), 
          bba_birdis_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_crozet_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          waal_birdis_example_gps.plot + theme(plot.margin = unit(c(0,0,1,0), 'lines')),
          ncol = 2, nrow = 2,
          align = "h")
dev.off()

```

## Figure S4: Bearings from the colony

```{r figS4_bearing_from_colony}

load("Data_outputs/bearings_data.RData")

bearings.df %<>% mutate(bearing_360 = ifelse(bearing_avg < 0, bearing_avg + 360, bearing_avg),
                        spec_col = case_when(
                            species == "bbal" & colony == "birdis" ~ "Black-browed albatrosses at Bird Island",
                            species == "bbal" & colony == "ker" ~ "Black-browed albatrosses at Kerguelen",
                            species == "waal" & colony == "birdis" ~ "Wandering albatrosses at Bird Island",
                            species == "waal" & colony == "cro" ~ "Wandering albatrosses at Crozet"))
breaks = seq(0, 360, 60)

png(filename = "Figures/MS Figures/FIGS4_bearings_plot.png", width = 16, height = 4.5, units = "in", res = 900)
ggplot(bearings.df, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks, 
                     labels = paste0(breaks, "\u00B0")) +
  labs(y = "Frequency") +
  ggh4x::facet_grid2(. ~ spec_col, scales = "free_y", independent = "y") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))
dev.off()

```


## Table S3: BA comparison statistics

```{r tableS3_BA_comparison, echo = FALSE}

load("Data_outputs/ba_model_summary.RData")

# Reformat the table
ba_summary %<>%
 mutate(Species = toupper(species),
        Colony = ifelse(colony == "birdis", "Bird Island",
                         ifelse(colony == "cro", "Crozet",
                                "Kerguelen")), 
        Term = case_when(
            Term == "(Intercept)"  ~ "Intercept",
            Term == "comp_seasonbetween_season" ~ "Across seasons",
            Term == "comp_indbetween_ring" ~ "Among individuals",
            Term == "comp_indbetween_ring:comp_seasonbetween_season" ~ "Among individuals * Across seasons",
            TRUE ~ "Inter-observation distance"),
         Estimate = paste0(signif(Estimate, 4), " \u00B1 ", signif(Std_Error, 4)),
         t_value = signif(t_value, 4),
         p_value = ifelse(p_value < 0.0001, "0", format(round(p_value, 4), nsmall = 4, scientific = FALSE))) %>%
  mutate(Term = factor(Term, levels = c(
    "Intercept",
    "Among individuals",
    "Across seasons",
    "Among individuals * Across seasons",
    "Inter-observation distance"))) %>%
  rename('t value' = t_value, 'p value' = p_value) %>%
  select(-c(species, colony, Std_Error)) %>%
  relocate(c(Species, Colony), .before = Term) %>%
  arrange(Species, Colony, Term) %>%
  mutate(CI = paste0("[", signif(CI_lower, 4), ", ", signif(CI_upper, 4), "]")) %>%
  relocate(CI, .before = "t value") %>%
  select(-c(CI_lower, CI_upper))
         
knitr::kable(ba_summary, format = "html", 
             caption = '<span style="font-size: 12px;">Table 3. Model outputs for BA comparisons. Bhattacharyyas Affinity measures overlap in two utilisation distributions.".</span>', digits = 3) %>%
  row_spec(which(ba_summary$`p value` < 0.05 & ba_summary$Term != "Intercept"), bold = TRUE) %>%
  collapse_rows(columns = c(1, 2)) %>%
  kable_styling(font_size = 9)


```


# Habitat 

## Used habitat

### Figure 3: Repeatability of used habitat

```{r fig3_rpt_used } 

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

used_variances <- rbind(bbal_birdis_var_used %>% mutate(spec_col = "BBAL Bird Island"),
                        bbal_ker_var_used %>% mutate(spec_col = "BBAL Kerguelen"),
                        waal_birdis_var_used %>% mutate(spec_col = "WAAL Bird Island"),
                        waal_cro_var_used %>% mutate(spec_col = "WAAL Crozet"))

# Wrangle the data into long form
used_variances %<>%
  select(-c(Effect, Variance, `Std. dev`)) %>%
  distinct() %>%
  group_by(spec_col) %>%
  pivot_longer(cols = matches("_(within|across)$"),
                   names_to = c(".value", "group"),
                   names_pattern = "(.*)_(within|across)")
# Change factors
used_variances %<>%
 mutate(Environment = recode(Environment,
      "bathy" = "Bathymetry",
      "chlA" = "Chlorophyll",
      "SST" = "SST")) %>%
  arrange(Environment) %>%
  mutate(Environment = factor(Environment, levels = rev(unique(Environment))))

rpt_used.plot <- ggplot(used_variances, aes(x = rpt, y = Environment, colour = group)) +
  geom_point(position = position_dodge(width = 0.4), size = 3) +
  geom_errorbarh(aes(xmin = LCL, xmax = UCL),
                height = 0.3,
                position = position_dodge(width = 0.4)) +
  labs(y = "", colour = "", x = "Repeatability (95% CI)") +
  facet_wrap(~spec_col, nrow = 2) +
  scale_colour_manual(
    values = c(col_between, col_within)) +
  theme_bw() +
  theme(legend.position = "none")


png(file = "Figures/MS Figures/FIG3_habitat_used_repeatability.png", width = 10, height = 10, units = "in", res = 300)
rpt_used.plot
dev.off()


```


### Figure S5: Random intercepts of habitat use

```{r figS5_plot_used}

load("Data_outputs/bbal_birdis_used_plot.RData")
load("Data_outputs/bbal_ker_used_plot.RData")
load("Data_outputs/waal_birdis_used_plot.RData")
load("Data_outputs/waal_cro_used_plot.RData")

png(file = "Figures/MS Figures/FIGS5_habitat_used_intercepts.png", width = 25, height = 17, units = "in", res = 300)
ggarrange(bbal_birdis_used_plot, bbal_ker_used_plot,
          waal_cro_used_plot, waal_birdis_used_plot,
          ncol = 2, nrow = 2,
          widths = c(1, 0.8))
dev.off()

```

### Table S4: Habitat used variances

```{r tableS4_habitat_used_variance}

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

var.table <- rbind(bbal_birdis_var_used %>% mutate(Species = "BBAL", Colony = "Bird Island"), 
                   bbal_ker_var_used %>% mutate(Species = "BBAL", Colony = "Kerguelen"), 
                   waal_birdis_var_used %>% mutate(Species = "WAAL", Colony = "Bird Island"), 
                   waal_birdis_var_used %>% mutate(Species = "WAAL", Colony = "Crozet")) %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorophyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(where(is.numeric), 
                ~ ifelse(.x < 1e-4 & .x > 0, 
                         "< 0.0001", 
                         formatC(.x, digits = 2, format = "fg")))) %>%
  filter(Effect == "ID_season" | Effect == "ID") %>%
  rename(Level = Effect) %>%
  relocate(c(Species, Colony), .before = Environment)

kable(var.table, format = "html", 
  caption = 'Table 1A. Variance associated with random intercepts from the habitat used model.') %>%
 # row_spec(c(6,11), extra_css = "border-top: 1px solid;") %>%
  collapse_rows(columns = c(1, 2, 3, 7, 8)) %>%
  kable_classic(font_size = 12)

```


## Habitat preference

### Figure 4: Preferred habitat

```{r Fig4_habitat_pref}

load("Data_outputs/bbal_birdis_selection_plot.RData")
load("Data_outputs/bbal_ker_selection_plot.RData")
load("Data_outputs/waal_birdis_selection_plot.RData")
load("Data_outputs/waal_cro_selection_plot.RData")

png(file = "Figures/MS Figures/FIG4_habitat_preference.png", width = 25, height = 15, units = "in", res = 300)
ggarrange(bbal_birdis_pref_plot, 
          bbal_ker_pref_plot,
          waal_birdis_pref_plot,
          waal_cro_pref_plot, 
          ncol = 2, nrow = 2,
          widths = c(1, 1))
dev.off()

```


### Table S5: Conditional repeatability of preference

```{r tableS5_rpt_preference}

load("Data_outputs/bbal_birdis_condR_results.RData")
load("Data_outputs/bbal_ker_condR_results.RData")
load("Data_outputs/waal_birdis_condR_results.RData")
load("Data_outputs/waal_cro_condR_results.RData")

rpt.table <- rbind(bbal_birdis_condR_results, 
                   bbal_ker_condR_results %>% rename(var_between = var_slope) %>% arrange(env), 
                   waal_birdis_condR_results, 
                   waal_cro_condR_results) %>%
  mutate(med_rpt = med_rpt / 100) %>%
  rename(Environment = env, 'Slope Variance' = var_between, "Median cond R" = med_rpt,
         Species = species, Colony = colony, Level = level) %>%
  relocate(Environment, .before = Level) %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chl" ~ "Chlorophyll-a",
    Environment == "sst" ~ "SST",
    Environment == "dist_shelf_binon_shelf" ~ "On shelf",
    Environment == "dist_shelf_binnear_break" ~ "Near shelf break",
    Environment == "dist_shelf_binoff_shelf" ~ "Off shelf"),
    Level = ifelse(Level == "Between", "Between", "Within")) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  select(-marginal_R2) %>%
  arrange(Species, Colony, Environment, Level)

kable(rpt.table, format = "html") %>% 
  collapse_rows(columns = c(1, 2, 3)) %>%
  kable_classic(font_size = 12)

```


# Heritability

## Figure 5: Plot heritability

```{r plot_heritability}

load("Data_outputs/heritability_posteriors_all.RData")

heritability_plot <- ggplot(posteriors_all, aes(x = h2, y = behaviour, fill = group)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  scale_fill_manual(values = c(
    "effort" = col_effort,
    "bearings" = col_bearing,
    "slopes" = col_slope,
    "used" = col_used)) +
  facet_wrap(~spec_col, nrow = 1, ncol = 3, scales = "free_x") +
  labs(x = "Posterior distribution for heritability (h²)", y = "") +
  theme_bw() +
  theme(legend.position = "none")


png(file = "Figures/MS Figures/FIG5_heritability.png", width = 15, height = 9, units = "in", res = 300)
heritability_plot
dev.off()

```

## Table S6: Posteriors of heritability

```{r tableS6_heritability}

load("Data_outputs/heritability_posteriors_all.RData")

heritability_table <- posteriors_all %>%
  group_by(spec_col, behaviour) %>%
  summarise(
    median = median(h2),
    lower = quantile(h2, 0.025),
    upper = quantile(h2, 0.975)) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2))) %>%
  mutate("Estimated h2" = paste0(median, " [", lower, ", ", upper, "]")) %>%
  select(c(spec_col, behaviour, 'Estimated h2')) %>%
  arrange(spec_col, behaviour) %>%
  separate(spec_col, into = c("Species", "Colony"), sep = " ", extra = "merge") %>%
  rename(Behaviour = behaviour)
 
heritability_table$Behaviour <- factor(
  heritability_table$Behaviour,
  levels = rev(levels(heritability_table$Behaviour)))

heritability_table %<>% arrange(Species, Colony, Behaviour)

kable(heritability_table, format = "html") %>% 
  collapse_rows(columns = c(1, 2)) %>%
  kable_classic(font_size = 12)

```


# Overall repeatability table

## Table S7: Repeatabilities summary

```{r table_rpt_overall}

# Effort ====

effort_rpt.full <- loadRData("Data_outputs/behavioural_repeatability_results.RData")

## Recode behaviour
effort_rpt.full %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

# Pivot data longer for faceted plotting
effort_rpt.summary <- effort_rpt.full %>%
  select(c(species, colony, behaviour, rpt_within, rpt_across)) %>%
  mutate(spec_col = paste(species, colony, sep = " ")) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  group_by(spec_col, repeatability_type, behaviour) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) ) %>%
  mutate(
    formatted = paste0(
      format_numbers(median_val), " [",
      format_numbers(lower_ci), ", ",
      format_numbers(upper_ci), "]")) %>%
  select(spec_col, behaviour, repeatability_type, formatted) %>%
  pivot_wider(
    names_from = repeatability_type,
    values_from = formatted  )


# Spatial ====

load("Data_outputs/bearing_repeatability_results.RData")

bearings_rpt.summary <- bearing_posteriors.df %>%
  select(c(spec_col, species, colony, rpt_within, rpt_across)) %>%
  group_by(spec_col) %>%
  pivot_longer(cols = c(rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  group_by(spec_col, repeatability_type) %>%
  summarise(
    median_val = median(rpt, na.rm = TRUE),
    lower_ci = quantile(rpt, 0.025, na.rm = TRUE),
    upper_ci = quantile(rpt, 0.975, na.rm = TRUE) ) %>%
  mutate(
    formatted = paste0(
      format_numbers(median_val), " [",
      format_numbers(lower_ci), ", ",
      format_numbers(upper_ci), "]"),
    behaviour = "Foraging bearing",
    spec_col = recode(spec_col,
    "bbal_birdis" = "BBAL Bird Island",
    "bbal_ker" = "BBAL Kerguelen",
    "waal_birdis" = "WAAL Bird Island",
    "waal_cro" = "WAAL Crozet")) %>%
  select(spec_col, behaviour, repeatability_type, formatted) %>%
  pivot_wider(
    names_from = repeatability_type,
    values_from = formatted)
  
  
# Habitat used ====

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")
load("Data_outputs/bbal_ker_habitat_used_variances.RData")
load("Data_outputs/waal_birdis_habitat_used_variances.RData")
load("Data_outputs/waal_cro_habitat_used_variances.RData")

used_rpt.summary <- rbind(bbal_birdis_var_used %>% mutate(Species = "BBAL", Colony = "Bird Island"), 
                   bbal_ker_var_used %>% mutate(Species = "BBAL", Colony = "Kerguelen"), 
                   waal_birdis_var_used %>% mutate(Species = "WAAL", Colony = "Bird Island"), 
                   waal_cro_var_used %>% mutate(Species = "WAAL", Colony = "Crozet")) %>%
  mutate(behaviour = case_when(
    Environment == "bathy" ~ "Bathymetry used",
    Environment == "chlA" ~ "Chlorophyll-a used",
    TRUE ~ "SST used" ),
    spec_col = paste0(Species, " ", Colony)) %>%
  mutate(rpt_across =  paste0(
        format_numbers(rpt_across), " [",
        format_numbers(LCL_across), ", ",
        format_numbers(UCL_across), "]"),
      rpt_within =  paste0(
        format_numbers(rpt_within), " [",
        format_numbers(LCL_within), ", ",
        format_numbers(UCL_within), "]")) %>%
  select(c(spec_col, behaviour, rpt_across, rpt_within)) %>%
  distinct()

# Habitat preference ====

load("Data_outputs/bbal_birdis_condR_results.RData")
load("Data_outputs/bbal_ker_condR_results.RData")
load("Data_outputs/waal_birdis_condR_results.RData")
load("Data_outputs/waal_cro_condR_results.RData")

preference_rpt.summary <- rbind(bbal_birdis_condR_results, 
                   bbal_ker_condR_results %>% arrange(env), 
                   waal_birdis_condR_results, 
                   waal_cro_condR_results) %>%
  mutate(med_rpt = ifelse(env == "chl" | env == "sst" | env == "bathy", 
                          med_rpt / 100,
                          med_rpt),
         spec_col = paste0(species, " ", colony),
         med_rpt = ifelse(is.na(med_rpt), R2, med_rpt),
         level = tolower(level),
         env = recode(env,
            "bathy" = "Bathymetry preference",
            "chl" = "Chlorophyll-a preference",
            "sst" = "SST preference",
            "dist_shelf_binon_shelf" = "On shelf preference",
            "dist_shelf_binnear_break" = "Near shelf break preference",
            "dist_shelf_binoff_shelf" = "Off shelf preference"),
         level = ifelse(level == "between", "across", "within")) %>%
  mutate(med_rpt =  paste0(
        format_numbers(med_rpt), " [",
        format_numbers(LCL_rpt), ", ",
        format_numbers(UCL_rpt), "]")) %>%
  select(c(spec_col, env, level, med_rpt)) %>%
  pivot_wider(
    names_from = level,
    names_prefix = "rpt_",
    values_from = med_rpt ) %>%
  rename(behaviour = env) %>%
  relocate(spec_col, behaviour, rpt_across, rpt_within)

# Bind all repeatability values

behaviour_order <- c("Flight time", "Flight:Rest ratio", "Landings", "Rest time", "Trip duration",
                     "Foraging bearing",
                     "Bathymetry used", "Chlorophyll-a used", "SST used",
                     "Bathymetry preference", "Chlorophyll-a preference", "SST preference", "Near shelf break preference",
                     "On shelf preference", "Off shelf preference")

rpt_all.summary <- rbind(effort_rpt.summary,
                         bearings_rpt.summary,
                         used_rpt.summary,
                         preference_rpt.summary) %>%
  mutate(behaviour = factor(behaviour, levels = behaviour_order)) %>%
  arrange(spec_col, behaviour)

```

```{r tableS7_all_rpt}

kable(rpt_all.summary, format = "html") %>%
  collapse_rows(columns = 1) %>%
  kable_classic(font_size = 12)

```
