---
title: "Repeatability in foraging effort results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

**Note:** Naming conventions for files:
  - Bird Island = birdis
  - Crozet = cro
  - Kerguelen = ker
  - Chlorophyll = chl
  
Inputs:
  - *species_colony*_individual_trips_summary.RData
  - ind_meta.RData 
  - *species_colony*_pedigree.csv

Outputs:
  - behavioural_repeatability_data.RData
  - Figure 1
  - Table 1
  - Figure SX

## Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4",  "gridExtra", "tidyr", "kableExtra", "stringr", "glmmTMB", "ggridges", "brms", "MCMCglmm", "igraph", "pheatmap", "MasterBayes")

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"

```

```{r process_data}

ind_files <- list.files("Data_inputs/", pattern = "individual_trips")
ind_list <- vector(mode = "list", length = 4)

for (i in 1:length(ind_files)) {
  
  my_data <- loadRData(paste0("Data_inputs/",ind_files[i]))
  ind_list[[i]] <- my_data
  
}

trip_data <- do.call("rbind", ind_list)

# Load metadata ====

load("Data_inputs/ind_meta.RData")

## Bind it in
trip_data <- merge(trip_data, all_ind.meta %>% mutate(species = tolower(species)), by = c("ring", "species"), all.x = T) %>%
  mutate(birthYr = as.numeric(birthYr),
         age = season - birthYr,
         age = ifelse(age < 0, NA, age)) 

# Filter data ====

# Remove any trips where transitions == 0 (not real foraging)
rep_data <- trip_data %>% 
  group_by(ring) %>% 
  mutate(n_seasons = n_distinct(season),
         ring_season = paste(ring, season, sep = "_")) %>%
  filter(n_seasons > 1 & !is.na(age) & transitions > 0) %>% 
  ungroup() %>%
  select(-c(n_seasons, sex)) %>%
  relocate(ring_season, .before = colony)


save(rep_data, file = "Data_inputs/behavioural_repeatability_data.RData")

```


### Sample sizes

We know age effects behaviour, and individuals are recorded at different ages, so this is included in the repeatability models and hence the dataset has been filtered to exclude individuals of unknown age. We don't worry about sex, as each individual can only be one sex. We are also only using individuals with more than one season of data, as otherwise we can't calculate across season repeatability.

```{r sample_sizes}

# Number of individuals with x number of seasons
trips.table <- rep_data %>%
  group_by(colony, species, ring) %>%
  summarise(n_seasons = n_distinct(season)) %>%
  ungroup() %>%
  mutate(rpt_category = case_when(
    n_seasons == 1 ~ "1 seasons",
    n_seasons == 2 ~ "2 seasons",
    n_seasons == 3 ~ "3 seasons",
    n_seasons > 3 ~ "4+ seasons"
  )) %>%
  group_by(colony, species, rpt_category) %>%
  summarise(num_birds = n()) %>%
  arrange(species, colony, rpt_category) %>%
  pivot_wider(
    names_from = rpt_category,       
    values_from = num_birds, 
    values_fill = 0) %>%
  mutate(species = ifelse(species == "bba", "BBAL", "WAAL"),
         colony = ifelse(colony == "kerguelen", "Kerguelen",
                         ifelse(colony == "crozet", "Crozet", "Bird Island"))) %>%
  relocate(species, .before = colony) %>%
  rename(Species = species, Colony = colony)


kable(trips.table, format = "html", 
      caption = 'Table 1. Sample sizes for each species and colony, showing the number of individuals for which we have 2, 3, or 4+ seasons of data, collected using GLS and/or GPS devices.') %>%
  collapse_rows(columns = c(1)) %>%
  row_spec(2, extra_css = "border-bottom: 1px solid;") %>%
  kable_classic(font_size = 12)


```

# Testing for repeatablity

This analysis looks at repeatability in foraging behaviours. I am looking at five behavioural variables:

1.  *duration.hrs* = Duration of the foraging trip in hours
2.  *transitions* = Number of transitions between rest and flight and vice versa
3.  *transit_time.pct* = Percentage of foraging trip spent in flight
4.  *rest_time.pct* = Percentage of foraging trip spent in rest
5.  *transit_to_rest.ratio* = Ratio of flight to rest time; \>1 = more flight than rest, \<1 = more rest than flight

All of the predictor values have positive-skewed distributions with heavy tails. As I'm not interested in predictions, a simple and stable approach to dealing with this is simply to **log-transform** $y$ in each model. The distribution of *rest_time.pct* is a little problematic - the values are right skewed with a very long tail towards high values and lots of values around 0. Possibly a beta model would be a better fit here. But for ease of computation, I'm pursuing the log transformation as the variances don't look worryingly different between this and a beta model.

```{r calculating_repeatability, eval = FALSE}

load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct, transit_to_rest.ratio) == 0) == 0)

# Get time since first trip for each bird/season
rep_data %<>%
  group_by(ring, season) %>%
  mutate(time_since_first = as.numeric(difftime(start_date, min(start_date), units = "days"))) %>%
  ungroup()

# Filter birds where have fewer than 3 repeats within a year
rep_data %<>% 
 group_by(ring_season) %>%
 mutate(n_trips = n_distinct(start_date)) %>%
 filter(n_trips >= 3) %>%
 select(-n_trips) %>%
 ungroup()

# Split data
rep_bbal <- rep_data %>% filter(species == "bba")
rep_waal <- rep_data %>% filter(species == "waal")

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")

rpt_output.list <- vector(mode = "list", length = length(cov))

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ phase + age + (1 | season) + (1 | ring ) + (1 + time_since_first | ring_season)")
  
  # Fit models
  
  ## WAAL ## ----
  print("Processing wanderers...")

  ### Crozet ### ----
  waal_cro.brm_rpt <- brm(
      formula = formula,
      data = rep_waal %>% filter(colony == "crozet"),
      family = gaussian(),  
      cores = 4, chains = 4, iter = 4000,
      control = list(adapt_delta = 0.95))
    
  # Calc repeatability
  posterior <- as_draws_df(waal_cro.brm_rpt)
  x_mean <- mean((rep_waal%>% filter(colony == "crozet"))$time_since_first)
  
  waal_cro.post_df <- data.frame(cbind(calc_rpt_brms(waal_cro.brm_rpt, rep_waal %>% filter(colony == "crozet"), x_mean)),
                             behaviour = cov[i],
                             species = "WAAL",
                             colony = "Crozet")
  
  mod_name <- paste0("waal_cro.brm_rpt_", cov[i])
  assign(mod_name, waal_cro.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/waal_cro_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  ### Bird Island ### ----
   waal_birdis.brm_rpt <- brm(
    formula = formula,
    data = rep_waal %>% filter(colony == "birdis"),
    family = gaussian(),  
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Calc repeatability
  posterior <- as_draws_df(waal_birdis.brm_rpt)
  x_mean <- mean((rep_waal %>% filter(colony == "birdis"))$time_since_first)
  
  waal_birdis.post_df <- data.frame(cbind(calc_rpt_brms(waal_birdis.brm_rpt, rep_waal %>% filter(colony == "birdis"), x_mean)),
                              behaviour = cov[i],
                              species = "WAAL",
                              colony = "Bird Island")
    
  mod_name <- paste0("waal_birdis.brm_rpt", cov[i])
  assign(mod_name, waal_birdis.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/waal_birdis_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  
  ## BBAL ## ----
  print("Processing black-browed...")

  ### Kerguelen ### ----
  bbal_ker.brm_rpt <- brm(
      formula = formula,
      data = rep_bbal %>% filter(colony == "kerguelen"),
      family = gaussian(),  
      cores = 4, chains = 4, iter = 4000,
      control = list(adapt_delta = 0.95))
    
  # Calc repeatability
  posterior <- as_draws_df(bbal_ker.brm_rpt)
  x_mean <- mean((rep_bbal%>% filter(colony == "kerguelen"))$time_since_first)
  
  bbal_ker.post_df <- data.frame(cbind(calc_rpt_brms(bbal_ker.brm_rpt, rep_bbal %>% filter(colony == "kerguelen"), x_mean)),
                             behaviour = cov[i],
                             species = "BBAL",
                             colony = "Kerguelen")
  
  mod_name <- paste0("bbal_ker.brm_rpt_", cov[i])
  assign(mod_name, bbal_ker.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/bbal_ker_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  ### Bird Island ### ----
  bbal_birdis.brm_rpt <- brm(
    formula = formula,
    data = rep_bbal %>% filter(colony == "birdis"),
    family = gaussian(),  
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Calc repeatability
  posterior <- as_draws_df(bbal_birdis.brm_rpt)
  x_mean <- mean((rep_bbal %>% filter(colony == "birdis"))$time_since_first)
  
  bbal_birdis.post_df <- data.frame(cbind(calc_rpt_brms(bbal_birdis.brm_rpt, rep_bbal %>% filter(colony == "birdis"), x_mean)),
                              behaviour = cov[i],
                              species = "BBAL",
                              colony = "Bird Island")
    
  mod_name <- paste0("bbal_birdis.brm_rpt", cov[i])
  assign(mod_name, bbal_birdis.brm_rpt)
  
  save(list = mod_name, file = paste0("Data_outputs/bbal_birdis_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  # Output summary
  rpt_output.list[[i]] <- rbind(waal_cro.post_df, waal_birdis.post_df,
                                bbal_ker.post_df, bbal_birdis.post_df)
   
}

rpt_output.full <- do.call("rbind", rpt_output.list)
rpt_output.full %<>% arrange(species)

# Get sample sizes and add them in
rep_samples <- rep_data %>% 
  group_by(colony, species) %>% 
  summarise(n_trips = n_distinct(start_date), n_birds = n_distinct(ring)) %>% 
  mutate(species = ifelse(species == "bba", "BBAL", "WAAL"),
         colony = case_when(colony == "birdis" ~ "Bird Island",
                            colony == "crozet" ~ "Crozet",
                            colony == "kerguelen" ~ "Kerguelen"))

rpt_output.full %<>% left_join(rep_samples, by = c("colony", "species"))

save(rpt_output.full, file = "Data_outputs/behavioural_repeatability_results.RData")

# Model checking 
# plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
# pp_check(bbal_birdis.brm_rpt)
# 
# pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

```


### Repeatability results

#### Table 1: Repeatability parameters

```{r repeatability_table}

load("Data_outputs/behavioural_repeatability_results.RData")

# Reorder for table
rpt_output.full$behaviour <- factor(rpt_output.full$behaviour, levels = rev(c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs")))

rpt_output.full %<>% arrange(species, behaviour) 

# Summarise output
rpt_summary <- rpt_output.full %>%
  group_by(species, colony, behaviour) %>%
  summarise(var_ringSeason_x = sprintf("%.2f [%.2f, %.2f]",
              mean(var_ringSeason_x),
              quantile(var_ringSeason_x, 0.025),
              quantile(var_ringSeason_x, 0.975)),
      
            var_ring = sprintf("%.2f [%.2f, %.2f]",
              mean(var_ring),
              quantile(var_ring, 0.025),
              quantile(var_ring, 0.975)),
      
            var_season = sprintf("%.2f [%.2f, %.2f]",
              mean(var_season),
              quantile(var_season, 0.025),
              quantile(var_season, 0.975)),
      
            var_resid = sprintf("%.2f [%.2f, %.2f]",
              mean(var_resid),
              quantile(var_resid, 0.025),
              quantile(var_resid, 0.975)),
      
            sd_slope = sprintf("%.2f [%.2f, %.2f]",
              mean(sd_slope),
              quantile(sd_slope, 0.025),
              quantile(sd_slope, 0.975)),
      
            rpt_within = sprintf("%.2f [%.2f, %.2f]",
              mean(rpt_within),
              quantile(rpt_within, 0.025),
              quantile(rpt_within, 0.975)),
      
            rpt_across = sprintf("%.2f [%.2f, %.2f]",
              mean(rpt_across),
              quantile(rpt_across, 0.025),
              quantile(rpt_across, 0.975)),
                  
            n_trips = n_trips[1],
            n_birds = n_birds[1]) %>%
  select(-sd_slope)

# Make table
kable(rpt_summary, format = "html", 
      caption = 'Table 2. Repeatability parameters for each species. Sample sizes are for trips, individuals.') %>%
  kable_styling("striped", full_width = FALSE) %>%
  column_spec(1, background = "white") %>%  # remove stripe from column 1
  column_spec(2, background = "white") %>%  # remove stripe from column 2
  collapse_rows(columns = c(1,2,10, 11)) %>%
  kable_classic(font_size = 12)


```

#### Figure 1: Repeatability by species & colony

```{r repeatability_figure, fig.cap = "Figure 1. Summary of repeatability values for each species.", fig.topcaption = TRUE}

## Recode behaviour
rpt_output.full %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

# Pivot data longer for faceted plotting
rpt_long <- rpt_output.full %>%
  mutate(rpt_within_total = rpt_within + rpt_across) %>%
  pivot_longer(cols = c(rpt_within_total, rpt_within, rpt_across),
               names_to = "repeatability_type",
               values_to = "rpt") %>%
  filter(repeatability_type != "rpt_within") %>%
  mutate(spec_col = paste(species, colony, sep = " "))


rptplot <- ggplot(rpt_long, aes(x = rpt, y = behaviour, fill = repeatability_type)) +
  geom_density_ridges(alpha = 0.6, scale = 1) +
  facet_wrap(~spec_col, nrow = 2, scales = "free_x") +
   scale_fill_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  labs(x = "Posterior distribution for repeatability", y = "") +
  theme_bw() +
  theme(legend.position = "none")


png(filename = "Figures/MS Figures/FIG1_behaviour_repeatability.png", width = 10, height = 10, units = "in", res = 900)
rptplot
dev.off()



######

# # Make dataframe long
# rpt_long <- rpt_output.full %>%
#   select(species, parameter,
#          within_R, between_R,
#          within_ci, between_ci) %>%
#   pivot_longer(
#     cols = starts_with(c("within_", "between_")),
#     names_to = c("type", ".value"),
#     names_pattern = "(.*)_(.*)"
#   ) %>%
#   separate(ci, into = c("ci_lower", "ci_upper"), sep = ", ", convert = TRUE) %>%
#   mutate(
#     ci_lower = as.numeric(str_remove(ci_lower, "\\[")),
#     ci_upper = as.numeric(str_remove(ci_upper, "\\]")),
#     R = as.numeric(str_extract(R, "\\d+\\.\\d+")),  # Extract point estimate only
#     type = factor(type, levels = c("between", "within"))  
#   )
# 
# ## Reorder for plotting
# rpt_long$parameter <- factor(rpt_long$parameter, levels = c(
#   "transit_to_rest.ratio",
#   "transit_time.pct",
#   "rest_time.pct",
#   "transitions",
#   "duration.hrs"
# ))
# 
# 
# # Set colours
# col_within <- "#91c5f2"
# col_between <- "#0b3860"
# 
# # Build plot
# rptplot <- 
#   ggplot(rpt_long, aes(x = R, y = parameter, color = type)) +
#   annotate("rect", xmin = -Inf, xmax = 0.1, ymin = -Inf, ymax = Inf,
#            fill = "grey80", alpha = 0.5) +
#   annotate("rect", xmin = 0.1, xmax = 0.3, ymin = -Inf, ymax = Inf,
#            fill = "grey80", alpha = 0.2) +
#   scale_colour_manual(values = c(col_between, col_within),
#                       labels = c("Between-season", "Within-season")) +
#   scale_y_discrete(labels = c(
#     "duration.hrs" = "Trip duration (hrs)",
#     "transitions" = "Number of landings",
#     "transit_time.pct" = "Flight time (% trip)",
#     "rest_time.pct" = "Rest time (% trip)",
#     "transit_to_rest.ratio" = "Ratio flight to rest (%)"
#   )) +
#   geom_point(position = position_dodge(width = 0.5)) +
#   geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper),
#                  position = position_dodge(width = 0.5), height = 0.2) +
#   facet_wrap(~ species) +
#   labs(x = "Repeatability (\u00B1 95% CI)") +
#   theme_bw() +
#   theme(legend.title = element_blank(),
#         legend.position = "top",
#         axis.title.y = element_blank())

```

## Testing for heritability

### * Black-browed albatrosses

#### * Bird Island

```{r bbal_birdis_behaviour_heritability }

# (1) Load and subsample the data  ----
load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct, transit_to_rest.ratio) == 0) == 0)
rep_data %<>% filter(species == "bba" & colony == "birdis")

pedigree.df <- read.csv("Data_inputs/bbal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)

# One sire encoded as '0'
pedigree.df$sire[pedigree.df$sire == "0"] <- NA

# (2) Check the pedigree  ----

## Are the ID in rep_data the same as id in pedigree.df?
pedigree.df$id <- as.character(pedigree.df$id)
rep_data$ring <- as.character(rep_data$ring)

model_ids <- unique(rep_data$ring)
pedigree_ids <- pedigree.df$id

### IDs in model AND pedigree
ids_in_both <- intersect(model_ids, pedigree_ids)

### IDs in model NOT in pedigree
ids_missing_from_pedigree <- setdiff(model_ids, pedigree_ids)


# (3) Subset the pedigree to include all ancestors  ----

## Check formatting
pedigree.df$id <- as.character(pedigree.df$id)
pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)
rep_data$ring <- as.character(rep_data$ring)

model_ids <- unique(rep_data$ring)
ped_ids_needed <- get_ancestors(model_ids, pedigree.df)
length(ped_ids_needed)

## Subset the pedigree
pedigree.pruned <- pedigree.df[pedigree.df$id %in% ped_ids_needed, ]

## Check for missing IDs - NONE MISSING
setdiff(model_ids, pedigree.pruned$id) 

# Fix birds that are recorded as both dams and sires
intersect(pedigree.pruned$sire[!is.na(pedigree.pruned$sire)], pedigree.pruned$dam[!is.na(pedigree.pruned$dam)])

na.omit(pedigree.pruned[pedigree.pruned$sire == "1318880" | pedigree.pruned$dam == "1318880", ])
pedigree.pruned$sire[pedigree.pruned$id == "1271009"] <- "1318880"
pedigree.pruned$dam[pedigree.pruned$id == "1271009"] <- "1318939"

## Chose randomly for this one
na.omit(pedigree.pruned[pedigree.pruned$sire == "1146056" | pedigree.pruned$dam == "1146056", ])
pedigree.pruned$sire[pedigree.pruned$id == "1434383"] <- "1139670"
pedigree.pruned$dam[pedigree.pruned$id == "1434383"] <- "1146056"


# (4) Visualise the pruned pedigree  ----

edges <- pedigree.pruned %>%
  tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
  dplyr::select(from = value, to = id)

## Build igraph object
ped_graph <- graph_from_data_frame(edges, directed = TRUE)

plot(ped_graph,
     vertex.size = 3,
     vertex.label = NA,
     edge.arrow.size = 0.2,
     layout = layout_with_sugiyama(ped_graph)$layout)


# (5) Look at the genetic matrix A  ----

ped <- pedigree.pruned
names(ped) <- c("animal", "dam", "sire")

# Deal with dams coming before offspring
ped <- orderPed(ped)

Ainv <- inverseA(ped)$Ainv
A <- as.matrix(solve(Ainv))

## Inspect sparse matrix
A_sparse <- as(A, "dgCMatrix")
summary(A_sparse)

### Check how many non-zero off-diagonal elements
nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")

pheatmap(as.matrix(A), show_rownames = FALSE, show_colnames = FALSE)


# (6) Fit the models ----

cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
rep_data$animal <- rep_data$ring

rep_data$animal_shuffled <- sample(rep_data$animal)

posteriors_list <- list()

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ 1 + phase + (1 | season) + (1 | ring) + (1 | ring_season) + (1 | gr(animal, cov = A))")
  
  mod_brm <- brm(
    formula = formula,
    data = rep_data,
    family = gaussian(),  
    data2 = list(A = A),  
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Get posteriors 
  posterior <- as_draws_df(mod_brm)
 
  sd_animal <- posterior$sd_animal__Intercept
  var_animal  <- posterior$sd_animal__Intercept^2
  var_season  <- posterior$sd_season__Intercept^2
  var_resid   <- posterior$sigma^2
  h2_post <- var_animal / (var_animal + var_season + var_resid)
  
  posterior.df <- data.frame(cbind(sd_animal, var_animal, var_season, var_resid, h2_post),
                             behaviour = cov[i])
  
  posteriors_list[[i]] <- posterior.df
 
  # Output model
  mod_name <- paste0("mod_brm_", cov[i])
  assign(mod_name, mod_brm)
  
  save(list = mod_name, file = paste0("Data_outputs/bbal_birdis_brms_behaviour_heritability_", cov[i], ".RData"))
    
}

posteriors.all <- do.call("rbind", posteriors_list)
save(posteriors.all, file = "Data_outputs/bbal_birdis_behaviour_heritability_posteriors.RData")


# (7) Visualise results  ----

# Model checking
plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
pp_check(mod_brm)

pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

# Heritability 
posteriors.all %<>%
  group_by(behaviour) %>%
  mutate(mean_h2 = mean(h2_post, na.rm = TRUE))

posteriors.all$behaviour <- reorder(posteriors.all$behaviour, posteriors.all$mean_h2)

## Recode behaviour
posteriors.all %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"))

ggplot(posteriors.all, aes(x = h2_post, y = behaviour)) +
  stat_density_ridges(fill = "skyblue",  colour = "skyblue4", linetype = "dashed", alpha = 0.7, quantile_lines = TRUE, quantiles = 2, vline_colour = "magenta4", vline_size = 1, vline_type = "dotted") +
  theme_bw() +
  labs(x = "Posterior heritability (h²)", y = "",
       title = "Posterior destributions of heritability for BBAL at Bird Island") 


```


### ** Wandering albatrosses

#### * Crozet

```{r waal_cro_behaviour_heritability }

# (1) Load and subsample the data  ----
load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct, transit_to_rest.ratio) == 0) == 0)
rep_data %<>% filter(species == "waal" & colony == "crozet")

pedigree.df <- read.csv("Data_inputs/waal_cro_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)


# (2) Check the pedigree  ----

## Are the ID in rep_data the same as id in pedigree.df?
pedigree.df$id <- as.character(pedigree.df$id)
rep_data$ring <- as.character(rep_data$ring)

model_ids <- unique(rep_data$ring)
pedigree_ids <- pedigree.df$id

### IDs in model AND pedigree
ids_in_both <- intersect(model_ids, pedigree_ids)

### IDs in model NOT in pedigree
ids_missing_from_pedigree <- setdiff(model_ids, pedigree_ids)

### IDs in pedigree NOT in model (optional info)
ids_unused_in_model <- setdiff(pedigree_ids, model_ids)


# (3) Subset the pedigree to include all ancestors  ----

## Check formatting
pedigree.df$id <- as.character(pedigree.df$id)
pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)
rep_data$ring <- as.character(rep_data$ring)

model_ids <- unique(rep_data$ring)
ped_ids_needed <- get_ancestors(model_ids, pedigree.df)
length(ped_ids_needed)

## Subset the pedigree
pedigree.pruned <- pedigree.df[pedigree.df$id %in% ped_ids_needed, ]

## Check for missing IDs
setdiff(model_ids, pedigree.pruned$id) 

## Add these as founders
missing_ids <- setdiff(model_ids, pedigree.pruned$id)
founders <- data.frame(id = missing_ids, dam = NA, sire = NA)

### Append to your original pedigree
pedigree.extended <- rbind(pedigree.df, founders)

### Now prune the extended version
ped_ids_needed <- get_ancestors(model_ids, pedigree.extended)
pedigree.pruned <- pedigree.extended[pedigree.extended$id %in% ped_ids_needed, ]

setdiff(model_ids, pedigree.pruned$id)


# (4) Visualise the pruned pedigree  ----

edges <- pedigree.pruned %>%
  tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
  dplyr::select(from = value, to = id)

## Build igraph object
ped_graph <- graph_from_data_frame(edges, directed = TRUE)

plot(ped_graph,
     vertex.size = 3,
     vertex.label = NA,
     edge.arrow.size = 0.2,
     layout = layout_with_sugiyama(ped_graph)$layout)


# (5) Look at the genetic matrix A  ----

ped <- pedigree.pruned
names(ped) <- c("animal", "dam", "sire")

Ainv <- inverseA(ped)$Ainv
A <- as.matrix(solve(Ainv))

## Inspect sparse matrix
A_sparse <- as(A, "dgCMatrix")
summary(A_sparse)

### Check how many non-zero off-diagonal elements
nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")

library(pheatmap)
pheatmap(as.matrix(A), show_rownames = FALSE, show_colnames = FALSE)



# (6) Fit the models ----

cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
rep_data$animal <- rep_data$ring

rep_data$animal_shuffled <- sample(rep_data$animal)

posteriors_list <- list()

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ 1 + age + (1 | season) + (1 | ring) + (1 | ring_season) + (1 | gr(animal, cov = A))")
  
  mod_brm <- brm(
    formula = formula,
    data = rep_data,
    family = gaussian(),  
    data2 = list(A = A),  # your pedigree relatedness matrix
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Get posteriors 
  posterior <- as_draws_df(mod_brm)
 
  sd_animal <- posterior$sd_animal__Intercept
  var_animal  <- posterior$sd_animal__Intercept^2
  var_season  <- posterior$sd_season__Intercept^2
  var_resid   <- posterior$sigma^2
  h2_post <- var_animal / (var_animal + var_season + var_resid)
  
  posterior.df <- data.frame(cbind(sd_animal, var_animal, var_season, var_resid, h2_post),
                             behaviour = cov[i])
  
  posteriors_list[[i]] <- posterior.df
 
  # Output model
  mod_name <- paste0("mod_brm_", cov[i])
  assign(mod_name, mod_brm)
  
  save(list = mod_name, file = paste0("Data_outputs/waal_crozet_brms_behaviour_heritability_", cov[i], ".RData"))
    
}

posteriors.all <- do.call("rbind", posteriors_list)
save(posteriors.all, file = "Data_outputs/waal_cro_behaviour_heritability_posteriors.RData")


# (7) Visualise results  ----

# Model checking
plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
pp_check(mod_brm)

pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

# Heritability 
posteriors.all %<>%
  group_by(behaviour) %>%
  mutate(mean_h2 = mean(h2_post, na.rm = TRUE))

posteriors.all$behaviour <- reorder(posteriors.all$behaviour, posteriors.all$mean_h2)

## Recode behaviour
posteriors.all %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"))

ggplot(posteriors.all, aes(x = h2_post, y = behaviour)) +
  stat_density_ridges(fill = "skyblue", alpha = 0.7, quantile_lines = TRUE, quantiles = 2, vline_colour = "magenta4", vline_size = 1, vline_type = "dotted") +
  theme_bw() +
  labs(x = "Posterior heritability (h²)", y = "",
       title = "Posterior destributions of heritability for WAAL at Crozet") +
  xlim(0, NA) 


```

#### * Bird Island

```{r waal_birdis_behaviour_heritability }

# (1) Load and subsample the data  ----
load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct, transit_to_rest.ratio) == 0) == 0)
rep_data %<>% filter(species == "waal" & colony == "birdis")

pedigree.df <- read.csv("Data_inputs/waal_birdis_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)


# (2) Check the pedigree  ----

## Are the ID in rep_data the same as id in pedigree.df?
pedigree.df$id <- as.character(pedigree.df$id)
rep_data$ring <- as.character(rep_data$ring)

model_ids <- unique(rep_data$ring)
pedigree_ids <- pedigree.df$id

### IDs in model AND pedigree
ids_in_both <- intersect(model_ids, pedigree_ids)

### IDs in model NOT in pedigree
ids_missing_from_pedigree <- setdiff(model_ids, pedigree_ids)

### IDs in pedigree NOT in model (optional info)
ids_unused_in_model <- setdiff(pedigree_ids, model_ids)


# (3) Subset the pedigree to include all ancestors  ----

## Check formatting
pedigree.df$id <- as.character(pedigree.df$id)
pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)
rep_data$ring <- as.character(rep_data$ring)

model_ids <- unique(rep_data$ring)
ped_ids_needed <- get_ancestors(model_ids, pedigree.df)
length(ped_ids_needed)

## Subset the pedigree
pedigree.pruned <- pedigree.df[pedigree.df$id %in% ped_ids_needed, ]

## Check for missing IDs - NONE MISSING
setdiff(model_ids, pedigree.pruned$id) 

# Fix birds that are recorded as both dams and sires
intersect(pedigree.pruned$sire[!is.na(pedigree.pruned$sire)], pedigree.pruned$dam[!is.na(pedigree.pruned$dam)])

na.omit(pedigree.pruned[pedigree.pruned$sire == "5116152" | pedigree.pruned$dam == "5116152", ])
pedigree.pruned$sire[pedigree.pruned$id == "4003739"] <- "5116152"
pedigree.pruned$dam[pedigree.pruned$id == "4003739"] <- "5142862"

na.omit(pedigree.pruned[pedigree.pruned$sire == "5110290" | pedigree.pruned$dam == "5110290", ])
pedigree.pruned$sire[pedigree.pruned$id == "4003674"] <- "4000720"
pedigree.pruned$dam[pedigree.pruned$id == "4003674"] <- "5110290"

na.omit(pedigree.pruned[pedigree.pruned$sire == "5097831" | pedigree.pruned$dam == "5097831", ])
pedigree.pruned$sire[pedigree.pruned$id == "5144330"] <- "5093486"
pedigree.pruned$dam[pedigree.pruned$id == "5144330"] <- "5097831"

# (4) Visualise the pruned pedigree  ----

edges <- pedigree.pruned %>%
  tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
  dplyr::select(from = value, to = id)

## Build igraph object
ped_graph <- graph_from_data_frame(edges, directed = TRUE)

plot(ped_graph,
     vertex.size = 3,
     vertex.label = NA,
     edge.arrow.size = 0.2,
     layout = layout_with_sugiyama(ped_graph)$layout)


# (5) Look at the genetic matrix A  ----

ped <- pedigree.pruned
names(ped) <- c("animal", "dam", "sire")

# Deal with dams coming before offspring
ped <- orderPed(ped)

Ainv <- inverseA(ped)$Ainv
A <- as.matrix(solve(Ainv))

## Inspect sparse matrix
A_sparse <- as(A, "dgCMatrix")
summary(A_sparse)

### Check how many non-zero off-diagonal elements
nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")

pheatmap(as.matrix(A), show_rownames = FALSE, show_colnames = FALSE)



# (6) Fit the models ----

cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
rep_data$animal <- rep_data$ring

rep_data$animal_shuffled <- sample(rep_data$animal)

posteriors_list <- list()

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ 1 + phase + (1 | season) + (1 | ring) + (1 | gr(animal, cov = A))")
  
  mod_brm <- brm(
    formula = formula,
    data = rep_data,
    family = gaussian(),  
    data2 = list(A = A), 
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Get posteriors 
  posterior <- as_draws_df(mod_brm)
 
  sd_animal <- posterior$sd_animal__Intercept
  var_animal  <- posterior$sd_animal__Intercept^2
  var_season  <- posterior$sd_season__Intercept^2
  var_resid   <- posterior$sigma^2
  h2_post <- var_animal / (var_animal + var_season + var_resid)
  
  posterior.df <- data.frame(cbind(sd_animal, var_animal, var_season, var_resid, h2_post),
                             behaviour = cov[i])
  
  posteriors_list[[i]] <- posterior.df
 
  # Output model
  mod_name <- paste0("mod_brm_", cov[i])
  assign(mod_name, mod_brm)
  
  save(list = mod_name, file = paste0("Data_outputs/waal_birdis_brms_behaviour_heritability_", cov[i], ".RData"))
    
}

posteriors.all <- do.call("rbind", posteriors_list)
save(posteriors.all, file = "Data_outputs/waal_birdis_behaviour_heritability_posteriors.RData")


# (7) Visualise results  ----

# Model checking
plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
pp_check(mod_brm)

pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

# Heritability 
posteriors.all %<>%
  group_by(behaviour) %>%
  mutate(mean_h2 = mean(h2_post, na.rm = TRUE))

posteriors.all$behaviour <- reorder(posteriors.all$behaviour, posteriors.all$mean_h2)

## Recode behaviour
posteriors.all %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"))

ggplot(posteriors.all, aes(x = h2_post, y = behaviour)) +
  stat_density_ridges(fill = "skyblue", alpha = 0.7, quantile_lines = TRUE, quantiles = 2, vline_colour = "magenta4", vline_size = 1, vline_type = "dotted") +
  theme_bw() +
  labs(x = "Posterior heritability (h²)", y = "",
       title = "Posterior destributions of heritability for WAAL at Crozet") +
  xlim(0, NA) 


```


### Plot heritability estimates

```{r plot_heritability}

bbal_birdis.posteriors <- loadRData("Data_outputs/bbal_birdis_behaviour_heritability_posteriors.RData") %>%
  mutate(spec_col = "BBAL Bird Island")
waal_birdis.posteriors <- loadRData("Data_outputs/waal_birdis_behaviour_heritability_posteriors.RData") %>%
  mutate(spec_col = "WAAL Bird Island")
waal_cro.posteriors <- loadRData("Data_outputs/waal_cro_behaviour_heritability_posteriors.RData") %>%
  mutate(spec_col = "WAAL Crozet")

# Process the data
posteriors.all <- rbind(bbal_birdis.posteriors, waal_birdis.posteriors, waal_cro.posteriors)

posteriors.all %<>%
  group_by(spec_col, behaviour) %>%
  mutate(mean_h2 = mean(h2_post, na.rm = TRUE)) %>%
  arrange(behaviour, mean_h2) %>%
  ungroup()

## Recode behaviour
posteriors.all %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "transitions" = "Landings",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time",
      "transit_to_rest.ratio" = "Flight:Rest ratio"),
      behaviour = factor(behaviour, levels = rev(c("Trip duration", "Landings", "Rest time", "Flight time", "Flight:Rest ratio"))))

## Add facet code for plotting
posteriors.all %<>%
   mutate(facet_code = case_when(
    spec_col == "BBAL Bird Island" ~ "A",
    spec_col == "WAAL Bird Island" ~ "C",
    spec_col == "WAAL Crozet" ~ "D"))


# Make the plot
design = "
A#
CD
"

facet_labels <- c(
  A = "BBAL Bird Island",
  C = "WAAL Bird Island",
  D = "WAAL Crozet"
)

heritability_plot <- ggplot(posteriors.all, aes(x = h2_post, y = behaviour)) +
  # stat_density_ridges(fill = "#1b9e77",  alpha = 0.6, scale = 1, quantile_lines = TRUE, quantiles = 2, vline_colour = "#08519c", vline_size = 1, vline_type = "dashed") +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.6, scale = 1) +
  theme_bw() +
   ggh4x::facet_manual(~facet_code, design = design, scales = "free_x",
                       labeller = as_labeller(facet_labels)) +
  labs(x = "Posterior distribution for heritability (h²)", y = "") 


png(filename = "Figures/behaviour_heritability.png", width = 10, height = 10, units = "in", res = 900)
heritability_plot
dev.off()

```

## Supplementary analyses

### How does repeatability change with the number of individual repeats?

I have a lot of variation in the number of seasons' worth of data are available per individual. To test whether this might be affecting my estimates, here I have rerun the analyses using data that have been sampled to individuals with more than or equal 2, 3, or 4 trip repeats (within a season).

To speed things up, I'm calculating repeatability using simple linear model versions of the brms models, as I'm not concerned about uncertainty here.

Repeatability values stabilise at >=3 repeats per individual-year so ensuring this filter is included in the main analysis.

```{r SUPPLEMENTARY_repeatability_by_repeats_models}

load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct, transit_to_rest.ratio) == 0) == 0)

# Get time since first trip for each bird/season
rep_data %<>%
  group_by(ring, season) %>%
  mutate(time_since_first = as.numeric(difftime(start_date, min(start_date), units = "days"))) %>%
  ungroup()

# Split data
rep_bbal <- rep_data %>% filter(species == "bba")
rep_waal <- rep_data %>% filter(species == "waal")

cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")

n_reps <- c(2,3,4,5)

rep_testing.output <- list()
index <- 1 

for (i in 1:length(cov)) {
  
  for (n in 1:length(n_reps)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + phase + age")
  
  # Subset to n_reps
  rep_bbal.sub <- rep_bbal %>% 
    group_by(ring_season) %>%
    mutate(n_trips = n_distinct(start_date)) %>%
    filter(n_trips >= n_reps[n])
  
  rep_waal.sub <- rep_waal %>% 
    group_by(ring_season) %>%
    mutate(n_trips = n_distinct(start_date)) %>%
    filter(n_trips >= n_reps[n])
  
  ## BBAL ##
  bbal_birdis.lmm <- lmer(formula, rep_bbal.sub %>% filter(colony == "birdis"))
  bbal_ker.lmm <- lmer(formula, rep_bbal.sub %>% filter(colony == "kerguelen"))
  
  bbal_birdis.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(bbal_birdis.lmm),
                                species = "bbal",
                                colony = "birdis")
  
  bbal_ker.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(bbal_ker.lmm),
                                species = "bbal",
                                colony = "kerguelen")
  
  ## WAAL ##
  waal_birdis.lmm <- lmer(formula, rep_waal.sub %>% filter(colony == "birdis"))
  waal_cro.lmm <- lmer(formula, rep_waal.sub %>% filter(colony == "crozet"))
  
  waal_birdis.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(waal_birdis.lmm),
                                species = "waal",
                                colony = "birdis")
  
  waal_cro.rpt <- data.frame(level = c("annual", "within_season", "between_season"),
                                rpt = extract_rpt(waal_cro.lmm),
                                species = "waal",
                                colony = "crozet")
 
  rep_all <- rbind(bbal_birdis.rpt, bbal_ker.rpt, waal_birdis.rpt, waal_cro.rpt) %>%
                mutate(behaviour = cov[i],
                       n_trips = n_reps[n]) %>%
                relocate(species, colony, behaviour, n_trips, level, rpt)
  
  rep_testing.output[[index]] <- rep_all
  index <- index + 1

  }
   
}

rep_testing.all <- bind_rows(rep_testing.output)
rep_testing.all %<>% arrange(species)

save(rep_testing.all, file = "Data_outputs/SUPP_behaviour_rpt_test_Ntrips.RData")

```

#### Figure SX: Repeatability by repeat trips

```{r SUPPLEMENTARY_repeatability_by_repeats_figure}

load("Data_outputs/SUPP_behaviour_rpt_test_Ntrips.RData")

rep_testing.all %<>%
  mutate(spec_col = case_when( colony == "kerguelen" ~ "BBAL Kerguelen",
                               colony == "crozet" ~ "WAAL Crozet",
                               colony == "birdis" & species == "bbal" ~ "BBAL Birdisland",
                               TRUE ~ "WAAL Bird Island"),
         behaviour = recode(behaviour,
                "duration.hrs" = "Trip duration",
                "transitions" = "Landings",
                "transit_time.pct" = "Flight time",
                "rest_time.pct" = "Rest time",
                "transit_to_rest.ratio" = "Flight:Rest ratio"),
        behaviour = factor(behaviour),              
        behaviour = forcats::fct_rev(behaviour)) %>%
  filter(level != "annual")

rpt_test.plot <- ggplot(rep_testing.all, aes(x = n_trips, y = rpt, group = level, col = level)) +
  geom_point() +
  geom_line() +
  scale_colour_manual(values = c(col_between, col_within),
                   labels = c("Between season", "Within-season")) +
  labs(x = "Minimum number of trips per individual-year", y = "Estimated repeatability") +
  facet_grid(spec_col ~ behaviour) +
  theme_bw() +
  theme(legend.position = "none")


png(filename = "Figures/MS Figures/FIGSx_repeatability_by_sample.png", width = 12, height = 10, units = "in", res = 900)
rpt_test.plot
dev.off()


```

```{r supplementary_subsample_by_repeats}

load("Data_inputs/behavioural_repeatability_data.RData")

rep_data.bba <- rep_data %>% filter(species == "bba")
rep_data.waal <- rep_data %>% filter(species == "waal")

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
n_seas = c(1,2,3)

rpt_output.list <- list()

for (n in n_seas) {

  print(paste0("Processing birds with >", n_seas[n], " seasons."))
  
  # Subsample to individuals with >= 3 or >=4 seasons here
  rep_data.bba.sub <- rep_data.bba %>% group_by(ring) %>% mutate(n_seasons = n_distinct(season)) %>% filter(n_seasons > n_seas[n])
  rep_data.waal.sub <- rep_data.waal %>% group_by(ring) %>% mutate(n_seasons = n_distinct(season)) %>% filter(n_seasons > n_seas[n])  
  
  for (i in 1:length(cov)) {
    
    print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
    
    cov_value <- ifelse(i == 3, cov[i], paste0("log(", cov[i], ")"))
    formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + phase + colony + age")
    
    # Bootstrap repeatability
    tok <- Sys.time() 
    
    print("Bootstrapping BBAL repeatability...")
    
     ## BBAL ##
     rpt.bba <- lmer(formula, rep_data.bba.sub)
     vars.bba <- extract_var(rpt.bba)
     repeatability_output.bba <- suppressWarnings(boot_rpt(1000, rep_data.bba.sub, formula))
   
    print("Bootstrapping WAAL repeatability...")
    
     ## WAAL ##
     rpt.waal <- lmer(formula, rep_data.waal.sub)
     vars.waal <- extract_var(rpt.waal)
     repeatability_output.waal <- suppressWarnings(boot_rpt(1000, rep_data.waal.sub, formula))
  
    tik <- Sys.time()
    tik-tok 
    
    ### Summarise repeatability output ----
  
    rpt_summary.bba <- cbind(vars.bba, repeatability_output.bba)
  
    rpt_summary.bba$group <- rownames(rpt_summary.bba)
    rpt_summary.bba$group = gsub("_var", "", rpt_summary.bba$group)
    rownames(rpt_summary.bba) <- NULL
  
    rpt_summary.bba %<>%
      mutate(species = "BBAL",
             parameter = cov[i],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),#, p),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci, 
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.bba.sub), ", ", n_distinct(rep_data.bba.sub$ring)),
             seasons = n_seas[n]+1)
  
    # WAAL
    
    rpt_summary.waal <- cbind(vars.waal, repeatability_output.waal)
    rpt_summary.waal$group <- rownames(rpt_summary.waal)
    rpt_summary.waal$group = gsub("_var", "", rpt_summary.waal$group)
    rownames(rpt_summary.waal) <- NULL
  
    rpt_summary.waal %<>%
      mutate(species = "WAAL",
             parameter = cov[i],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),#, p),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.waal.sub), ", ", n_distinct(rep_data.waal.sub$ring)),
             seasons = n_seas[n]+1)
    
    # Output summary
    rpt_output.list[[length(rpt_output.list) + 1]] <- rbind(rpt_summary.bba, rpt_summary.waal)

  }

}

rpt_output.seasons <- do.call("rbind", rpt_output.list)
rpt_output.seasons %<>% arrange(species, seasons)

save(rpt_output.seasons, file = "Data_outputs/behavioural_repeatability_results_by_season.RData")


```

```{r supplementary_plot_by_season, fig.dim = c(6,8)}

load("Data_outputs/behavioural_repeatability_results_by_season.RData")

# Make dataframe long
rpt_long.seasons <- rpt_output.seasons %>%
  select(seasons, species, parameter, within_R, between_R, within_ci, between_ci) %>%
  pivot_longer(
    cols = starts_with(c("within_", "between_")),
    names_to = c("type", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  separate(ci, into = c("ci_lower", "ci_upper"), sep = ", ", convert = TRUE) %>%
  mutate(
    ci_lower = as.numeric(str_remove(ci_lower, "\\[")),
    ci_upper = as.numeric(str_remove(ci_upper, "\\]")),
    R = as.numeric(str_extract(R, "\\d+\\.\\d+")),  
    type = factor(type, levels = c("between", "within"))  )

## Reorder for plotting
rpt_long.seasons$parameter <- factor(rpt_long.seasons$parameter, levels = c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs" ))


# Set colours
col_within <- "#91c5f2"
col_between <- "#0b3860"

# Build plot
rptplot.seasons <- 
  ggplot(rpt_long.seasons, aes(x = R, y = parameter, color = type)) +
  annotate("rect", xmin = -Inf, xmax = 0.1, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.5) +
  annotate("rect", xmin = 0.1, xmax = 0.3, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  scale_y_discrete(labels = c(
    "duration.hrs" = "Trip duration (hrs)",
    "transitions" = "Number of landings",
    "transit_time.pct" = "Flight time (% trip)",
    "rest_time.pct" = "Rest time (% trip)",
    "transit_to_rest.ratio" = "Ratio flight to rest (%)"
  )) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper),
                 position = position_dodge(width = 0.5), height = 0.2) +
  facet_grid(rows = vars(seasons), cols = vars(species),
             labeller = labeller(seasons = as_labeller(
               c(`3` = "3+ seasons", `4` = "4+ seasons")))) +
  labs(x = "Repeatability (\u00B1 95% CI)") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "top",
        axis.title.y = element_blank())

save_plot(rptplot.seasons, "Figures/behavioural_rpt_results_bySeason.png", 6, 6)
rptplot.seasons

```

### Does repeatability depend on breeding phase?

Behaviour may be very different between incubation and brooding. Although I account for breeding phase in my models, it's important to check whether there are major differences between the breeding phases that could preclude fitting them in a single model.

```{r supplementary_repeatability_by_phase}

load("Data_inputs/behavioural_repeatability_data.RData")

rep_data.bba <- rep_data %>% filter(species == "bba")
rep_data.waal <- rep_data %>% filter(species == "waal")

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
phases <- unique(rep_data.bba$phase) 

rpt_output.list <- list()

for (ph in 1:length(phases)) {
  
  print(paste0("Processing the ", phases[ph], " phase."))
  
  rep_data.bba.sub <- rep_data.bba %>% filter(phase == phases[ph])
  rep_data.waal.sub <- rep_data.waal %>% filter(phase == phases[ph])
  
  for (i in 1:length(cov)) {
    
    print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), 
                 " Processing the ", phases[ph], " phase, covariate ", i, " of ", length(cov), ": ", cov[i]))
    
    cov_value <- ifelse(i == 3, cov[i], paste0("log(", cov[i], ")"))
    formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + colony + age")
    
    # Bootstrap repeatability
    print("Bootstrapping repeatability...")
    tok <- Sys.time() 
    
     ## BBAL ##
     rpt.bba <- lmer(formula, rep_data.bba.sub)
     vars.bba <- extract_var(rpt.bba)
     repeatability_output.bba <- suppressWarnings(boot_rpt(1000, rep_data.bba.sub, formula, nophase = TRUE))
   
     ## WAAL ##
     rpt.waal <- lmer(formula, rep_data.waal.sub)
     vars.waal <- extract_var(rpt.waal)
     repeatability_output.waal <- suppressWarnings(boot_rpt(1000, rep_data.waal.sub, formula, nophase = TRUE))
  
   
    tik <- Sys.time()
    tik-tok 
    
    ### Summarise repeatability output ----
  
    rpt_summary.bba <- cbind(vars.bba, repeatability_output.bba)
  
    rpt_summary.bba$group <- rownames(rpt_summary.bba)
    rpt_summary.bba$group = gsub("_var", "", rpt_summary.bba$group)
    rownames(rpt_summary.bba) <- NULL
  
    rpt_summary.bba %<>%
      mutate(species = "BBAL",
             parameter = cov[i],
             phase = phases[ph],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci, 
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.bba.sub), ", ", n_distinct(rep_data.bba.sub$ring)))
  
    # WAAL
    
    rpt_summary.waal <- cbind(vars.waal, repeatability_output.waal)
    rpt_summary.waal$group <- rownames(rpt_summary.waal)
    rpt_summary.waal$group = gsub("_var", "", rpt_summary.waal$group)
    rownames(rpt_summary.waal) <- NULL
  
    rpt_summary.waal %<>%
      mutate(species = "WAAL",
             phase = phases[ph],
             parameter = cov[i],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),#, p),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.waal.sub), ", ", n_distinct(rep_data.waal.sub$ring)))
    
    # Output summary
    rpt_output.list[[length(rpt_output.list) + 1]] <- rbind(rpt_summary.bba, rpt_summary.waal)

  }

}

rpt_output.phase <- do.call("rbind", rpt_output.list)
rpt_output.phase %<>% arrange(species, phase)

save(rpt_output.phase, file = "Data_outputs/behavioural_repeatability_results_by_phase.RData")


```

```{r supplementary_plot_by_season, fig.dim = c(6,6)}

load("Data_outputs/behavioural_repeatability_results_by_phase.RData")

# Make dataframe long
rpt_long.phase <- rpt_output.phase %>%
  select(phase, species, parameter, within_R, between_R, within_ci, between_ci) %>%
  pivot_longer(
    cols = starts_with(c("within_", "between_")),
    names_to = c("type", ".value"),
    names_pattern = "(.*)_(.*)") %>%
  separate(ci, into = c("ci_lower", "ci_upper"), sep = ", ", convert = TRUE) %>%
  mutate(
    ci_lower = as.numeric(str_remove(ci_lower, "\\[")),
    ci_upper = as.numeric(str_remove(ci_upper, "\\]")),
    R = as.numeric(str_extract(R, "\\d+\\.\\d+")),  
    type = factor(type, levels = c("between", "within"))  )

## Reorder for plotting
rpt_long.phase$parameter <- factor(rpt_long.phase$parameter, levels = c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs" ))

rpt_long.phase$phase <- factor(rpt_long.phase$phase, levels = c(
  "incubation", "brooding" ))

# Set colours
col_within <- "#91c5f2"
col_between <- "#0b3860"

# Build plot
rptplot.phase <- 
  ggplot(rpt_long.phase, aes(x = R, y = parameter, color = type)) +
  annotate("rect", xmin = -Inf, xmax = 0.1, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.5) +
  annotate("rect", xmin = 0.1, xmax = 0.3, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  scale_y_discrete(labels = c(
    "duration.hrs" = "Trip duration (hrs)",
    "transitions" = "Number of landings",
    "transit_time.pct" = "Flight time (% trip)",
    "rest_time.pct" = "Rest time (% trip)",
    "transit_to_rest.ratio" = "Ratio flight to rest (%)"
  )) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper),
                 position = position_dodge(width = 0.5), height = 0.2) +
  facet_grid(rows = vars(phase), cols = vars(species),
             labeller = labeller(phase = as_labeller(
               c("incubation" = "Incubation", "brooding" = "Brooding")))) +
  labs(x = "Repeatability (\u00B1 95% CI)") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "top",
        axis.title.y = element_blank())

save_plot(rptplot.seasons, "Figures/behavioural_rpt_results_byPhase.png", 6, 6)
rptplot.seasons

```

# Appendix

```{r foraging_effort_bertille, eval = FALSE}

rep_data %<>% mutate(landings_per_day = landing/duration.days)

# Get mean landings per day per season per individual

summary_landings <- rep_data %>% group_by(species, colony, ring, season) %>% summarise(mean_landings = mean(landings_per_day))

lmm1 <- lmer(landings_per_day ~ colony + species + (1|ring), data = rep_data)

ggplot(summary_landings, aes(x = factor(season), y = mean_landings, col = colony)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~ species) +
  labs(y = "Mean landings per day of trip") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank())


# Plot random intercepts
ranefs <- broom.mixed::tidy(lmm1, effects = "ran_vals", conf.int = TRUE)
ring_ranefs <- ranefs[ranefs$group == "ring", ] %>% rename(ring = level)

## Add colony/species back
meta <- rep_data %>% select(c(ring, colony, species)) %>% distinct()
ring_ranefs <- merge(ring_ranefs, meta, by = "ring", all.x = T) %>%
  mutate(spec_col = paste(species, colony, sep = "_"))

ggplot(ring_ranefs, aes(x = estimate, y = reorder(ring, estimate))) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  facet_wrap(~ spec_col, scales = "free_y") +
  labs(x = "Random intercept estimate", y = "Ring ID") +
  theme_bw() +
  theme(axis.text.y = element_blank())


# Get summarised data
behav.df <- rep_data %>%
              select(c(species, colony, ring, season, landing, duration.days, duration.hrs, phase))

write.csv(behav.df, file = "Data_outputs/behaviour_summarised.csv", row.names = F)

```