---
title: "Behavioural repeatability results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4",  "gridExtra", "tidyr", "kableExtra", "stringr", "glmmTMB")

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select


# Load & process data ==========================================================

ind_files <- list.files("Data_inputs/", pattern = "individual_trips")
ind_list <- vector(mode = "list", length = 4)

for (i in 1:length(ind_files)) {
  
  my_data <- loadRData(paste0("Data_inputs/",ind_files[i]))
  ind_list[[i]] <- my_data
  
}

trip_data <- do.call("rbind", ind_list)

# Load metadata ---------------

load("Data_inputs/ind_meta.RData")

## Bind it in
trip_data <- merge(trip_data, all_ind.meta %>% mutate(species = tolower(species)), by = c("ring", "species"), all.x = T) %>%
  mutate(birthYr = as.numeric(birthYr),
         age = season - birthYr,
         age = ifelse(age < 0, NA, age)) 

# Filter data ---------------

# Remove any trips where transitions == 0 (not real foraging)
rep_data <- trip_data %>% 
  group_by(ring) %>% 
  mutate(n_seasons = n_distinct(season),
         ring_season = paste(ring, season, sep = "_")) %>%
  filter(n_seasons > 1 & !is.na(age) & transitions > 0) %>% 
  ungroup() %>%
  select(-c(n_seasons, sex)) %>%
  relocate(ring_season, .before = colony)


save(rep_data, file = "Data_inputs/behavioural_repeatability_data.RData")


```

```{r foraging_effort_bertille, eval = FALSE}

rep_data %<>% mutate(landings_per_day = landing/duration.days)

# Get mean landings per day per season per individual

summary_landings <- rep_data %>% group_by(species, colony, ring, season) %>% summarise(mean_landings = mean(landings_per_day))

lmm1 <- lmer(landings_per_day ~ colony + species + (1|ring), data = rep_data)

ggplot(summary_landings, aes(x = factor(season), y = mean_landings, col = colony)) +
  geom_jitter(width = 0.2, alpha = 0.5) +
  facet_wrap(~ species) +
  labs(y = "Mean landings per day of trip") +
  theme_bw() +
  theme(axis.title.x = element_blank(),
        axis.ticks.x = element_blank(),
        axis.text.x = element_blank())


# Plot random intercepts
ranefs <- broom.mixed::tidy(lmm1, effects = "ran_vals", conf.int = TRUE)
ring_ranefs <- ranefs[ranefs$group == "ring", ] %>% rename(ring = level)

## Add colony/species back
meta <- rep_data %>% select(c(ring, colony, species)) %>% distinct()
ring_ranefs <- merge(ring_ranefs, meta, by = "ring", all.x = T) %>%
  mutate(spec_col = paste(species, colony, sep = "_"))

ggplot(ring_ranefs, aes(x = estimate, y = reorder(ring, estimate))) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  facet_wrap(~ spec_col, scales = "free_y") +
  labs(x = "Random intercept estimate", y = "Ring ID") +
  theme_bw() +
  theme(axis.text.y = element_blank())


# Get summarised data
behav.df <- rep_data %>%
              select(c(species, colony, ring, season, landing, duration.days, duration.hrs, phase))

write.csv(behav.df, file = "Data_outputs/behaviour_summarised.csv", row.names = F)

```


### Sample sizes

We know age effects behaviour, and individuals are recorded at different ages, so this is included in the repeatability models and hence the dataset has been filtered to exclude individuals of unknown age. We don't worry about sex, as each individual can only be one sex. We are also only using individuals with more than one season of data, as otherwise we can't calculate across season repeatability.

```{r sample_sizes}

# Number of individuals with x number of seasons
trips.table <- rep_data %>%
  group_by(colony, species, ring) %>%
  summarise(n_seasons = n_distinct(season)) %>%
  ungroup() %>%
  mutate(rpt_category = case_when(
    n_seasons == 1 ~ "1 seasons",
    n_seasons == 2 ~ "2 seasons",
    n_seasons == 3 ~ "3 seasons",
    n_seasons > 3 ~ "4+ seasons"
  )) %>%
  group_by(colony, species, rpt_category) %>%
  summarise(num_birds = n()) %>%
  arrange(species, colony, rpt_category) %>%
  pivot_wider(
    names_from = rpt_category,       
    values_from = num_birds, 
    values_fill = 0) %>%
  mutate(species = ifelse(species == "bba", "BBAL", "WAAL"),
         colony = ifelse(colony == "kerguelen", "Kerguelen",
                         ifelse(colony == "crozet", "Crozet", "Bird Island"))) %>%
  relocate(species, .before = colony) %>%
  rename(Species = species, Colony = colony)


kable(trips.table, format = "html", 
      caption = 'Table 1. Sample sizes for each species and colony, showing the number of individuals for which we have 2, 3, or 4+ seasons of data, collected using GLS and/or GPS devices.') %>%
  collapse_rows(columns = c(1)) %>%
  row_spec(2, extra_css = "border-bottom: 1px solid;") %>%
  kable_classic(font_size = 12)


```

## Testing for repeatablity

This analysis looks at repeatability in foraging behaviours. I am looking at five behavioural variables:

1.  *duration.hrs* = Duration of the foraging trip in hours
2.  *transitions* = Number of transitions between rest and flight and vice versa
3.  *transit_time.pct* = Percentage of foraging trip spent in flight
4.  *rest_time.pct* = Percentage of foraging trip spent in rest
5.  *transit_to_rest.ratio* = Ratio of flight to rest time; \>1 = more flight than rest, \<1 = more rest than flight

I dropped the distance metrics, as I did not have sufficient repeats across the datasets to produce reliable estimates of repeatability.

I have then written my own repeatability functions, which explicitly partition repeatability into **annual** repeatability (season level), **within-season** repeatabilty (ring:season level) and **between-season** repeatability (ring level). This is because *rptR* doesn't natively handle this kind of decomposition - it can only estimate repeatabilty for a single grouping factor at a time.

In my functions, I manually boostrap confidence intervals and SEs around repeatability by resampling the data and refitting models:

*To estimate uncertainty around repeatability estimates, we implemented a nonparametric bootstrap procedure with 1000 iterations. In each iteration, the dataset was resampled with replacement, the linear mixed-effects model was refitted, and variance components were extracted. We calculated repeatability for each iteration as the proportion of variance explained at the relevant grouping level (ring, season, or ring:season). From the resulting distributions, we computed mean estimates, standard errors, and 95% confidence intervals based on the 2.5th and 97.5th percentiles.*

The basic formula used in this analysis is:

$$y = \beta_0 + \beta_1 \cdot \text{phase} + \beta_2 \cdot \text{age} + \beta_3 \cdot \text{colony} + U_{\text{ring}} + U_{\text{season}} + U_{\text{ring:season}} + \varepsilon$$

where:

-   $y$ represents each of the 5 behavioural response variables,

-   $\beta_0$ is the intercept,

-   $\beta_1$ - $\beta_{3}$ are fixed effect coefficients for $phase$, $age$, and $colony$,

-   $U_{\text{ring}}$, $U_{\text{season}}$, and $U_{\text{ring:season}}$ are random intercepts for ring, season, and ring within season respectively,

    $$U_{\text{ring}} \sim \mathcal{N}(0, \sigma^2_{\text{ring}})$$

    $$U_{\text{season}} \sim \mathcal{N}(0, \sigma^2_{\text{season}})$$

    $$U_{\text{ring_season}} \sim \mathcal{N}(0, \sigma^2_{\text{ring_season}})$$

-   $ε$ is the residual error.

    $$ε \sim \mathcal{N}(0, \sigma^2_{\text{res}})$$

Several of the predictor values have positive-skewed distributions with heavy tails. As I'm not interested in predictions, a simple and stable approach to dealing with this is simply to **log-transform** $y$ in each model. The only exception for this is *transit_time.pct*, which, even though it is a percentage, it is relatively normally distributed, the residuals look good, and the values are comfortably away from the 0 to 100 boundaries (i.e. mostly between 20 and 80). The distribution of *rest_time.pct* is a little problematic - the values are right skewed with a very long tail towards high values and lots of values around 0. Possibly a beta model would be a better fit here. But for ease of computation, I'm pursuing the log transformation for now. The variances don't look super different between this and a beta model.

```{r calculating_repeatability, eval = FALSE}

load("Data_inputs/behavioural_repeatability_data.RData")

# Remove zero values
rep_data %<>% filter(rowSums(select(., transitions, transit_time.pct, transit_to_rest.ratio) == 0) == 0)

rep_bba <- rep_data %>% filter(species == "bba")
rep_waal <- rep_data %>% filter(species == "waal")

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")

rpt_output.list <- vector(mode = "list", length = length(cov))

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  cov_value <- paste0("log(", cov[i], ")")
  formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + phase + colony + age")
  
  # Bootstrap repeatability
  print("Bootstrapping repeatability...")
  tok <- Sys.time() 
  
   ## BBAL ##
   rpt.bba <- lmer(formula, rep_bba)
   vars.bba <- extract_var(rpt.bba)
   repeatability_output.bba <- suppressWarnings(boot_rpt(1000, rep_bba, formula))
 
   ## WAAL ##
   rpt.waal <- lmer(formula, rep_waal)
   vars.waal <- extract_var(rpt.waal)
   repeatability_output.waal <- suppressWarnings(boot_rpt(1000, rep_waal, formula))

 
  tik <- Sys.time()
  tik-tok 
  
  ### Summarise repeatability output ----

  rpt_summary.bba <- cbind(vars.bba, repeatability_output.bba)

  rpt_summary.bba$group <- rownames(rpt_summary.bba)
  rpt_summary.bba$group = gsub("_var", "", rpt_summary.bba$group)
  rownames(rpt_summary.bba) <- NULL

  rpt_summary.bba %<>%
    mutate(species = "BBAL",
           parameter = cov[i],
           R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
           group = gsub("se_", "", group)) %>%
    relocate(c(group, R), .before = se) %>%
    select(-c(est, se)) %>%
    pivot_wider(
      names_from = group,
      values_from = c(R, ci, var),
      names_glue = "{group}_{.value}"
    ) %>%
    relocate(species, parameter, annual_var, 
             within_var, within_R, within_ci, 
             between_var, between_R, between_ci) %>%
    relocate(species, parameter, annual_var, 
             within_var, within_R, within_ci,
             between_var, between_R, between_ci) %>%
    mutate(sample_size = paste0(nrow(rep_bba), ", ", n_distinct(rep_bba$ring)))

  # WAAL
  
  rpt_summary.waal <- cbind(vars.waal, repeatability_output.waal)
  
  rpt_summary.waal$group <- rownames(rpt_summary.waal)
  rpt_summary.waal$group = gsub("_var", "", rpt_summary.waal$group)
  rownames(rpt_summary.waal) <- NULL

  rpt_summary.waal %<>%
    mutate(species = "WAAL",
           parameter = cov[i],
           R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
           group = gsub("se_", "", group)) %>%
    relocate(c(group, R), .before = se) %>%
    select(-c(est, se)) %>%
    pivot_wider(
      names_from = group,
      values_from = c(R, ci, var),
      names_glue = "{group}_{.value}"
    ) %>%
    relocate(species, parameter, annual_var, 
             within_var, within_R, within_ci,
             between_var, between_R, between_ci) %>%
    mutate(sample_size = paste0(nrow(rep_waal), ", ", n_distinct(rep_waal$ring)))
  
  # Output summary
  rpt_output.list[[i]] <- rbind(rpt_summary.bba, rpt_summary.waal)
   
}

rpt_output.full <- do.call("rbind", rpt_output.list)
rpt_output.full %<>% arrange(species)

save(rpt_output.full, file = "Data_outputs/behavioural_repeatability_results.RData")

```

### Repeatability results

```{r repeatability_table}

load("Data_outputs/behavioural_repeatability_results.RData")
rpt_output.full %<>% filter(parameter != "max_distance.km" & parameter != "total_distance.km")

# Reorder for table
rpt_output.full$parameter <- factor(rpt_output.full$parameter, levels = rev(c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs"
)))

rpt_output.full <- rpt_output.full %>%
  arrange(species, parameter) 


kable(rpt_output.full, format = "html", 
      caption = 'Table 2. Repeatability parameters for each species. Sample sizes are for trips, individuals.',
      digits = 2) %>%
  row_spec(5, extra_css = "border-bottom: 1px solid;") %>%
  column_spec(2, extra_css = "border-left: 1px solid;") %>%
  column_spec(c(3,4,7,10), extra_css = "border-left: 1px dashed;") %>%
  collapse_rows(columns = c(1,10)) %>%
  kable_classic(font_size = 12)

```

```{r repeatability_figure, fig.cap = "Figure 1. Summary of bootstrapped repeatability values for each species. Grey intervals indicate R < 0.1 (darker interval), where repeatability is deemed to be neglibile, and 0.1 < R < 0.3, where repeatabiltiy is deemed to be low-modest. Repeatability values greater than this may be considered biologically interesting." , fig.topcaption = TRUE}

# Make dataframe long
rpt_long <- rpt_output.full %>%
  select(species, parameter,
         within_R, between_R,
         within_ci, between_ci) %>%
  pivot_longer(
    cols = starts_with(c("within_", "between_")),
    names_to = c("type", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  separate(ci, into = c("ci_lower", "ci_upper"), sep = ", ", convert = TRUE) %>%
  mutate(
    ci_lower = as.numeric(str_remove(ci_lower, "\\[")),
    ci_upper = as.numeric(str_remove(ci_upper, "\\]")),
    R = as.numeric(str_extract(R, "\\d+\\.\\d+")),  # Extract point estimate only
    type = factor(type, levels = c("between", "within"))  
  )

## Reorder for plotting
rpt_long$parameter <- factor(rpt_long$parameter, levels = c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs"
))


# Set colours
col_within <- "#91c5f2"
col_between <- "#0b3860"

# Build plot
rptplot <- 
  ggplot(rpt_long, aes(x = R, y = parameter, color = type)) +
  annotate("rect", xmin = -Inf, xmax = 0.1, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.5) +
  annotate("rect", xmin = 0.1, xmax = 0.3, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  scale_y_discrete(labels = c(
    "duration.hrs" = "Trip duration (hrs)",
    "transitions" = "Number of landings",
    "transit_time.pct" = "Flight time (% trip)",
    "rest_time.pct" = "Rest time (% trip)",
    "transit_to_rest.ratio" = "Ratio flight to rest (%)"
  )) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper),
                 position = position_dodge(width = 0.5), height = 0.2) +
  facet_wrap(~ species) +
  labs(x = "Repeatability (\u00B1 95% CI)") +
  theme_bw() +
  theme(legend.title = element_blank(),
        legend.position = "top",
        axis.title.y = element_blank())

png(rptplot, filename = "Figures/MS Figures/FIGx_behaviour_repeatability.png", width = 8, height = 6, units = "in", res = 900)
rptplot
dev.off()


```

### Supplementary analyses

#### How does repeatability change with the number of individual repeats?

I have a lot of variation in the number of seasons' worth of data are available per individual. To test whether this might be affecting my estimates, here I have rerun the analyses using data that have been sampled to individuals with more than or equal to 3, or 4 repeat seasons respectively.

```{r supplementary_subsample_by_repeats}

load("Data_inputs/behavioural_repeatability_data.RData")

rep_data.bba <- rep_data %>% filter(species == "bba")
rep_data.waal <- rep_data %>% filter(species == "waal")

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
n_seas = c(1,2,3)

rpt_output.list <- list()

for (n in n_seas) {

  print(paste0("Processing birds with >", n_seas[n], " seasons."))
  
  # Subsample to individuals with >= 3 or >=4 seasons here
  rep_data.bba.sub <- rep_data.bba %>% group_by(ring) %>% mutate(n_seasons = n_distinct(season)) %>% filter(n_seasons > n_seas[n])
  rep_data.waal.sub <- rep_data.waal %>% group_by(ring) %>% mutate(n_seasons = n_distinct(season)) %>% filter(n_seasons > n_seas[n])  
  
  for (i in 1:length(cov)) {
    
    print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
    
    cov_value <- ifelse(i == 3, cov[i], paste0("log(", cov[i], ")"))
    formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + phase + colony + age")
    
    # Bootstrap repeatability
    tok <- Sys.time() 
    
    print("Bootstrapping BBAL repeatability...")
    
     ## BBAL ##
     rpt.bba <- lmer(formula, rep_data.bba.sub)
     vars.bba <- extract_var(rpt.bba)
     repeatability_output.bba <- suppressWarnings(boot_rpt(1000, rep_data.bba.sub, formula))
   
    print("Bootstrapping WAAL repeatability...")
    
     ## WAAL ##
     rpt.waal <- lmer(formula, rep_data.waal.sub)
     vars.waal <- extract_var(rpt.waal)
     repeatability_output.waal <- suppressWarnings(boot_rpt(1000, rep_data.waal.sub, formula))
  
    tik <- Sys.time()
    tik-tok 
    
    ### Summarise repeatability output ----
  
    rpt_summary.bba <- cbind(vars.bba, repeatability_output.bba)
  
    rpt_summary.bba$group <- rownames(rpt_summary.bba)
    rpt_summary.bba$group = gsub("_var", "", rpt_summary.bba$group)
    rownames(rpt_summary.bba) <- NULL
  
    rpt_summary.bba %<>%
      mutate(species = "BBAL",
             parameter = cov[i],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),#, p),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci, 
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.bba.sub), ", ", n_distinct(rep_data.bba.sub$ring)),
             seasons = n_seas[n]+1)
  
    # WAAL
    
    rpt_summary.waal <- cbind(vars.waal, repeatability_output.waal)
    rpt_summary.waal$group <- rownames(rpt_summary.waal)
    rpt_summary.waal$group = gsub("_var", "", rpt_summary.waal$group)
    rownames(rpt_summary.waal) <- NULL
  
    rpt_summary.waal %<>%
      mutate(species = "WAAL",
             parameter = cov[i],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),#, p),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.waal.sub), ", ", n_distinct(rep_data.waal.sub$ring)),
             seasons = n_seas[n]+1)
    
    # Output summary
    rpt_output.list[[length(rpt_output.list) + 1]] <- rbind(rpt_summary.bba, rpt_summary.waal)

  }

}

rpt_output.seasons <- do.call("rbind", rpt_output.list)
rpt_output.seasons %<>% arrange(species, seasons)

save(rpt_output.seasons, file = "Data_outputs/behavioural_repeatability_results_by_season.RData")


```

```{r supplementary_plot_by_season, fig.dim = c(6,8)}

load("Data_outputs/behavioural_repeatability_results_by_season.RData")

# Make dataframe long
rpt_long.seasons <- rpt_output.seasons %>%
  select(seasons, species, parameter, within_R, between_R, within_ci, between_ci) %>%
  pivot_longer(
    cols = starts_with(c("within_", "between_")),
    names_to = c("type", ".value"),
    names_pattern = "(.*)_(.*)"
  ) %>%
  separate(ci, into = c("ci_lower", "ci_upper"), sep = ", ", convert = TRUE) %>%
  mutate(
    ci_lower = as.numeric(str_remove(ci_lower, "\\[")),
    ci_upper = as.numeric(str_remove(ci_upper, "\\]")),
    R = as.numeric(str_extract(R, "\\d+\\.\\d+")),  
    type = factor(type, levels = c("between", "within"))  )

## Reorder for plotting
rpt_long.seasons$parameter <- factor(rpt_long.seasons$parameter, levels = c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs" ))


# Set colours
col_within <- "#91c5f2"
col_between <- "#0b3860"

# Build plot
rptplot.seasons <- 
  ggplot(rpt_long.seasons, aes(x = R, y = parameter, color = type)) +
  annotate("rect", xmin = -Inf, xmax = 0.1, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.5) +
  annotate("rect", xmin = 0.1, xmax = 0.3, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  scale_y_discrete(labels = c(
    "duration.hrs" = "Trip duration (hrs)",
    "transitions" = "Number of landings",
    "transit_time.pct" = "Flight time (% trip)",
    "rest_time.pct" = "Rest time (% trip)",
    "transit_to_rest.ratio" = "Ratio flight to rest (%)"
  )) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper),
                 position = position_dodge(width = 0.5), height = 0.2) +
  facet_grid(rows = vars(seasons), cols = vars(species),
             labeller = labeller(seasons = as_labeller(
               c(`3` = "3+ seasons", `4` = "4+ seasons")))) +
  labs(x = "Repeatability (\u00B1 95% CI)") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "top",
        axis.title.y = element_blank())

save_plot(rptplot.seasons, "Figures/behavioural_rpt_results_bySeason.png", 6, 6)
rptplot.seasons

```

#### Does repeatability depend on breeding phase?

Behaviour may be very different between incubation and brooding. Although I account for breeding phase in my models, it's important to check whether there are major differences between the breeding phases that could preclude fitting them in a single model.

```{r supplementary_repeatability_by_phase}

load("Data_inputs/behavioural_repeatability_data.RData")

rep_data.bba <- rep_data %>% filter(species == "bba")
rep_data.waal <- rep_data %>% filter(species == "waal")

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "transitions", "transit_time.pct", "rest_time.pct", "transit_to_rest.ratio")
phases <- unique(rep_data.bba$phase) 

rpt_output.list <- list()

for (ph in 1:length(phases)) {
  
  print(paste0("Processing the ", phases[ph], " phase."))
  
  rep_data.bba.sub <- rep_data.bba %>% filter(phase == phases[ph])
  rep_data.waal.sub <- rep_data.waal %>% filter(phase == phases[ph])
  
  for (i in 1:length(cov)) {
    
    print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), 
                 " Processing the ", phases[ph], " phase, covariate ", i, " of ", length(cov), ": ", cov[i]))
    
    cov_value <- ifelse(i == 3, cov[i], paste0("log(", cov[i], ")"))
    formula <- paste0(cov_value, " ~ (1 | ring) + (1 | season) + (1 | ring_season) + colony + age")
    
    # Bootstrap repeatability
    print("Bootstrapping repeatability...")
    tok <- Sys.time() 
    
     ## BBAL ##
     rpt.bba <- lmer(formula, rep_data.bba.sub)
     vars.bba <- extract_var(rpt.bba)
     repeatability_output.bba <- suppressWarnings(boot_rpt(1000, rep_data.bba.sub, formula, nophase = TRUE))
   
     ## WAAL ##
     rpt.waal <- lmer(formula, rep_data.waal.sub)
     vars.waal <- extract_var(rpt.waal)
     repeatability_output.waal <- suppressWarnings(boot_rpt(1000, rep_data.waal.sub, formula, nophase = TRUE))
  
   
    tik <- Sys.time()
    tik-tok 
    
    ### Summarise repeatability output ----
  
    rpt_summary.bba <- cbind(vars.bba, repeatability_output.bba)
  
    rpt_summary.bba$group <- rownames(rpt_summary.bba)
    rpt_summary.bba$group = gsub("_var", "", rpt_summary.bba$group)
    rownames(rpt_summary.bba) <- NULL
  
    rpt_summary.bba %<>%
      mutate(species = "BBAL",
             parameter = cov[i],
             phase = phases[ph],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci, 
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.bba.sub), ", ", n_distinct(rep_data.bba.sub$ring)))
  
    # WAAL
    
    rpt_summary.waal <- cbind(vars.waal, repeatability_output.waal)
    rpt_summary.waal$group <- rownames(rpt_summary.waal)
    rpt_summary.waal$group = gsub("_var", "", rpt_summary.waal$group)
    rownames(rpt_summary.waal) <- NULL
  
    rpt_summary.waal %<>%
      mutate(species = "WAAL",
             phase = phases[ph],
             parameter = cov[i],
             R = paste0(signif(est, 2), " \u00B1 ", signif(se, 2)),
             group = gsub("se_", "", group)) %>%
      relocate(c(group, R), .before = se) %>%
      select(-c(est, se)) %>%
      pivot_wider(
        names_from = group,
        values_from = c(R, ci, var),#, p),
        names_glue = "{group}_{.value}"
      ) %>%
      relocate(species, parameter, annual_var, annual_R, annual_ci,
               within_var, within_R, within_ci,
               between_var, between_R, between_ci) %>%
      mutate(sample_size = paste0(nrow(rep_data.waal.sub), ", ", n_distinct(rep_data.waal.sub$ring)))
    
    # Output summary
    rpt_output.list[[length(rpt_output.list) + 1]] <- rbind(rpt_summary.bba, rpt_summary.waal)

  }

}

rpt_output.phase <- do.call("rbind", rpt_output.list)
rpt_output.phase %<>% arrange(species, phase)

save(rpt_output.phase, file = "Data_outputs/behavioural_repeatability_results_by_phase.RData")


```

```{r supplementary_plot_by_season, fig.dim = c(6,6)}

load("Data_outputs/behavioural_repeatability_results_by_phase.RData")

# Make dataframe long
rpt_long.phase <- rpt_output.phase %>%
  select(phase, species, parameter, within_R, between_R, within_ci, between_ci) %>%
  pivot_longer(
    cols = starts_with(c("within_", "between_")),
    names_to = c("type", ".value"),
    names_pattern = "(.*)_(.*)") %>%
  separate(ci, into = c("ci_lower", "ci_upper"), sep = ", ", convert = TRUE) %>%
  mutate(
    ci_lower = as.numeric(str_remove(ci_lower, "\\[")),
    ci_upper = as.numeric(str_remove(ci_upper, "\\]")),
    R = as.numeric(str_extract(R, "\\d+\\.\\d+")),  
    type = factor(type, levels = c("between", "within"))  )

## Reorder for plotting
rpt_long.phase$parameter <- factor(rpt_long.phase$parameter, levels = c(
  "transit_to_rest.ratio",
  "transit_time.pct",
  "rest_time.pct",
  "transitions",
  "duration.hrs" ))

rpt_long.phase$phase <- factor(rpt_long.phase$phase, levels = c(
  "incubation", "brooding" ))

# Set colours
col_within <- "#91c5f2"
col_between <- "#0b3860"

# Build plot
rptplot.phase <- 
  ggplot(rpt_long.phase, aes(x = R, y = parameter, color = type)) +
  annotate("rect", xmin = -Inf, xmax = 0.1, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.5) +
  annotate("rect", xmin = 0.1, xmax = 0.3, ymin = -Inf, ymax = Inf,
           fill = "grey80", alpha = 0.2) +
  scale_colour_manual(values = c(col_between, col_within),
                      labels = c("Between-season", "Within-season")) +
  scale_y_discrete(labels = c(
    "duration.hrs" = "Trip duration (hrs)",
    "transitions" = "Number of landings",
    "transit_time.pct" = "Flight time (% trip)",
    "rest_time.pct" = "Rest time (% trip)",
    "transit_to_rest.ratio" = "Ratio flight to rest (%)"
  )) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = ci_lower, xmax = ci_upper),
                 position = position_dodge(width = 0.5), height = 0.2) +
  facet_grid(rows = vars(phase), cols = vars(species),
             labeller = labeller(phase = as_labeller(
               c("incubation" = "Incubation", "brooding" = "Brooding")))) +
  labs(x = "Repeatability (\u00B1 95% CI)") +
  theme_minimal() +
  theme(legend.title = element_blank(),
        legend.position = "top",
        axis.title.y = element_blank())

save_plot(rptplot.seasons, "Figures/behavioural_rpt_results_byPhase.png", 6, 6)
rptplot.seasons

```

<!-- ## Changes to the approach -->

<!-- ### Permutation test -->

<!-- I originally conducted a permutation test to examine whether repeatability estimates are greater than expected by chance, under the null hypothesis that individual identity has no consistent effect within or across seasons. This was achieved my shuffling ring labels within seasons, to maintain seasonal structure (so no mixing of behaviours across year) and generate a null distribution: -->

<!-- *To assess the statistical significance of repeatability estimates, we performed a permutation test with 1000 iterations. In each iteration, individual identities (`ring`) were randomly reassigned within each season, breaking any consistent among-individual patterns while preserving seasonal structure. The linear mixed-effects model was refitted to each permuted dataset, and repeatability values were recalculated. P-values were computed as the proportion of permuted values greater than or equal to the observed repeatability, for annual, within-season, and between-season components.* -->

<!-- I ultimately decided to remove this in favour of just assessing the confidence intervals. I think my approach wasn't significantly conservative, and the p values weren't necessarily informative - a mean repeatability of 0.1 might be significantly different to 0, but if the lower bound of the CI is 0.001, I don't think it's of biological interest. I can come back to this and think about cleverer ways to conduct this test but for now I'm going to drop it (particularly as it's computationally intense). -->
