---
title: 'RPT: Supplementary'
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

```{r setup, echo = FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
# Load functions, packages, & data ---------------------------------------------

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "ggpubr", "tidyr", "sf", "ecmwfr", "ggridges", "rnaturalearth", "purrr", "geosphere", "lmerTest", "sp", "kableExtra", "emmeans", "data.table", "dtw") 

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

```

## Bearing calculations

As part of the spatial repeatability analyses, I comparing birds' bearings during their foraging trips within and between individuals, and within and between seasons.

Originally I computed the bearings as the bearing between the distal point on a bird's foraging trip to the colony. However, this can lead to small but important perturbations - for example the distal point might represent a short section of the foraging track that deviates substantially from the rest of the track.

To test for the importance of this, and to determine the most appropriate bearing measure, I calculated three options for birds' bearings:

-   Distal bearing - the bearing between the distal point and the colony

-   Average bearing - the average bearing on birds' foraging trips

-   Average foraging bearing - the average bearing for sections of birds' trips marked as 'foraging'

Below I summarise the results. Overall, the choice of bearing did not change the results substantially. I settled on the average bearing, as this is more defensible and easier to explain.

```{r PREP_AND_MODEL_comparing_bearing_measures, eval = TRUE}

# * PREPARE THE DATA * -------

# Set parameters
spec_col <- c("bba_birdis", "bba_kerguelen", "waal_birdis", "waal_crozet")

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

for (i in 1:length(spec_col)) {
  
  if (grepl("ker", spec_col[i])) {
    col_coords <- ker_coords
  } else if (grepl("birdis", spec_col[i])) {
    col_coords <- bi_coords
  } else if (grepl("cro", spec_col[i])) {
    col_coords <- cro_coords }
  
  # Load GPS data
  gps_files <- list.files("Data_inputs/", pattern = paste0(spec_col[i], "_gps"))
  my_file <- gps_files[grepl("labelled", gps_files)]
  
  my_gps <- loadRData(paste0("Data_inputs/", my_file)) 
  
  my_gps %<>%
    mutate(ID = paste0(ring, "_", season, "_", boutID),
           season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                           as.numeric(format(as.Date(datetime), "%Y")),
                           as.numeric(format(as.Date(datetime), "%Y")) + 1 )) %>%
    filter(loc == "trip")
  
  ## Correct issues with breeding phase
  trips_to_fix <- my_gps %>%
    group_by(tripID) %>%
    summarise(n_phases = n_distinct(phase)) %>%
    filter(n_phases > 1) %>%
    pull(tripID)
  
  my_gps %<>%
        mutate(phase = if_else(tripID %in% trips_to_fix & phase == "brooding", "incubation", phase))

  # Calculate bearing measures -------
  
  ## Average bearing -----
  trip_bearings <- my_gps %>% 
     mutate(bearing_calc = bearing(
                p1 = matrix(c(col_coords$longitude, col_coords$latitude), ncol = 2, byrow = TRUE),
                p2 = matrix(c(longitude, latitude), ncol = 2) )) %>%
    group_by(ID) %>%
    mutate(bearing_col = mean(bearing_calc),
           species = strsplit(spec_col[i], "_")[[1]][1],
           colony = strsplit(spec_col[i], "_")[[1]][2])
  
  ## Distal bearing -----
  
  ## Nest the data by ID
  nested_trips <- my_gps %>%
    filter(loc == "trip") %>%
    group_by(ID) %>%
    nest()

  ## Apply bearing function to each trip
  trip_bearings_OG <- nested_trips %>%
    mutate(bearing_OG = map_dbl(data, ~ bearing_from_col(.x, col_coords$longitude, col_coords$latitude))) %>%
    select(ID, bearing_OG)

  ## Average foraging bearing
  load(paste0("Data_outputs/", strsplit(spec_col[i], "_")[[1]][1], "_", colony = strsplit(spec_col[i], "_")[[1]][2], "_labelledHMM.RData"))

  my_foraging <- my_gps %>% filter(State == "Search")
  my_foraging %<>%
     mutate(bearing_for = bearing(
                p1 = matrix(c(col_coords$longitude, col_coords$latitude), ncol = 2, byrow = TRUE),
                p2 = matrix(c(longitude, latitude), ncol = 2) )) %>%
    group_by(ID) %>%
    mutate(bearing_for = mean(bearing_for, na.rm = T)) %>%
    select(ID, bearing_for) %>%
    distinct()

  
  # Bind the options together
  trip_bearings <- merge(trip_bearings, trip_bearings_OG, by = "ID", all.x = T)
  trip_bearings.all <- merge(trip_bearings, my_foraging, by = "ID", all.x = T)
  
  # OPTIONAL: Plot the bearings -----  
  # IDs <- unique(trip_bearings$ID)
  # 
  # for (j in 1:50) {
  #   
  #   plotdat <- subset(trip_bearings, ID == IDs[j]) %>% arrange(datetime)
  #   
  #   # Convert bearings to radians
  #   bearing_col_rad <- plotdat$bearing_col[1] * pi / 180
  #   bearing_OG_rad  <- plotdat$bearing_OG[1]  * pi / 180
  #   bearing_for_rad <- plotdat$bearing_for[1] * pi / 180  # new line
  #   
  #   # Start point (first GPS location)
  #   start_lon <- plotdat$longitude[1]
  #   start_lat <- plotdat$latitude[1]
  #   
  #   # Calculate dynamic arrow length
  #   lon_range <- range(plotdat$longitude, na.rm = TRUE)
  #   lat_range <- range(plotdat$latitude, na.rm = TRUE)
  #   diag_deg <- sqrt(diff(lon_range)^2 + diff(lat_range)^2)
  #   arrow_length <- diag_deg * 0.2
  #   
  #   # End point for bearing_col (red arrow)
  #   end_lon_col <- start_lon + arrow_length * sin(bearing_col_rad) / cos(start_lat * pi / 180)
  #   end_lat_col <- start_lat + arrow_length * cos(bearing_col_rad)
  #   
  #   # End point for bearing_OG (green arrow)
  #   end_lon_OG <- start_lon + arrow_length * sin(bearing_OG_rad) / cos(start_lat * pi / 180)
  #   end_lat_OG <- start_lat + arrow_length * cos(bearing_OG_rad)
  #   
  #   # End point for bearing_for (blue arrow) - new lines
  #   end_lon_for <- start_lon + arrow_length * sin(bearing_for_rad) / cos(start_lat * pi / 180)
  #   end_lat_for <- start_lat + arrow_length * cos(bearing_for_rad)
  #   
  #   # Plot
  #   p <- ggplot(plotdat, aes(x = longitude, y = latitude)) +
  #     geom_path() +
  #     geom_segment(aes(x = start_lon, y = start_lat, xend = end_lon_col, yend = end_lat_col),
  #                  arrow = arrow(type = "closed", length = unit(0.3, "cm")),
  #                  colour = "red", linewidth = 1) +
  #     geom_segment(aes(x = start_lon, y = start_lat, xend = end_lon_OG, yend = end_lat_OG),
  #                  arrow = arrow(type = "closed", length = unit(0.3, "cm")),
  #                  colour = "forestgreen", linewidth = 1) +
  #     geom_segment(aes(x = start_lon, y = start_lat, xend = end_lon_for, yend = end_lat_for),  
  #                  arrow = arrow(type = "closed", length = unit(0.3, "cm")),
  #                  colour = "blue", linewidth = 1) +
  #     coord_fixed() +
  #     labs(title = plotdat$ID[1],
  #          caption = "Red = average bearing; green = distal bearing; blue = foraging bearing.") +
  #     theme_minimal()
  # 
  #   filename <- paste0("Figures/Bearings/", plotdat$species[1], "_", plotdat$colony[1], "_", plotdat$ID[1], ".png")
  #   
  #   png(filename = filename, width = 6, height = 5, units = "in", res = 300) 
  #   print(p)
  #   dev.off()
  # 
  # }
  
  # Isolate bearing per trip
  trip_bearings <- trip_bearings.all %<>% 
    mutate(boutID = paste0(season, "_", boutID)) %>%
    select(c(ID, ring, season, phase, boutID, bearing_col, bearing_OG, bearing_for)) %>%
    distinct()
  
  trip_bearings$species <- strsplit(spec_col[i], "_")[[1]][1]
  trip_bearings$colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  # Compute the bearing differences ----------
  
  # Within-individual bearings #
  
  # Compute differences and remove any that occur in different breeding phases
  col_within  <- calculate_within_bearing_diffs.comparison(trip_bearings, "bearing_col", "bearing_col.diff")
  OG_within   <- calculate_within_bearing_diffs.comparison(trip_bearings, "bearing_OG", "bearing_OG.diff")
  for_within  <- calculate_within_bearing_diffs.comparison(trip_bearings, "bearing_for", "bearing_for.diff")

  bearings_within <- reduce(
  list(col_within, OG_within, for_within),
  ~ left_join(.x, .y, by = c("ring_1", "tripID_1", "season_1", 
                             "ring_2", "tripID_2", "season_2", 
                             "comp_ind", "comp_season")) ) %>%
    relocate(c(bearing_OG.diff, bearing_for.diff), .before = comp_ind)
  
  # Among-individual differences #
  
  # Make the dataframe of possible comparisons to start
  bearings_between.df <- expand.grid(
    ID1 = unique(trip_bearings$ID),
    ID2 = unique(trip_bearings$ID)) %>%
    separate(ID1, into = c("ring_1", "season_1", "dummy1", "tripID_1"), sep = "_", remove = F) %>%
    separate(ID2, into = c("ring_2", "season_2", "dummy2", "tripID_2"), sep = "_", remove = F) %>%
    select(-c(dummy1, dummy2)) %>%
    mutate(tripID_1 = paste0(season_1, "_trip_", tripID_1),
           tripID_2 = paste0(season_2, "_trip_", tripID_2),
           # Remove duplicate comparisons
           tripID_1.tmp = paste0(ring_1, tripID_1),
           tripID_2.tmp = paste0(ring_2, tripID_2),
           comp_ID = paste0(pmin(tripID_1.tmp, tripID_2.tmp), "_", pmax(tripID_1.tmp, tripID_2.tmp))) %>%
    distinct(comp_ID, .keep_all = TRUE) %>%
    select(-c(comp_ID, tripID_1.tmp, tripID_2.tmp)) %>%
    filter(ring_1 != ring_2) %>%
    # Get phase back and filter for same breeding phase
    left_join(trip_bearings %>% select(ring, boutID, phase, bearing_col) %>%
                rename(ring_1 = ring, tripID_1 = boutID),
              by = c("ring_1", "tripID_1")) %>%
    rename(phase_1 = phase, bearing_1 = bearing_col) %>%
    left_join(trip_bearings %>% select(ring, boutID, phase, bearing_col) %>%
                rename(ring_2 = ring, tripID_2 = boutID),
              by = c("ring_2", "tripID_2")) %>%
    rename(phase_2 = phase, bearing_2 = bearing_col) %>%
    mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
    filter(phase_comp == "same_phase") %>%
    select(-c(phase_comp, phase_1, phase_2))
  
  
  bearings_between.df <- calculate_between_bearing_diffs.comparison(trip_bearings)

  # Combine results #
  bearings_between <- bearings_between.df %>% select(c(ring_1, tripID_1, season_1, ring_2, tripID_2, season_2, bearing_col.diff, bearing_OG.diff, bearing_for.diff)) %>% mutate(comp_ind = "between_ring", comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))
  
  bearing_differences <- rbind(bearings_within, bearings_between)
  
  save(bearing_differences, file = paste0("Data_outputs/", spec_col[i], "_bearing_differences_comparisons.RData"))

}


# * MODEL THE DIFFERENCES * -----

# Output containers
bearings_summary.list <- list()
bearings_means.list <- list()

for (i in seq_along(spec_col)) {
  
  # Load data and prep species/colony labels
  load(paste0("Data_outputs/", spec_col[i], "_bearing_differences_comparisons.RData"))
  
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  meta_info <- tibble(species, colony)
  
  # Make dyad ID and subsample to ensure each comparison appears only once (circumvents random effects issues)
  set.seed(817) 
  
  bearing_differences %<>%
    mutate(ring_season_1 = paste(ring_1, season_1, sep = "_"),
           ring_season_2 = paste(ring_2, season_2, sep = "_"),
           dyad = if_else(ring_season_1 < ring_season_2,
                          paste(ring_season_1, ring_season_2, sep = "_"),
                          paste(ring_season_2, ring_season_1, sep = "_"))) %>%
    group_by(dyad) %>%
    slice_sample(n = 1) %>%
    ungroup() %>%
    select(-c(ring_season_1, ring_season_2))
  
  ## Relevel factors for comparison
  bearing_differences %<>% mutate(comp_ind = as.factor(comp_ind), comp_season = as.factor(comp_season))
  
  bearing_differences$comp_ind <- relevel(bearing_differences$comp_ind, ref = "within_ring")
  bearing_differences$comp_season <- relevel(bearing_differences$comp_season, ref = "within_season")

  
  ## LOOP THROUGH DIFFERENT BEARING OPTIONS
  
  # Define bearing types
  bearing_vars <- c("bearing_col.diff", "bearing_OG.diff", "bearing_for.diff")
  
  # Create empty lists to store results
  model_list <- list()
  emmeans_list <- list()
  model_output_list <- list()
  
  # Loop over each bearing type
  for (bearing_var in bearing_vars) {
    
    # Build model
    m <- lm(as.formula(paste0(bearing_var, " ~ comp_season * comp_ind")), data = bearing_differences)
    
    # Store model
    model_list[[bearing_var]] <- m
    
    # Extract estimated marginal means
    em <- data.frame(emmeans(m, ~comp_ind * comp_season, pbkrtest.limit = 7000)) %>%
      mutate(
        species = species,
        colony = colony,
        bearing_type = bearing_var,
        comp_ind = factor(comp_ind, levels = c("within_ring", "between_ring"))
      ) %>%
      relocate(c(species, colony, bearing_type), .before = comp_ind) %>%
      arrange(comp_ind)
    
    emmeans_list[[bearing_var]] <- em
    
    # Extract model summary output
    out <- extract_output.lm(m) %>%
      mutate(
        species = species,
        colony = colony,
        bearing_type = bearing_var
      ) %>%
      relocate(c(species, colony, bearing_type), .before = Term)
    
    model_output_list[[bearing_var]] <- out
  }

  all_emmeans <- bind_rows(emmeans_list)
  all_model_outputs <- bind_rows(model_output_list)
  
  # Store results
  bearings_summary.list[[i]] <- all_model_outputs
  bearings_means.list[[i]] <- all_emmeans
  
}

# Combine and save
bearings_summary <- bind_rows(bearings_summary.list)
bearings_means   <- bind_rows(bearings_means.list)

save(bearings_summary, file = "Data_outputs/trajectory_models_comparison_summary.RData")
save(bearings_means, file = "Data_outputs/trajectory_comparison_means.RData")


```

### Plot of bearing differences for each metric

For most of the species/colony combinations, which metric is used makes little difference to the overall results. The exception is for WAAL at Bird Island, which seem to be a bit of a mess regardless of the metric used. Possibly this is because I have few repeats for this population, or they are simply not very consistent by any measure. Based on these results, intepretation, and interpretability, I settled on the *average bearing* measure.

```{r PLOT_bearing_differences_comparison, fig.dim = c(16, 12)}

load("Data_outputs/trajectory_comparison_means.RData")

bearings_plot <- bearings_means %>%
  mutate(
    species = case_when(
      species == "bba" ~ "BBAL",
      species == "waal" ~ "WAAL",
      TRUE ~ species
    ),
    colony = case_when(
      colony == "birdis" ~ "Bird Island",
      colony == "kerguelen" ~ "Kerguelen",
      colony == "crozet" ~ "Crozet",
      TRUE ~ colony
    ),
    spec_col = paste(species, colony, sep = ", "),
    comp = case_when(
      comp_season == "within_season" & comp_ind == "within_ring" ~ "Within-individual, within season",
      comp_season == "between_season" & comp_ind == "within_ring" ~ "Within-individual, across seasons",
      comp_season == "within_season" & comp_ind == "between_ring" ~ "Among-individual, within seasons",
      comp_season == "between_season" & comp_ind == "between_ring" ~ "Among-individual, across season",
      TRUE ~ paste(comp_season, comp_ind, sep = ", ")
    )
  )

comparison_levels <- c(
  "Within-individual, within season",
  "Within-individual, across seasons",
  "Among-individual, within seasons",
  "Among-individual, across season"
)

bearings_plot$comp <- factor(bearings_plot$comp, levels = comparison_levels)

# Plot
#png(filename = "Figures/bearings_comparison_diffMeasures.png", width = 16, height = 12, units = "in", res = 900)
ggplot(bearings_plot, aes(x = comp_ind, y = emmean, colour = comp_season)) +
  geom_point() +
  geom_errorbar(aes(ymin = lower.CL, ymax = upper.CL), width = 0.1) +
  geom_line(aes(group = comp_season)) +
  labs(
    x = "", 
    colour = "", 
    y = "Angular bearing differences",
    caption = "Point and line range indicate mean and 95% confidence intervals."
  ) +
  facet_wrap(~bearing_type + spec_col, scales = "free") +
  scale_colour_viridis_d(labels = c("Within season", "Across seasons")) +
  scale_x_discrete(labels = c("Within individual", "Among individual")) +
  theme_bw() +
  theme(
    legend.position = c(0.35, 0.62),
    legend.background = element_rect(fill = "transparent", colour = NA),
    legend.key = element_rect(fill = "transparent", colour = NA)  )
#dev.off()

```
