---
title: "Spatial repeatability analysis"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

In this part of the project, I am looking at **spatial repeatability** in birds' foraging tracks. I have isolated GPS tracks that represent foraging trips, and I am comparing the consistency in individuals' trips through two methods:

1.  Comparing the difference in trip trajectories within- versus between- individuals. *Trip trajectory* in this context is the bearing between the colony and the distal point in the forage trip - this has previously been found to be a good representative of birds' bearings over the trip as a whole.

2.  Comparing the dissimilarity between pairs of trips using *Dynamic time warping*. This quantifies the dissimilarity between two spatial tracks: higher values indicate greater dissimilarity.

In all cases, I am comparing individuals within the same breeding phase (i.e. incubation or chick-rearing).

I started developing these methods using *Kerguelen black-browed albatrosses*. Interrogation of the results for the later colonies comes at the end of the document.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, include = FALSE)
# Load functions, packages, & data ---------------------------------------------

# Functions for GPS processing and plotting
source("RPT_functions.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "ggpubr", "tidyr", "sf", "ecmwfr", "ggridges", "rnaturalearth", "DHARMa","geosphere", "purrr", "lmerTest", "sp", "dtw") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
# if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

```

```{r load_gps_data}

# Set parameters
colony <- "ker" # cro bi ker
my_species <- "bba" # waal bba

colony_exp <- ifelse(colony == "ker", "kerguelen",
                     ifelse(colony == "cro", "crozet",
                            "birdis"))

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

if (colony == "ker") {
  col_coords <- ker_coords
} else if (colony == "bi") {
  col_coords <- bi_coords
} else if (colony == "cro") {
  col_coords <- cro_coords }


# Load the data -----------------------------------------------------------

gps_files <- list.files("Data_inputs/", pattern = paste0(my_species, "_", colony_exp, "_gps"))
my_file <- gps_files[grepl("labelled", gps_files)]

my_gps <- loadRData(paste0("Data_inputs/", my_file)) 

my_trip_meta <- my_gps %>% select(c(ring, season, datetime, dist_col, phase, boutID)) %>%
  group_by(ring, season) %>% mutate(max_range = max(dist_col, na.rm = T)) %>% ungroup()

my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))

## Filter out individuals with only one trip
my_gps %<>% 
  filter(loc == "trip") %>%
  group_by(ring, season) %>%
  mutate(n_trips = n_distinct(ID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Make a metadata dataframe
my_meta <- my_gps %>%
  group_by(ID) %>%
  select(c(ring, season, boutID, duration.mins, duration.hours, ID, phase)) %>%
  distinct()


### Calculate bearing differences ------

# Calculate bearings from colony to distal point 

## Nest the data by ID
nested_trips <- my_gps %>%
  filter(loc == "trip") %>%
  group_by(ID) %>%
  nest()

## Apply bearing function to each trip
trip_bearings <- nested_trips %>%
  mutate(bearing_col = map_dbl(data, ~ bearing_from_col(.x, col_coords$longitude, col_coords$latitude))) %>%
  select(ID, bearing_col) 

## Put metadata back in
trip_bearings <- merge(trip_bearings, my_meta, by = "ID", all.x = T)

# Compute the bearing differences ----------

# Within-individual bearings #

# Make sure boutIDs are unique to season
trip_bearings %<>% mutate(boutID = paste0(season, "_", boutID))

# Compute differences and remove any that occur in different breeding phases
bearings_within <- trip_bearings %>%
  group_by(ring) %>%
  summarise(trip_comparisons = list(combn(boutID, 2, paste, collapse = " vs ")),  # Store trip comparisons
            bearing_diff = list(combn(bearing_col, 2, function(x) angle_diff(x[1], x[2]))),
            .groups = "drop") %>%
  unnest(cols = c(trip_comparisons, bearing_diff)) %>%
  separate(trip_comparisons, into = c("tripID_1", "tripID_2"), sep = " vs ") %>%
  left_join(trip_bearings %>% select(ring, boutID, season, phase), 
            by = c("ring", "tripID_1" = "boutID")) %>%
  rename(season_1 = season, phase_1 = phase) %>%
  left_join(trip_bearings %>% select(ring, boutID, season, phase), 
            by = c("ring", "tripID_2" = "boutID")) %>%
  rename(season_2 = season, phase_2 = phase) %>%
  mutate(comp_ind = "within_ring",
         comp_season = ifelse(season_1 == season_2, "within_season", "between_season"),
         comp_phase = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
  arrange(ring, season_1, season_2) %>%
  filter(comp_phase == "same_phase") %>%
  select(-c(phase_1, phase_2, comp_phase)) %>%
  # Add ring 2 for compatability with larger df
  mutate(ring_2 = ring) %>%
  rename(ring_1 = ring) %>%
  relocate(ring_1, tripID_1, season_1, ring_2, tripID_2, season_2, bearing_diff, comp_ind, comp_season) %>%
  # Remove duplicate comparisons
  mutate(tripID_1.tmp = paste0(ring_1, tripID_1),
         tripID_2.tmp = paste0(ring_2, tripID_2),
         comp_ID = paste0(pmin(tripID_1.tmp, tripID_2.tmp), "_", pmax(tripID_1.tmp, tripID_2.tmp))) %>%
  distinct(comp_ID, .keep_all = TRUE) %>%
  select(-c(comp_ID, tripID_1.tmp, tripID_2.tmp)) 



# Among-individual differences #

# Make the dataframe of possible comparisons to start
bearings_between.df <- expand.grid(
  ID1 = unique(trip_bearings$ID),
  ID2 = unique(trip_bearings$ID)) %>%
  separate(ID1, into = c("ring_1", "season_1", "dummy1", "tripID_1"), sep = "_", remove = F) %>%
  separate(ID2, into = c("ring_2", "season_2", "dummy2", "tripID_2"), sep = "_", remove = F) %>%
  select(-c(dummy1, dummy2)) %>%
  mutate(tripID_1 = paste0(season_1, "_trip_", tripID_1),
         tripID_2 = paste0(season_2, "_trip_", tripID_2),
         # Remove duplicate comparisons
         tripID_1.tmp = paste0(ring_1, tripID_1),
         tripID_2.tmp = paste0(ring_2, tripID_2),
         comp_ID = paste0(pmin(tripID_1.tmp, tripID_2.tmp), "_", pmax(tripID_1.tmp, tripID_2.tmp))) %>%
  distinct(comp_ID, .keep_all = TRUE) %>%
  select(-c(comp_ID, tripID_1.tmp, tripID_2.tmp)) %>%
  filter(ring_1 != ring_2) %>%
  # Get phase back and filter for same breeding phase
  left_join(trip_bearings %>% select(ring, boutID, phase, bearing_col) %>% 
              rename(ring_1 = ring, tripID_1 = boutID),
            by = c("ring_1", "tripID_1")) %>%
  rename(phase_1 = phase, bearing_1 = bearing_col) %>%
  left_join(trip_bearings %>% select(ring, boutID, phase, bearing_col) %>% 
              rename(ring_2 = ring, tripID_2 = boutID),
            by = c("ring_2", "tripID_2")) %>%
  rename(phase_2 = phase, bearing_2 = bearing_col) %>%
  mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
  filter(phase_comp == "same_phase") %>%
  select(-c(phase_comp, phase_1, phase_2))


## Downsample the comparison data and calculate the differences
set.seed(817)
n = 10

bearings_between <- bearings_between.df %>% 
  sample_n(size = n * nrow(bearings_within), replace = FALSE)

## Calculate comparisons
bearings_between %<>% 
  mutate(bearing_diff = angle_diff(bearing_1, bearing_2),
         comp_ind = "between_ring",
         comp_season = ifelse(season_1 == season_2, "within_season", "between_season")) %>%
  select(c(ring_1, tripID_1, season_1, ring_2, tripID_2, season_2, bearing_diff, comp_ind, comp_season)) %>%
  relocate(ring_1, tripID_1, season_1, ring_2, tripID_2, season_2, bearing_diff, comp_ind, comp_season)

# Combine results #
bearing_differences <- rbind(bearings_within, bearings_between)

save(bearing_differences, file = paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

```

## Approach 1: Trip trajectories

I'm starting by comparing trip trajectories. For this, I calculated the bearing between the colony and the distal point of each trip, and then calculated the difference between these bearings within and between individuals, and also within and between years.

First, let's visualise the GPS tracks from the colony and the trip bearings across the entire sample (not the prettiest plots, but I'll work on it!).

```{r GPS_plotting, include = FALSE}
# Plot the GPS tracks -----------------------------------------------------

# Set projections (use same for Ker + Cro)
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

if(colony == "bi") {
  proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
} else {
  proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
}

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
world2 <- sf::st_transform(world, crs = proj.utm)

# Project GPS data
my_gps.proj <- my_gps %>% arrange(ringYr, datetime)
coordinates(my_gps.proj) <- ~ longitude+latitude
proj4string(my_gps.proj) <- proj.dec

my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
my_gps.proj.df <- as.data.frame(my_gps.proj.sp)


# Build the plot

## Set min/max limits
y_min = min(my_gps.proj.df$coords.x2 - 10000)
y_max = max(my_gps.proj.df$coords.x2 + 10000)
x_min = min(my_gps.proj.df$coords.x1 - 10000)
x_max = max(my_gps.proj.df$coords.x1 + 10000)


gps_tracks.plot <- ggplot(data = world2) + 
  geom_sf(fill = "cadetblue", colour = "cadetblue") +
  coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
           label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
  geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
            dat = my_gps.proj.df, group = "identity") +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
  scale_colour_viridis_d() +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "grey80"),
        panel.border = element_rect(colour = "black", fill = NA),
        axis.text.x.bottom = element_blank(),
        axis.title.x.bottom = element_blank(),
        axis.text.y.right = element_blank(),
        axis.title.y.right = element_blank(),
        axis.title.y.left = element_blank(),
        axis.text.y.left = element_text(size = 10),
        axis.text.x.top = element_text(size = 10),
        legend.position = "none")

```

```{r Bearings_plotting, include = FALSE}
### Plot colony bearings in circular plot -----------------------------------

trip_bearings %<>% mutate(bearing_360 = ifelse(bearing_col < 0, bearing_col + 360, bearing_col))

breaks = seq(0, 360, 60)

bearings.plot <- ggplot(trip_bearings, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks,
                     labels = paste0(breaks, "°")) +
  labs(y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))

```

```{r Plot_tracks_bearings, echo = FALSE}
ggarrange(gps_tracks.plot, bearings.plot, ncol = 2)
```

It's important to note here that birds' trajectories tend to clump in either a NE or a SW direction from the colony. This has implications later on but also explains some of the histograms of my results.

After calculating my trajectories, I then compute the **absolute differences between pairs of bearings**. I do this (i) within individual trajectories, (ii) between individual trajectories, (iii) within seasons, and (iv) between seasons. When I compute these differences, the variable is no longer circular but is rather a linear measure of angular difference that ranges from 0 to 180°. I can therefore analyse within- and between- individual/season differences using a **linear approach**.

Obviously, if I compare every individual with every individual then my comparative approach becomes really unwieldy. To account for this, I downsampled the between-individual comparisons to be 10x the size of the within-individual comparisons.

### Within-individual, seasonal differences

First, I want to consider whether or not individuals' trajectories differ across seasons. To do this, I'm isolating data where I extract only the within-individual comparisons (i.e. compare trajectories for different trips of the same individual) and examine whether they differ within- and between-seasons. I am fitting a random intercept for ring_1 (I assume throughout that the first ring is the comparative individual, though of course here ring_1 and ring_2 are identical) as I expect some individuals may be more similar than others.

**I have also included a random slope here**. This is to allow individuals to be more or less similar between seasons - i.e. an individual that doesn't differ between seasons would have a flat slope, one that differed a lot would have a steep slope.

```{r bearings_within_individual_model, results='hide'}
bearings_within.lmm <- lmer(bearing_diff ~ comp_season + (1+comp_season|ring_1), data = bearing_differences %>% filter(comp_ind == "within_ring"))
```

```{r bearings_within_individual_setup, include = FALSE}

## Extract mean differences from lmm
bearings_within.summary <- summary(bearings_within.lmm)

bearings_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = c(bearings_within.summary$coefficients[1,1], bearings_within.summary$coefficients[1,1] + bearings_within.summary$coefficients[2,1]),
                         sd = c(bearings_within.summary$coefficients[1,2], bearings_within.summary$coefficients[2,2]))

# Plot #
within_ind.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_ind == "within_ring"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_within.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_within.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within Season",
    "between_season" = "Between Season" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


```

```{r bearings_within_individual_output, echo = FALSE}
summary(bearings_within.lmm)

within_ind.plot
```

Overall, it appears that birds' trajectories are as similar within as between seasons (t = `r signif(bearings_within.summary$coefficients[2,4], 2)`, p =`r signif(bearings_within.summary$coefficients[2,5], 2)`. This is also well-illustrated in the histogram comparison.

#### Random effects

```{r bearings_within_individual_reaptability, include = FALSE}

# Calculate individual repeatability
bearings_within.varcor <- as.data.frame(VarCorr(bearings_within.lmm))

ind_rpt.intercept <- bearings_within.varcor$vcov[1] / (bearings_within.varcor$vcov[1] + bearings_within.varcor$vcov[4])

ind_rpt.slope <- bearings_within.varcor$vcov[2] / (bearings_within.varcor$vcov[2] + bearings_within.varcor$vcov[4])


```

There is evidence for individual repeatability in both slope and intercept. Repeatability in slope is `r signif(ind_rpt.slope, 2)` while repeatability in intercept is `signif(ind_rpt.intercept, 2)` .

The high repeatability in the intercept suggests that about `paste0(signif(ind_rpt.intercept*100, 2),"%")` of the variation in baseline is attributable to variation among individuals. In other words, individuals are moderately consistent in how different their trip trajectories are.

The slope repeatability is fairly low in comparison, and suggests that the way that bearing differences vary within- versus between-season is not consistent across individuals.

```{r bearings_within_individual_random_effects, include = FALSE}

within_individual.random <- ranef(bearings_within.lmm)$ring_1

within_individual.slopes <- within_individual.random$comp_seasonwithin_season
within_individual.intercept <- within_individual.random$`(Intercept)`

```

### Between-individual differences

Next, I want to know whether individuals have more similar trajectories than you would expect by chance. To do this, I will compare their trajectories to different individuals, both wtihin- and between- seasons.

#### Between-seasons

First, I used a histogram to visualise within- and between-individual differences in bearings. Interestingly, there is a clear bimodal distribution with birds tending to take trajectories that are very similar to one another (0-20°) or opposite to one another (160-180°). I suspect this is because birds' trip trajectories tend to cluster towards the NE or SW; i.e. they don't vary continuously around the range of possible departure locations. **This is an important point as it may change how we model these differences, and it may influence how we consider habitat selection.**

```{r bearings_bearing_differences_plot_between, echo = FALSE}
ggplot() + 
  geom_histogram(data = bearing_differences, aes(x = bearing_diff)) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0") +
  theme_bw()
```

Next, I fit the model using the same structure as for the within-individual analysis. Due to the down-sampling of my dataset, I end up with not having many repeats per individual, and so I have dropped the random effect for ring (if I try to fit with a random intercept, I get a singular boundary and 0 variance for ring).

```{r bearings_between_individuals_model}
bearings_between.lm <- lm(bearing_diff ~ comp_ind, data = bearing_differences)
```

```{r bearings_between_individuals_setup, include = FALSE}

bearings_between.summary <- summary(bearings_between.lm)

# Mean bearing differences
bearings_between.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                             mean = c(bearings_between.summary$coefficients[1,1], bearings_between.summary$coefficients[1,1] + bearings_between.summary$coefficients[2,1]),
                             sd = c(bearings_between.summary$coefficients[1,2], bearings_between.summary$coefficients[2,2]))

# Plot #
bearings_between.plot <- ggplot() +
  geom_histogram(data = bearing_differences, aes(x = bearing_diff)) +
  geom_segment(data = bearings_between.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model.")  +
  theme_bw()

```

```{r bearings_between_individuals_results, echo = FALSE}
summary(bearings_between.lm)

bearings_between.plot

```

Here, I find good evidence that birds' trajectories on their own trips are more similar than to those of other birds (t = `r signif(bearings_between.summary$coefficients[2,3], 2)`, p =`r signif(bearings_within.summary$coefficients[2,4], 2)`. This suggests that birds go on trips with similar trajectories.

#### Within-seasons

Finally, I looked at similarities in birds' trips that fell within the same season, to ensure that differences in trip trajectories could not be explained by seasonal effects. As it's possible that we would have a seasonal effect e.g. if in one year all birds went to the same place, in this model I included a **random intercept for season** (i.e., year): this will allow the average similarity in a given year to vary.

```{r bearings_between_individuals_season_model}
bearings_between_season.lmm <- lmer(bearing_diff ~ comp_ind + (1|season_1), data = bearing_differences %>% filter(comp_season == "within_season"))

```

```{r bearings_between_individuals_season_setup, include = FALSE}

## Extract mean differences from lmm
bearings_between_season.summary <- summary(bearings_between_season.lmm)

bearings_between_season.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                         mean = c(bearings_between_season.summary$coefficients[1,1], bearings_between_season.summary$coefficients[1,1] + bearings_between_season.summary$coefficients[2,1]),
                         sd = c(bearings_between_season.summary$coefficients[1,2], bearings_between_season.summary$coefficients[2,2]))

# Plot #
bearings_between_season.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_season == "within_season"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_between_season.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between_season.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between_season.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


```

```{r bearings_between_individuals_season_results, echo = FALSE}
summary(bearings_between_season.lmm)

bearings_between_season.plot
```

### Summary

Overall, I think we can conclude the following:

1.  Individuals take more similar trip trajectories than would be expected by chance.

    -   This holds even when considering only trips that occur within the same season.

2.  Individuals do not take more similar trips within seasons than between seasons.

3.  The degree to which individual trip trajectories vary is repeatable (by around `paste0(signif(ind_rpt.intercept*100, 2),"%")` ).

## Approach 2: Dynamic time warping

*Quick note*: originally I tried the Earth mover's distance (EMD): this calculates the effort required to transform one image to another. In the context of GPS tracks, it quantifies the work needed to transpose one track to another. Therefore, a low EMD indicates high similarity between tracks, and a high EMD indicates low similarity. However, **this was extremely computationally intensive**, especially for long tracks.

As a result, I switched to an alternative method called **dynamic time warping**. This is much faster, and also retains the sequential structure of the track so is probably better for movement data. There is even a clustering package available called 'dtwclust' if I need to speed things up more quickly.

Briefly, DTW recursively compares all point combinations between two trajectories and identifies the smallest path between them. This does make it sensitive to noise, as it will also compare outliers, but this is the case for EMD too. The DTW value essentially measures the 'warping' needed to aling to two trajectories. As a result, a high DTW suggests low similarity, and a low DTW suggests high similarity.

There is a good write-up on DTW in [Cleasby et al. 2019, Behav Ecol Socio](https://link.springer.com/article/10.1007/s00265-019-2761-1#Sec18) that shows it performing well in comparison to several measures of trajectory similarity.

For this analysis, I'll take a similar modelling approach that used to the trip trajectories analysis, but using DTW as my response. As with the trip trajectories comparison, there are a very high number of possible comparisons, so I will down sample the dataset to n \* 10 where n is the number of within-individual comparisons.

```{r dtw_comparison_code_not-run, include = FALSE, eval = FALSE}

## The dtw comparison takes a little while so this code is not run here. 

### Calculate DTW -------------------------------

# Generate all possible trip combinations 
trip_combinations <- combn(unique(my_gps$ID), 2, simplify = FALSE)
trip_combinations.df <- data.frame(do.call("rbind", trip_combinations))
colnames(trip_combinations.df) <- c("tripID_1", "tripID_2")

# Within-individual dataset
dtw_within.df <- trip_combinations.df %>%
  separate(tripID_1, into = "ring_1", sep = "_", remove = FALSE) %>%
  separate(tripID_2, into = "ring_2", sep = "_", remove = FALSE) %>%
  filter(ring_1 == ring_2) %>%
  select(-c(ring_1, ring_2)) %>%
  # Ensure comparisons are within same phase
  left_join(my_meta %>% select(ID, phase) %>% 
            rename(tripID_1 = ID),
          by = "tripID_1") %>%
  rename(phase_1 = phase) %>%
  left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_2 = ID),
            by = "tripID_2") %>%
  rename(phase_2 = phase) %>%
  mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
  filter(phase_comp == "same_phase") %>%
  select(-c(phase_comp, phase_1, phase_2))

## Run through the comparisons
dtw_within.list <- vector(mode = "list", length = nrow(dtw_within.df))

for (i in 1:nrow(dtw_within.df)) {
  
  print(paste0("Processing iteration ", i, " of ", nrow(dtw_within.df), "."))
  
  trip1 <- convert_to_spdf(subset(my_gps, ID == dtw_within.df$tripID_1[i]), proj.dec)
  trip2 <- convert_to_spdf(subset(my_gps, ID == dtw_within.df$tripID_2[i]), proj.dec)
  
  trip1_mat <- as.matrix(trip1@coords)
  trip2_mat <- as.matrix(trip2@coords)
  
  dtw_out <- dtw(trip1_mat, trip2_mat, dist.method = "Euclidean")$distance
  
  output <- data.frame(trip1 = dtw_within.df$tripID_1[i],
                       trip2 = dtw_within.df$tripID_2[i],
                       dtw = dtw_out)

  dtw_within.list[[i]] <- output 
  
}

dtw_within.output <- do.call("rbind", dtw_within.list)

dtw_within.output %<>% 
  separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
  separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
  select(-loc) %>%
  mutate(tripID_1 = paste0("trip_", tripID_1),
         tripID_2 = paste0("trip_", tripID_2),
         comp_ind = "within_ring",
         comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))


# Between-individual dataset

dtw_between.df <- trip_combinations.df %>%
  separate(tripID_1, into = "ring_1", sep = "_", remove = FALSE) %>%
  separate(tripID_2, into = "ring_2", sep = "_", remove = FALSE) %>%
  filter(ring_1 != ring_2) %>%
  select(-c(ring_1, ring_2)) %>%
  # Get phase and filter for same breeding phase
  left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_1 = ID),
            by = "tripID_1") %>%
  rename(phase_1 = phase) %>%
  left_join(my_meta %>% select(ID, phase) %>% 
              rename(tripID_2 = ID),
            by = "tripID_2") %>%
  rename(phase_2 = phase) %>%
  mutate(phase_comp = ifelse(phase_1 == phase_2, "same_phase", "diff_phase")) %>%
  filter(phase_comp == "same_phase") %>%
  select(-c(phase_comp, phase_1, phase_2))


## Downsample the between-individual data
set.seed(817)
n = 10

dtw_between <- dtw_between.df %>% 
  sample_n(size = n * nrow(dtw_within.df), replace = FALSE)


## Calculate the comparisons
dtw_between.list <- vector(mode = "list", length = nrow(dtw_between))

for (i in 1:nrow(dtw_between)) {
  
  print(paste0("Processing iteration ", i, " of ", nrow(dtw_between), "."))
  
  trip1 <- convert_to_spdf(subset(my_gps, ID == dtw_between$tripID_1[i]), proj.dec)
  trip2 <- convert_to_spdf(subset(my_gps, ID == dtw_between$tripID_2[i]), proj.dec)
  
  trip1_mat <- as.matrix(trip1@coords)
  trip2_mat <- as.matrix(trip2@coords)
  
  dtw_out <- dtw(trip1_mat, trip2_mat, dist.method = "Euclidean")$distance
  
  output <- data.frame(trip1 = dtw_between$tripID_1[i],
                       trip2 = dtw_between$tripID_2[i],
                       dtw = dtw_out)
  
  dtw_between.list[[i]] <- output 
  
}

dtw_between.output <- do.call("rbind", dtw_between.list)

dtw_between.output %<>% 
  separate(trip1, into = c("ring_1", "season_1", "loc", "tripID_1"), sep = "_") %>%
  separate(trip2, into = c("ring_2", "season_2", "loc", "tripID_2"), sep = "_") %>%
  select(-loc) %>%
  mutate(tripID_1 = paste0("trip_", tripID_1),
         tripID_2 = paste0("trip_", tripID_2),
         comp_ind = "between_ring",
         comp_season = ifelse(season_1 == season_2, "within_season", "between_season"))


# Bind the two dataframes together
dtw.df <- rbind(dtw_within.output, dtw_between.output)

save(dtw.df, file = paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


```

### Within-individual, seasonal differences

The analysis methods here will be the same as for the trip trajectory comparison; that is: comparing within individual, between individual, and between individual but within seasons. There is a heavy right skew in the DTW distances, so to account for this I've log-transformed the response so it better approximates a normal distribution.

```{r dtw_transformation, echo = FALSE}
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))

par(mfrow=c(1,2))
hist(dtw.df$dtw)
hist(log(dtw.df$dtw))

```

```{r dtw_within_individual_model, results = 'hide'}
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))
  
dtw_within.loglmm <- lmer(log(dtw) ~ comp_season + (1+comp_season|ring_1), data = dtw.df %>% filter(comp_ind == "within_ring"))

```

```{r dtw_model_diagnostics, include = FALSE}

dtw_within.lmm <- lmer(dtw ~ comp_season + (1+comp_season|ring_1), data = dtw.df %>% filter(comp_ind == "within_ring"))

dtw_within.loglmm <- lmer(log(dtw) ~ comp_season + (1+comp_season|ring_1), data = dtw.df %>% filter(comp_ind == "within_ring")) 

library(glmmTMB)

dtw_within.glmm <- glmmTMB(dtw ~ comp_season + (1+comp_season|ring_1), family = Gamma(link = "log"), data = dtw.df %>% filter(comp_ind == "within_ring"))

plot(residuals(dtw_within.glmm))

# AIC & BIC comparisons: log LMMM has smallest values
AIC(dtw_within.lmm, dtw_within.loglmm, dtw_within.glmm) 
BIC(dtw_within.lmm, dtw_within.loglmm, dtw_within.glmm) 
```

```{r dtw_within_individual_analysis, echo = FALSE}

## Extract mean differences and confint from log LMM
dtw_within.summary <- summary(dtw_within.loglmm)

ci_between <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + fixef(dtw_within.loglmm)["comp_seasonwithin_season"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["comp_seasonwithin_season", "Std. Error"])

dtw_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = exp(c(dtw_within.summary$coefficients[1,1], dtw_within.summary$coefficients[1,1] + dtw_within.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_within_ind.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_ind == "within_ring"), aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within seasons",
    "between_season" = "Between seasons" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))

summary(dtw_within.loglmm)

dtw_within_ind.plot
```

Overall, it appears that birds' trajectories are as similar within as between seasons (t = `r signif(dtw_within.summary$coefficients[2,4], 2)`, p =`r signif(dtw_within.summary$coefficients[2,5], 2)`). This is also illustrated in the histogram comparison.

#### Random effects

```{r dtw_within_individual_reaptability, echo = FALSE}

# Calculate individual repeatability
dtw_within.varcor <- as.data.frame(VarCorr(dtw_within.loglmm))

ind_rpt.intercept <- dtw_within.varcor$vcov[1] / (dtw_within.varcor$vcov[1] + dtw_within.varcor$vcov[4])
ind_rpt.slope <- dtw_within.varcor$vcov[2] / (dtw_within.varcor$vcov[2] + dtw_within.varcor$vcov[4])

dtw_within.varcor

```

The repeatability for individual intercept is reasonably high at `r signif(ind_rpt.intercept, 2)` , while conversely the repeatability in slope is essentially 0(`signif(ind_rpt.slope, 2)`). This suggests that about `paste0(signif(ind_rpt.intercept*100, 2),"%")` of the variation in baseline DTW is attributable to variation among individuals. This probably means that individuals are consistent in how much their trips differ: i.e. some birds will always have high DTW differences, some will always be low. However, the amount they differ within- versus between-seasons is always the same.

### Between-individual differences

#### Between-seasons

The model structures remain the same for the between-individual differences comparisons.

```{r dtw_between_individual_model}
dtw_between.loglm <- lm(log(dtw) ~ comp_ind, data = dtw.df)
```

```{r dtw_between_individual, echo = FALSE}
dtw_between.summary <- summary(dtw_between.loglm)

ci_log <- confint(dtw_between.loglm)
ci_between <- exp(ci_log["(Intercept)",])
ci_within <- exp(ci_log["(Intercept)", ] + ci_log["comp_indwithin_ring", ])

dtw_between.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = c(exp(coef(dtw_between.loglm)["(Intercept)"]),
                                  exp(coef(dtw_between.loglm)["(Intercept)"] +    coef(dtw_between.loglm)["comp_indwithin_ring"])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_ind == "within_ring"), aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within Season",
    "between_season" = "Between Season" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


summary(dtw_between.loglm)
dtw_between.plot
```

Individuals have much more similar tracks to themselves than to other birds (t = `r signif(dtw_between.summary$coefficients[2,3], 2)`, p =`r signif(dtw_within.summary$coefficients[2,4], 2)`, suggesting that birds' foraging tracks better resemble their own tracks than would be expected by chance.

#### Within-seasons

```{r dtw_between_individuals_withinSeason_model}
dtw_between_withinseason.loglmm <- lmer(log(dtw) ~ comp_ind + (1|season_1), data = dtw.df %>% filter(comp_season == "within_season"))

```

```{r dtw_between_individuals_season_setup, echo = FALSE}

## Extract mean differences and confint from log LMM
dtw_between_withinseason.summary <- summary(dtw_between_withinseason.loglmm)

ci_between <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + fixef(dtw_between_withinseason.loglmm)["comp_indwithin_ring"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["comp_indwithin_ring", "Std. Error"])

dtw_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                         mean = exp(c(dtw_between_withinseason.summary$coefficients[1,1], dtw_between_withinseason.summary$coefficients[1,1] + dtw_between_withinseason.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between_withinseason.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_season == "within_season"), aes(x = dtw)) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "between_ring" = "Between individuals",
    "within_ring" = "Within individuals" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))

summary(dtw_between_withinseason.loglmm)
dtw_between_withinseason.plot

```

Again, individuals' trips are more similar to their own trips than would be expected by chance; this holds even when considering only trips that occurred within the same season.

### Summary

Overall, the results of the DTW comparison shows similar results as for the bearing differences: individuals have more similar tracks to themselves than to other birds and they are not particularly consistent within (versus between) seasons.

## Results for other species and colonies

In this section I will use the methods above to produce results for each of the other colonies and species, and interpret these.

### BBAL at Bird Island

(1) Look at the GPS tracks and the bearings to the distal points of the colony.

```{r bba_bi_setup, include = FALSE}
# Each of the outputs have been run separately in the main R script, so here I'm just calling in the outputs.

# Set parameters
colony <- "bi" # cro bi ker
my_species <- "bba" # waal bba

colony_exp <- ifelse(colony == "ker", "kerguelen",
                     ifelse(colony == "cro", "crozet",
                            "birdis"))

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

if (colony == "ker") {
  col_coords <- ker_coords
} else if (colony == "bi") {
  col_coords <- bi_coords
} else if (colony == "cro") {
  col_coords <- cro_coords }


# Load the data -----------------------------------------------------------

gps_files <- list.files("Data_inputs/", pattern = paste0(my_species, "_", colony_exp, "_gps"))
my_file <- gps_files[grepl("labelled", gps_files)]

my_gps <- loadRData(paste0("Data_inputs/", my_file)) 

my_trip_meta <- my_gps %>% select(c(ring, season, datetime, dist_col, phase, boutID)) %>%
  group_by(ring, season) %>% mutate(max_range = max(dist_col, na.rm = T)) %>% ungroup()

my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))

## Filter out individuals with only one trip
my_gps %<>% 
  filter(loc == "trip") %>%
  group_by(ring, season) %>%
  mutate(n_trips = n_distinct(ID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Make a metadata dataframe
my_meta <- my_gps %>%
  group_by(ID) %>%
  select(c(ring, season, boutID, duration.mins, duration.hours, ID, phase)) %>%
  distinct()


### Calculate bearing differences ------

# Calculate bearings from colony to distal point 

## Nest the data by ID
nested_trips <- my_gps %>%
  filter(loc == "trip") %>%
  group_by(ID) %>%
  nest()

## Apply bearing function to each trip
trip_bearings <- nested_trips %>%
  mutate(bearing_col = map_dbl(data, ~ bearing_from_col(.x, col_coords$longitude, col_coords$latitude))) %>%
  select(ID, bearing_col) 

## Put metadata back in
trip_bearings <- merge(trip_bearings, my_meta, by = "ID", all.x = T)

# Load the bearing differences ----------
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

# Load the DTW values ---------
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


```

```{r bba_bi_build_plots, echo = FALSE}

# Plot bearings -----------------------------------

trip_bearings %<>% mutate(bearing_360 = ifelse(bearing_col < 0, bearing_col + 360, bearing_col))

breaks = seq(0, 360, 60)

bearings.plot <- ggplot(trip_bearings, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks, 
                     labels = paste0(breaks, "\u00B0")) +
  labs(y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))


# Plot GPS --------------

# Set projections (use same for Ker + Cro)
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

if(colony == "bi") {
  proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
} else {
  proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
}

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
world2 <- sf::st_transform(world, crs = proj.utm)

# Project GPS data
my_gps.proj <- my_gps %>% arrange(ringYr, datetime)
coordinates(my_gps.proj) <- ~ longitude+latitude
proj4string(my_gps.proj) <- proj.dec

my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

# Project the colony
coordinates(col_coords) <- ~longitude+latitude
proj4string(col_coords) <- proj.dec
colony.sp <- spTransform(col_coords, CRS(proj.utm))
colony.df <- as.data.frame(colony.sp)


# Build the plot

## Set min/max limits
y_min = min(my_gps.proj.df$coords.x2 - 100000)
y_max = max(my_gps.proj.df$coords.x2 + 100000)
x_min = min(my_gps.proj.df$coords.x1 - 100000)
x_max = max(my_gps.proj.df$coords.x1 + 100000)
 
gps_tracks.plot <- ggplot(data = world2) + 
  geom_sf(fill = "cadetblue", colour = "cadetblue") +
  coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
           label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
  geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
            dat = my_gps.proj.df, group = "identity") +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
  scale_colour_viridis_d() +
  annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "grey80"),
        panel.border = element_rect(colour = "black", fill = NA),
        axis.text.x.bottom = element_blank(),
        axis.title.x.bottom = element_blank(),
        axis.text.y.right = element_blank(),
        axis.title.y.right = element_blank(),
        axis.title.y.left = element_blank(),
        axis.text.y.left = element_text(size = 10),
        axis.text.x.top = element_text(size = 10),
        legend.position = "none")


ggarrange(gps_tracks.plot, bearings.plot, ncol = 2)

```

#### Trajectory differences

Next, let's model the bearing differences.

```{r bba_bi_bearing_diffs_setup, include = FALSE }
### Model bearing differences --------------------------------------------------------

load(paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

#### Within-individual, across seasons ---------

# Fit the model
bearings_within.lmm <- lmer(bearing_diff ~ comp_season + (1+comp_season|ring_1), 
                            data = bearing_differences %>% filter(comp_ind == "within_ring"))

summary(bearings_within.lmm)


# Summarise means
bearings_within.summary <- summary(bearings_within.lmm)

bearings_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                                    mean = c(bearings_within.summary$coefficients[1,1], bearings_within.summary$coefficients[1,1] + bearings_within.summary$coefficients[2,1]),
                                    sd = c(bearings_within.summary$coefficients[1,2], bearings_within.summary$coefficients[2,2]))


# Plot #
bearings_within.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_ind == "within_ring"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_within.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_within.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within Season",
    "between_season" = "Between Season" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


#### Between-individual, across seasons  -------------------------------

# Fit the model
bearings_between.lm <- lm(bearing_diff ~ comp_ind, data = bearing_differences)
summary(bearings_between.lm)

## Summarise mean differences
bearings_between.summary <- summary(bearings_between.lm)
bearings_between.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                                     mean = c(bearings_between.summary$coefficients[1,1], bearings_between.summary$coefficients[1,1] + bearings_between.summary$coefficients[2,1]),
                                     sd = c(bearings_between.summary$coefficients[1,2], bearings_between.summary$coefficients[2,2]))

# Plot #
bearings_between.plot <- ggplot() +
  geom_histogram(data = bearing_differences, aes(x = bearing_diff)) +
  geom_segment(data = bearings_between.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model.")  +
  theme_bw()

#### Between-individual, within seasons  -------------------------------

# Fit the model
bearings_between_withinseason.lmm <- lmer(bearing_diff ~ comp_ind + (1|season_1), data = bearing_differences %>% filter(comp_season == "within_season"))
summary(bearings_between_withinseason.lmm)

## Extract mean differences from lmm
bearings_between_withinseason.summary <- summary(bearings_between_withinseason.lmm)

bearings_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                                            mean = c(bearings_between_withinseason.summary$coefficients[1,1], bearings_between_withinseason.summary$coefficients[1,1] + bearings_between_withinseason.summary$coefficients[2,1]),
                                            sd = c(bearings_between_withinseason.summary$coefficients[1,2], bearings_between_withinseason.summary$coefficients[2,2]))

# Plot #
bearings_between_withinseason.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_season == "within_season"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


## Build the overall plot ##
bearings_within.plot2 <- bearings_within.plot +
  theme(plot.caption = element_blank())

bearings_between.plot2 <- bearings_between.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, between seasons",
    "between_ring" = "Between individual, between seasons" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

bearings_between_withinseason.plot2 <- bearings_between_withinseason.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, within season",
    "between_ring" = "Between individual, within season" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

```

The overall differences look like this:

```{r bba_bi_bearing_diffs_plot, echo = FALSE}

ggarrange(bearings_within.plot2, bearings_between.plot2, bearings_between_withinseason.plot2,
          ncol = 3)

```

Output of within-individual model:

```{r bba_bi_bearing_diffs_within_model, echo = FALSE}
bearings_within.summary

```

Output of between-individual models (between- and then within-seasons):

```{r bba_bi_bearing_diffs_between_model, echo = FALSE}
bearings_between.summary
bearings_between_withinseason.summary 
```

#### DTW differences

Let's now look at track similarity across individuals and seasons.

```{r bba_bi_dtw differences, echo = FALSE}
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


## WITHIN INDIVIDUAL -----------

dtw_within.loglmm <- lmer(log(dtw) ~ comp_season + (1+comp_season|ring_1), data = dtw.df %>% filter(comp_ind == "within_ring"))

## Extract mean differences and confint from log LMM
dtw_within.summary <- summary(dtw_within.loglmm)

ci_between <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + fixef(dtw_within.loglmm)["comp_seasonwithin_season"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["comp_seasonwithin_season", "Std. Error"])

dtw_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = exp(c(dtw_within.summary$coefficients[1,1], dtw_within.summary$coefficients[1,1] + dtw_within.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_within.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_ind == "within_ring"), aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within seasons",
    "between_season" = "Between seasons" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


## BETWEEN INDIVIDUAL ---------

dtw_between.loglm <- lm(log(dtw) ~ comp_ind, data = dtw.df)

dtw_between.summary <- summary(dtw_between.loglm)

ci_log <- confint(dtw_between.loglm)
ci_between <- exp(ci_log["(Intercept)",])
ci_within <- exp(ci_log["(Intercept)", ] + ci_log["comp_indwithin_ring", ])

dtw_between.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = c(exp(coef(dtw_between.loglm)["(Intercept)"]),
                                  exp(coef(dtw_between.loglm)["(Intercept)"] +    coef(dtw_between.loglm)["comp_indwithin_ring"])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between.plot <- ggplot() +
  geom_histogram(data = dtw.df, aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within ring",
    "between_ring" = "Between ring" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


## BETWEEN INDIVIDUAL, WITHIN-SEASON -----------

dtw_between_withinseason.loglmm <- lmer(log(dtw) ~ comp_ind + (1|season_1), data = dtw.df %>% filter(comp_season == "within_season"))

## Extract mean differences and confint from log LMM
dtw_between_withinseason.summary <- summary(dtw_between_withinseason.loglmm)

ci_between <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + fixef(dtw_between_withinseason.loglmm)["comp_indwithin_ring"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["comp_indwithin_ring", "Std. Error"])

dtw_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                         mean = exp(c(dtw_between_withinseason.summary$coefficients[1,1], dtw_between_withinseason.summary$coefficients[1,1] + dtw_between_withinseason.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between_withinseason.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_season == "within_season"), aes(x = dtw)) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "between_ring" = "Between individuals",
    "within_ring" = "Within individuals" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))

### Build the overall plot -------
dtw_within.plot2 <- dtw_within.plot +
  theme(plot.caption = element_blank())

dtw_between.plot2 <- dtw_between.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, between seasons",
    "between_ring" = "Between individual, between seasons" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

dtw_between_withinseason.plot2 <- dtw_between_withinseason.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, within season",
    "between_ring" = "Between individual, within season" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())



```

The overall differences look like this:

```{r bba_bi_dtw_plot, echo = FALSE}

ggarrange(dtw_within.plot2, dtw_between.plot2, dtw_between_withinseason.plot2,
          ncol = 3)

```

Output of within-individual model:

```{r bba_bi_dtw_within_model, echo = FALSE}
dtw_within.summary

```

Output of between-individual models (between- and then within-seasons):

```{r bba_bi_dtw_between_model, echo = FALSE}
dtw_between.summary
dtw_between_withinseason.summary 
```

### WAAL at Bird Island

(1) Look at the GPS tracks and the bearings to the distal points of the colony.

```{r waal_bi_setup, include = FALSE}
# Each of the outputs have been run separately in the main R script, so here I'm just calling in the outputs.

# Set parameters
colony <- "bi" # cro bi ker
my_species <- "waal" # waal bba

colony_exp <- ifelse(colony == "ker", "kerguelen",
                     ifelse(colony == "cro", "crozet",
                            "birdis"))

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

if (colony == "ker") {
  col_coords <- ker_coords
} else if (colony == "bi") {
  col_coords <- bi_coords
} else if (colony == "cro") {
  col_coords <- cro_coords }


# Load the data -----------------------------------------------------------

gps_files <- list.files("Data_inputs/", pattern = paste0(my_species, "_", colony_exp, "_gps"))
my_file <- gps_files[grepl("labelled", gps_files)]

my_gps <- loadRData(paste0("Data_inputs/", my_file)) 

my_trip_meta <- my_gps %>% select(c(ring, season, datetime, dist_col, phase, boutID)) %>%
  group_by(ring, season) %>% mutate(max_range = max(dist_col, na.rm = T)) %>% ungroup()

my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))

## Filter out individuals with only one trip
my_gps %<>% 
  filter(loc == "trip") %>%
  group_by(ring, season) %>%
  mutate(n_trips = n_distinct(ID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Make a metadata dataframe
my_meta <- my_gps %>%
  group_by(ID) %>%
  select(c(ring, season, boutID, duration.mins, duration.hours, ID, phase)) %>%
  distinct()


### Calculate bearing differences ------

# Calculate bearings from colony to distal point 

## Nest the data by ID
nested_trips <- my_gps %>%
  filter(loc == "trip") %>%
  group_by(ID) %>%
  nest()

## Apply bearing function to each trip
trip_bearings <- nested_trips %>%
  mutate(bearing_col = map_dbl(data, ~ bearing_from_col(.x, col_coords$longitude, col_coords$latitude))) %>%
  select(ID, bearing_col) 

## Put metadata back in
trip_bearings <- merge(trip_bearings, my_meta, by = "ID", all.x = T)

# Load the bearing differences ----------
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

# Load the DTW values ---------
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


```

```{r waal_bi_build_plots, echo = FALSE}

# Plot bearings -----------------------------------

trip_bearings %<>% mutate(bearing_360 = ifelse(bearing_col < 0, bearing_col + 360, bearing_col))

breaks = seq(0, 360, 60)

bearings.plot <- ggplot(trip_bearings, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks, 
                     labels = paste0(breaks, "\u00B0")) +
  labs(y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))


# Plot GPS --------------

# Set projections (use same for Ker + Cro)
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

if(colony == "bi") {
  proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
} else {
  proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
}

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
world2 <- sf::st_transform(world, crs = proj.utm)

# Project GPS data
my_gps.proj <- my_gps %>% arrange(ringYr, datetime)
coordinates(my_gps.proj) <- ~ longitude+latitude
proj4string(my_gps.proj) <- proj.dec

my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

# Project the colony
coordinates(col_coords) <- ~longitude+latitude
proj4string(col_coords) <- proj.dec
colony.sp <- spTransform(col_coords, CRS(proj.utm))
colony.df <- as.data.frame(colony.sp)


# Build the plot

## Set min/max limits
y_min = min(my_gps.proj.df$coords.x2 - 100000)
y_max = max(my_gps.proj.df$coords.x2 + 100000)
x_min = min(my_gps.proj.df$coords.x1 - 100000)
x_max = max(my_gps.proj.df$coords.x1 + 100000)
 
gps_tracks.plot <- ggplot(data = world2) + 
  geom_sf(fill = "cadetblue", colour = "cadetblue") +
  coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
           label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
  geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
            dat = my_gps.proj.df, group = "identity") +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
  scale_colour_viridis_d() +
  annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "grey80"),
        panel.border = element_rect(colour = "black", fill = NA),
        axis.text.x.bottom = element_blank(),
        axis.title.x.bottom = element_blank(),
        axis.text.y.right = element_blank(),
        axis.title.y.right = element_blank(),
        axis.title.y.left = element_blank(),
        axis.text.y.left = element_text(size = 10),
        axis.text.x.top = element_text(size = 10),
        legend.position = "none")


ggarrange(gps_tracks.plot, bearings.plot, ncol = 2)

```

#### Trajectory differences

Next, let's model the bearing differences.

```{r waal_bi_bearing_diffs_setup, include = FALSE }
### Model bearing differences --------------------------------------------------------

load(paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

#### Within-individual, across seasons ---------

# Fit the model
bearings_within.lmm <- lmer(bearing_diff ~ comp_season + (1+comp_season|ring_1), 
                            data = bearing_differences %>% filter(comp_ind == "within_ring"))

summary(bearings_within.lmm)


# Summarise means
bearings_within.summary <- summary(bearings_within.lmm)

bearings_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                                    mean = c(bearings_within.summary$coefficients[1,1], bearings_within.summary$coefficients[1,1] + bearings_within.summary$coefficients[2,1]),
                                    sd = c(bearings_within.summary$coefficients[1,2], bearings_within.summary$coefficients[2,2]))


# Plot #
bearings_within.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_ind == "within_ring"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_within.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_within.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within Season",
    "between_season" = "Between Season" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


#### Between-individual, across seasons  -------------------------------

# Fit the model
bearings_between.lm <- lm(bearing_diff ~ comp_ind, data = bearing_differences)
summary(bearings_between.lm)

## Summarise mean differences
bearings_between.summary <- summary(bearings_between.lm)
bearings_between.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                                     mean = c(bearings_between.summary$coefficients[1,1], bearings_between.summary$coefficients[1,1] + bearings_between.summary$coefficients[2,1]),
                                     sd = c(bearings_between.summary$coefficients[1,2], bearings_between.summary$coefficients[2,2]))

# Plot #
bearings_between.plot <- ggplot() +
  geom_histogram(data = bearing_differences, aes(x = bearing_diff)) +
  geom_segment(data = bearings_between.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model.")  +
  theme_bw()

#### Between-individual, within seasons  -------------------------------

# Fit the model
bearings_between_withinseason.lmm <- lmer(bearing_diff ~ comp_ind + (1|season_1), data = bearing_differences %>% filter(comp_season == "within_season"))
summary(bearings_between_withinseason.lmm)

## Extract mean differences from lmm
bearings_between_withinseason.summary <- summary(bearings_between_withinseason.lmm)

bearings_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                                            mean = c(bearings_between_withinseason.summary$coefficients[1,1], bearings_between_withinseason.summary$coefficients[1,1] + bearings_between_withinseason.summary$coefficients[2,1]),
                                            sd = c(bearings_between_withinseason.summary$coefficients[1,2], bearings_between_withinseason.summary$coefficients[2,2]))

# Plot #
bearings_between_withinseason.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_season == "within_season"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


## Build the overall plot ##
bearings_within.plot2 <- bearings_within.plot +
  theme(plot.caption = element_blank())

bearings_between.plot2 <- bearings_between.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, between seasons",
    "between_ring" = "Between individual, between seasons" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

bearings_between_withinseason.plot2 <- bearings_between_withinseason.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, within season",
    "between_ring" = "Between individual, within season" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

```

The overall differences look like this:

```{r waal_bi_bearing_diffs_plot, echo = FALSE}

ggarrange(bearings_within.plot2, bearings_between.plot2, bearings_between_withinseason.plot2,
          ncol = 3)

```

Output of within-individual model:

```{r waal_bi_bearing_diffs_within_model, echo = FALSE}
bearings_within.summary

```

Output of between-individual models (between- and then within-seasons):

```{r waal_bi_bearing_diffs_between_model, echo = FALSE}
bearings_between.summary
bearings_between_withinseason.summary 
```

#### DTW differences

Let's now look at track similarity across individuals and seasons.

```{r waal_bi_dtw differences, echo = FALSE}
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


## WITHIN INDIVIDUAL -----------

dtw_within.loglmm <- lmer(log(dtw) ~ comp_season + (1+comp_season|ring_1), data = dtw.df %>% filter(comp_ind == "within_ring"))

## Extract mean differences and confint from log LMM
dtw_within.summary <- summary(dtw_within.loglmm)

ci_between <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + fixef(dtw_within.loglmm)["comp_seasonwithin_season"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["comp_seasonwithin_season", "Std. Error"])

dtw_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = exp(c(dtw_within.summary$coefficients[1,1], dtw_within.summary$coefficients[1,1] + dtw_within.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_within.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_ind == "within_ring"), aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within seasons",
    "between_season" = "Between seasons" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


## BETWEEN INDIVIDUAL ---------

dtw_between.loglm <- lm(log(dtw) ~ comp_ind, data = dtw.df)

dtw_between.summary <- summary(dtw_between.loglm)

ci_log <- confint(dtw_between.loglm)
ci_between <- exp(ci_log["(Intercept)",])
ci_within <- exp(ci_log["(Intercept)", ] + ci_log["comp_indwithin_ring", ])

dtw_between.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = c(exp(coef(dtw_between.loglm)["(Intercept)"]),
                                  exp(coef(dtw_between.loglm)["(Intercept)"] +    coef(dtw_between.loglm)["comp_indwithin_ring"])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_ind == "within_ring"), aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within ring",
    "between_ring" = "Between ring" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


## BETWEEN INDIVIDUAL, WITHIN-SEASON -----------

dtw_between_withinseason.loglmm <- lmer(log(dtw) ~ comp_ind + (1|season_1), data = dtw.df %>% filter(comp_season == "within_season"))

## Extract mean differences and confint from log LMM
dtw_between_withinseason.summary <- summary(dtw_between_withinseason.loglmm)

ci_between <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + fixef(dtw_between_withinseason.loglmm)["comp_indwithin_ring"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["comp_indwithin_ring", "Std. Error"])

dtw_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                         mean = exp(c(dtw_between_withinseason.summary$coefficients[1,1], dtw_between_withinseason.summary$coefficients[1,1] + dtw_between_withinseason.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between_withinseason.plot <- ggplot() +
  geom_histogram(data = dtw.df, aes(x = dtw)) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "between_ring" = "Between individuals",
    "within_ring" = "Within individuals" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))

### Build the overall plot -------
dtw_within.plot2 <- dtw_within.plot +
  theme(plot.caption = element_blank())

dtw_between.plot2 <- dtw_between.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, between seasons",
    "between_ring" = "Between individual, between seasons" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

dtw_between_withinseason.plot2 <- dtw_between_withinseason.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, within season",
    "between_ring" = "Between individual, within season" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())



```

The overall differences look like this:

```{r waal_bi_dtw_plot, echo = FALSE}

ggarrange(dtw_within.plot2, dtw_between.plot2, dtw_between_withinseason.plot2,
          ncol = 3)

```

Output of within-individual model:

```{r waal_bi_dtw_within_model, echo = FALSE}
dtw_within.summary

```

Output of between-individual models (between- and then within-seasons):

```{r waal_bi_dtw_between_model, echo = FALSE}
dtw_between.summary
dtw_between_withinseason.summary 
```

### WAAL at Crozet

(1) Look at the GPS tracks and the bearings to the distal points of the colony.

```{r waal_cro_setup, include = FALSE}
# Each of the outputs have been run separately in the main R script, so here I'm just calling in the outputs.

# Set parameters
colony <- "cro" # cro bi ker
my_species <- "waal" # waal bba

colony_exp <- ifelse(colony == "ker", "kerguelen",
                     ifelse(colony == "cro", "crozet",
                            "birdis"))

# Set colony coordinates
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

if (colony == "ker") {
  col_coords <- ker_coords
} else if (colony == "bi") {
  col_coords <- bi_coords
} else if (colony == "cro") {
  col_coords <- cro_coords }


# Load the data -----------------------------------------------------------

gps_files <- list.files("Data_inputs/", pattern = paste0(my_species, "_", colony_exp, "_gps"))
my_file <- gps_files[grepl("labelled", gps_files)]

my_gps <- loadRData(paste0("Data_inputs/", my_file)) 

my_trip_meta <- my_gps %>% select(c(ring, season, datetime, dist_col, phase, boutID)) %>%
  group_by(ring, season) %>% mutate(max_range = max(dist_col, na.rm = T)) %>% ungroup()

my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))

## Filter out individuals with only one trip
my_gps %<>% 
  filter(loc == "trip") %>%
  group_by(ring, season) %>%
  mutate(n_trips = n_distinct(ID)) %>%
  filter(n_trips > 1) %>%
  select(-n_trips)

## Make a metadata dataframe
my_meta <- my_gps %>%
  group_by(ID) %>%
  select(c(ring, season, boutID, duration.mins, duration.hours, ID, phase)) %>%
  distinct()


### Calculate bearing differences ------

# Calculate bearings from colony to distal point 

## Nest the data by ID
nested_trips <- my_gps %>%
  filter(loc == "trip") %>%
  group_by(ID) %>%
  nest()

## Apply bearing function to each trip
trip_bearings <- nested_trips %>%
  mutate(bearing_col = map_dbl(data, ~ bearing_from_col(.x, col_coords$longitude, col_coords$latitude))) %>%
  select(ID, bearing_col) 

## Put metadata back in
trip_bearings <- merge(trip_bearings, my_meta, by = "ID", all.x = T)

# Load the bearing differences ----------
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

# Load the DTW values ---------
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


```

```{r waal_cro_build_plots, echo = FALSE}

# Plot bearings -----------------------------------

trip_bearings %<>% mutate(bearing_360 = ifelse(bearing_col < 0, bearing_col + 360, bearing_col))

breaks = seq(0, 360, 60)

bearings.plot <- ggplot(trip_bearings, aes(x = bearing_360)) +
  geom_histogram(binwidth = 10,
                 fill = "skyblue", color = "black", alpha = 0.7) + 
  coord_polar(start = 0) +
  scale_x_continuous("", limits = c(0, 360), 
                     breaks = breaks, 
                     labels = paste0(breaks, "\u00B0")) +
  labs(y = "Frequency") +
  theme_minimal() +
  theme(plot.title = element_text(size = 12, hjust = 0.5),
        axis.ticks.y = element_line(color = "black"))


# Plot GPS --------------

# Set projections (use same for Ker + Cro)
proj.dec <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"

if(colony == "bi") {
  proj.utm = "+proj=laea +lat_0=-54 +lon_0=-38.03 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
} else {
  proj.utm = "+proj=laea +lat_0=-46.358639 +lon_0=51.706972 +x_0=0 +y_0=0 +datum=WGS84 +units=m +no_defs +ellps=WGS84 +towgs84=0,0,0"
}

# Get the base world plot
world <- ne_countries(scale = "medium", returnclass = "sf")
world2 <- sf::st_transform(world, crs = proj.utm)

# Project GPS data
my_gps.proj <- my_gps %>% arrange(ringYr, datetime)
coordinates(my_gps.proj) <- ~ longitude+latitude
proj4string(my_gps.proj) <- proj.dec

my_gps.proj.sp <- spTransform(my_gps.proj, CRS(proj.utm))
my_gps.proj.df <- as.data.frame(my_gps.proj.sp)

# Project the colony
coordinates(col_coords) <- ~longitude+latitude
proj4string(col_coords) <- proj.dec
colony.sp <- spTransform(col_coords, CRS(proj.utm))
colony.df <- as.data.frame(colony.sp)


# Build the plot

## Set min/max limits
y_min = min(my_gps.proj.df$coords.x2 - 100000)
y_max = max(my_gps.proj.df$coords.x2 + 100000)
x_min = min(my_gps.proj.df$coords.x1 - 100000)
x_max = max(my_gps.proj.df$coords.x1 + 100000)
 
gps_tracks.plot <- ggplot(data = world2) + 
  geom_sf(fill = "cadetblue", colour = "cadetblue") +
  coord_sf(crs = proj.utm, xlim = c(x_min, x_max), ylim = c(y_min, y_max),
           label_axes = list(top = "E", left = "N", bottom = "E", right = "N")) +
  geom_path(aes(x = coords.x1 , y = coords.x2, col = ringYr), size = 0.8,
            dat = my_gps.proj.df, group = "identity") +
  scale_x_continuous(breaks = seq(-180, 180, by = 5)) +
  scale_y_continuous(breaks = seq(0, -90, by = -5)) + 
  scale_colour_viridis_d() +
  annotate("point", shape = 17, size = 3, (colony.df$coords.x1 + 100), colony.df$coords.x2) +
  theme(panel.background = element_rect(fill = "white"),
        panel.grid.major = element_line(colour = "grey80"),
        panel.border = element_rect(colour = "black", fill = NA),
        axis.text.x.bottom = element_blank(),
        axis.title.x.bottom = element_blank(),
        axis.text.y.right = element_blank(),
        axis.title.y.right = element_blank(),
        axis.title.y.left = element_blank(),
        axis.text.y.left = element_text(size = 10),
        axis.text.x.top = element_text(size = 10),
        legend.position = "none")


ggarrange(gps_tracks.plot, bearings.plot, ncol = 2)

```

#### Trajectory differences

Next, let's model the bearing differences.

```{r waal_cro_bearing_diffs_setup, include = FALSE }
### Model bearing differences --------------------------------------------------------

load(paste0("Data_outputs/", my_species, "_", colony_exp, "_bearing_differences.RData"))

#### Within-individual, across seasons ---------

# Fit the model
bearings_within.lmm <- lmer(bearing_diff ~ comp_season + (1+comp_season|ring_1), 
                            data = bearing_differences %>% filter(comp_ind == "within_ring"))

summary(bearings_within.lmm)


# Summarise means
bearings_within.summary <- summary(bearings_within.lmm)

bearings_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                                    mean = c(bearings_within.summary$coefficients[1,1], bearings_within.summary$coefficients[1,1] + bearings_within.summary$coefficients[2,1]),
                                    sd = c(bearings_within.summary$coefficients[1,2], bearings_within.summary$coefficients[2,2]))


# Plot #
bearings_within.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_ind == "within_ring"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_within.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_within.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within Season",
    "between_season" = "Between Season" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


#### Between-individual, across seasons  -------------------------------

# Fit the model
bearings_between.lm <- lm(bearing_diff ~ comp_ind, data = bearing_differences)
summary(bearings_between.lm)

## Summarise mean differences
bearings_between.summary <- summary(bearings_between.lm)
bearings_between.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                                     mean = c(bearings_between.summary$coefficients[1,1], bearings_between.summary$coefficients[1,1] + bearings_between.summary$coefficients[2,1]),
                                     sd = c(bearings_between.summary$coefficients[1,2], bearings_between.summary$coefficients[2,2]))

# Plot #
bearings_between.plot <- ggplot() +
  geom_histogram(data = bearing_differences, aes(x = bearing_diff)) +
  geom_segment(data = bearings_between.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model.")  +
  theme_bw()

#### Between-individual, within seasons  -------------------------------

# Fit the model
bearings_between_withinseason.lmm <- lmer(bearing_diff ~ comp_ind + (1|season_1), data = bearing_differences %>% filter(comp_season == "within_season"))
summary(bearings_between_withinseason.lmm)

## Extract mean differences from lmm
bearings_between_withinseason.summary <- summary(bearings_between_withinseason.lmm)

bearings_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                                            mean = c(bearings_between_withinseason.summary$coefficients[1,1], bearings_between_withinseason.summary$coefficients[1,1] + bearings_between_withinseason.summary$coefficients[2,1]),
                                            sd = c(bearings_between_withinseason.summary$coefficients[1,2], bearings_between_withinseason.summary$coefficients[2,2]))

# Plot #
bearings_between_withinseason.plot <- ggplot() +
  geom_histogram(data = bearing_differences %>% filter(comp_season == "within_season"), aes(x = abs(bearing_diff))) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean+sd, xend = mean+sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = bearings_between_withinseason.means, aes(x = mean-sd, xend = mean-sd, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual",
    "between_ring" = "Between individual" ))) +
  labs(y = "Frequency", x = "Angular bearing differences, \u00B0",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model.") +
  theme_bw()


## Build the overall plot ##
bearings_within.plot2 <- bearings_within.plot +
  theme(plot.caption = element_blank())

bearings_between.plot2 <- bearings_between.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, between seasons",
    "between_ring" = "Between individual, between seasons" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

bearings_between_withinseason.plot2 <- bearings_between_withinseason.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, within season",
    "between_ring" = "Between individual, within season" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

```

The overall differences look like this:

```{r waal_cro_bearing_diffs_plot, echo = FALSE}

ggarrange(bearings_within.plot2, bearings_between.plot2, bearings_between_withinseason.plot2,
          ncol = 3)

```

Output of within-individual model:

```{r waal_cro_bearing_diffs_within_model, echo = FALSE}
bearings_within.summary

```

Output of between-individual models (between- and then within-seasons):

```{r waal_cro_bearing_diffs_between_model, echo = FALSE}
bearings_between.summary
bearings_between_withinseason.summary 
```

#### DTW differences

Let's now look at track similarity across individuals and seasons.

```{r waal_cro_dtw differences, echo = FALSE}
load(paste0("Data_outputs/", my_species, "_", colony_exp, "_dtw_comparison.RData"))


## WITHIN INDIVIDUAL -----------

dtw_within.loglmm <- lmer(log(dtw) ~ comp_season + (1+comp_season|ring_1), data = dtw.df %>% filter(comp_ind == "within_ring"))

## Extract mean differences and confint from log LMM
dtw_within.summary <- summary(dtw_within.loglmm)

ci_between <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_within.loglmm)["(Intercept)"] + fixef(dtw_within.loglmm)["comp_seasonwithin_season"] + c(-1, 1) * 1.96 * summary(dtw_within.loglmm)$coefficients["comp_seasonwithin_season", "Std. Error"])

dtw_within.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = exp(c(dtw_within.summary$coefficients[1,1], dtw_within.summary$coefficients[1,1] + dtw_within.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_within.plot <- ggplot() +
  geom_histogram(data = dtw.df, aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_season, ncol = 1, labeller = labeller(comp_season = c(
    "within_season" = "Within seasons",
    "between_season" = "Between seasons" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


## BETWEEN INDIVIDUAL ---------

dtw_between.loglm <- lm(log(dtw) ~ comp_ind, data = dtw.df)

dtw_between.summary <- summary(dtw_between.loglm)

ci_log <- confint(dtw_between.loglm)
ci_between <- exp(ci_log["(Intercept)",])
ci_within <- exp(ci_log["(Intercept)", ] + ci_log["comp_indwithin_ring", ])

dtw_between.means <- data.frame(comp_season = c("between_season", "within_season"),
                         mean = c(exp(coef(dtw_between.loglm)["(Intercept)"]),
                                  exp(coef(dtw_between.loglm)["(Intercept)"] +    coef(dtw_between.loglm)["comp_indwithin_ring"])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_ind == "within_ring"), aes(x = dtw)) +
  geom_segment(data = dtw_within.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_within.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_within.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within ring",
    "between_ring" = "Between ring" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))


## BETWEEN INDIVIDUAL, WITHIN-SEASON -----------

dtw_between_withinseason.loglmm <- lmer(log(dtw) ~ comp_ind + (1|season_1), data = dtw.df %>% filter(comp_season == "within_season"))

## Extract mean differences and confint from log LMM
dtw_between_withinseason.summary <- summary(dtw_between_withinseason.loglmm)

ci_between <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["(Intercept)", "Std. Error"])
ci_within <- exp(fixef(dtw_between_withinseason.loglmm)["(Intercept)"] + fixef(dtw_between_withinseason.loglmm)["comp_indwithin_ring"] + c(-1, 1) * 1.96 * summary(dtw_between_withinseason.loglmm)$coefficients["comp_indwithin_ring", "Std. Error"])

dtw_between_withinseason.means <- data.frame(comp_ind = c("between_ring", "within_ring"),
                         mean = exp(c(dtw_between_withinseason.summary$coefficients[1,1], dtw_between_withinseason.summary$coefficients[1,1] + dtw_between_withinseason.summary$coefficients[2,1])),
                         ci_low = c(ci_between[1], ci_within[1]),
                         ci_high = c(ci_between[2], ci_within[2]))

# Plot #
dtw_between_withinseason.plot <- ggplot() +
  geom_histogram(data = dtw.df %>% filter(comp_season == "within_season"), aes(x = dtw)) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = mean, xend = mean, y = 0, yend = Inf), 
               color = "red", linewidth = 1) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_low, xend = ci_low, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  geom_segment(data = dtw_between_withinseason.means, aes(x = ci_high, xend = ci_high, y = 0, yend = Inf), 
               color = "red", linetype = "dashed", linewidth = 0.8) +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "between_ring" = "Between individuals",
    "within_ring" = "Within individuals" ))) +
  labs(y = "Frequency", x = "DTW distance (track dissimilarity)",
       caption = "Red lines indicate mean (solid) \u00B1 standard error (dashed) taken from linear mixed effects model with logged response.") +
  theme_bw() +
  theme(plot.caption = element_text(hjust = 0),
        strip.background = element_rect(fill ="grey90"),
        strip.text = element_text(face = "bold"))

### Build the overall plot -------
dtw_within.plot2 <- dtw_within.plot +
  theme(plot.caption = element_blank())

dtw_between.plot2 <- dtw_between.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, between seasons",
    "between_ring" = "Between individual, between seasons" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())

dtw_between_withinseason.plot2 <- dtw_between_withinseason.plot +
  facet_wrap(.~comp_ind, ncol = 1, labeller = labeller(comp_ind = c(
    "within_ring" = "Within individual, within season",
    "between_ring" = "Between individual, within season" ))) +
  theme(plot.caption = element_blank(),
        axis.title.y = element_blank())



```

The overall differences look like this:

```{r waal_cro_dtw_plot, echo = FALSE}

ggarrange(dtw_within.plot2, dtw_between.plot2, dtw_between_withinseason.plot2,
          ncol = 3)

```

Output of within-individual model:

```{r waal_cro_dtw_within_model, echo = FALSE}
dtw_within.summary

```

Output of between-individual models (between- and then within-seasons):

```{r waal_cro_dtw_between_model, echo = FALSE}
dtw_between.summary
dtw_between_withinseason.summary 
```

## Overall summary

-   Kerguelen BBAL:
    -   Both trajectories and trips are more similar within than between individuals. There is no seasonal difference
    -   Individuals tend to go either NE or SW of the colony, which has implications for interpreting the results
-   Bird Island BBAL:
    -   Trajectories are more similar within than between individuals, and are slightly (but not significantly) more similar within than between seasons.
    -   Trips are more similar within individuals than between, but the similarity is much smaller than for Kerguelen. When considering trips within the same season, trips are more similar within- than between-individuals, suggesting that there is a seasonal component to these relationships.
-   Bird Island WAAL:
    -   Trajectories are slightly more similar within- than between-individuals when considering all trips. However, when considering only trips that occur within the same season, there is no effect of individual, suggesting that similarity is driven primarily by seasonal similarities. Given that there is no difference in trajectory similarity with season for within-individual trips, this suggests that birds go to similar places within the same season, but individuals do not repeatedly visit the same places.

    -   There is no effect of individual on the trip structure itself. This suggests that even though birds might go to similar places within a given season, they are not covering similar paths at either an individual or seasonal level.
-   Crozet WAAL:
    -   Individuals do not take more similar trajectories within- versus between-seasons. They do take more similar trajectories to themselves than other birds both within- and across- seasons.
    -   Similarly, individuals go on similarly structured trips both across and within seasons.
    -   **May have to consider sex effects here as this could explain some of this variation.** I quickly modelled sex effects on the trajectories difference and there is a clear effect in males, but the within-individual difference still holds. May have to be an interactive effect? Think carefully about this.
