---
title: "Habitat selection analyses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

**Note:** Naming conventions for files:
  - Bird Island = birdis
  - Crozet = cro
  - Kerguelen = ker
  - Chlorophyll = chl

## SETUP

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)
# Functions for GPS processing and plotting
source("RPT_functions.R")
source("condR.R")

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "tidyr", "lubridate", "DHARMa", "data.table", "ggpubr",  "kableExtra", "ggridges", "corrplot", "pROC", "ggeffects", "glmmTMB", "brms") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

# Make sure using dplyr select
select <- dplyr::select

# Scale function
scale_2SD <- function(variable) { (variable - mean(variable, na.rm = TRUE)) / sd(variable, na.rm = TRUE) * 2 }

# Set colours
bathy_col = "#1E88E5"
chl_col = "#004D40"
sst_col = "#D81B60"

```

At this point I have (1) randomly sampled available habitat on a ratio of 6 available to 1 used points; and (2) matched up all of these points with environmental data - specifically, SST, bathymetry, SSH, and chlorophyll.

To ensure appropriate model fit, I'll be working on each species/colony separately. I can think about integrating them later, if the residual distributions etc are similar.

## EXPLORATION: Black-browed albatrosses

### Bird Island

First, I'm exploring the environmental variation experienced by individuals. Here, for a subsample of 20 birds (for readability), I'm plotting the environmental variability used by each individual.

```{r bbal_birdis_plot_env, fig.cap="**Figure 1.1** *Individual density histograms for use of SST, bathymetry, chlorophyll, and sea surface height. Random sample of 20 individuals.*", fig.dim = c(8, 10), fig.topcaption = TRUE}

load("Data_inputs/bba_birdis_sst-bathy-chl-ssh_data.RData")

subset_ids <- sample(unique(pnts_all$ring), 30)

#### ~# SST # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_sst = mean(SST, na.rm = T)) %>%
  ungroup() %>%
  mutate(ring = factor(ring, levels = unique(ring[order(mean_sst)]))) 

plot.sst <- ggplot(pnts_subset, 
       aes(x = SST, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#0D0887FF", "#CC4678FF", "#F0F921FF"),
    name = "SST (°C)" ) +
  labs(x = "Sea surface temperature (°C)", y= "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())

png(filename = "Figures/sst_use.png", width = 4.5, height = 6, units = "in", res = 600)
plot.sst
dev.off()

#### ~# Bathymetry # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_bathy = mean(bathy, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_bathy)]))) 

plot.bathy <- ggplot(pnts_subset, 
       aes(x = bathy, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_viridis_c( name = "Sea floor depth (m below sea level)" ) +
  labs(x = "Sea floor depth (m below sea level)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())

png(filename = "Figures/bathy_use.png", width = 4.5, height = 6, units = "in", res = 600)
plot.bathy
dev.off()


#### ~# Chlorophyll # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_chl = mean(chl, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_chl)]))) 

plot.chl <- ggplot(pnts_subset, 
       aes(x = chl, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1A6D2A", "#6CDC82", "#D3F5D9"),
    name = expression("Chlorophyll (" * mg ~ Chl ~ m^-3 * ")") ) +
  labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"), y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


png(filename = "Figures/chl_use.png", width = 4.5, height = 6, units = "in", res = 600)
plot.chl
dev.off()

#### ~# Sea surface height # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_ssh = mean(SSH, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_ssh)]))) 

plot.ssh <- ggplot(pnts_subset, 
       aes(x = SSH, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1a2290", "#B2B6F1", "#D8DCE0"),
    name = "SSH (m)") +
  labs(x = "Sea surface height (m)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


ggarrange(plot.sst, plot.bathy + theme(axis.title.y = element_blank()),
          plot.chl, plot.ssh + theme(axis.title.y = element_blank()), 
          ncol = 2, nrow = 2)

```

Next, I looked at potential covariation between the different variables.

```{r bbal_birdis_correlations, fig.cap = "**Figure 1.2.** *Correlations between chosen environmental variables.*", fig.dim = c(5, 5), fig.topcaption = TRUE}

corr_matrix <- cor(na.omit(pnts_all[, c(8, 12, 13, 15, 16)]))
corrplot.mixed(corr_matrix, order = "AOE")

```

Surprisingly, SST and SSH are extremely correlated - it turns out that they do indeed have a strong spatial correlation, particularly due to e.g. El Nino events, e.g. see:

Leuliette, E. W., & Wahr, J. M. (1999). Coupled pattern analysis of sea surface temperature and TOPEX/Poseidon sea surface height. *Journal of Physical Oceanography, 29*(4), 599–611. [https://doi.org/10.1175/1520-0485(1999)029\<0599:CPAOSS\>2.0.CO;2](https://doi.org/10.1175/1520-0485(1999)029%3C0599:CPAOSS%3E2.0.CO;2)

Apart from this, the correlations don't seem very worrying. In response to these correlations, I dropped SSH.

```{r bbal_birdis_process_data, eval = TRUE}

# Do some data cleaning
load("Data_inputs/bba_birdis_sst-bathy-chl-ssh_data.RData")

scale_2SD <- function(variable) { (variable - mean(variable, na.rm = TRUE)) / sd(variable, na.rm = TRUE) * 2 }

## Scale the variables
pnts_all %<>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         #ssh_scaled = scale_2SD(chl),
         chl_scaled = scale_2SD(chl),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
                     bathy_scaled2 = bathy_scaled^2) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID)) 

## Remove birds with only a couple of points
pnts_all %<>% 
  group_by(ID, year) %>%
  mutate(n_recs = n(),
         flag = ifelse(n_recs < 20, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-flag)

## Remove NAs and set distance to km
pnts_all %<>% filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chl)) %>%
  mutate(dist_col.km = dist_col/1000)

## Calculate within-individual centered SST
# pnts_all %<>%
#   group_by(ID) %>%
#   mutate(sst_within = SST - mean(SST, na.rm = TRUE),
#          sst_within2 = sst_within^2) 


```

#### Average preference

Here I'm looking at differences in average habitat choice amongst individuals. Here, I fit simple linear models where the average environmental variable experienced is predicted by distance from the colony, breeding phase, and - critically - the random effect of individual.

From this I can extract random intercepts as the variation in environment use attributable to individual - e.g. average value of SST for each individual - and then the repeatability of individual choice for the environmental variable.

The model I fitted was:

$$y = \beta_0 + \beta_1 \cdot \text{dist_col} + \beta_2 \cdot \text{phase} + U_{\text{ID}} + U_{\text{tripID}}$$

where:

-   $\beta_0$ is the intercept,
-   $\beta_1$ is the fixed effect of the distance from the colony (to control for environmental variation associated with this),
-   $\beta_2$ is the fixed effects of breeding phase,
-   $U_{\text{ID}}$ is the individual-level random intercept,
-   $U_{\text{ID}}$ is the random intercept for trip.

```{r average_habitat}

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chl")
rpt_df <- data.frame(env_var = character(), rpt_R = numeric())

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1 | ID) + (1 | tripID)")), data = pnts.used)
  
  # Extract random effects for ID
  random_effects <- ranef(mod)$ID
  ind_intercepts <- random_effects[, 1]
  
  # Extract fixed effects for the intercept
  ind_avg <- ind_intercepts + fixef(mod)[1]
  assign(paste0("ind_avg.", env_vars[i]), ind_avg)
  
  # Calculate variance components
  var_components <- as.data.frame(VarCorr(mod))
  var_id <- var_components[var_components$grp == "ID", "vcov"]
  var_residual <- attr(VarCorr(mod), "sc")^2
  
  # Total variance = variance from ID + residual variance
  total_var <- var_id + var_residual
  
  # Calculate repeatability
  repeatability <- var_id / total_var
  
  # Output
  rpt_df <- rbind(rpt_df, data.frame(variable = env_vars[i], repeatability = repeatability))
 
}

## Plot outputs
ind_avg.plot <- data.frame(SST = ind_avg.sst,
                           bathy = ind_avg.bathy,
                           chl = ind_avg.chl)

ind_avg_long <- ind_avg.plot %>%
  pivot_longer(cols = everything(), names_to = "variable", values_to = "value") %>%
  arrange(variable) %>%
  mutate(dummy_x = 1)

# Add y position for plot: SST - bathy - chl
rpt_df$y_position = c(8, 1200, 4)

ind_avg_long_with_rpt <- ind_avg_long %>%
  left_join(rpt_df, by = "variable")

# Plot
label_map <- c(
  bathy = "Bathymetry (m below sea level)",
  chl = "Chlorophyll (mg Chl m^-3)",
  SST = "SST (°C)")

png(filename = "Figures/avg_environment.png", width = 10, height = 5, units = "in", res = 600)
ggplot(ind_avg_long_with_rpt, aes(x = dummy_x, y = value, fill = variable)) +
  geom_violin(trim = FALSE, colour = "black") +
  geom_jitter(width = 0.1, alpha = 0.5, size = 1) +
   # Add the repeatability text
  geom_text(aes(x = 1, y = y_position, label = paste("Repeatability: ", round(repeatability, 3))),
            fontface = "plain", colour = "black", size = 3) +
  facet_wrap(~ variable, scales = "free_y", 
             labeller = labeller(variable = label_map)) + 
  scale_fill_manual(values = c("#37D5AC", "#339F48", "#A0307E")) +
  theme_minimal(base_size = 14) +
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank())
dev.off()



```

#### Habitat selection models

To understand individual variation in habitat preference, I am fitting habitat selection models. These are binomial models that look at differences in environmental conditions between 'used' and 'available' habitats. I have fitted a random slope to allow indiviudal selection to vary.

After having issues fitting each environmental variable in turn (specifically, the AUC was very low, suggesting poor predictive performance), I realised that a maximal model including all the predictors might be better. Indeed, I get better AUC and fit (see below).

This current model is only looking at between season individual variation, I still need to look at within season.

The generalized linear mixed model (GLMM) used in this analysis is:

$$\text{logit}(\Pr(\text{used} = 1)) = \beta_0 + \beta_1 \cdot \text{sst} + \beta_2 \cdot \text{sst}^2 + \beta_3 \cdot \text{bathy} + \beta_4 \cdot \text{bathy}^2 + \beta_5 \cdot \text{chl} + \beta_6 \cdot \text{dist col} + \beta_7 \cdot \text{phase} + U_{\text{ID}} + U_{\text{ID}}^{\text{sst}} \cdot \text{sst} + U_{\text{ID}}^{\text{bathy}} \cdot \text{bathy} + U_{\text{ID}}^{\text{chl}} \cdot \text{chl}$$

where:

-   $\beta_0$ is the intercept,
-   $\beta_1$–$\beta_{7}$ are the fixed effects of each environmental variable ($sst$, $bathy$, $chl$), breeding phase ($phase$), and distance from the colony in km ($dist col$)
-   $U_{\text{ID}}$ is the individual-level random intercept,
-   $U_{\text{ID}}^{\text{sst}}$, $U_{\text{ID}}^{\text{bathy}}$, and
-   $U_{\text{ID}}^{\text{chl}}$ are **individual-level random slopes** for **SST, bathymetry, and chlorophyll**, allowing individuals to have **different responses** to these variables.

Both SST and bathymetry were fitted with quadratic effects based on exploratory modelling. For SST, it looks like birds select medium-low temperatures; for bathymetry it looks like birds prefer shallow and deep areas (possibly confounded by colony distance, but we'll get on to that). Models fitted with quadratic SST and bathymetry are a significiantly better fit based on AIC and an ANOVA comparison. The binned residual plot also looks considerably better, and the AUC improves (marginally, as it was reasonably good to begin with). Most importantly, I have no convergence issues with this model structure.

I have not included quadratic random slopes as this gets very complicated. So for now, we are assuming that all birds have the same quadratic preference (same slope shape) but this can 'shift'.

```{r bbal_birdis_models, eval = FALSE}

pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

## currently working here 13/05/2025
mod_linear <- glmmTMB(
  used ~ sst_scaled + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

mod_quadratic <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

AIC(mod_linear, mod_quadratic)
anova(mod_linear, mod_quadratic)

tok <- Sys.time()
mod_brms.quad <- brm(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase + 
        (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_all,
  family = bernoulli(),
  chains = 4,       
  cores = 4,        
  iter = 2000,      
  warmup = 1000     
)
tik <- Sys.time()

save(mod_brms.quad, file = "Data_outputs/mod_brms_quad.RData")




mod_between.bbal_bi.sst <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

summary(mod_between.bbal_bi.sst)
save(mod_between.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_between_sst.RData")

# Within-season ----

mod_within.bbal_bi.sst <- glmmTMB(
  used ~ poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial()
)

save(mod_within.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_within_sst.RData")

mod_within.bbal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.bbal_bi.bathy, file = "Data_outputs/bba_birdis_habitat_model_within_bathy.RData")

mod_within.bbal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.bbal_bi.chl, file = "Data_outputs/bba_birdis_habitat_model_within_chl.RData")


# Between-season ----

mod_between.bbal_bi.sst <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled + sst_scaled2 | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

summary(mod_between.bbal_bi.sst)
save(mod_between.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_between_sst.RData")


mod_between.bbal_bi.sst <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.bbal_bi.sst, file = "Data_outputs/bba_birdis_habitat_model_between_sst.RData")

mod_between.bbal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 +bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.bbal_bi.bathy, file = "Data_outputs/bba_birdis_habitat_model_between_bathy.RData")

mod_between.bbal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.bbal_bi.chl, file = "Data_outputs/bba_birdis_habitat_model_between_chl.RData")

```

```{r plot_individual_slopes}

env_variables <- c("sst", "bathy", "chl")
env_colours <- c(
  "sst" = "#A0307E",
  "bathy" = "#37D5AC",
  "chl" = "#339F48")

env_labels <- c(
  "sst" = "Scaled SST",
  "bathy" = "Scaled bathymetry",
  "chl" = "Scaled chlorophyll conc.")

rpt_list <- list() 
within = TRUE

for (i in seq_along(env_variables)) {
  
  if (within == TRUE) {
    mod_loaded <-loadRData(paste0("Data_outputs/mod_brms_quad_", env_variables[i], "_within.RData")) } else {
      mod_loaded <-loadRData(paste0("Data_outputs/mod_brms_quad_", env_variables[i], ".RData"))
    }
  
  env_scaled <- paste0(env_variables[i], "_scaled")
  
  # Get variances
  posterior <- as_draws_df(mod_loaded)
  var_col <- ifelse(within == TRUE, paste0("sd_ID_season__", env_scaled), paste0("sd_ID__", env_scaled))
  var_among <- posterior[[var_col]]^2 
  var_resid <- pi^2 / 3
  
  # Calculate repeatability per posterior sample
  repeatability <- var_among / (var_among + var_resid)
  repeatability_summary <- data.frame(
    var = env_variables[i],
    mean = mean(repeatability),
    median = median(repeatability),
    CI_lower = quantile(repeatability, 0.025),
    CI_upper = quantile(repeatability, 0.975))
  
  rpt_list[[i]] <- repeatability_summary
  
  # Extract random & fixed effects
  re <- ranef(mod_loaded)[[1]]
  fixed <- fixef(mod_loaded)
  
  # Build data frame of individual slopes
  slopes <- data.frame(
    ID = rownames(re),
    intercept = fixed["Intercept", "Estimate"] + re[, "Estimate", "Intercept"],
    slope = fixed[env_scaled, "Estimate"] + re[, "Estimate", env_scaled])
  
  if(env_variables[i] == "chl") {
    slopes$quad = NA
  } else {
    slopes$quad = fixed[paste0(env_scaled, 2), "Estimate"] + re[, "Estimate", paste0(env_scaled, 2)] 
  }
  
   
  png(file = ifelse(within == TRUE, 
                    paste0("Figures/individual_", env_variables[i], "_within_slopes.png"), 
                    paste0("Figures/individual_", env_variables[i], "_between_slopes.png")),
                    width = 5, height = 4, units = "in", res = 600)
  ggplot(slopes, aes(x = reorder(ID, slope), y = slope)) +
    geom_point() +
    geom_hline(yintercept = fixed[env_scaled, "Estimate"], linetype = "dashed", colour = "red") +
    coord_flip() +
    labs(y = "Individual slope", x = "Individual ID") +
    theme_bw() +
    theme(axis.text.y = element_blank(),
          axis.ticks.y = element_blank()) 
  dev.off()
  
  png(file =ifelse(within == TRUE, 
                    paste0("Figures/histogram_", env_variables[i], "_within_slopes.png"), 
                    paste0("Figures/histogram_", env_variables[i], "_between_slopes.png")),
      width = 5, height = 4, units = "in", res = 600)
  ggplot(slopes, aes(x = slope)) +
    geom_histogram() +
    labs(x = "Individual slope", y = "Count") +
    theme_bw() 
  dev.off()
  
  # Plot individual curves
  ## Create a sequence of environmental variables
  env_seq <- seq(-3, 3, length.out = 100)
  
  # Expand to a grid: one row per individual per sst_scaled value
  curve_data <- slopes %>%
    mutate(ID = as.character(ID)) %>%
    slice(rep(1:n(), each = length(env_seq))) %>%
    mutate(env_scaled = rep(env_seq, times = nrow(slopes)),
           env_scaled2 = env_scaled^2)
  
  if(env_variables[i] == "chl") {
    curve_data %<>%
      mutate(logit_p = intercept + slope * env_scaled,
             prob = plogis(logit_p))
  } else {
    curve_data %<>%
      mutate(logit_p = intercept + slope * env_scaled + quad * env_scaled2,
             prob = plogis(logit_p))
  }

  colour_choice <- env_colours[env_variables[i]]
  x_label <- env_labels[env_variables[i]]
  
  png(file = ifelse(within == TRUE, 
                    paste0("Figures/individual_", env_variables[i], "_within_curves.png"), 
                    paste0("Figures/individual_", env_variables[i], "_between_curves.png")), 
      width = 6.5, height = 6, units = "in", res = 600)
  ggplot(curve_data, aes(x = env_scaled, y = prob, group = ID)) +
    geom_line(alpha = 0.3, colour = colour_choice) +
    labs(x = x_label, y = "P|used") +
    theme_bw()
  dev.off()

}


```

```{r code from Bertille}

load("Data_outputs/mod_brms.quad.Rdata")
summary(mod_brms.quad)

set.seed(817)
pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

# Onset of senescence
df_rawAge_randomSlopesAndInterceptsIDlevel$OnsetSenescence = - df_rawAge_randomSlopesAndInterceptsIDlevel$Slope / (2*df_rawAge_randomSlopesAndInterceptsIDlevel$Slope2 )


# extract rate of senescence
# Function that calculates the derivative of 
# exp(quad*xval^2+lin*xval+const)/(1+exp(quad*xval^2+lin*xval+const))
deriv_logit_quad<-function(xval,quad,lin,const)
{
  output<-(2*quad*xval+lin)*exp(quad*xval^2+lin*xval+const)/((1+exp(quad*xval^2+lin*xval+const))^2)
  return(output)
}
 
# Find minimum of the derivative 
xinfl<-c()
gradmins<-c()
for(index in 1:length(df_rawAge_randomSlopesAndInterceptsIDlevel$Slope2))
{
  minval<-optimize(deriv_logit_quad,c(0,100),df_rawAge_randomSlopesAndInterceptsIDlevel$Slope2[index],df_rawAge_randomSlopesAndInterceptsIDlevel$Slope[index],df_rawAge_randomSlopesAndInterceptsIDlevel$Intercept[index]) 
  xinfl<-c(xinfl,minval$minimum)
  gradmins<-c(gradmins,minval$objective)
}
 
# Add to df_rawAge_randomSlopesAndInterceptsIDlevelframe and put out to CSV
df_rawAge_randomSlopesAndInterceptsIDlevel$SenesceRate<-gradmins
df_rawAge_randomSlopesAndInterceptsIDlevel$ageInfl<-xinfl

```

```{r bbal_birdis_diagnostics, eval = FALSE}

# Need to expand to all models #

# load("Data_outputs/bba_birdis_habitat_model_within_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_chl.RData")

load("Data_outputs/bba_birdis_habitat_model_between_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_chl.RData")


### Diagnostics ----

model_info <- list(
  within_sst = list(model = mod_within.bbal_bi.sst, data = pnts_subset),
  within_bathy = list(model = mod_within.bbal_bi.bathy, data = pnts_subset),
  within_chl = list(model = mod_within.bbal_bi.chl, data = pnts_subset)#,
  # between_sst = list(model = mod_between.bbal_bi.sst, data = pnts_all),
  # between_bathy = list(model = mod_between.bbal_bi.bathy, data = pnts_all),
  # between_chl = list(model = mod_between.bbal_bi.chl, data = pnts_all)
)

auc_results <- data.frame(
  model = character(),
  auc = numeric(),
  stringsAsFactors = FALSE
)

# Loop through models and compute AUC
for (model_name in names(model_info)) {
  this_model <- model_info[[model_name]]$model
  this_data <- model_info[[model_name]]$data
  
  preds <- predict(this_model, type = "response")
  auc_value <- as.numeric(auc(this_data$used, preds))
  
  auc_results <- rbind(auc_results, data.frame(model = model_name, auc = auc_value))
}



# AUC around 0.5 = essential random; closer to 1 = better fit
preds <- predict(mod_between.bbal_bi, type = "response")  
auc.mod <- as.numeric(auc(pnts_all$used, preds))

# Check for correlations
vif.mod <- car::vif(mod_between.bbal_bi)

# Check for patterns and that most points are within 95% confidence intervals
plot_binnedplot(mod_between.bbal_bi)

# Check ratio is close to 1
overdisp.mod <- overdisp_fun(mod_between.bbal_bi)

# Check ratio: >1 may indicate zero inflation (more 0s than expected); < 1 zero deflation
# Small P = evidence to suggest zero-inflation/deflation
zeroinf.mod <- testZeroInflation(mod_between.bbal_bi, plot = FALSE)

```

#### Model diagnostics

-   In the most simple model, season explained essentially no variation, so I dropped this.
-   The AUC value was \> 0.5, which suggests the model separates 'used' vs. 'not used' locations very well, or in other words, the predictors are strongly informative.
-   The VIF values are all \< 2, so there is no concerning multicollinearity.
-   The overdispersion function is slightly high. This could indicate moderate overdispersion (i.e. variance is higher than expected in a binomial model), but probably not a major concern.
-   There is some evidence for zero-inflation. However, this might be to expected, as zero counts are an inherent part of the process I am trying to model, and so a zero-inflated approach is likely unnecessary.

I am getting some convergence warnings, which, coupled with the high variance in my slopes, might suggest that the random effects can't be trusted. This could be because I have too many complex random effects, or sparse data per group. I should probably try to fit the random effects separately...

#### Model predictions

```{r model_outputs, fig.cap = "**Figure 1.3.** *Predicted probabilities of use for each fixed effect.*", fig.topcaption = TRUE}

# load("Data_outputs/bba_birdis_habitat_model_within_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_within_chl.RData")

load("Data_outputs/bba_birdis_habitat_model_between_sst.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_bathy.RData")
# load("Data_outputs/bba_birdis_habitat_model_between_chl.RData")

# Plot effects
pred_plot.sst <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "sst_scaled [all]"))
pred_plot.bathy <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "bathy_scaled [all]"))
pred_plot.chl <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "chl_scaled [all]"))
pred_plot.dist <- plot(ggeffects::ggpredict(mod_within.bbal_bi.sst, terms = "dist_col_scaled [all]"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(plot.title = element_blank()), 
          pred_plot.dist + labs(x = "Scaled distance from colony (km)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          ncol = 2, nrow = 2)

all_plots

```

The results show that in general, birds select for median SST, deeper waters, higher chlorophyll concentrations, and locations closer to the colony. There is a high degree of individual variation, especially in the use of SST and bathymetry.

#### Individual effects

```{r bbal_birdis_individual_within, fig.cap = "**Figure 1.4.** *Within-season, individual level responses to changes in the environmental variables. Individual slopes are plotted for the extent of each individual's range. Black line and grey interval shows population-level response and associated confidence interval.*", fig.topcaption = TRUE}

# Build plots
load("Data_outputs/bba_birdis_habitat_model_within_sst.RData")

env_vars = c("sst_scaled", "bathy_scaled", "chl_scaled")

for (i in 1:length(env_vars)) {

  habitat_var = env_vars[i]
  rnd_eff <- get_rnd_effects(mod_within.bbal_bi, habitat_var)
  
  # Get range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_sst = min(!!sym(habitat_var), na.rm = TRUE),
              max_sst = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(habitat_val)) 
  
  predictions.ind <- merge(predictions.ind, rnd_eff, by = "ID", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Get global predictions
  predictions.global <- data.frame(ggpredict(mod_within.bbal_bi, terms = paste0(habitat_var, " [all]")))
  
  ## Plot
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, colour = abs(slope), group = ID)) +
      scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      geom_line(data = predictions.global, aes(x = x, y = predicted), col = "black",
                 linetype = "dashed", linewidth = 1) +
      geom_ribbon(data = predictions.global, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "grey90", alpha = 0.5, col = "grey10") +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

ggarrange(sst_scaled_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) , 
          bathy_scaled_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.y = element_blank()), 
          chl_scaled_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")),
          ncol = 2, nrow = 2)


```

```{r bbal_birdis_individual_between, fig.cap = "**Figure 1.5.** *Between-season, individual level responses to changes in the environmental variables. Individual slopes are plotted for the extent of each individual's range. Black line and grey interval shows population-level response and associated confidence interval.*", fig.topcaption = TRUE}

# Build plots

env_vars = c("sst_scaled", "bathy_scaled", "chl_scaled")

for (i in 1:length(env_vars)) {

  habitat_var = env_vars[i]
  rnd_eff <- get_rnd_effects(mod_between.bbal_bi, habitat_var)
  
  # Get range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_sst = min(!!sym(habitat_var), na.rm = TRUE),
              max_sst = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(habitat_val)) 
  
  predictions.ind <- merge(predictions.ind, rnd_eff, by = "ID", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Get global predictions
  predictions.global <- data.frame(ggpredict(mod_between.bbal_bi, terms = paste0(habitat_var, " [all]")))
  
  ## Plot
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, colour = abs(slope), group = ID)) +
      scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      geom_line(data = predictions.global, aes(x = x, y = predicted), col = "black",
                 linetype = "dashed", linewidth = 1) +
      geom_ribbon(data = predictions.global, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "grey90", alpha = 0.5, col = "grey10") +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

ggarrange(sst_scaled_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) , 
          bathy_scaled_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.y = element_blank()), 
          chl_scaled_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")),
          ncol = 2, nrow = 2)


```

```{r bbal_birdis_calc_repeatability}

# Within season -----
summary(mod_within.bbal_bi)

## manual test
vc <- VarCorr(mod_within.bbal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.within <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.within <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.within <- V_ID_chl / (V_ID_chl + V_residual)



# Between season -----
summary(mod_between.bbal_bi)

## manual test
vc <- VarCorr(mod_between.bbal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.between <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.between <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.between <- V_ID_chl / (V_ID_chl + V_residual)


```

# Understanding within-individual centering

Currently, the model is fitted with within-population centering of the environmental variables - i.e., each variable has been scaled to the mean and 2SD of the entire datasets. Conversely, within-individual centering removes between-individual variation, allowing me to focus just on within-individual responses (i.e. plasticity).

Here I am trying a few different model structures to better understand how within-individual centering, random slopes, and quadratic random slopes change the predictions of my models. For simplicity, I have removed the distance effect and the phase effect.

```{r exploring_structures, eval = FALSE}

#set.seed(817)
#pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

#pnts_subset <- pnts_all

# Population centering, linear slope ----

mod_pop_linear <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled +
         (1 + sst_scaled | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

summary(mod_pop_linear)
save(mod_pop_linear, file = "Data_outputs/bba_birdis_habitat_model_pop_linear.RData")

# Population centering, quadratic slope ----

mod_pop_quad <- glmmTMB(
  used ~ sst_scaled + sst_scaled2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (sst_scaled + sst_scaled2 | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

summary(mod_pop_quad)
save(mod_pop_quad, file = "Data_outputs/bba_birdis_habitat_model_pop_quad.RData")


# Within-individual centering, linear slope ----

mod_ind_linear <- glmmTMB(
  used ~ sst_within + sst_within2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_within | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

save(mod_ind_linear, file = "Data_outputs/bba_birdis_habitat_model_ind_linear.RData")

# Within-individual centering, quadratic slope ----

mod_ind_quad <- glmmTMB(
  used ~ sst_within + sst_within2 + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (sst_within + sst_within2 | ID_season) + (1 | tripID),
  data = pnts_subset,
  family = binomial()
)

save(mod_ind_quad, file = "Data_outputs/bba_birdis_habitat_model_ind_quad.RData")


```

```{r exploring_structures_plots, eval = TRUE}

load("Data_outputs/bba_birdis_habitat_model_pop_linear.RData")
load("Data_outputs/bba_birdis_habitat_model_pop_quad.RData")
load("Data_outputs/bba_birdis_habitat_model_ind_linear.RData")
load("Data_outputs/bba_birdis_habitat_model_ind_quad.RData")

# set.seed(817)
# pnts_subset <- pnts_all[sample(nrow(pnts_all), 50000),]

# Get individual range
ind_ranges <- pnts_subset %>%
    mutate(ID_season = paste(ring, season, sep = "_")) %>%
    group_by(ID_season) %>%
    summarise(min_sst = min(sst_scaled, na.rm = TRUE),
              max_sst = max(sst_scaled, na.rm = TRUE)) 

predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(sst_scaled = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(sst_scaled)) %>%
    mutate(sst_scaled2 = sst_scaled^2)

predictions.ind_within <- pnts_subset %>% group_by(ID_season) %>% select(ID_season, sst_within, sst_within2)

# Population centering, linear slope ----
rnd_eff <- ranef(mod_pop_linear)
rnd_eff <- rnd_eff$cond$ID_season
rnd_eff$ID_season <- rownames(rnd_eff)
rownames(rnd_eff) <- NULL
rnd_eff %<>% rename(slope = sst_scaled, intercept = '(Intercept)')

predictions.ind_pop.linear <- merge(predictions.ind, rnd_eff, by = "ID_season", all.x = TRUE)
predictions.ind_pop.linear$logit <- predictions.ind_pop.linear$intercept + predictions.ind_pop.linear$slope * predictions.ind_pop.linear$sst_scaled
predictions.ind_pop.linear$probability <- 1 / (1 + exp(-predictions.ind_pop.linear$logit))  
  
## Plot
predictions.ind_pop.linear %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_pop.linear <- ggplot() +
    geom_line(data = predictions.ind_pop.linear, aes(x = sst_scaled, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
  scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use", x = "SST, scaled", title = "Linear random slope, population-level environmental centering")


# Population centering, quadratic slope ----

rnd_eff_pop_quad <- ranef(mod_pop_quad)
rnd_eff_pop_quad <- rnd_eff_pop_quad$cond$ID_season
rnd_eff_pop_quad$ID_season <- rownames(rnd_eff_pop_quad)
rownames(rnd_eff_pop_quad) <- NULL
rnd_eff_pop_quad %<>% rename(slope_linear = sst_scaled, slope_quadratic = sst_scaled2, intercept = '(Intercept)')

predictions.ind_pop.quad <- merge(predictions.ind, rnd_eff_pop_quad, by = "ID_season", all.x = TRUE)
predictions.ind_pop.quad$logit <- predictions.ind_pop.quad$intercept + 
                                  predictions.ind_pop.quad$slope_linear * predictions.ind_pop.quad$sst_scaled + 
                                  predictions.ind_pop.quad$slope_quadratic * predictions.ind_pop.quad$sst_scaled2
predictions.ind_pop.quad$probability <- 1 / (1 + exp(-predictions.ind_pop.quad$logit))

## Plot
predictions.ind_pop.quad %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_pop.quad <- ggplot() +
    geom_line(data = predictions.ind_pop.quad, aes(x = sst_scaled, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
    scale_colour_gradient(low = "#f4edff", high = "#340088") +  
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Probability of use", x = "SST, scaled", title = "Quadratic random slope, population-level environmental centering")


# Within-individual centering, linear slope ----

rnd_eff_ind_linear <- ranef(mod_ind_linear)
rnd_eff_ind_linear <- rnd_eff_ind_linear$cond$ID_season
rnd_eff_ind_linear$ID_season <- rownames(rnd_eff_ind_linear)
rownames(rnd_eff_ind_linear) <- NULL
rnd_eff_ind_linear %<>% rename(slope = sst_within, intercept = '(Intercept)')

predictions.ind_within.linear <- merge(predictions.ind_within, rnd_eff_ind_linear, by = "ID_season", all.x = TRUE)
predictions.ind_within.linear$logit <- predictions.ind_within.linear$intercept + predictions.ind_within.linear$slope * predictions.ind_within.linear$sst_within
predictions.ind_within.linear$probability <- 1 / (1 + exp(-predictions.ind_within.linear$logit))

## Plot
predictions.ind_within.linear %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_ind.linear <- ggplot() +
    geom_line(data = predictions.ind_within.linear, aes(x = sst_within, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
    scale_colour_gradient(low = "#f4edff", high = "#340088") +  
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Probability of use", x = "SST, within-individual centered", title = "Linear random slope, within-individual centering")

# Within-individual centering, quadratic slope ----

rnd_eff_ind_quad <- ranef(mod_ind_quad)
rnd_eff_ind_quad <- rnd_eff_ind_quad$cond$ID_season
rnd_eff_ind_quad$ID_season <- rownames(rnd_eff_ind_quad)
rownames(rnd_eff_ind_quad) <- NULL
rnd_eff_ind_quad %<>% rename(slope_linear = sst_within, slope_quadratic = sst_within2, intercept = '(Intercept)')

predictions.ind_within.quad <- merge(predictions.ind_within, rnd_eff_ind_quad, by = "ID_season", all.x = TRUE)
predictions.ind_within.quad$logit <- predictions.ind_within.quad$intercept + 
                                     predictions.ind_within.quad$slope_linear * predictions.ind_within.quad$sst_within + 
                                     predictions.ind_within.quad$slope_quadratic * predictions.ind_within.quad$sst_within2
predictions.ind_within.quad$probability <- 1 / (1 + exp(-predictions.ind_within.quad$logit))

## Plot
predictions.ind_within.quad %<>% separate(ID_season, into = c("ring", "season"), sep = "_", remove = FALSE) %>% mutate(ring_num = as.numeric(as.factor(ring)))

plot_ind.quad <- ggplot() +
    geom_line(data = predictions.ind_within.quad, aes(x = sst_within, y = probability, colour = ring_num, group = ID_season), alpha = 0.3) +
    scale_colour_gradient(low = "#f4edff", high = "#340088") +  
    theme_bw() +
    theme(legend.position = "none") +
    labs(y = "Probability of use", x = "SST, within-individual centered", title = "Quadratic random slope, within-individual centering")


# Arrange plots ----------

ggarrange(plot_pop.linear, plot_pop.quad,
          plot_ind.linear, plot_ind.quad,
          ncol = 2, nrow = 2)

#knitr::knit_exit()

```

# MAIN ANALYSIS

- Models need to have uncorrelated random slopes, as the intercept measures the sampling ratio. 

# Wandering albatrosses

### Bird Island

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than two day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID))  %>%
  group_by(ID) %>%
  mutate(n_recs = n_distinct(date_hourly),
         flag = ifelse(n_recs < 48, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-c(flag, n_recs)) 

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(ID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()


## Scale the variables
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
                     bathy_scaled2 = bathy_scaled^2)

# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/waal_birdis_habitat_data.RData")


```

#### Habitat use

The purpose of these analyses is to investigate whether individuals tend to vary in the average habitat they experience when foraging. This might be interpreted as their habitat 'preference'.

##### Fit the models

```{r waal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_birdis_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

# # To get a summary of trips per season per individual
# pnts.used.summary <- pnts.used %>% group_by(tripID) %>% filter(row_number() == 1)
# pnts_used.table <- table(pnts.used.summary$ID, pnts.used.summary$season)
# pnts_used.table <- cbind(pnts_used.table, rowSums(pnts_used.table))

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID)")), data = pnts.used)
  
  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID", ] %>% 
    rename(ring = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  ## Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_across <- var_across / var_total 
  
  # Outputs
  rpt_df <- rbind(rpt_df, data.frame(variable = env_vars[i], rpt_across = rpt_across))
  
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i]) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

var.df_all <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(var.df_all, file = "Data_outputs/waal_birdis_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/waal_birdis_habitat_used_ranef.RData")


```

##### Plot habitat use

```{r waal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/waal_birdis_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

make_plot <- function(df) {
  df <- df %>%
    mutate(ring_ordered = forcats::fct_reorder(ring, estimate))

  ggplot(df, aes(x = estimate, y = ring_ordered, col = env)) +
    geom_point() +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
    scale_color_manual(values = env_colours, guide = "none") +
    labs(title = unique(df$env),
      x = "Random intercept estimate",
      y = "Ring ID") +
    theme_bw() +
    theme(axis.text.y = element_blank())}

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
waal_birdis_used_plot <- annotate_figure(waal_birdis_used_plot, top = text_grob("WAAL - Bird Island", face = "bold", size = 14))
save(waal_birdis_used_plot, file = "Data_outputs/waal_birdis_used_plot.RData")

png(filename = "Figures//waal_birdis_habitat_used_ranef_plot.png", width = 14, height = 6, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()



```

```{r waal_birdis_table_variance_habitat_used}

load("Data_outputs/waal_birdis_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chl" ~ "Chlorphyll-a",
    TRUE ~ "SST" ))

kable(var.table, format = "html", 
  caption = 'Table 1A. Variance associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px solid;") %>%
  collapse_rows(columns = c(1)) %>%
  kable_classic(font_size = 12)

```

```{r table_rpt_habitat_used}

load("Data_outputs/habitat_used_repeatability.RData")

rpt.table <- rpt_df %>%
  rename(Variable = variable, "Within-season" = rpt_within, "Across-season" = rpt_across) %>%
  mutate(across(where(is.numeric), ~ round(.x, 2)))
  

kable(rpt.table, format = "html", 
      caption = 'Table 1B. Repeatability of habitat varible use, within- and across seasons.') %>%
      kable_classic(font_size = 12)

```

#### Habitat selection

##### Fit the models

There are only 7 individual WAAL at Bird Island with \> 1 season of data, so I dropped the ID:season models.

```{r waal_birdis_fit_models, eval = FALSE}

tok <- Sys.time()
waal_birdis_glmm.between_sst <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
         (1 | ID ) + (0 + sst_scaled | ID ) + (1 | tripID),
  data = pnts_all,
  family = binomial() )
tik <- Sys.time()

tik-tok

save(waal_birdis_glmm.between_sst, file = "Data_outputs/waal_birdis_glmm_between_sst.RData")
rm(waal_birdis_glmm.between_sst); gc()

waal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID ) + (0 + bathy_scaled | ID ) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.between_bathy, file = "Data_outputs/waal_birdis_glmm_between_bathy.RData")
rm(waal_birdis_glmm.between_bathy); gc()


waal_birdis_glmm.between_chl <-  glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID ) + (0 + chl_scaled | ID ) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.between_chl, file = "Data_outputs/waal_birdis_glmm_between_chl.RData")
rm(waal_birdis_glmm.between_chl); gc()

```

##### Diagnostics

```{r waal_birdis_diagnostics, eval = FALSE}

load("Data_outputs/waal_birdis_glmm_between_sst.RData")
load("Data_outputs/waal_birdis_glmm_between_bathy.RData")
load("Data_outputs/waal_birdis_glmm_between_chl.RData")


### Diagnostics ----

model_info <- list(
  between_sst =   list(model = waal_birdis_glmm.between_sst, data = pnts_all),
  between_bathy = list(model = waal_birdis_glmm.between_bathy, data = pnts_all),
  between_chl =   list(model = waal_birdis_glmm.between_chl, data = pnts_all),
)

auc_results <- data.frame(
  model = character(),
  auc = numeric()  )

# Loop through models and compute AUC, output diagnostics
for (model_name in names(model_info)) {

  this_model <- model_info[[model_name]]$model
  this_data <- model_info[[model_name]]$data
  
  print(paste0("Next model: ", model_name, " ------------------------------------"))
  
  print("Covariance structure: ")
  print(VarCorr(this_model))
  
  print("Collinearity: ")
  print(performance::check_collinearity(this_model))
  
  print(paste0("DONE. ---------------------------"))
  
  preds <- predict(this_model, type = "response")
  auc_value <- as.numeric(auc(this_data$used, preds))
  
  auc_results <- rbind(auc_results, data.frame(model = model_name, auc = auc_value))
}

auc_results %<>% mutate(species = "WAAL", colony = "birdis") %>% relocate(c(species, colony), .before = model)


```

##### Fixed effects

```{r waal_birdis_fixed_effects_predictions }

load("Data_outputs/waal_birdis_glmm_between_sst.RData")
load("Data_outputs/waal_birdis_glmm_between_bathy.RData")
load("Data_outputs/waal_birdis_glmm_between_chl.RData")

# Plot effects
pred_plot.sst <- plot(ggeffects::ggpredict(waal_birdis_glmm.between_sst, terms = "sst_scaled [all]", type = "fixed"))
pred_plot.bathy <- plot(ggeffects::ggpredict(waal_birdis_glmm.between_bathy, terms = "bathy_scaled [all]", type = "fixed"))
pred_plot.chl <- plot(ggeffects::ggpredict(waal_birdis_glmm.between_chl, terms = "chl_scaled [all]", type = "fixed"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)", y = "Used")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)", y = "") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")", y = "")) + theme(plot.title = element_blank()),
          ncol = 3, nrow = 1)

png("Figures/waal_birdis_habitat_fixed_effects.png", width = 14, height = 6, units = "in", res = 600)
all_plots
dev.off()

rm(pred_plot.sst); rm(pred_plot.bathy); rm(pred_plot.chl); rm(all_plots)
gc()

```


##### Conditional repeatability

Repeatability at the edges of chlorophyll is exceptionally high, but almost certainly not real (indeed, how could it be \> 100?). The computation of condition repeatability means that as you move away from the mean, the x2 term grows, but if there are no data points at that x, it's extrapolating a parabola that has no real support i.e. the slope is being estimated from the small range in which birds do forage, and then extrapolating where they don't.

```{r waal_birdis_calc_repeatability}

load("Data_inputs/waal_birdis_habitat_data.RData")

# Calculate conditional repeatability ------

## Store envronmental variables
env_vars = c("sst", "bathy", "chl")

sst_scaled <- unlist(pnts_all$sst_scaled)
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)

var_map <- list(
  sst = sst_scaled,
  bathy = bathy_scaled,
  chl = chl_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_between = get(paste0("waal_birdis_glmm.between_", env_vars[i]))
  
  # Calculate conditional repeatability parameters
  
  env_current <- var_map[[env_vars[i]]]
  
  ## Between ##
 
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current) )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Bird Island", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- res_between.summary %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = "#FFC107", size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = "#f3a20c") +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_between)
  rm(list = paste0("waal_birdis_glmm.between_", env_vars[i]))
  gc()
  
}

condR_results <- do.call("rbind", cond_res.list)
save(condR_results, file = "Data_outputs/waal_birdis_condR_results.RData")

# Plot individual slopes 

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/waal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_between = get(paste0("waal_birdis_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <-rnd_eff.between
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}


png("Figures/waal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
ggarrange(bathy_pred_plot + labs(x = "Scaled bathymetry (m below sea level)"), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank()),
          bathy_condR_plot + labs(x = "Scaled bathymetry (m below sea level)"),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"))+ theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)
dev.off()

```

### Crozet

Chlorophyll values are heavily skewed, with extremely high concentrations occurring near coastlines in the ‘available’ dataset. To prevent these large outliers from unduly influencing the model results, I apply a log-transformation to chlorophyll.

```{r waal_cro_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_crozet_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than two day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID))  %>%
  group_by(ID) %>%
  mutate(n_recs = n_distinct(date_hourly),
         flag = ifelse(n_recs < 48, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-c(flag, n_recs)) 

### TESTING - MAP HIGH CHLOROPHYLL

# temp<-subset(pnts_all, chlA > 3)
# 
# ggplot(temp, aes(x = longitude, y = latitude, color = chlA)) +
#   geom_point(alpha = 0.5) +
#   scale_color_viridis_c(option = "inferno", na.value = "grey90") +
#   labs(
#     color = "Chl-a (mg/m³)",
#     title = "Where are high chlorophyll values?"
#   ) +
#   theme_minimal()
##

## Scale the variables
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         
         bathy_scaled = scale_2SD(bathy),
         chl_scaled_log = scale(log(chlA + 0.001)),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
                     bathy_scaled2 = bathy_scaled^2) 

save(pnts_all, file = "Data_inputs/waal_cro_habitat_data.RData")

```

#### Habitat use

##### Fit the models

```{r waal_cro_model_average_habitat_used, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

# # To get a summary of trips per season per individual
# pnts.used.summary <- pnts.used %>% group_by(tripID) %>% filter(row_number() == 1)
# pnts_used.table <- table(pnts.used.summary$ID, pnts.used.summary$season)
# pnts_used.table <- cbind(pnts_used.table, rowSums(pnts_used.table))

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.used)
  
  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"]
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

var.df_all <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(var.df_all, file = "Data_outputs/waal_cro_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/waal_cro_habitat_used_ranef.RData")

```

##### Plot habitat use

```{r waal_cro_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/waal_cro_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

make_plot <- function(df) {
  df <- df %>%
    mutate(ring_ordered = forcats::fct_reorder(ID, estimate))

  ggplot(df, aes(x = estimate, y = ring_ordered, col = env)) +
    geom_point() +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
    scale_color_manual(values = env_colours, guide = "none") +
    facet_wrap(~group, scales = "free") +
    labs(title = unique(df$env),
      x = "Random intercept estimate",
      y = "ID") +
    theme_bw() +
    theme(axis.text.y = element_blank())}

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_cro_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
waal_cro_used_plot <- annotate_figure(waal_cro_used_plot, top = text_grob("WAAL - Crozet", face = "bold", size = 14))
save(waal_cro_used_plot, file = "Data_outputs/waal_cro_used_plot.RData")

png(filename = "Figures/waal_cro_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r waal_cro_table_variance_habitat_used}

load("Data_outputs/waal_cro_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

# Add whitespace to avoid collapsing bathymetry & chlorophyll together
var.table$`Across season rpt`[6:10] <- paste0(var.table$`Across season rpt`[6:10], ".00")

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

#### Habitat selection

##### Fit the models

The random slope for chlorophyll-a was retained to account for potential individual variation in response to environmental conditions. However, the estimated correlation between the random intercept and slope for chl_scaled was extremely high (0.99), suggesting that these components are statistically confounded in the current model. This may reflect either collinearity in the data or insufficient within-individual variation in chlorophyll exposure to reliably separate baseline use from individual responses.

As a result, I have droppped the polynomial terms as I think the models are unable to cope with this degree of complexity.

```{r waal_cro_fit_models, eval = FALSE}

load("Data_inputs/waal_cro_habitat_data.RData")

# try logging chl-A

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

# Within season #

tok <- Sys.time()
waal_cro_glmm.within_sst <- glmmTMB(
   used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase + 
         (1 | ID_season) + (0 + sst_scaled | ID_season ) + (1 | tripID),
  data = pnts_all,
  family = binomial() )
tik <- Sys.time()

tik-tok

save(waal_cro_glmm.within_sst, file = "Data_outputs/waal_cro_glmm_within_sst.RData")
rm(waal_cro_glmm.within_sst); gc()

waal_cro_glmm.within_bathy <- glmmTMB(
   used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase + 
        (1 | ID_season) + (0 + bathy_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_bathy, file = "Data_outputs/waal_cro_glmm_within_bathy.RData")
rm(waal_cro_glmm.within_bathy); gc()


waal_cro_glmm.within_chl <-  glmmTMB(
   used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase +  
        (1 | ID_season) + (0 + chl_scaled_log | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_chl, file = "Data_outputs/waal_cro_glmm_within_chl.RData")
rm(waal_cro_glmm.within_chl); gc()

# Between season #

waal_cro_glmm.between_sst <-  glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase +
        (1 | ID) + (0  + sst_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_sst, file = "Data_outputs/waal_cro_glmm_between_sst.RData")
rm(waal_cro_glmm.between_sst); gc()

waal_cro_glmm.between_bathy <- glmmTMB(
   used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase + 
        (1 | ID) + (0 + bathy_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_bathy, file = "Data_outputs/waal_cro_glmm_between_bathy.RData")
rm(waal_cro_glmm.between_bathy); gc()


waal_cro_glmm.between_chl <- glmmTMB(
   used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase + 
        (1 | ID) + (0 + chl_scaled_log | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_chl, file = "Data_outputs/waal_cro_glmm_between_chl.RData")

# waal_cro_glmm.between_chl.log <- glmmTMB(
#    used ~ sst_scaled + bathy_scaled + chl_scaled_log + dist_col_scaled + phase + 
#         (1 | ID) + (0 + chl_scaled_log | ID) + (1 | tripID),
#   data = pnts_all,
#   family = binomial() )
# 
# save(waal_cro_glmm.between_chl.log, file = "Data_outputs/waal_cro_glmm_between_chl_log.RData")

rm(waal_cro_glmm.between_chl); gc()

```

#### Diagnostics

```{r waal_cro_diagnostics, eval = FALSE}

load("Data_outputs/waal_cro_glmm_within_sst.RData")
load("Data_outputs/waal_cro_glmm_within_bathy.RData")
load("Data_outputs/waal_cro_glmm_within_chl.RData")

load("Data_outputs/waal_cro_glmm_between_sst.RData")
load("Data_outputs/waal_cro_glmm_between_bathy.RData")
load("Data_outputs/waal_cro_glmm_between_chl.RData")


model_info <- list(
  within_sst =    list(model = waal_cro_glmm.within_sst, data = pnts_all),
  within_bathy =  list(model = waal_cro_glmm.within_bathy, data = pnts_all),
  within_chl =    list(model = waal_cro_glmm.within_chl, data = pnts_all),
  between_sst =   list(model = waal_cro_glmm.between_sst, data = pnts_all),
  between_bathy = list(model = waal_cro_glmm.between_bathy, data = pnts_all),
  between_chl =   list(model = waal_cro_glmm.between_chl, data = pnts_all)
)

auc_results <- data.frame(
  model = character(),
  auc = numeric()  )

# Loop through models and compute AUC, output diagnostics
for (model_name in names(model_info)) {

  this_model <- model_info[[model_name]]$model
  this_data <- model_info[[model_name]]$data
  
  print(paste0("Next model: ", model_name, " ------------------------------------"))
  
  print("Covariance structure: ")
  print(VarCorr(this_model))
  
  print("Collinearity: ")
  print(performance::check_collinearity(this_model))
  
  print(paste0("DONE. ---------------------------"))
  
  preds <- predict(this_model, type = "response")
  auc_value <- as.numeric(auc(this_data$used, preds))
  
  auc_results <- rbind(auc_results, data.frame(model = model_name, auc = auc_value))
}

auc_results %<>% mutate(species = "WAAL", colony = "crozet") %>% relocate(c(species, colony), .before = model)


```

#### Fixed effects

```{r waal_cro_fixed_effects_predictions }

load("Data_outputs/waal_cro_glmm_between_sst.RData")
load("Data_outputs/waal_cro_glmm_between_bathy.RData")
load("Data_outputs/waal_cro_glmm_between_chl.RData")


# Plot effects
pred_plot.sst <-   plot(ggeffects::ggpredict(waal_cro_glmm.between_sst, terms = "sst_scaled [all]", type = "fixed"))
pred_plot.bathy <- plot(ggeffects::ggpredict(waal_cro_glmm.between_bathy, terms = "bathy_scaled [all]", type = "fixed"))
pred_plot.chl <-  plot(ggeffects::ggpredict(waal_cro_glmm.between_chl, terms = "chl_scaled [all]", type = "fixed"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)", y = "Used")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)", y = "Used") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")", y = "Used")) + theme(plot.title = element_blank()),
          ncol = 2, nrow = 2)

png("Figures/waal_cro_habitat_fixed_effects.png")
all_plots
dev.off()

rm(pred_plot.sst); rm(pred_plot.bathy); rm(pred_plot.chl); rm(all_plots)
gc()

```

#### Conditional repeatability

```{r waal_cro_conditional_repeatability}

load("Data_inputs/waal_cro_habitat_data.RData")

env_vars = c("sst", "bathy", "chl")

# Calculate conditional repeatability ------

## Store envronmental variables
sst_scaled <- unlist(pnts_all$sst_scaled)
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled_log)

var_map <- list(
  sst = sst_scaled,
  bathy = bathy_scaled,
  chl = chl_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))
  
  # Calculate conditional repeatability parameters
  
  env_current <- var_map[[env_vars[i]]]
  
  ## Within ##
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
  nrow = 2) 
   
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var_between = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Crozet", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
   
  ## Between ##
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "WAAL",
                            colony = "Crozet", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = "#FFC107", size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = "#D81B60", size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = "#f3a20c") +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = "#b31650") +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("waal_cro_glmm.within_", env_vars[i]))
  rm(list = paste0("waal_cro_glmm.between_", env_vars[i]))
  gc()
  
}

condR_results <- do.call("rbind", cond_res.list)
save(condR_results, file = "Data_outputs/waal_cro_condR_results.RData")

# Plot individual slopes 

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Data_outputs/waal_cro_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/waal_cro_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between)
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}


png("Figures/waal_cro_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
ggarrange(bathy_pred_plot + labs(x = "Scaled bathymetry (m below sea level)"), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank()),
          bathy_condR_plot + labs(x = "Scaled bathymetry (m below sea level)"),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled & logged chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"))+ theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)
dev.off()

```

#### Heritability of habitat use

```{r waal_cro_heritability }

library(MCMCglmm)
library(brms) 
library(igraph)

# (1) Load and subsample the data
load("Data_inputs/waal_cro_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)
pnts.sampled <- sample_n(pnts.used, 200000)

pedigree.df <- read.csv("Data_inputs/waal_crozet_pedigree.csv")
pedigree.df %<>% dplyr::select(-Year)


# (2) Check the pedigree

## Are the ID in pnts.used the same as id in pedigree.df?
pedigree.df$id <- as.character(pedigree.df$id)
pnts.used$ID <- as.character(pnts.used$ID)

model_ids <- unique(pnts.used$ID)
pedigree_ids <- pedigree.df$id

### IDs in model AND pedigree
ids_in_both <- intersect(model_ids, pedigree_ids)

### IDs in model NOT in pedigree
ids_missing_from_pedigree <- setdiff(model_ids, pedigree_ids)

### IDs in pedigree NOT in model (optional info)
ids_unused_in_model <- setdiff(pedigree_ids, model_ids)


# (3) Subset the pedigree to include all ancestors

## Check formatting
pedigree.df$id <- as.character(pedigree.df$id)
pedigree.df$dam <- as.character(pedigree.df$dam)
pedigree.df$sire <- as.character(pedigree.df$sire)
pnts.used$ID <- as.character(pnts.used$ID)

## Function to recursively get ancestors
get_ancestors <- function(ids, pedigree) {
  all_ids <- ids
  new_ids <- ids
  
  repeat {
    # Get dams and sires of current new_ids
    parents <- pedigree[pedigree$id %in% new_ids, c("dam", "sire")]
    parents <- unique(na.omit(unlist(parents)))
    
    # Stop if no new parents found
    parents <- setdiff(parents, all_ids)
    if (length(parents) == 0) break
    
    # Add new parents to the list
    all_ids <- c(all_ids, parents)
    new_ids <- parents
  }
  
  return(unique(all_ids))
}

model_ids <- unique(pnts.used$ID)
ped_ids_needed <- get_ancestors(model_ids, pedigree.df)
length(ped_ids_needed)

## Subset the pedigree
pedigree.pruned <- pedigree.df[pedigree.df$id %in% ped_ids_needed, ]

## Check for missing IDs
setdiff(model_ids, pedigree.pruned$id) 

## Add these as founders
missing_ids <- setdiff(model_ids, pedigree.pruned$id)
founders <- data.frame(id = missing_ids, dam = NA, sire = NA)

### Append to your original pedigree
pedigree.extended <- rbind(pedigree.df, founders)

### Now prune the extended version
ped_ids_needed <- get_ancestors(model_ids, pedigree.extended)
pedigree.pruned <- pedigree.extended[pedigree.extended$id %in% ped_ids_needed, ]

setdiff(model_ids, pedigree.pruned$id)

# (4) Visualise the pruned pedigree

edges <- pedigree.pruned %>%
  tidyr::pivot_longer(cols = c("sire", "dam"), values_drop_na = TRUE) %>%
  dplyr::select(from = value, to = id)

## Build igraph object
ped_graph <- graph_from_data_frame(edges, directed = TRUE)

plot(ped_graph,
     vertex.size = 3,
     vertex.label = NA,
     edge.arrow.size = 0.2,
     layout = layout_with_sugiyama(ped_graph)$layout)


# (5) Look at the genetic matrix A

ped <- pedigree.pruned
names(ped) <- c("animal", "dam", "sire")

Ainv <- inverseA(ped)$Ainv
A <- as.matrix(solve(Ainv))

## Inspect sparse matrix
A_sparse <- as(A, "dgCMatrix")
summary(A_sparse)

### Check how many non-zero off-diagonal elements
nnzero_offdiag <- sum(A_sparse != 0) - nrow(A_sparse)
density_offdiag <- nnzero_offdiag / (nrow(A_sparse)^2 - nrow(A_sparse))
cat("Off-diagonal density:", round(density_offdiag * 100, 4), "%\n")

library(pheatmap)
pheatmap(as.matrix(A), show_rownames = FALSE, show_colnames = FALSE)



# (6) Fit the model
pnts.used$animal <- pnts.used$ID

mod_brm <- brm(
  formula = bf(
    sst_scaled ~ dist_col + phase +
      (1 | season) +
      (1 | tripID) +
      (1 | gr(animal, cov = A)) ),
  data = pnts.used,
  family = gaussian(),  
  data2 = list(A = A),  # your pedigree relatedness matrix
  cores = 4, chains = 4, iter = 4000,
  control = list(adapt_delta = 0.95))

mod_noA <- brm(
  sst_scaled ~ dist_col + phase + (1 | season) + (1 | tripID) + (1 | animal),
  data = pnts.heritable,
  family = gaussian(),
  chains = 4, cores = 4, iter = 4000)


mod_simple <- brm(
  sst_scaled ~ (1 | gr(animal, cov = A)),
  data = pnts.heritable,
  family = gaussian(),
  data2 = list(A = A) )

# subsampled version took 24 hours

save(mod_brm, file = "Data_outputs/TEMP_waal_cro_hertiability.RData")
save(mod_noA, file = "Data_outputs/TEMP_waal_cro_hertiability_noPedigree.RData")
save(mod_simple, file = "Data_outputs/TEMP_waal_cro_hertiability_simple.RData")
# to do this on a slope would require
# (1 + sst_scaled | gr(animal, cov = A))

plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
pp_check(mod_brm)

pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

posterior <- as_draws_df(mod_brm)
ggplot(posterior, aes(x = sd_animal__Intercept)) +
  geom_density(fill = "skyblue", alpha = 0.5) +
  labs(
    title = "Posterior for additive genetic SD",
    x = "SD(animal)",
    y = "Density" )


```

## Black-browed albatrosses

### Bird Island

```{r bbal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/bbal_birdis_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points (i.e. fewer than a day's worth of data)
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID))  %>%
  group_by(ID_season) %>%
  mutate(n_recs = n_distinct(date_hourly),
         flag = ifelse(n_recs < 24, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-c(flag, n_recs)) 

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(ID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()

## Scale the variables
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
                     bathy_scaled2 = bathy_scaled^2) 

used_avail_summary <- pnts_all %>%
  group_by(ID) %>%
  summarise(
    used_count = sum(used == 1),
    avail_count = sum(used == 0),
    total = n(),
    used_ratio = used_count / total )

hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_birdis_habitat_data.RData")


```

#### Habitat use

##### Fit the models

```{r bbal_birdis_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) ")), data = pnts.used)
  
  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level, intercept = estimate) %>%
    mutate(env = env_vars[i],
           estimate = fixef(mod)[1] + intercept,
           season_type = ifelse(group == "ID_season", "Within season", "Between seasons")) %>%
    separate(ID, into = c("ring", "season"), sep = "_", remove = FALSE) %>%
    select(-c(term, effect)) %>%
    relocate(estimate, .before = intercept)
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_within <- var_components[var_components$grp == "ID_season", "vcov"]
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_within <- var_within / var_total 
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_within = rpt_within,
            rpt_across = rpt_across) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

var.df_all <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(var.df_all, file = "Data_outputs/bbal_birdis_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/bbal_birdis_habitat_used_ranef.RData")

```

##### Plot habitat use

```{r bbal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

make_plot <- function(df) {
  df <- df %>%
    mutate(ring_ordered = forcats::fct_reorder(ID, intercept))

  ggplot(df, aes(x = intercept, y = ring_ordered, col = env)) +
    geom_point() +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
    scale_color_manual(values = env_colours, guide = "none") +
    facet_wrap(~group, scales = "free") +
    labs(title = unique(df$env),
      x = "Random intercept estimate",
      y = "ID") +
    theme_bw() +
    theme(axis.text.y = element_blank())}

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_birdis_used_plot <- annotate_figure(bbal_birdis_used_plot, top = text_grob("BBAL - Bird Island", face = "bold", size = 14))
save(bbal_birdis_used_plot, file = "Data_outputs/bbal_birdis_used_plot.RData")

png(filename = "Figures/bbal_birdis_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r bbal_birdis_table_variance_habitat_used}

load("Data_outputs/bbal_birdis_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

#### Habitat selection

##### Fit the models

```{r bbal_birdis_fit_models, eval = FALSE}

load("Data_inputs/bbal_birdis_habitat_data.RData")

# Within season #

bbal_birdis_glmm.within_sst <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
         (1 | ID_season) + (0 + sst_scaled | ID_season ) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_sst, file = "Data_outputs/bbal_birdis_glmm_within_sst.RData")
rm(bbal_birdis_glmm.within_sst); gc()


bbal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID_season) + (0 + bathy_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_bathy, file = "Data_outputs/bbal_birdis_glmm_within_bathy.RData")
rm(bbal_birdis_glmm.within_bathy); gc()


bbal_birdis_glmm.within_chl <-  glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID_season) + (0 + chl_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_chl, file = "Data_outputs/bbal_birdis_glmm_within_chl.RData")
rm(bbal_birdis_glmm.within_chl); gc()

# Between season #

bbal_birdis_glmm.between_sst <-  glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID) + (0 + sst_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_birdis_glmm.between_sst, file = "Data_outputs/bbal_birdis_glmm_between_sst.RData")
rm(bbal_birdis_glmm.between_sst); gc()


bbal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID) + (0 + bathy_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_bathy, file = "Data_outputs/bbal_birdis_glmm_between_bathy.RData")
rm(bbal_birdis_glmm.between_bathy); gc()


bbal_birdis_glmm.between_chl <- glmmTMB(
  used ~ sst_scaled + bathy_scaled + chl_scaled + dist_col_scaled + phase +
        (1 | ID) + (0 + chl_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_chl, file = "Data_outputs/bbal_birdis_glmm_between_chl.RData")
rm(bbal_birdis_glmm.between_chl); gc()

```

##### Diagnostics

```{r bbal_birdis_diagnostics, eval = FALSE}

load("Data_outputs/bbal_birdis_glmm_within_sst.RData")
load("Data_outputs/bbal_birdis_glmm_within_bathy.RData")
load("Data_outputs/bbal_birdis_glmm_within_chl.RData")

load("Data_outputs/bbal_birdis_glmm_between_sst.RData")
load("Data_outputs/bbal_birdis_glmm_between_bathy.RData")
load("Data_outputs/bbal_birdis_glmm_between_chl.RData")


model_info <- list(
  within_sst =    list(model = bbal_birdis_glmm.within_sst, data = pnts_all),
  within_bathy =  list(model = bbal_birdis_glmm.within_bathy, data = pnts_all),
  within_chl =    list(model = bbal_birdis_glmm.within_chl, data = pnts_all),
  between_sst =   list(model = bbal_birdis_glmm.between_sst, data = pnts_all),
  between_bathy = list(model = bbal_birdis_glmm.between_bathy, data = pnts_all),
  between_chl =   list(model = bbal_birdis_glmm.between_chl, data = pnts_all)
)

auc_results <- data.frame(
  model = character(),
  auc = numeric()  )

# Loop through models and compute AUC, output diagnostics
for (model_name in names(model_info)) {

  this_model <- model_info[[model_name]]$model
  this_data <- model_info[[model_name]]$data
  
  print(paste0("Next model: ", model_name, " ------------------------------------"))
  
  print("Covariance structure: ")
  print(VarCorr(this_model))
  
  print("Collinearity: ")
  print(performance::check_collinearity(this_model))
  
  print(paste0("DONE. ---------------------------"))
  
  preds <- predict(this_model, type = "response")
  auc_value <- as.numeric(auc(this_data$used, preds))
  
  auc_results <- rbind(auc_results, data.frame(model = model_name, auc = auc_value))
}

auc_results %<>% mutate(species = "BBAL", colony = "birdis") %>% relocate(c(species, colony), .before = model)


```

##### Fixed effects

```{r bbal_birdis_fixed_effects_predictions }

load("Data_outputs/bbal_birdis_glmm_between_sst.RData")
load("Data_outputs/bbal_birdis_glmm_between_bathy.RData")
load("Data_outputs/bbal_birdis_glmm_between_chl.RData")

# Plot effects
pred_plot.sst <-   plot(ggeffects::ggpredict(bbal_birdis_glmm.between_sst, terms = "sst_scaled [all]", type = "fixed"))
pred_plot.bathy <- plot(ggeffects::ggpredict(bbal_birdis_glmm.between_bathy, terms = "bathy_scaled [all]", type = "fixed"))
pred_plot.chl <-  plot(ggeffects::ggpredict(bbal_birdis_glmm.between_chl, terms = "chl_scaled [all]", type = "fixed"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)", y = "Used")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)", y = "Used") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")", y = "Used")) + theme(plot.title = element_blank()),
          ncol = 3, nrow = 1)

png("Figures/bbal_birdis_habitat_fixed_effects.png", width = 9, height = 4, units = "in", res = 600)
all_plots
dev.off()

rm(pred_plot.sst); rm(pred_plot.bathy); rm(pred_plot.chl); rm(all_plots)
gc()

```

##### Conditional repeatability

```{r bbal_birdis_cond_repeatability}

load("Data_inputs/bbal_birdis_habitat_data.RData")

# Calculate conditional repeatability ------

## Store envronmental variables
env_vars = c("sst", "bathy", "chl")

sst_scaled <- unlist(pnts_all$sst_scaled)
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)

var_map <- list(
  sst = sst_scaled,
  bathy = bathy_scaled,
  chl = chl_scaled
)

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars[i]))
  
  # Calculate conditional repeatability parameters
  
  env_current <- var_map[[env_vars[i]]]
  
  ## Within ##
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_within)$cond$ID_season), 0,
    0, as.numeric(VarCorr(mod_within)$cond$ID_season.1)),
    nrow = 2) 
  
  res_within <- condR(
    betaX = fixef(mod_within)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_within)$cond$tripID),  
    xrange = range(env_current)
  )
  
  res_within.summary <- data.frame(var_between = res_within$Vi,
                            R2 = res_within$R2s, 
                            marginal_R2 = res_within$Rmar,
                            med_rpt = median(res_within$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Bird Island", 
                            level = "within")
  
  rownames(res_within.summary) <- NULL
   
  ## Between ##
  
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current)  )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Bird Island", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- rbind(res_within.summary, res_between.summary) %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_line(data = res_between$condR, aes(x=x, y=condR/100), col = "#FFC107", size = 1) +
    geom_line(data = res_within$condR, aes(x=x, y=condR/100), col = "#D81B60", size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = "#f3a20c") +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = "#b31650") +
    labs(y = "Conditional repeatability") +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_birdis_glmm.within_", env_vars[i]))
  rm(list = paste0("bbal_birdis_glmm.between_", env_vars[i]))
  gc()
  
}

condR_results <- do.call("rbind", cond_res.list)
save(condR_results, file = "Data_outputs/bbal_birdis_condR_results.RData")

# Plot individual slopes ----

rnd_effs.list <- list()

for (i in 1:length(env_vars)) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_birdis_glmm_within_", env_vars[i], ".RData"))
  load(paste0("Data_outputs/bbal_birdis_glmm_between_", env_vars[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)
save(rnd_effs.out, file = "Data_outputs/bbal_birdis_habitat_selection.RData")

# Make the plot ----
png("Figures/bbal_birdis_habitat_conditional_repeatability.png", width = 14, height = 9, units = "in", res = 600)
ggarrange(bathy_pred_plot + labs(x = "Scaled bathymetry (m below sea level)"), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank()),
          bathy_condR_plot + labs(x = "Scaled bathymetry (m below sea level)"),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"))+ theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)
dev.off()

```

#### Extract individual estimates

```{r bbal_birdis_extract_ind_estimates}

load("Data_outputs/bbal_birdis_habitat_selection.RData")
load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")

ranefs.df2 <- ranefs.df %>% 
  select(c(season_type, env, season, ring, estimate, intercept)) %>%
  rename(used_estimate = estimate,
         used_intercept = intercept) %>%
  mutate(env = case_when( env == "SST" ~ "sst",
                          env == "chlA" ~ "chl",
                          env == "bathy" ~ "bathy"))

rnd_effs.out2 <- rnd_effs.out %>%
  select(c(season_type, env, season, ring, intercept, slope)) %>%
  rename(selection_intercept = intercept,
         selection_slope = slope)

ind_estimates <- merge(ranefs.df2, rnd_effs.out2, by = c("ring", "season", "season_type", "env"), all = T)
write.csv(ind_estimates, file = "Data_outputs/bbal_birdis_habitat_individual_estimates.csv", row.names = FALSE)

```

### Kerguelen

For Kerguelen especially, birds' preference for shelf zones makes the bathymetry model fall apart. To tackle this, I have calculated distance from the shelf, and split preferences into on shelf, shelf break, and off shelf. This seems to resolve my problems. The shelf break is delimited at 1000m following Hindell et al. 2011.

```{r bbal_ker_process_data, eval = FALSE}

load("Data_inputs/bbal_kerguelen_sst-bathy-chla_data_population_level.RData")

## Remove NAs and set distance to km; remove birds with too few data points
pnts_all %<>% 
  filter(!is.na(SST) & !is.na(used) & !is.na(bathy) &!is.na(chlA) & !is.na(phase)) %>%
  mutate(dist_col.km = dist_col/1000) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID))  %>%
  group_by(ID_season) %>%
  mutate(n_recs = n_distinct(date_hourly),
         flag = ifelse(n_recs < 24, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-c(flag, n_recs)) 

## Resample to maintain used:available ratio
pnts_all %<>%
  group_by(ID, date_hourly) %>%
  group_modify(~ {
    used_pts <- .x %>% filter(used == 1)
    avail_pts <- .x %>% filter(used == 0)
    # How many used?
    n_used <- nrow(used_pts)
    target_avail <- n_used * 6
    # If enough available points
    if (nrow(avail_pts) >= target_avail) {
      avail_pts <- avail_pts %>% slice_sample(n = target_avail)
      bind_rows(used_pts, avail_pts)
    } else {
      # Not enough: drop the whole group
      tibble()
    }
  }) %>%
  ungroup()


## Scale the variables
pnts_all %<>%
  ungroup() %>%
  mutate(sst_scaled = scale_2SD(SST),
         bathy_scaled = scale_2SD(bathy),
         chl_scaled = scale_2SD(chlA),
         dist_col_scaled = scale_2SD(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 ),
         sst_scaled2 = sst_scaled^2,
                     bathy_scaled2 = bathy_scaled^2)

## Add binomial bathymetry variable
pnts_all %<>%
  mutate(dist_shelf_bin = case_when(
            dist_shelf_break_signed > 20  ~ "on_shelf",       # >20 km inside shelf
            dist_shelf_break_signed >= -20 & dist_shelf_break_signed <= 20 ~ "near_break", # ±20 km of break
            dist_shelf_break_signed < -20  ~ "off_shelf"       # >20 km offshore (deeper)
    ) )

pnts_all$dist_shelf_bin <- factor(pnts_all$dist_shelf_bin, levels = c("on_shelf", "near_break", "off_shelf"))


# used_avail_summary <- pnts_all %>%
#   group_by(ID) %>%
#   summarise(
#     used_count = sum(used == 1),
#     avail_count = sum(used == 0),
#     total = n(),
#     used_ratio = used_count / total )
# 
# hist(used_avail_summary$used_ratio)

save(pnts_all, file = "Data_inputs/bbal_ker_habitat_data.RData")


pnts_subset <- pnts_all[sample(nrow(pnts_all), 20000),]

ggplot() +
  geom_histogram(data = pnts_subset, aes(x = bathy, fill = dist_shelf_bin)) +
  scale_fill_manual(values = c("lightblue", "blue", "darkblue")) +
  labs(x = "Bathymetry (m below sea level)") +
  theme_bw() +
  theme(legend.position = c(0.1, 0.85),
        legend.background = element_blank())

```


#### Habitat use

##### Fit the models

```{r bbal_ker_model_average_habitat_used, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

pnts.used <- pnts_all %>% filter(used == 1)

# Visualising bathymetry
ggplot(pnts.used, aes(x = bathy, fill = dist_shelf_bin)) +
  geom_histogram(col = "black", bins = 50) +
  scale_fill_manual(values = c("darkblue", "#86B8B8", "#B7DCEB")) +
  theme_bw() +
  theme(legend.position = c(0.1, 0.85),
        legend.background = element_blank())


# # To get a summary of trips per season per individual
# pnts.used.summary <- pnts.used %>% group_by(tripID) %>% filter(row_number() == 1)
# pnts_used.table <- table(pnts.used.summary$ID, pnts.used.summary$season)
# pnts_used.table <- cbind(pnts_used.table, rowSums(pnts_used.table))

env_vars <- c("SST", "bathy", "chlA")
rpt_df <- data.frame(env_var = character(), rpt_within = numeric(), rpt_across = numeric())
var.list <- list()
ranef.list <- list()

for (i in 1:length(env_vars)) {
  
  # Fit the linear mixed model
  mod <- lmer(as.formula(paste(env_vars[i], "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID)")), data = pnts.used)
  
  ## Extract random intercepts
  ranefs <- broom.mixed::tidy(mod, effects = "ran_vals", conf.int = TRUE)
  ring_ranefs <- ranefs[ranefs$group == "ID_season" | ranefs$group == "ID", ] %>% 
    rename(ID = level) %>%
    mutate(env = env_vars[i])
  
  ranef.list[[i]] <- ring_ranefs
  
  # Calculate repeatability
  var_components <- as.data.frame(VarCorr(mod))
  var_across <- var_components[var_components$grp == "ID", "vcov"]
  var_total <- sum(var_components$vcov)
  
  ## Calculate repeatability
  rpt_across <- var_across / var_total 
  
  # Outputs
  var_df <- var_components %>%
     select(grp = grp, name = var1, variance = vcov, std_dev = sdcor) %>%
     mutate(across(where(is.numeric), ~ round(.x, 3)),
            Environment = env_vars[i],
            rpt_across = rpt_across) %>%
     rename(Effect = grp, Variance = variance, "Std. dev" = std_dev) %>%
     select(-name) %>%
    relocate(Environment, .before = Effect)

  var.list[[i]] <- var_df
  
}

var.df_all <- do.call("rbind", var.list)
ranefs.df <- do.call("rbind", ranef.list)

save(var.df_all, file = "Data_outputs/bbal_ker_habitat_used_variances.RData")
save(ranefs.df, file = "Data_outputs/bbal_ker_habitat_used_ranef.RData")

```

##### Plot habitat use

```{r bbal_ker_plot_habitat_used }

load("Data_outputs/bbal_ker_habitat_used_ranef.RData")

ranefs.df %<>%
 mutate(env = case_when(
  env == "bathy" ~ "Bathymetry (m below sea level)",
  env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
  env == "SST"   ~ "SST (°C)" ))

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


split_df <- ranefs.df %>% group_by(env) %>% group_split()

make_plot <- function(df) {
  df <- df %>%
    mutate(ring_ordered = forcats::fct_reorder(ID, estimate))

  ggplot(df, aes(x = estimate, y = ring_ordered, col = env)) +
    geom_point() +
    geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
    scale_color_manual(values = env_colours, guide = "none") +
    labs(title = unique(df$env),
      x = "Random intercept estimate",
      y = "ID") +
    theme_bw() +
    theme(axis.text.y = element_blank())}

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_ker_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_ker_used_plot <- annotate_figure(bbal_ker_used_plot, top = text_grob("BBAL - Kerguelen", face = "bold", size = 14))
save(bbal_ker_used_plot, file = "Data_outputs/bbal_ker_used_plot.RData")

png(filename = "Figures/bbal_ker_habitat_used_ranef_plot.png", width = 10, height = 5, units = "in", res = 600)
ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()


```

```{r bbal_ker_table_variance_habitat_used}

load("Data_outputs/bbal_ker_habitat_used_variances.RData")

var.table <- var.df_all %>%
  mutate(Environment = case_when(
    Environment == "bathy" ~ "Bathymetry",
    Environment == "chlA" ~ "Chlorphyll-a",
    TRUE ~ "SST" )) %>%
  mutate(across(3:6, ~ round(.x, 2))) %>%
  rename("Within-season rpt" = rpt_within,
         "Across season rpt" = rpt_across)

kable(var.table, format = "html", 
  caption = 'Table X. Variances and repeatability associated with random intercepts from the habitat used model.') %>%
  row_spec(c(6,11), extra_css = "border-top: 1px dotted;") %>%
  collapse_rows(columns = c(1, 5, 6)) %>%
  kable_classic(font_size = 12)

```

### Habitat selection

#### Fit the models

```{r bbal_ker_fit_model, eval = FALSE}

load("Data_inputs/bbal_ker_habitat_data.RData")

# Between season #

bbal_ker_glmm.between_sst <- glmmTMB(
  used ~ sst_scaled + dist_shelf_bin + chl_scaled + dist_col_scaled + phase + 
         (1 | ID) + (0 + sst_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_sst, file = "Data_outputs/bbal_ker_glmm_between_sst.RData")
rm(bbal_ker_glmm.between_sst); gc()

s
bbal_ker_glmm.between_bathy <- glmmTMB(
  used ~ sst_scaled + dist_shelf_bin + chl_scaled + dist_col_scaled + phase + 
         (1 | ID) + (0 + dist_shelf_bin | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_ker_glmm.between_bathy, file = "Data_outputs/bbal_ker_glmm_between_bathy.RData")
rm(bbal_ker_glmm.between_bathy); gc()


bbal_ker_glmm.between_chl <- glmmTMB(
  used ~ sst_scaled + dist_shelf_bin + chl_scaled + dist_col_scaled + phase + 
         (1 | ID) + (0 + chl_scaled | ID) + (1 | tripID),
  data = pnts_all,
  family = binomial() )

save(bbal_ker_glmm.between_chl, file = "Data_outputs/bbal_ker_glmm_between_chl.RData")
rm(bbal_ker_glmm.between_chl); gc()

```

#### Diagnostics

```{r bbal_ker_diagnostics, eval = FALSE}

load("Data_outputs/bbal_ker_glmm_within_sst.RData")
load("Data_outputs/bbal_ker_glmm_within_bathy.RData")
load("Data_outputs/bbal_ker_glmm_within_chl.RData")

load("Data_outputs/bbal_ker_glmm_between_sst.RData")
load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
load("Data_outputs/bbal_ker_glmm_between_chl.RData")


model_info <- list(
  within_sst =    list(model = bbal_ker_glmm.within_sst, data = pnts_all),
  within_bathy =  list(model = bbal_ker_glmm.within_bathy, data = pnts_all),
  within_chl =    list(model = bbal_ker_glmm.within_chl, data = pnts_all),
  between_sst =   list(model = bbal_ker_glmm.between_sst, data = pnts_all),
  between_bathy = list(model = bbal_ker_glmm.between_bathy, data = pnts_all),
  between_chl =   list(model = bbal_ker_glmm.between_chl, data = pnts_all)
)

auc_results <- data.frame(
  model = character(),
  auc = numeric()  )

# Loop through models and compute AUC, output diagnostics
for (model_name in names(model_info)) {

  this_model <- model_info[[model_name]]$model
  this_data <- model_info[[model_name]]$data
  
  print(paste0("Next model: ", model_name, " ------------------------------------"))
  
  print("Covariance structure: ")
  print(VarCorr(this_model))
  
  print("Collinearity: ")
  print(performance::check_collinearity(this_model))
  
  print(paste0("DONE. ---------------------------"))
  
  preds <- predict(this_model, type = "response")
  auc_value <- as.numeric(auc(this_data$used, preds))
  
  auc_results <- rbind(auc_results, data.frame(model = model_name, auc = auc_value))
}

auc_results %<>% mutate(species = "BBAL", colony = "kerguelen") %>% relocate(c(species, colony), .before = model)


```

#### Fixed effects

```{r bbal_ker_fixed_effects_predictions }

load("Data_outputs/bbal_ker_glmm_between_sst.RData")
load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
load("Data_outputs/bbal_ker_glmm_between_chl.RData")

# Plot effects
pred_plot.sst <-   plot(ggeffects::ggpredict(bbal_ker_glmm.between_sst, terms = "sst_scaled [all]", type = "fixed"))
pred_plot.bathy <- plot(ggeffects::ggpredict(bbal_ker_glmm.between_bathy, terms = "bathy_scaled [all]", type = "fixed"))
pred_plot.chl <-   plot(ggeffects::ggpredict(bbal_ker_glmm.between_chl, terms = "chl_scaled [all]", type = "fixed"))

all_plots <- ggarrange(pred_plot.sst + 
                         labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)", y = "Used")) +
                         theme(plot.title = element_blank()), 
                       pred_plot.bathy + 
                         labs(x = "Scaled bathymetry (m below sea level)", y = "Used") + 
                         theme(plot.title = element_blank(),
                               axis.title.y = element_blank()),
                      pred_plot.chl + 
                        labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")", y = "Used")) +
                        theme(plot.title = element_blank(),
                              axis.title.y = element_blank()),
          ncol = 3, nrow = 1)

png("Figures/bbal_ker_habitat_fixed_effects.png", width = 9, height = 4, units = "in", res = 600)
all_plots
dev.off()

rm(pred_plot.sst); rm(pred_plot.bathy); rm(pred_plot.chl); rm(all_plots)
gc()

```


#### Conditional repeatability

Using the custom function provided by Schielzeth and Nakagawa 2022 Methods Ecol Evol.

Conditional repeatability essential asks: 'How does repeatability change depend on the covariate?'. This is important because individual responses depend on the environment they're measured in - this can be seen clearly in the random slopes plots, which often show a steep change in the response at certain levels of the predictor, and plateaus either side.

One sticking point with conditional repeatability is that it only works with linear responses (though I am likely to drop all my polynomial responses going forward).


```{r bbal_ker_cond_repeatability}

load("Data_inputs/bbal_ker_habitat_data.RData")

## Store envronmental variables
env_vars = c("sst", "chl", "bathy")

sst_scaled <- unlist(pnts_all$sst_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)

var_map <- list(
  sst = sst_scaled,
  chl = chl_scaled )

# Calculate conditional repeatability ------

cond_res.list <- vector(mode = "list", length = length(env_vars))

for (i in 1:2) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))

  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))
  
  # Calculate conditional repeatability parameters
  
  env_current <- var_map[[env_vars[i]]]
  
  ## Between ##
  my_sigma = matrix( c(
    as.numeric(VarCorr(mod_between)$cond$ID), 0,
    0, as.numeric(VarCorr(mod_between)$cond$ID.1)),
  nrow = 2) 
  
  res_between <- condR(
    betaX = fixef(mod_between)$cond[habitat_var],
    meanX =  mean(env_current),
    Vx = var(env_current),
    Vr = (pi^2)/3,
    Sigma = my_sigma,    
    Vo =  as.numeric(VarCorr(mod_between)$cond$tripID),  
    xrange = range(env_current) )
  
  res_between.summary <- data.frame(var_between = res_between$Vi,
                            R2 = res_between$R2s, 
                            marginal_R2 = res_between$Rmar,
                            med_rpt = median(res_between$condR$condR),
                            env = env_vars[i],
                            species = "BBAL",
                            colony = "Kerguelen", 
                            level = "Between")
  
  rownames(res_between.summary) <- NULL
  
  res.summary <- res_between.summary %>% 
    relocate(c(species, colony, level, env), .before = var_between)
  
  cond_res.list[[i]] <- res.summary
  
  condR_plot <- ggplot() +
    geom_line(data = res_between$condR, aes(x = x, y = condR/100), col = "#FFC107", size = 1) +
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt/100), linetype = "dashed", col = "#f3a20c") +
    labs(y = "Conditional repeatability") +
    expand_limits(y = 0) +
    theme_bw() 

  assign(paste0(env_vars[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_ker_glmm.between_", env_vars[i]))
  gc()
  
}

condR_results <- do.call("rbind", cond_res.list)


## Bathymetry only ----

load("Data_outputs/bbal_ker_glmm_between_bathy.RData")
mod_between = get("bbal_ker_glmm.between_bathy")
  
# Extract variance-covariance matrices of random effects
vc <- VarCorr(mod_between)$cond
V_ID_intercept <- as.numeric(vc$ID[1,1])
V_tripID <- as.numeric(vc$tripID[1,1])
V_slopes_matrix <- vc$ID.1 

# Extract variances (diagonal) for each bathymetry level slope
V_slope <- diag(V_slopes_matrix)
names(V_slope) <- colnames(V_slopes_matrix)
  
# Residual variance for binomial logistic link
V_residual <- (pi^2) / 3

# Total random variance per level
V_random <- V_ID_intercept + V_slope + V_tripID
V_total <- V_ID_intercept + V_slope + V_tripID + V_residual

# Conditional repeatability per level
R_cond <- V_random / V_total

R_cond.df <- data.frame(R_cond) 
R_cond.df <- cbind(rownames(R_cond.df), data.frame(R_cond.df, row.names = NULL))
R_cond.df %<>%
  mutate(species = "BBAL",
         colony = "Kerguelen",
         level = "Between", 
         var_between = V_slope,
         R2 = NA,
         marginal_R2 = NA) %>%
  rename(med_rpt = R_cond, env = `rownames(R_cond.df)`) %>%
  relocate(c(species, colony, level, env, var_between, R2, marginal_R2, med_rpt))

R_cond.df$env <- factor(R_cond.df$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

condR_results <- rbind(condR_results, R_cond.df)

save(condR_results, file = "Data_outputs/bbal_ker_condR_results.RData")

## Make the plot

bathy_labels <- c("Off shelf", "Near shelf break", "On shelf")

condR_plot <- ggplot() +
    geom_point(data = R_cond.df, aes(x= env, y = med_rpt), col = "#FFC107", size = 3) +
    labs(y = "Conditional repeatability") +
    scale_x_discrete(labels = bathy_labels) +
    expand_limits(y = 0) +
    theme_bw() 

assign("bathy_condR_plot", condR_plot)

# Plot individual slopes ---- 

rnd_effs.list <- list()

for (i in 1:2) {

  habitat_var = paste0(env_vars[i], "_scaled")
  
  # Get models
  load(paste0("Data_outputs/bbal_ker_glmm_between_", env_vars[i], ".RData"))

  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars[i]))
  
  # Get environmental range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rnd_eff.between
  rnd_effs.list[[i]] <- rnd_eff.all %>% mutate(env = env_vars[i])
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)


## Bathymetry only ----

fixef_vals <- fixef(bbal_ker_glmm.between_bathy)$cond
levels_bathy <- levels(pnts_all$dist_shelf_bin)

# Prepare a data frame to hold predictions
ranef_slopes <- ranef(bbal_ker_glmm.between_bathy)$cond$ID
pred_df <- expand.grid(ID = rownames(ranef_slopes),
                       dist_shelf_bin = levels_bathy)

# Calculate linear predictor per bird & category:
pred_df %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes[ID, "dist_shelf_binnear_break"],
                      ranef_slopes[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()

# Plot lines connecting points per bird
pred_plot <- ggplot(pred_df, aes(x = dist_shelf_bin, y = prob, group = ID)) +
  geom_line(col = bathy_col, alpha = 0.2) +
  geom_point(col = bathy_col, alpha = 0.2) +
  scale_x_discrete(labels = c("On shelf", "Near shelf break", "Off shelf")) +
  labs(x = "Bathymetry category",
       y = "Probability of use") +
  theme_bw()

assign("bathy_pred_plot", pred_plot)

save(rnd_effs.out, file = "Data_outputs/bbal_birdis_habitat_selection.RData")





# Make the plot ----
png("Figures/bbal_ker_habitat_conditional_repeatability2.png", width = 14, height = 9, units = "in", res = 600)
ggarrange(bathy_pred_plot + ylim(0, 1) + labs(x = "Proximity to shelf"), 
          sst_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(axis.title.y = element_blank()),
          bathy_condR_plot + labs(x = "Proximity to shelf"),
          sst_condR_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(axis.title.y = element_blank()), 
          chl_condR_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"))+ theme(axis.title.y = element_blank()),
          ncol = 3, nrow = 2)
dev.off()




```

#### Extract individual estimates

```{r bbal_ker_extract_ind_estimates}

load("Data_outputs/bbal_birdis_habitat_used_ranef.RData")
ranefs.df2 <- ranefs.df %>% 
  select(c(season_type, env, season, ring, estimate, intercept)) %>%
  rename(used_estimate = estimate,
         used_intercept = intercept) %>%
  mutate(env = case_when( env == "SST" ~ "sst",
                          env == "chlA" ~ "chl",
                          env == "bathy" ~ "bathy"))

rnd_effs.out2 <- rnd_effs.out %>%
  select(c(season_type, env, season, ring, intercept, slope)) %>%
  rename(selection_intercept = intercept,
         selection_slope = slope)

ind_estimates <- merge(ranefs.df2, rnd_effs.out2, by = c("ring", "season", "season_type", "env"), all = T)
write.csv(ind_estimates, file = "Data_outputs/bbal_birdis_habitat_individual_estimates.csv", row.names = FALSE)


```

# Final results

For the manuscript, we want to show:

-   Differences in average habitat use by species
-   Individual curves for each habitat type by species
-   Repeatability summaries for each species


## Figure SX - Plot used habitat

```{r plot_used}

load("Data_outputs/bbal_birdis_used_plot.RData")
load("Data_outputs/bbal_ker_used_plot.RData")
load("Data_outputs/waal_birdis_used_plot.RData")
load("Data_outputs/waal_cro_used_plot.RData")

png(file = "Figures/MS Figures/FigureSX.png", width = 22, height = 15, units = "in", res = 300)
ggarrange(bbal_birdis_used_plot, bbal_ker_used_plot,
          waal_cro_used_plot, waal_birdis_used_plot,
          ncol = 2, nrow = 2,
          widths = c(1, 0.8))
dev.off()

```


<!-- #### Habitat preference -->

<!-- After discussion with Sam & Al it was suggested that I could also look at mean habitat use across individuals; essentially I want to look at whether birds are specialists or generalists. Then, the habitat modelling could tell me whether individuals vary in plasticity i.e. are they flexible or specific.  -->

<!-- If I just want to see whether individuals have different preferences, I can do a simple LMM, e.g.:  -->

<!-- $$y = \beta_0 + \beta_1 \cdot \text{phase} + \beta_2 \cdot \text{age} + U_{\text{ring}} + U_{\text{season}} + U_{\text{ring:season}} + \varepsilon$$ -->

<!-- where: -->

<!-- -   $y$ represents each of the 3 environmental variables, -->

<!-- -   $\beta_0$ is the intercept, -->

<!-- -   $\beta_1$ - $\beta_{2}$ are fixed effect coefficients for $phase$ and $age$, -->

<!-- -   $U_{\text{ring}}$, $U_{\text{season}}$, and $U_{\text{ring:season}}$ are random intercepts for ring, season, and ring within season respectively, -->

<!--     $$U_{\text{ring}} \sim \mathcal{N}(0, \sigma^2_{\text{ring}})$$ -->

<!--     $$U_{\text{season}} \sim \mathcal{N}(0, \sigma^2_{\text{season}})$$ -->

<!--     $$U_{\text{ring_season}} \sim \mathcal{N}(0, \sigma^2_{\text{ring_season}})$$ -->

<!-- -   $ε$ is the residual error. -->

<!--     $$ε \sim \mathcal{N}(0, \sigma^2_{\text{res}})$$ -->

<!-- But if I want to know whether some individuals are more selective than others, I would need to use DHGLMMs. This would allow me to model the variance in selection as a function of individual identity. However, this requires some thought - I tried running a simple DHGLMM, and after 3 days it had only completed 30% of chains. Come back to this. -->

```{r bba_birdis_habitat_preferences, eval = FALSE}

# Integrate meta
load("Data_inputs/ind_meta.RData")
all_ind.meta %<>% filter(species == "BBA") %>% mutate(birthYr = as.numeric(birthYr)) %>% rename(ID = ring)

pnts_all <- merge(pnts_all, all_ind.meta, by = "ID", all.x = T) %>% 
  mutate(ID_season = paste(ID, season, sep = "_"), 
         age = season - birthYr)

pnts_all %<>% filter(!is.na(age))

# Build models

## SST ##
pnts_subset <- pnts_all[sample(nrow(pnts_all), 20000),]

library(brms)

sst_dhglmm <- brm(
  bf(SST ~ phase + age + (1 | ID) + (1 | season)) +
  lf(sigma ~ 0 + ID),
  iter = 2000, warmup = 1000, thin = 10,
  data = pnts_subset)

summary(sst_dhglmm)
save(sst_dhglmm, file = "Data_outputs/test_dhglmm.RData")


sst_pref <- lmer(SST ~ (1 | ID) + (1 | season) + (1 | ID_season) + phase + age,
                 data = pnts_all)

summary(sst_pref)

## Diagnostics
  # plot(m1, resid(., scaled=TRUE) ~ fitted(.), abline = 0)
  # qqnorm(resid(m1))
  # qqline(resid(m1))
  # hist(resid(m1), breaks = 50, main = "Histogram of residuals")

```

<!-- ## Wandering albatrosses -->

<!-- ### Bird Island -->

<!-- Again, for a subsample of 20 birds (for readability), I'm plotting the environmental variability used by each individual. -->

```{r waal_birdis_plot_env, eval = FALSE, fig.cap="**Figure 2.1.** *Individual density histograms for use of SST, bathymetry, chlorophyll, and sea surface height. Random sample of 20 individuals.*", fig.dim = c(8, 10), fig.topcaption = TRUE}

load("Data_inputs/waal_birdis_sst-bathy-chl-ssh_data.RData")

subset_ids <- sample(unique(pnts_all$ring), 30)

#### ~# SST # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_sst = mean(SST, na.rm = T)) %>%
  ungroup() %>%
  mutate(ring = factor(ring, levels = unique(ring[order(mean_sst)]))) 

plot.sst <- ggplot(pnts_subset, 
       aes(x = SST, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#0D0887FF", "#CC4678FF", "#F0F921FF"),
    name = "SST (°C)" ) +
  labs(x = "Sea surface temperature (°C)", y= "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())

#### ~# Bathymetry # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
   filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_bathy = mean(bathy, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_bathy)]))) 

plot.bathy <- ggplot(pnts_subset, 
       aes(x = bathy, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_viridis_c( name = "Sea floor depth (m below sea level)" ) +
  labs(x = "Sea floor depth (m below sea level)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


#### ~# Chlorophyll # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
   filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_chl = mean(chl, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_chl)]))) 

plot.chl <- ggplot(pnts_subset, 
       aes(x = chl, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1A6D2A", "#6CDC82", "#D3F5D9"),
    name = expression("Chlorophyll (" * mg ~ Chl ~ m^-3 * ")") ) +
  labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")"), y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


#### ~# Sea surface height # ----------

pnts_subset <- pnts_all %>%
  filter(ring %in% subset_ids) %>%
  filter(used == 1) %>%
  group_by(ring) %>%
  mutate(mean_ssh = mean(SSH, na.rm = T)) %>%
  ungroup() %>%
  mutate(ID = factor(ring, levels = unique(ring[order(mean_ssh)]))) 

plot.ssh <- ggplot(pnts_subset, 
       aes(x = SSH, y = ring)) +
  geom_density_ridges_gradient(aes(fill = after_stat(x)), scale = 2) +
  scale_fill_gradientn(
    colours = c("#1a2290", "#B2B6F1", "#D8DCE0"),
    name = "SSH (m)") +
  labs(x = "Sea surface height (m)", y = "Individual") +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.y = element_blank())


ggarrange(plot.sst, plot.bathy + theme(axis.title.y = element_blank()),
          plot.chl, plot.ssh + theme(axis.title.y = element_blank()), 
          ncol = 2, nrow = 2)

```

<!-- #### Modelling -->

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data_inputs/waal_birdis_sst-bathy-chl-ssh_data.RData")

## Scale the variables
pnts_all %<>%
  mutate(sst_scaled = scale(SST),
         bathy_scaled = scale(bathy),
         ssh_scaled = scale(SSH),
         chl_scaled = scale(chl),
         dist_col_scaled = scale(dist_col),
         season = ifelse(as.numeric(month(date_hourly)) < 9,
                         as.numeric(year(date_hourly)),
                         as.numeric(year(date_hourly)) + 1 )) %>%
  rename(ID = ring) %>%
  select(-index) %>%
  mutate(ID_season = paste0(ID, "_", season),
         tripID = paste0(ID, "_", tripID)) 

## Remove birds with only a couple of points
pnts_all %<>% 
  group_by(ID, year) %>%
  mutate(n_recs = n(),
         flag = ifelse(n_recs < 20, "flag", "fine")) %>%
  filter(flag == "fine") %>%
  select(-flag)

## Remove NAs and set distance to km
pnts_all %<>% filter(!is.na(SST) & !is.na(used) & !is.na(bathy) & !is.na(SSH) &!is.na(chl)) %>%
  mutate(dist_col.km = dist_col/1000)


```

```{r waal_birdis_models, eval = FALSE}

# Within-season ----

mod_within.waal_bi.sst <- glmmTMB(
  used ~ poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1 | tripID),
  data = pnts_all,
  family = binomial()
)

save(mod_within.waal_bi.sst, file = "Data_outputs/waal_birdis_habitat_model_within_sst.RData")

mod_within.waal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.waal_bi.bathy, file = "Data_outputs/waal_birdis_habitat_model_within_bathy.RData")

mod_within.waal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_within.waal_bi.chl, file = "Data_outputs/waal_birdis_habitat_model_within_chl.RData")


# Between-season ----

mod_between.waal_bi.sst <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + sst_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.waal_bi.sst, file = "Data_outputs/waal_birdis_habitat_model_between_sst.RData")

mod_between.waal_bi.bathy <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 +bathy_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.waal_bi.bathy, file = "Data_outputs/waal_birdis_habitat_model_between_bathy.RData")

mod_between.waal_bi.chl <- glmmTMB(
  used ~  poly(sst_scaled, 2) + poly(bathy_scaled, 2) + chl_scaled + dist_col_scaled + phase +
         (1 + chl_scaled | ID_season) + (1|tripID), 
  data = pnts_all, 
  family = binomial()
)

save(mod_between.waal_bi.chl, file = "Data_outputs/waal_birdis_habitat_model_between_chl.RData")

```

```{r waal_birdis_diagnostics, eval = FALSE}

load("Data_outputs/waal_birdis_habitat_model_between.RData")

### Diagnostics ----

# AUC around 0.5 = essential random; closer to 1 = better fit
preds <- predict(mod_between.waal_bi, type = "response")  
auc.mod <- as.numeric(auc(pnts_all$used, preds))

# Check for correlations
vif.mod <- car::vif(mod_between.waal_bi)

# Check for patterns and that most points are within 95% confidence intervals
plot_binnedplot(mod_between.waal_bi)

# Check ratio is close to 1
overdisp.mod <- overdisp_fun(mod_between.waal_bi)

# Check ratio: >1 may indicate zero inflation (more 0s than expected); < 1 zero deflation
# Small P = evidence to suggest zero-inflation/deflation
zeroinf.mod <- testZeroInflation(mod_between.waal_bi, plot = FALSE)

```

<!-- #### Diagnostics -->

<!-- -   AUC = `r signif(as.numeric(auc.mod), 2)`, which suggests the model separates 'used' vs. 'not used' locations well. -->

<!-- -   The VIF values are all \< 2, so there is no concerning multicollinearity. -->

<!-- -   The overdispersion function returns `r as.numeric(overdisp.mod[2])`, which is slightly high. This could indicate moderate overdispersion (i.e. variance is higher than expected in a binomial model), but probably not a major concern. -->

<!-- -   There is some evidence for zero-inflation - p = `r zeroinf.mod$p.value`. However, this might be to expected, as zero counts are an inherent part of the process I am trying to model, and so a zero-inflated approach is likely unncessary. -->

<!-- #### Predictions -->

```{r waal_birdis_model_outputs, eval = FALSE, fig.cap = "**Figure 2.3.** *Predicted probabilities of use for each fixed effect.*", fig.topcaption = TRUE}

# Plot effects
pred_plot.sst <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "sst_scaled [all]"))
pred_plot.bathy <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "bathy_scaled [all]"))
pred_plot.chl <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "chl_scaled [all]"))
pred_plot.dist <- plot(ggeffects::ggpredict(mod_between.waal_bi, terms = "dist_col_scaled [all]"))

all_plots <- ggarrange(pred_plot.sst + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) + theme(plot.title = element_blank()), 
          pred_plot.bathy + labs(x = "Scaled bathymetry (m below sea level)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          pred_plot.chl + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + theme(plot.title = element_blank()), 
          pred_plot.dist + labs(x = "Scaled distance from colony (km)") + theme(plot.title = element_blank(),
                                  axis.title.y = element_blank()),
          ncol = 2, nrow = 2)

all_plots

```

<!-- The results show that in general, birds select for median SST, deeper waters, higher chlorophyll concentrations, and locations closer to the colony. There is a high degree of individual variation, especially in the use of SST and bathymetry. -->

<!-- #### Individual effects -->

```{r waal_birdis_individual, eval = FALSE, fig.cap = "**Figure 2.4.** *Individual level responses to changes in the environmental variables. Individual slopes are plotted for the extent of each individual's range. Black line and grey interval shows population-level response and associated confidence interval.*", fig.topcaption = TRUE}

# Build plots

env_vars = c("sst_scaled", "bathy_scaled", "chl_scaled")

for (i in 1:length(env_vars)) {

  habitat_var = env_vars[i]
  rnd_eff <- get_rnd_effects(mod_between.bbal_bi, habitat_var)
  
  # Get range
  ind_ranges <- pnts_all %>%
    group_by(ID) %>%
    summarise(min_sst = min(!!sym(habitat_var), na.rm = TRUE),
              max_sst = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_sst, max_sst, length.out = 100))) %>%
    unnest(cols = c(habitat_val)) 
  
  predictions.ind <- merge(predictions.ind, rnd_eff, by = "ID", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Get global predictions
  predictions.global <- data.frame(ggpredict(mod_between.bbal_bi, terms = paste0(habitat_var, " [all]")))
  
  ## Plot
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, colour = abs(slope), group = ID)) +
      scale_colour_gradient(low = "#f4edff", high = "#340088") +  
      geom_line(data = predictions.global, aes(x = x, y = predicted), col = "black",
                 linetype = "dashed", linewidth = 1) +
      geom_ribbon(data = predictions.global, aes(x = x, ymin = conf.low, ymax = conf.high), fill = "grey90", alpha = 0.5, col = "grey10") +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0(env_vars[i], "_pred_plot"), pred_plot)
  
}

ggarrange(sst_scaled_pred_plot + labs(x = expression("Scaled sea surface temperature (" * ~ degree ~ "C)")) , 
          bathy_scaled_pred_plot + labs(x = "Scaled bathymetry (m below sea level)") + theme(axis.title.y = element_blank()), 
          chl_scaled_pred_plot + labs(x = expression("Scaled chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")),
          ncol = 2, nrow = 2)


```

```{r waal_birdis_calc_repeatability, eval = FALSE}

# Within season -----
summary(mod_within.waal_bi)

## manual test
vc <- VarCorr(mod_within.waal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.within <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.within <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.within <- V_ID_chl / (V_ID_chl + V_residual)



# Between season -----
summary(mod_between.waal_bi)

## manual test
vc <- VarCorr(mod_between.waal_bi)

# Variance components 
V_ID_intercept <- attr(vc$ID, "stddev")[1]^2  
V_ID_sst <- attr(vc$ID, "stddev")[2]^2        
V_ID_bathy <- attr(vc$ID, "stddev")[3]^2      
V_ID_chl <- attr(vc$ID, "stddev")[4]^2        

# Residual variance
V_residual <- (pi^2) / 3  # Residual variance for binary GLMM with logit link

R_sst.between <- V_ID_sst / (V_ID_sst + V_residual)
R_bathy.between <- V_ID_bathy / (V_ID_bathy + V_residual)
R_bathy.between <- V_ID_chl / (V_ID_chl + V_residual)


```
