---
title: "Repeatability in at-sea activity"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---

# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_MasterBayes(v.2.59)_, _kinship2(v.1.9.6.2)_, _quadprog(v.1.5-8)_, _genetics(v.1.3.8.1.3)_, _mvtnorm(v.1.3-3)_, _MASS(v.7.3-60)_, _gtools(v.3.9.5)_, _gdata(v.3.0.1)_, _combinat(v.0.0-8)_, _igraph(v.2.1.4)_, _MCMCglmm(v.2.36)_, _ape(v.5.8-1)_, _coda(v.0.19-4.1)_, _brms(v.2.23.0)_, _Rcpp(v.1.1.1)_, _ggridges(v.0.5.7)_, _glmmTMB(v.1.1.11)_, _stringr(v.1.6.0)_, _kableExtra(v.1.4.0)_, _tidyr(v.1.3.2)_, _gridExtra(v.2.3)_, _lme4(v.2.0-1)_, _Matrix(v.1.5-4.1)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_

## Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

source("R/0_setup.R")

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4",  "gridExtra", "tidyr", "kableExtra", "stringr", "glmmTMB", "ggridges", "brms", "MCMCglmm", "igraph", "MasterBayes")

# Load packages
invisible(lapply(packages, library, character.only = TRUE))


# Set colours
col_between <- "#91c5f2"
col_within <- "#0b3860"


```

```{r load_data}

ind_files <- list.files("Data/Intermediate/", pattern = "individual_gls_trips")
ind_list <- vector(mode = "list", length = 4)

for (i in 1:length(ind_files)) {
  
  my_data <- loadRData(paste0("Data/Intermediate/", ind_files[i]))
  ind_list[[i]] <- my_data
  
}

rep_data <- do.call("rbind", ind_list)

save(rep_data, file = "Data/Processed/behavioural_repeatability_data.RData") 

```

### Sample sizes

```{r sample_sizes}

trip_sum <- rep_data %>%
  group_by(colony, species, ring) %>%
  summarise(n_trips = n(),
            n_seasons = n_distinct(season),
            .groups = "drop") %>%
  group_by(colony, species) %>%
  summarise(N = n_distinct(ring),
            'Trips per bird' = mean(n_trips),
            'Seasons per bird' = mean(n_seasons)) %>%
  mutate(species = ifelse(species == "bbal", "BBAL", "WAAL"),
         colony = ifelse(colony == "ker", "Kerguelen",
                         ifelse(colony == "cro", "Crozet", "Bird Island")))

ind_per_seasons <- rep_data %>%
  group_by(colony, species, ring) %>%
  summarise(n_seasons = n_distinct(season)) %>%
  ungroup() %>%
  mutate(rpt_category = case_when(
    n_seasons == 1 ~ "1 seasons",
    n_seasons == 2 ~ "2 seasons",
    n_seasons == 3 ~ "3 seasons",
    n_seasons > 3 ~ "4+ seasons"
  )) %>%
  group_by(colony, species, rpt_category) %>%
  summarise(num_birds = n()) %>%
  arrange(species, colony, rpt_category) %>%
  pivot_wider(
    names_from = rpt_category,       
    values_from = num_birds, 
    values_fill = 0) %>%
  mutate(species = ifelse(species == "bbal", "BBAL", "WAAL"),
         colony = ifelse(colony == "ker", "Kerguelen",
                         ifelse(colony == "cro", "Crozet", "Bird Island"))) %>%
  relocate(species, .before = colony) 

summary.table <- merge(ind_per_seasons, trip_sum, by = c("species", "colony")) %>%
  rename(Species = species, Colony = colony)


kable(summary.table, format = "html", 
      caption = 'Table 1. Sample sizes for each species and colony, showing the number of individuals for which we have 2, 3, or 4+ seasons of data, collected using GLS and/or GPS devices.') %>%
  collapse_rows(columns = c(1)) %>%
  row_spec(2, extra_css = "border-bottom: 1px solid;") %>%
  kable_classic(font_size = 12)




```

# Estimate repeatablity

```{r estimate_repeatability, eval = FALSE}

load("Data/Processed/behavioural_repeatability_data.RData") 

# Split data
rep_bbal <- rep_data %>% filter(species == "bbal")
rep_waal <- rep_data %>% filter(species == "waal")

# Set weakly informative priors
priors <- c(
  prior(normal(0, 10), class = "Intercept"), # Intercept
  prior(normal(0, 2), class = "b"), # Fixed effects
  prior(student_t(3, 0, 5), class = "sd"), # Group-level SDs
  prior(student_t(3, 0, 10), class = "sigma"))

# Trying different response variables and examining repeatability
cov <- c("duration.hrs", "landings_per_day", "transit_time.pct", "rest_time.pct")

rpt_output.list <- vector(mode = "list", length = length(cov))

for (i in 1:length(cov)) {
  
  print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", i, " of ", length(cov), ": ", cov[i]))
  
  if(grepl("landings", cov[i])) { cov_value <- cov[i] } else {
  cov_value <- paste0("log(", cov[i], ")") }
  formula <- paste0(cov_value, " ~ phase + (1 | season) + (1 | ring ) + (1 | ring_season)") 
  
  # Fit models
  
  
  ## BBAL ## ----
  print("Processing black-browed...")

  
  ### Bird Island ### ----
  bbal_birdis.brm_rpt <- brm(
    formula = formula,
    data = rep_bbal %>% filter(colony == "birdis"),
    family = gaussian(),  
    prior = priors,
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Calc repeatability
  posterior <- as_draws_df(bbal_birdis.brm_rpt)
  
  bbal_birdis.post_df <- data.frame(cbind(calc_rpt_brms_slope(bbal_birdis.brm_rpt, rep_bbal %>% filter(colony == "birdis"))),
                              behaviour = cov[i],
                              species = "BBAL",
                              colony = "Bird Island")
    
  mod_name <- paste0("bbal_birdis.brm_rpt", cov[i])
  assign(mod_name, bbal_birdis.brm_rpt)
  
  save(list = mod_name, file = paste0("Outputs/Models/bbal_birdis_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  
  ### Kerguelen ### ----
  bbal_ker.brm_rpt <- brm(
      formula = formula,
      data = rep_bbal %>% filter(colony == "ker"),
      family = gaussian(),  
      prior = priors,
      cores = 4, chains = 4, iter = 4000,
      control = list(adapt_delta = 0.95))
    
  # Calc repeatability
  posterior <- as_draws_df(bbal_ker.brm_rpt)

  bbal_ker.post_df <- data.frame(cbind(calc_rpt_brms_slope(bbal_ker.brm_rpt, rep_bbal %>% filter(colony == "ker"))),
                             behaviour = cov[i],
                             species = "BBAL",
                             colony = "Kerguelen")
  
  mod_name <- paste0("bbal_ker.brm_rpt_", cov[i])
  assign(mod_name, bbal_ker.brm_rpt)
  
  save(list = mod_name, file = paste0("Outputs/Models/bbal_ker_brms_behaviour_repeatability_", cov[i], ".RData")) 
  
  
  ## WAAL ## ----
  print("Processing wanderers...")

  ### Bird Island ### ----
   waal_birdis.brm_rpt <- brm(
    formula = formula,
    data = rep_waal %>% filter(colony == "birdis"),
    family = gaussian(),  
    prior = priors,
    cores = 4, chains = 4, iter = 4000,
    control = list(adapt_delta = 0.95))
  
  # Calc repeatability
  posterior <- as_draws_df(waal_birdis.brm_rpt)
 
  waal_birdis.post_df <- data.frame(cbind(calc_rpt_brms_slope(waal_birdis.brm_rpt, rep_waal %>% filter(colony == "birdis"))),
                              behaviour = cov[i],
                              species = "WAAL",
                              colony = "Bird Island")
    
  mod_name <- paste0("waal_birdis.brm_rpt", cov[i])
  assign(mod_name, waal_birdis.brm_rpt)
  
  save(list = mod_name, file = paste0("Outputs/Models/waal_birdis_brms_behaviour_repeatability_", cov[i], ".RData"))
  
  ### Crozet ### ----
  waal_cro.brm_rpt <- brm(
      formula = formula,
      data = rep_waal %>% filter(colony == "cro"),
      family = gaussian(),  
      prior = priors,
      cores = 4, chains = 4, iter = 4000,
      control = list(adapt_delta = 0.95))
    
  # Calc repeatability
  posterior <- as_draws_df(waal_cro.brm_rpt)
  
  waal_cro.post_df <- data.frame(cbind(calc_rpt_brms_slope(waal_cro.brm_rpt, rep_waal %>% filter(colony == "cro"))),
                             behaviour = cov[i],
                             species = "WAAL",
                             colony = "Crozet")
  
  mod_name <- paste0("waal_cro.brm_rpt_", cov[i])
  assign(mod_name, waal_cro.brm_rpt)
  
  save(list = mod_name, file = paste0("Outputs/Models/waal_cro_brms_behaviour_repeatability_", cov[i], ".RData")) 
  
  
  # Output summary
  rpt_output.list[[i]] <- rbind(waal_cro.post_df, waal_birdis.post_df,
                                bbal_ker.post_df, bbal_birdis.post_df)
   
}

rpt_output.full <- do.call("rbind", rpt_output.list)
rpt_output.full %<>% arrange(species)

# Get sample sizes and add them in
rep_samples <- rep_data %>% 
  group_by(colony, species) %>% 
  summarise(n_trips = n_distinct(start_date), n_birds = n_distinct(ring)) %>% 
  mutate(species = ifelse(species == "bbal", "BBAL", "WAAL"),
         colony = case_when(colony == "birdis" ~ "Bird Island",
                            colony == "cro" ~ "Crozet",
                            colony == "ker" ~ "Kerguelen"))

rpt_output.full %<>% left_join(rep_samples, by = c("species", "colony"))

save(rpt_output.full, file = "Outputs/Posteriors/behavioural_repeatability_posteriors.RData")  
# Model checking 
# plot(mod_brm, variable = c("sd_animal__Intercept", "sigma"))
# pp_check(bbal_birdis.brm_rpt)
# 
# pairs(mod_brm, variable = c("sd_animal__Intercept", "sigma"))

```


# Estimate heritability

```{r effort_heritability}

# Set priors
effort_priors <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(normal(0, 1), class = "b"),
  prior(student_t(3, 0, 1), class = "sd", group = "ring_pe"),
  prior(exponential(1), class = "sd", group = "animal"),
  prior(student_t(3, 0, 1), class = "sigma"))

# Load the data
load("Data/Processed/behavioural_repeatability_data.RData") 

## Remove zero values
rep_data %<>% filter(rowSums(select(., landings_per_day, transit_time.pct) == 0) == 0)

## New variable names; scale covariates
rep_data %<>% mutate(animal = ring, ring_pe = ring) 

# Loop through species and colonies
spec_col <- c("bbal_birdis", "waal_birdis", "waal_cro")
posteriors = list()

cov <- c("duration.hrs", "landings_per_day", "transit_time.pct", "rest_time.pct")

for (i in 1:length(spec_col)) {

  my_species <- strsplit(spec_col[i], "_")[[1]][1]
  my_colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  mod_data <- rep_data %>% filter(species == my_species & colony == my_colony)

  # Load the pedigree info
  pedigree.df <- read.csv(paste0("Data/Raw/", my_species, "_", my_colony, "_pedigree.csv"))[,1:3] #Data/Raw/
  pedigree.df[,c(1:3)] <- lapply(pedigree.df[,c(1:3)], as.character)
  
  ## Create the genetic matrix
  Amat <- make_Amat(mod_data, pedigree.df)
  
  # Fit the models
  posteriors_list <- list()
  
  for (c in 1:length(cov)) {
    
    print(paste0(format(Sys.time(), "%Y-%m-%d %H:%M"), " Processing covariate ", c, " of ", length(cov), ": ", cov[c]))
    
    if(grepl("landings", cov[c])) { cov_value <- paste0(cov[c], "_scaled") } else {
         cov_value <- paste0(cov[c], "_log_scaled") }
    
    formula <- paste0(cov_value, " ~ 1 + phase + (1 | season) + (1 | ring_pe) + (1 | ring_season) + (1 | gr(animal, cov = Amat))")

    # Scale and log the covariate
    
    if(grepl("landings", cov[c])) { 
      
      scaled_col <- paste0(cov[c], "_scaled")
      mod_data[[scaled_col]] <- scale(mod_data[[cov[c]]])
      
    } else {
         
      log_col <- paste0(cov[c], "_log")
      scaled_col <- paste0(cov[c], "_log_scaled")

      mod_data[[log_col]] <- log(mod_data[[cov[c]]])
      mod_data[[scaled_col]] <- scale(mod_data[[log_col]])[,1]
    }
    
    # Fit the model
    mod_brm <- brm(
      formula = formula,
      prior = effort_priors,
      data = mod_data,
      family = gaussian(),
      data2 = list(Amat = Amat),
      chains = 4, cores = 4, iter = 8000, warmup = 4000,
      control = list(max_treedepth = 15, adapt_delta = 0.99))
    
    ## Get posteriors 
    posterior <- as_draws_df(mod_brm)
    var_season  <- posterior$sd_season__Intercept^2
    var_ring_pe <- posterior$sd_ring_pe__Intercept^2
    var_ring_season <- posterior$sd_ring_season__Intercept^2
    var_animal  <- posterior$sd_animal__Intercept^2
    var_resid   <- posterior$sigma^2
    
    h2_post <- var_animal / (var_animal + var_season + var_ring_pe + var_ring_season + var_resid)
    
    posterior.df <- data.frame(cbind(var_animal, var_season, var_resid, h2_post),
                               behaviour = cov[c])
    
    posteriors_list[[c]] <- posterior.df
   
    # Output model
    save(mod_brm, file = paste0("Outputs/Models/",
                                my_species, "_", my_colony,
                                "_brms_behaviour_heritability_", cov[c], ".RData"))
    }
     
  posteriors[[spec_col[i]]] <- bind_rows(posteriors_list)
  
  }


posteriors_df <- bind_rows(posteriors, .id = "spec_col")
save(posteriors_df, file = "Outputs/Posteriors/behaviour_heritability_posteriors.RData")

```

### Process heritability estimates

```{r plot_effort_heritability}

load("Outputs/Posteriors/behaviour_heritability_posteriors.RData") 

posteriors_df %<>%
  mutate(behaviour = recode(behaviour,
      "duration.hrs" = "Trip duration",
      "landings_per_day" = "Landings per day",
      "transit_time.pct" = "Flight time",
      "rest_time.pct" = "Rest time"),
      behaviour = factor(behaviour),              
      behaviour = forcats::fct_rev(behaviour))

posteriors_df %<>%
  group_by(spec_col, behaviour) %>%
  mutate(mean_h2 = mean(h2_post, na.rm = TRUE)) %>%
  arrange(behaviour, mean_h2) %>%
  ungroup()

## Add facet code for plotting
posteriors_df %<>%
   mutate(facet_code = case_when(
    spec_col == "bbal_birdis" ~ "A",
    spec_col == "waal_birdis" ~ "C",
    spec_col == "waal_cro" ~ "D"))


# Make the plot
design = "
A#
CD
"

facet_labels <- c(
  A = "BBAL Bird Island",
  C = "WAAL Bird Island",
  D = "WAAL Crozet"
)

heritability_plot <- ggplot(posteriors_df, aes(x = h2_post, y = behaviour)) +
  # stat_density_ridges(fill = "#1b9e77",  alpha = 0.6, scale = 1, quantile_lines = TRUE, quantiles = 2, vline_colour = "#08519c", vline_size = 1, vline_type = "dashed") +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.6, scale = 1) +
  theme_bw() +
   ggh4x::facet_manual(~facet_code, design = design, scales = "free_x",
                       labeller = as_labeller(facet_labels)) +
  labs(x = "Posterior distribution for heritability (h²)", y = "") 


```
