---
title: "Habitat selection analyses"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
editor_options: 
  markdown: 
    wrap: 72
---

# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_nadiv(v.2.18.0)_, _pheatmap(v.1.0.13)_, _MasterBayes(v.2.59)_, _kinship2(v.1.9.6.2)_, _quadprog(v.1.5-8)_, _genetics(v.1.3.8.1.3)_, _mvtnorm(v.1.3-3)_, _gtools(v.3.9.5)_, _gdata(v.3.0.1)_, _combinat(v.0.0-8)_, _igraph(v.2.1.4)_, _MCMCglmm(v.2.36)_, _ape(v.5.8-1)_, _coda(v.0.19-4.1)_, _MASS(v.7.3-60)_, _brms(v.2.23.0)_, _Rcpp(v.1.1.1)_, _glmmTMB(v.1.1.11)_, _ggeffects(v.2.3.2)_, _pROC(v.1.19.0.1)_, _corrplot(v.0.95)_, _ggridges(v.0.5.7)_, _kableExtra(v.1.4.0)_, _ggpubr(v.0.6.3)_, _data.table(v.1.18.2.1)_, _DHARMa(v.0.4.7)_, _lubridate(v.1.9.4)_, _tidyr(v.1.3.2)_, _lme4(v.2.0-1)_, _Matrix(v.1.5-4.1)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_


# Set-up

```{r setup_packages_functions, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

source("R/0_setup.R")

## MasterBayes installation
# install.packages(c('genetics', 'gtools', 'kinship2'))
# install.packages("https://cran.r-project.org/src/contrib/Archive/MasterBayes/MasterBayes_2.58.tar.gz", repos = NULL, type = "source", dependencies = T)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "lme4", "tidyr", "lubridate", "DHARMa", "data.table", "ggpubr",  "kableExtra", "ggridges", "corrplot", "pROC", "ggeffects", "glmmTMB", "brms", "MASS", "MCMCglmm", "igraph", "MasterBayes", "pheatmap", "nadiv") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))
source("R/Functions/condR.R")


```

```{r setup_parameters}

# Set colours
bathy_col = "#0072B2" 
chl_col = "#009E73" 
sst_col = "#F0E442" 

col_between <- "#91c5f2"
col_within <- "#0b3860"

# Heritability priors
priors_animal <- c(
  prior(normal(0, 1), class = "Intercept"),
  prior(exponential(1), class = "sd", group = "ring_pe"), 
  prior(exponential(1), class = "sd", group = "animal"),
  prior(exponential(1), class = "sigma") )

# Sim number for bootstrapping CIs 
n_sim = 500

# Environmental variable names
env_vars.used <- c("bathy", "chlA", "SST")
env_vars.pref = c("bathy", "chl", "sst")

env_colours <- c(
  "Bathymetry (m below sea level)" = bathy_col,
  "Chlorophyll (mg Chl m^-3)" = chl_col,
  "SST (°C)" = sst_col)


```

# Black-browed albatrosses

## Bird Island

```{r bbal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Data/Intermediate/bbal_birdis_sst-bathy-chla_data_population_level.RData")

pnts_all <- process_habitat_data(pnts_all, 6, 50) # ratio used:avail; min recs per individual

save(pnts_all, file = "Data/Processed/bbal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r bbal_birdis_model_average_habitat_used, eval = FALSE}

load("Data/Processed/bbal_birdis_habitat_data.RData")

bbal_birdis_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) "

bbal_birdis_vars <- model_habitat_use(bbal_birdis_pnts_all, env_vars.used, "BBAL", "Bird Island", n_sim, my_mod)
bbal_birdis_var_used <- bbal_birdis_vars$variance
bbal_birdis_ranefs_used <- bbal_birdis_vars$ranef

save(bbal_birdis_var_used, file = "Outputs/Posteriors/bbal_birdis_habitat_used_variances.RData")
save(bbal_birdis_ranefs_used, file = "Outputs/Posteriors/bbal_birdis_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Outputs/Posteriors/bbal_birdis_habitat_used_ranef.RData")

bbal_birdis_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" )) 

split_df <- bbal_birdis_ranefs_used  %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
bbal_birdis_used_plot <- annotate_figure(bbal_birdis_used_plot, top = text_grob("BBAL Bird Island", face = "bold", size = 14))
save(bbal_birdis_used_plot, file = "Outputs/Plots/bbal_birdis_used_plot.RData")

png(filename = "Outputs/Figures/bbal_birdis_habitat_used_ranef_plot.png", width = 14, height = 5, units = "in", res = 600)
 ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
dev.off()

```

### Habitat selection

#### Fit the models

```{r bbal_birdis_fit_models, eval = FALSE}

load("Data/Processed/bbal_birdis_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Fit the models ====

# Within season #

bbal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_bathy, file = "Outputs/Models/bbal_birdis_glmm_within_bathy.RData")
rm(bbal_birdis_glmm.within_bathy); gc()


bbal_birdis_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_chl, file = "Outputs/Models/bbal_birdis_glmm_within_chl.RData")
rm(bbal_birdis_glmm.within_chl); gc()


bbal_birdis_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),                  
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.within_sst, file = "Outputs/Models/bbal_birdis_glmm_within_sst.RData")
rm(bbal_birdis_glmm.within_sst); gc()


# Between season #

bbal_birdis_glmm.between_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID) + (0 + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_bathy, file = "Outputs/Models/bbal_birdis_glmm_between_bathy.RData")
rm(bbal_birdis_glmm.between_bathy); gc()


bbal_birdis_glmm.between_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(bbal_birdis_glmm.between_chl, file = "Outputs/Models/bbal_birdis_glmm_between_chl.RData")
rm(bbal_birdis_glmm.between_chl); gc()


bbal_birdis_glmm.between_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled + sst_scaled + dist_col_scaled + phase +
        (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_birdis_glmm.between_sst, file = "Outputs/Models/bbal_birdis_glmm_between_sst.RData")
rm(bbal_birdis_glmm.between_sst); gc()


```

#### Conditional repeatability

We will calculate metrics of repeatability from our preference models.
We measure both raw variation and derive conditional repeatability
across the environmental covariates.

Raw variation includes: - *slope_var*: Among-individual variance in
plasticity (measured as individual preference for different
environmental levels) - *marginal_R2*: Average repeatability across the
entire environmental range - *R2*: Proportion of total variance
attributed to the random slopes

Conditional repeatability metrics include: - *med_rpt*: Median
repeatability. Since repeatability varies across the 90% range of
chlorophyll, condR calculates it at many points. med_rpt is the average
(median) repeatability across that whole environmental range

```{r bbal_birdis_cond_repeatability}

# [ 1 ] Calculate conditional repeatability ----

load("Data/Processed/bbal_birdis_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = length(env_vars.pref))

for (i in 1:length(env_vars.pref)) {
  
  env_current <- var_map[[env_vars.pref[i]]]
  
  # Load models
  load(paste0("Outputs/Models/bbal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Outputs/Models/bbal_birdis_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_between.summary <- extract_pref_metrics(mod_between, env_vars.pref[i], "between", env_current, "BBAL", "Bird Island")
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "BBAL", "Bird Island")
  
  res.summary <- rbind(res_between.summary, res_within.summary)
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.between <- bootstrap_condR(mod_between, env_vars.pref[i], "between", n_sim, env_current, "BBAL")
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current, "BBAL")

  # Make the conditional repeatability plot
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.between, aes(x = x, ymin = lower, ymax = upper), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty.within, aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_between, size = 1) +
    geom_line(data = condR_uncertainty.within, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("bbal_birdis_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_birdis_glmm.within_", env_vars.pref[i]))
  rm(list = paste0("bbal_birdis_glmm.between_", env_vars.pref[i]))
  gc()
  
}

bbal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(bbal_birdis_condR_results, file = "Outputs/Posteriors/bbal_birdis_condR_results.RData")

```

```{r bbal_birdis_plot_slopes}

# [ 2 ] Plot individual slopes ----

rnd_effs.list <- list()

for (i in 1:length(env_vars.pref)) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  
  # Get models
  load(paste0("Outputs/Models/bbal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Outputs/Models/bbal_birdis_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("bbal_birdis_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_birdis_glmm.between_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars.pref[i])
 
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  

  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("bbal_birdis_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}


```

```{r bbal_birdis_preference_figures}

# [ 3 ] Final plots ----

## Conditional repeatability ----

bbal_birdis_habitat_condR_plot <- ggarrange(
  bbal_birdis_bathy_condR_plot +
    ylim(0, 1) +
    labs(x = "Bathymetry (m below sea level)"),
  bbal_birdis_chl_condR_plot +
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    ylim(0, 1) +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  bbal_birdis_sst_condR_plot +
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  ncol = 3, 
  nrow = 1,
  align = "hv")

bbal_birdis_habitat_condR_plot <- annotate_figure(bbal_birdis_habitat_condR_plot,
                top = text_grob("BBAL Bird Island", size = 14))

save(bbal_birdis_habitat_condR_plot, file = "Outputs/Plots/bbal_birdis_habitat_condR_plot.RData")


bbal_birdis_habitat_slopes_plot <- ggarrange(
  bbal_birdis_bathy_pred_plot +
    labs(x = "Bathymetry (m below sea level)"), 
  bbal_birdis_chl_pred_plot +
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()),
  bbal_birdis_sst_pred_plot +
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()), 
  ncol = 3, 
  nrow = 1,
  align = "hv")

bbal_birdis_habitat_slopes_plot <- annotate_figure(bbal_birdis_habitat_slopes_plot,
                top = text_grob("BBAL Bird Island", size = 14))

save(bbal_birdis_habitat_slopes_plot, file = "Outputs/Plots/bbal_birdis_habitat_slopes_plot.RData")

```

### Heritability

#### Habitat used

```{r bbal_birdis_heritability_used }

# Load & prep data ====
load("Outputs/Posteriors/bbal_birdis_habitat_used_ranef.RData")
bbal_birdis_pedigree.df <- read.csv("Data/Raw/bbal_birdis_pedigree.csv") %>% select(-Year)

data_prep <- prep_used_heritability_data(bbal_birdis_ranefs_used, bbal_birdis_pedigree.df)
bbal_birdis_used_h2.df <- data_prep$data
bbal_birdis_used_h2.Amat <- data_prep$Amat


# Fit models ====
models_list <- lapply(env_vars.used, function(e) {
  fit_habitat_heritability_model("estimate", e, bbal_birdis_used_h2.df, bbal_birdis_used_h2.Amat, priors_animal, "Outputs/Models/bbal_birdis_")
})
names(models_list) <- env_vars.used


# Extract h2 ==== 
bbal_birdis_h2_used.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chlA"   = "Chlorophyll",
      "SST"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) %>%
  rename(h2 = var1)

rownames(bbal_birdis_h2_used.posteriors) <- NULL

save(bbal_birdis_h2_used.posteriors, file = "Outputs/Posteriors/bbal_birdis_habitat_used_h2_posteriors.RData")

```

#### Habitat preference

```{r bbal_birdis_heritability_pref }

# Extract slopes ====

load("Data/Processed/bbal_birdis_habitat_data.RData")
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

bbal_birdis_habitat_slopes <- extract_habitat_slopes(env_vars.pref, "bbal_birdis", pnts_meta)
save(bbal_birdis_habitat_slopes, file = "Outputs/Posteriors/bbal_birdis_habitat_slopes.RData")


# Estimate heritability ===

## Prep the data
bbal_birdis_pedigree.df <- read.csv("Data/Raw/bbal_birdis_pedigree.csv") %>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
bbal_birdis_pedigree.df$sire[bbal_birdis_pedigree.df$id == "1271009"] <- "1318880"
bbal_birdis_pedigree.df$dam[bbal_birdis_pedigree.df$id == "1271009"] <- "1318939"

data_prep <- prep_slope_heritability_data(bbal_birdis_habitat_slopes, bbal_birdis_pedigree.df)
bbal_birdis_pref_h2.df <- data_prep$data
bbal_birdis_pref_h2.Amat <- data_prep$Amat


# Fit models ====

models_list <- lapply(env_vars.pref, function(e) {
  fit_habitat_heritability_model("slope", e, bbal_birdis_pref_h2.df, bbal_birdis_pref_h2.Amat, priors_animal, "Outputs/Models/bbal_birdis_")
})
names(models_list) <- env_vars.pref


# Extract h2 ==== 
bbal_birdis_h2_slope.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chl"   = "Chlorophyll",
      "sst"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) %>%
  rename(h2 = var1)

rownames(bbal_birdis_h2_slope.posteriors) <- NULL

save(bbal_birdis_h2_slope.posteriors, file = "Outputs/Posteriors/bbal_birdis_habitat_pref_h2_posteriors.RData")

```

## Kerguelen

For Kerguelen especially, birds' preference for shelf zones makes the
bathymetry model fall apart. To tackle this, I have calculated distance
from the shelf, and split preferences into on shelf, shelf break, and
off shelf. This seems to resolve my problems. The shelf break is
delimited at 1000m following Hindell et al. 2011.

```{r bbal_ker_process_data, eval = FALSE}

load("Data/Intermediate/bbal_ker_sst-bathy-chla_data_population_level.RData")
pnts_all <- process_habitat_data(pnts_all, 6, 50)

save(pnts_all, file = "Data/Processed/bbal_ker_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r bbal_ker_model_average_habitat_used, eval = FALSE}

load("Data/Processed/bbal_ker_habitat_data.RData")

bbal_ker_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) "

bbal_ker_vars <- model_habitat_use(bbal_ker_pnts_all, env_vars.used, "BBAL", "Kerguelen", n_sim, my_mod)
bbal_ker_var_used <- bbal_ker_vars$variance
bbal_ker_ranefs_used <- bbal_ker_vars$ranef

save(bbal_ker_var_used, file = "Outputs/Posteriors/bbal_ker_habitat_used_variances.RData")
save(bbal_ker_ranefs_used, file = "Outputs/Posteriors/bbal_ker_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r bbal_ker_plot_habitat_used }

load("Outputs/POsteriors/bbal_ker_habitat_used_ranef.RData")

bbal_ker_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" ))

split_df <- bbal_ker_ranefs_used %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

bbal_ker_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot,
          ncol = 3)
bbal_ker_used_plot <- annotate_figure(bbal_ker_used_plot, top = text_grob("BBAL Kerguelen", face = "bold", size = 14))
save(bbal_ker_used_plot, file = "Outputs/Plots/bbal_ker_used_plot.RData")

```

### Habitat selection

#### Fit the models

```{r bbal_ker_fit_model, eval = FALSE}

load("Data/Processed/bbal_ker_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Fit the models ====

# Within season #

bbal_ker_glmm.within_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID_season) + (0 + dist_shelf_bin | ID_season),              
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_bathy, file = "Outputs/Models/bbal_ker_glmm_within_bathy.RData")
rm(bbal_ker_glmm.within_bathy); gc()


bbal_ker_glmm.within_chl <-  glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID_season) + (0 + chl_scaled | ID_season),           
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_chl, file = "Outputs/Models/bbal_ker_glmm_within_chl.RData")
rm(bbal_ker_glmm.within_chl); gc()


bbal_ker_glmm.within_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),       
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.within_sst, file = "Outputs/Models/bbal_ker_glmm_within_sst.RData")
rm(bbal_ker_glmm.within_sst); gc()


# Between season #

bbal_ker_glmm.between_bathy <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + dist_shelf_bin | tripID) + (1 | ID) + (0 + dist_shelf_bin | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_bathy, file = "Outputs/Models/bbal_ker_glmm_between_bathy.RData")
rm(bbal_ker_glmm.between_bathy); gc()


bbal_ker_glmm.between_chl <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + chl_scaled | tripID) + (1 | ID) + (0 + chl_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_chl, file = "Outputs/Models/bbal_ker_glmm_between_chl.RData")
rm(bbal_ker_glmm.between_chl); gc()


bbal_ker_glmm.between_sst <- glmmTMB(
  used ~ dist_shelf_bin + chl_scaled + sst_scaled +  dist_col_scaled + phase +
         (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID) + (0 + sst_scaled | ID),
  data = pnts_all,
  family = binomial(),
  control = glmmTMBControl(optCtrl = list(iter.max=1e5, eval.max=1e5)) )

save(bbal_ker_glmm.between_sst, file = "Outputs/Models/bbal_ker_glmm_between_sst.RData")
rm(bbal_ker_glmm.between_sst); gc()


```

#### Conditional repeatability

```{r bbal_ker_cond_repeatability}

load("Data/Processed/bbal_ker_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = 2)

# [ 1A ] Calculate conditional repeatability - Chl & SST ------

for (i in 2:3) { 

  env_current <- var_map[[env_vars.pref[i]]]
  print(env_vars.pref[i])
  
  # Load models
  load(paste0("Outputs/Models/bbal_ker_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Outputs/Models/bbal_ker_glmm_between_", env_vars.pref[i], ".RData"))
  
  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_between.summary <- extract_pref_metrics(mod_between, env_vars.pref[i], "between", env_current, "BBAL", "Kerguelen")
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "BBAL", "Kerguelen")
  
  res.summary <- rbind(res_between.summary, res_within.summary)
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.between <- bootstrap_condR(mod_between, env_vars.pref[i], "between", n_sim, env_current, "BBAL")
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current, "BBAL")
  
  condR_uncertainty <- rbind(condR_uncertainty.between, condR_uncertainty.within)
  
  # Make the conditional repeatability plot
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "between"), aes(x = x, ymin = lower, ymax = upper), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty %>% filter(level == "within"), aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_between, size = 1) +
    geom_line(data = condR_uncertainty.within, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("bbal_ker_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("bbal_ker_glmm.within_", env_vars.pref[i]))
  rm(list = paste0("bbal_ker_glmm.between_", env_vars.pref[i]))
  gc()
}

condR_results <- do.call("rbind", cond_res.list)


# [ 1B ] Calculate conditional repeatability - Bathymetry ------

# Between ====
  
## Calculate repeatability -----
load("Outputs/Models/bbal_ker_glmm_between_bathy.RData")
mod_between = get("bbal_ker_glmm.between_bathy")

load("Outputs/Models/bbal_ker_glmm_within_bathy.RData")
mod_within = get("bbal_ker_glmm.within_bathy")

R_cond.df_between <- process_categorical_slopes(mod_between, "between", n_sim)
R_cond.df_between$env <- factor(R_cond.df_between$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

R_cond.df_within <- process_categorical_slopes(mod_within, "within", n_sim)
R_cond.df_within$env <- factor(R_cond.df_within$env, levels = c("dist_shelf_binon_shelf", "dist_shelf_binnear_break", "dist_shelf_binoff_shelf"))

bbal_ker_condR_results <- rbind(condR_results, R_cond.df_between, R_cond.df_within)
save(bbal_ker_condR_results, file = "Outputs/Posteriors/bbal_ker_condR_results.RData")


# Make the plot ====
bathy_labels <- c("On shelf", "Near shelf break", "Off shelf")

condR_plot <- ggplot() +
    geom_point(data = R_cond.df_between, aes(x= env, y = med_rpt), col = col_between, size = 3) +
    geom_errorbar(data = R_cond.df_between, aes(x = env, ymin = LCL_rpt, ymax = UCL_rpt), width = 0.1, colour = col_between) +
    geom_point(data = R_cond.df_within, aes(x= env, y = med_rpt), col = col_within, size = 3) +
    geom_errorbar(data = R_cond.df_within, aes(x = env, ymin = LCL_rpt, ymax = UCL_rpt), width = 0.1, colour = col_within) +
    labs(y = "Conditional repeatability") +
    scale_x_discrete(labels = bathy_labels) +
    expand_limits(y = c(0, 0.2)) +
    theme_bw() 

assign("bbal_ker_bathy_condR_plot", condR_plot)

```

```{r bbal_ker_plot_slopes}

# [ 2 ] Plot individual slopes ---- 

env_vars.pref = c("chl", "sst", "bathy")

rnd_effs.list <- list()

for (i in 1:2) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  
  # Get models
  load(paste0("Outputs/Models/bbal_ker_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Outputs/Models/bbal_ker_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("bbal_ker_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("bbal_ker_glmm.between_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between) %>% mutate(env = env_vars.pref[i])
  
  rnd_effs.list[[i]] <- rnd_eff.all
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring")
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("bbal_ker_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

rnd_effs.out <- do.call("rbind", rnd_effs.list)


## Bathymetry only ====

levels_bathy <- levels(mod_df$dist_shelf_bin)

fixef_vals.between <- fixef(bbal_ker_glmm.between_bathy)$cond
fixef_vals.within <- fixef(bbal_ker_glmm.within_bathy)$cond

# Prepare a data frame to hold predictions
ranef_slopes.between <- ranef(bbal_ker_glmm.between_bathy)$cond$ID
ranef_slopes.within  <- ranef(bbal_ker_glmm.within_bathy)$cond$tripID

pred_df.between <- expand.grid(ID = rownames(ranef_slopes.between),
                       dist_shelf_bin = levels_bathy)
pred_df.within <- expand.grid(ID = rownames(ranef_slopes.within),
                       dist_shelf_bin = levels_bathy)

# Calculate linear predictor per bird & category: - STOPPED HERE
pred_df.between %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.between[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.between[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.between[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.between[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.between["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.between["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.between["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()


pred_df.within %<>%
  rowwise() %>%
  mutate(rand_int = ranef_slopes.within[ID, "(Intercept)"],
         rand_slope = if_else(dist_shelf_bin == "on_shelf", ranef_slopes.within[ID, "dist_shelf_binon_shelf"],
                      if_else(dist_shelf_bin == "near_break", ranef_slopes.within[ID, "dist_shelf_binnear_break"],
                      ranef_slopes.within[ID, "dist_shelf_binoff_shelf"])),
    fixef_int = fixef_vals.within["(Intercept)"],
    fixef_slope = if_else(dist_shelf_bin == "off_shelf", fixef_vals.within["dist_shelf_binoff_shelf"],
                  if_else(dist_shelf_bin == "near_break", fixef_vals.within["dist_shelf_binnear_break"],
                  0)),
    eta = fixef_int + rand_int + fixef_slope + rand_slope,
    prob = plogis(eta)) %>%
  ungroup()

## Bind together
pred_df.bathy <- rbind(pred_df.between %>% mutate(season_type = "Between seasons"),
                       pred_df.within %>% mutate(season_type = "Within season"))

# Plot lines connecting points per bird
pred_plot <- ggplot() +
  geom_line(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  geom_point(data = pred_df.bathy, aes(x = dist_shelf_bin, y = prob, group = ID), col = bathy_col, alpha = 0.5) +
  facet_grid(~season_type) +
  scale_x_discrete(labels = c("On shelf", "Near shelf", "Off shelf")) +
  labs(x = "Bathymetry category",
       y = "Probability of use") +
  theme_bw()

assign("bbal_ker_bathy_pred_plot", pred_plot)

```

```{r bbal_ker_preference_figures}

# [ 3 ] Final plots ----

## Conditional repeatability ----

bbal_ker_habitat_condR_plot <- ggarrange(
  bbal_ker_bathy_condR_plot +
    labs(x = "Bathymetry category") + 
    ylim(0, 1),
  bbal_ker_chl_condR_plot +
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  bbal_ker_sst_condR_plot +
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()), 
  ncol = 3, 
  nrow = 1,
  align = "hv")

bbal_ker_habitat_condR_plot <- annotate_figure(bbal_ker_habitat_condR_plot,
                top = text_grob("BBAL Kerguelen", size = 14))

save(bbal_ker_habitat_condR_plot, file = "Outputs/Plots/bbal_ker_habitat_condR_plot.RData")


## Slopes ----

bbal_ker_habitat_slopes_plot <- ggarrange(
  bbal_ker_bathy_pred_plot +
    ylim(0, 1), 
  bbal_ker_chl_pred_plot +
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()),
  bbal_ker_sst_pred_plot +
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()), 
  ncol = 3, 
  nrow = 1,
  align = "hv")

bbal_ker_habitat_slopes_plot <- annotate_figure(bbal_ker_habitat_slopes_plot, top = text_grob("BBAL Kerguelen", size = 14))

save(bbal_ker_habitat_slopes_plot, file = "Outputs/Plots/bbal_ker_habitat_slopes_plot.RData")


```

# Wandering albatrosses

## Bird Island

```{r waal_birdis_process_data, eval = FALSE}

# Do some data cleaning
load("Outputs/Intermediate/waal_birdis_sst-bathy-chla_data_population_level.RData")

pnts_all <- process_habitat_data(pnts_all, 6, 50)

save(pnts_all, file = "Data/Processed/waal_birdis_habitat_data.RData")


```

### Habitat use

#### Fit the models

```{r waal_birdis_model_average_habitat_used, eval = FALSE}

load("Data/Processed/waal_birdis_habitat_data.RData")

waal_birdis_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID_season) "

waal_birdis_vars <- model_habitat_use(waal_birdis_pnts_all, env_vars.used, "WAAL", "Bird Island", n_sim, my_mod)
waal_birdis_var_used <- waal_birdis_vars$variance
waal_birdis_ranefs_used <- waal_birdis_vars$ranef

save(waal_birdis_var_used, file = "Outputs/Posteriors/waal_birdis_habitat_used_variances.RData")
save(waal_birdis_ranefs_used, file = "Outputs/Posteriors/waal_birdis_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r waal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Outputs/Posteriors/waal_birdis_habitat_used_ranef.RData")

waal_birdis_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" )) 

split_df <- waal_birdis_ranefs_used  %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_birdis_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
waal_birdis_used_plot <- annotate_figure(waal_birdis_used_plot, top = text_grob("WAAL Bird Island", face = "bold", size = 14))
save(waal_birdis_used_plot, file = "Outputs/Plots/waal_birdis_used_plot.RData")


```

### Habitat selection

#### Fit the models

```{r waal_birdis_fit_models, eval = FALSE}

load("Data/Processed/waal_birdis_habitat_data.RData")

# Data processing ====

## Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Add a small constant to avoid log(0)
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )


# Fit the models ====

# Within season #

waal_birdis_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_bathy, file = "Outputs/Models/waal_birdis_glmm_within_bathy.RData")
rm(waal_birdis_glmm.within_bathy); gc()


waal_birdis_glmm.within_chl <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),        
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_chl, file = "Outputs/Models/waal_birdis_glmm_within_chl.RData")
rm(waal_birdis_glmm.within_chl); gc()


waal_birdis_glmm.within_sst <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),      
  data = pnts_all,
  family = binomial() )

save(waal_birdis_glmm.within_sst, file = "Outputs/Models/waal_birdis_glmm_within_sst.RData")
rm(waal_birdis_glmm.within_sst); gc()

```

#### Conditional repeatability

```{r waal_birdis_conditional_repeatability}

# [ 1 ] Calculate conditional repeatability ----

load("Data/Processed/waal_birdis_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = length(env_vars.pref))

for (i in 1:length(env_vars.pref)) {
  
  env_current <- var_map[[env_vars.pref[i]]]
  
  # Load models
  load(paste0("Outputs/Models/waal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "WAAL", "Bird Island")
  
  res.summary <- res_within.summary
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current, "WAAL")

  # Make the conditional repeatability plot
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.within, aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.within, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("waal_birdis_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within)
  rm(list = paste0("waal_birdis_glmm.within_", env_vars.pref[i]))
  gc()
  
}

waal_birdis_condR_results <- do.call("rbind", cond_res.list)
save(waal_birdis_condR_results, file = "Outputs/Posteriors/waal_birdis_condR_results.RData")

```

```{r waal_birdis_plot_slopes}

load("Data/Processed/waal_birdis_habitat_data.RData")
env_vars.pref = c("bathy", "chl", "sst")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars.pref)) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Outputs/Models/waal_birdis_glmm_within_", env_vars.pref[i], ".RData"))
  
  mod_within = get(paste0("waal_birdis_glmm.within_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  
  rnd_eff.all <- rnd_eff.within
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID), alpha = 0.2, colour = line_colour) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
   
  assign(paste0("waal_birdis_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_birdis_preference_figures}

# [ 3 ] Final plot ----

## Conditional repeatability ----

waal_birdis_habitat_condR_plot <- ggarrange(
  waal_birdis_bathy_condR_plot +
    labs(x = "Bathymetry (m below sea level)") + 
    ylim(0, 1),
  waal_birdis_chl_condR_plot + 
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) +
    ylim(0, 1) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  waal_birdis_sst_condR_plot + 
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()), 
   ncol = 3, 
   nrow = 1,
   align = "hv")

waal_birdis_habitat_condR_plot <- annotate_figure(waal_birdis_habitat_condR_plot, top = text_grob("WAAL Bird Island", size = 14))
save(waal_birdis_habitat_condR_plot, file = "Outputs/Plots/waal_birdis_habitat_condR_plot.RData")


## Slopes plot -----

waal_birdis_habitat_slopes_plot <- ggarrange(
  waal_birdis_bathy_pred_plot +
    labs(x = "Bathymetry (m below sea level)"), 
  waal_birdis_chl_pred_plot + 
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()),
  waal_birdis_sst_pred_plot +
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()), 
  ncol = 3,
  nrow = 1, 
  align = "hv")

waal_birdis_habitat_slopes_plot <- annotate_figure(waal_birdis_habitat_slopes_plot,
                top = text_grob("WAAL Bird Island", size = 14))

save(waal_birdis_habitat_slopes_plot, file = "Outputs/Plots/waal_birdis_habitat_slopes_plot.RData")

```

### Heritability

#### Habitat used

```{r waal_birdis_heritability_used }

# Load & prep data ====
load("Outputs/Posteriors/waal_birdis_habitat_used_ranef.RData")
waal_birdis_pedigree.df <- read.csv("Data/Raw/waal_birdis_pedigree.csv") %>% select(-Year)

data_prep <- prep_used_heritability_data(waal_birdis_ranefs_used, waal_birdis_pedigree.df)
waal_birdis_used_h2.df <- data_prep$data
waal_birdis_used_h2.Amat <- data_prep$Amat


# Fit models ====
models_list <- lapply(env_vars.used, function(e) {
  fit_habitat_heritability_model("estimate", e, waal_birdis_used_h2.df, waal_birdis_used_h2.Amat, priors_animal, "Outputs/Models/waal_birdis_")
})
names(models_list) <- env_vars.used


# Extract h2 ==== 
waal_birdis_h2_used.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chlA"   = "Chlorophyll",
      "SST"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_birdis_h2_used.posteriors) <- NULL

save(waal_birdis_h2_used.posteriors, file = "Outputs/Posteriors/waal_birdis_habitat_used_h2_posteriors.RData")

```

#### Habitat preference

```{r waal_birdis_heritability_pref }

# Extract slopes ====

load("Data/Processed/waal_birdis_habitat_data.RData")
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

waal_birdis_habitat_slopes <- extract_habitat_slopes(env_vars.pref, "waal_birdis", pnts_meta)
save(waal_birdis_habitat_slopes, file = "Outputs/Posteriors/waal_birdis_habitat_slopes.RData")


# Estimate heritability ===

## Prep the data
waal_birdis_pedigree.df <- read.csv("Data/Raw/waal_birdis_pedigree.csv") %>% dplyr::select(-Year)

# Fix birds that are recorded as both dams and sires
waal_birdis_pedigree.df$sire[waal_birdis_pedigree.df$id == "1271009"] <- "1318880"
waal_birdis_pedigree.df$dam[waal_birdis_pedigree.df$id == "1271009"] <- "1318939"

data_prep <- prep_slope_heritability_data(waal_birdis_habitat_slopes, waal_birdis_pedigree.df)
waal_birdis_pref_h2.df <- data_prep$data
waal_birdis_pref_h2.Amat <- data_prep$Amat


# Fit models ====

models_list <- lapply(env_vars.pref, function(e) {
  fit_habitat_heritability_model("slope", e, waal_birdis_pref_h2.df, waal_birdis_pref_h2.Amat, priors_animal, "Outputs/Models/waal_birdis_")
})
names(models_list) <- env_vars.pref


# Extract h2 ==== 
waal_birdis_h2_slope.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chl"   = "Chlorophyll",
      "sst"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_birdis_h2_slope.posteriors) <- NULL

save(waal_birdis_h2_slope.posteriors, file = "Outputs/Posteriors/waal_birdis_habitat_pref_h2_posteriors.RData")

```

## Crozet

Chlorophyll values are heavily skewed, with extremely high
concentrations occurring near coastlines in the 'available' dataset. To
prevent these large outliers from unduly influencing the model results,
I apply a log-transformation to chlorophyll.

```{r waal_cro_process_data, eval = FALSE}

# Do some data cleaning
load("Data/Intermediate/waal_cro_sst-bathy-chla_data_population_level.RData")

## Fix some rings
pnts_all %<>%
  mutate(ring = recode(ring,
                       "11376" = "BS11376",
                       "11378" = "BS11378",
                       "25076" = "BS25076",
                       "8196" = "BS28916",
                       "8337" = "BS28337",
                       "BS8337" = "BS28337",
                       "BB21651" = "BS81651",
                       "BS7625" = "BS27625",
                       "BS8402" = "BS28402",
                       "BS8404" = "BS28404",
                       "BS9084" = "BS19084",
                       "BS9444" = "BS19444"))

pnts_all <- process_habitat_data(pnts_all, 6, 50)

save(pnts_all, file = "Data/Processed/waal_cro_habitat_data.RData")

```

### Habitat use

#### Fit the models

```{r waal_cro_model_average_habitat_used, eval = FALSE}

load("Data/Processed/waal_cro_habitat_data.RData")

waal_cro_pnts_all <- prep_used_data(pnts_all, "1 hours")

my_mod <- "~ dist_col + phase + (1| season) + (1 | tripID) + (1 | ID) + (1 | ID_season) "

waal_cro_vars <- model_habitat_use(waal_cro_pnts_all, env_vars.used, "WAAL", "Crozet", n_sim, my_mod)
waal_cro_var_used <- waal_cro_vars$variance
waal_cro_ranefs_used <- waal_cro_vars$ranef

save(waal_cro_var_used, file = "Outputs/Posteriors/waal_cro_habitat_used_variances.RData")
save(waal_cro_ranefs_used, file = "Outputs/Posteriors/waal_cro_habitat_used_ranef.RData")

```

#### Plot habitat use

```{r waal_birdis_plot_habitat_used, fig.dim = c(10,5)}

load("Outputs/Posteriors/waal_cro_habitat_used_ranef.RData")

waal_cro_ranefs_used %<>%
  filter(group %in% c("ID", "ID_season")) %>%
  mutate(env = case_when(
    env == "bathy" ~ "Bathymetry (m below sea level)",
    env == "chlA"  ~ "Chlorophyll (mg Chl m^-3)",
    env == "SST"   ~ "SST (°C)" )) 

split_df <- waal_cro_ranefs_used  %>% group_by(env) %>% group_split()

lapply(split_df, make_plot)
plots <- lapply(split_df, make_plot)

bathy_ranef.plot <- plots[[1]] 
chl_ranef.plot <- plots[[2]] + 
  theme(axis.title.y = element_blank())
sst_ranef.plot <- plots[[3]] + 
  theme(axis.title.y = element_blank())

waal_cro_used_plot <- ggarrange(bathy_ranef.plot, chl_ranef.plot, sst_ranef.plot, ncol = 3)
waal_cro_used_plot <- annotate_figure(waal_cro_used_plot, top = text_grob("WAAL Crozet", face = "bold", size = 14))
save(waal_cro_used_plot, file = "Outputs/Plots/waal_cro_used_plot.RData")

```

### Habitat selection

#### Fit the models

```{r waal_cro_fit_models, eval = FALSE}

load("Data/Processed/waal_cro_habitat_data.RData")

# Add a small constant to avoid log(0) when logging chl
pnts_all %<>% mutate(chl_scaled_log = scale(log(chlA + 0.01)) )

# Thin to every hour to reduce autocorrelation
pnts_all %<>%
  group_by(tripID, date_hourly) %>%
  # first, pick 1 used point if there are any
  {
    used_point <- filter(., used == 1) %>%
      slice_sample(n = 1)

    avail_points <- filter(., used == 0) %>%
      slice_sample(n = 6)

    bind_rows(used_point, avail_points)
  } %>%
  ungroup()

# Within season #

waal_cro_glmm.within_bathy <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + bathy_scaled | tripID) + (1 | ID_season) + (0 + bathy_scaled | ID_season),    data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_bathy, file = "Outputs/Models/waal_cro_glmm_within_bathy.RData")
rm(waal_cro_glmm.within_bathy); gc()


waal_cro_glmm.within_chl <-  glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + chl_scaled_log | tripID) + (1 | ID_season) + (0 + chl_scaled_log | ID_season),         
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.within_chl, file = "Outputs/Models/waal_cro_glmm_within_chl.RData")
rm(waal_cro_glmm.within_chl); gc()


waal_cro_glmm.within_sst <- glmmTMB(
  used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
    (1 | tripID) + (0 + sst_scaled | tripID) + (1 | ID_season) + (0 + sst_scaled | ID_season),  
  data = pnts_all,
  family = binomial())

save(waal_cro_glmm.within_sst, file = "Outputs/Models/waal_cro_glmm_within_sst.RData")
rm(waal_cro_glmm.within_sst); gc()

# Between season #

waal_cro_glmm.between_bathy <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + bathy_scaled | tripID) + (1 | ID) + (0  + bathy_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_bathy, file = "Outputs/Models/waal_cro_glmm_between_bathy.RData")
rm(waal_cro_glmm.between_bathy); gc()


waal_cro_glmm.between_chl <- glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + chl_scaled_log | tripID) + (1 | ID) + (0  + chl_scaled_log | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_chl, file = "Outputs/Models/waal_cro_glmm_between_chl.RData")
rm(waal_cro_glmm.between_chl); gc()

waal_cro_glmm.between_sst <-  glmmTMB(
   used ~ bathy_scaled + chl_scaled_log + sst_scaled + dist_col_scaled + phase +
         (1 | tripID) + (0  + sst_scaled | tripID) + (1 | ID) + (0  + sst_scaled | ID),
  data = pnts_all,
  family = binomial() )

save(waal_cro_glmm.between_sst, file = "Outputs/Models/waal_cro_glmm_between_sst.RData")
rm(waal_cro_glmm.between_sst); gc()


```

#### Conditional repeatability

```{r waal_cro_conditional_repeatability}

# [ 1 ] Calculate conditional repeatability ----

load("Data/Processed/waal_cro_habitat_data.RData")

## Store envronmental variables
bathy_scaled <- unlist(pnts_all$bathy_scaled)
chl_scaled <- unlist(pnts_all$chl_scaled)
sst_scaled <- unlist(pnts_all$sst_scaled)

var_map <- list(bathy = bathy_scaled, chl = chl_scaled, sst = sst_scaled )

## Approximate degrees of freedom for uncertainty estimates:
df_ID <- length(unique(pnts_all$ID)) - 1
df_ID_season <- length(unique(pnts_all$ID_season)) - 1
df_tripID <- length(unique(pnts_all$tripID)) - 1

cond_res.list <- vector(mode = "list", length = length(env_vars.pref))

for (i in 1:length(env_vars.pref)) {
  
  env_current <- var_map[[env_vars.pref[i]]]
  
  # Load models
  load(paste0("Outputs/Models/waal_cro_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Outputs/Models/waal_cro_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars.pref[i]))
  
  # Estimate slope variances and conditional repeatability
  res_between.summary <- extract_pref_metrics(mod_between, env_vars.pref[i], "between", env_current, "WAAL", "Crozet")
  res_within.summary <- extract_pref_metrics(mod_within, env_vars.pref[i], "within", env_current, "WAAL", "Crozet")
  
  res.summary <- rbind(res_between.summary, res_within.summary)
  cond_res.list[[i]] <- res.summary
  
  # Boostrap uncertainty in conditional repeatability
  condR_uncertainty.between <- bootstrap_condR(mod_between, env_vars.pref[i], "between", n_sim, env_current, "WAAL")
  condR_uncertainty.within <- bootstrap_condR(mod_within, env_vars.pref[i], "within", n_sim, env_current, "WAAL")

  # Make the conditional repeatability plot
  #xrange_90 <- quantile(env_current, probs = c(0.05, 0.95), na.rm = TRUE)
  
  condR_plot <- ggplot() +
    geom_ribbon(data = condR_uncertainty.between, aes(x = x, ymin = lower, ymax = upper), fill = col_between, alpha = 0.4) +
    geom_ribbon(data = condR_uncertainty.within, aes(x = x, ymin = lower, ymax = upper), fill = col_within, alpha = 0.4) +
    
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_between, size = 1) +
    geom_line(data = condR_uncertainty.between, aes(x=x, y=median), col = col_within, size = 1) +
    
    geom_hline(data = res_between.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_between) +
    geom_hline(data = res_within.summary, aes(yintercept = med_rpt), linetype = "dashed", col = col_within) +
    
    labs(y = "Conditional repeatability") +
    theme_bw() 
  
  assign(paste0("waal_cro_", env_vars.pref[i], "_condR_plot"), condR_plot)
  
  rm(mod_within, mod_between)
  rm(list = paste0("waal_cro_glmm.within_", env_vars.pref[i]))
  rm(list = paste0("waal_cro_glmm.between_", env_vars.pref[i]))
  gc()
  
}

waal_cro_condR_results <- do.call("rbind", cond_res.list)
save(waal_cro_condR_results, file = "Outputs/Posteriors/waal_cro_condR_results.RData")

```

```{r waal_cro_plot_slopes}

env_vars.pref = c("bathy", "chl", "sst")

# [ 2 ] Plot individual slopes ---- 

for (i in 1:length(env_vars.pref)) {

  habitat_var = paste0(env_vars.pref[i], "_scaled")
  if(grepl("chl", habitat_var)) { habitat_var = "chl_scaled_log"}
  
  # Get models
  load(paste0("Outputs/Models/waal_cro_glmm_within_", env_vars.pref[i], ".RData"))
  load(paste0("Outputs/Models/waal_cro_glmm_between_", env_vars.pref[i], ".RData"))

  mod_within = get(paste0("waal_cro_glmm.within_", env_vars.pref[i]))
  mod_between = get(paste0("waal_cro_glmm.between_", env_vars.pref[i]))
  
  mod_df <- mod_within$frame %>% separate(ID_season, into = c("ID", "season"), sep = "_", remove = F) %>% filter(used == 1)
  
  # Get environmental range
  ind_ranges <- mod_df %>%
    group_by(ID) %>%
    summarise(min_env = min(!!sym(habitat_var), na.rm = TRUE),
              max_env = max(!!sym(habitat_var), na.rm = TRUE)) 
  
  predictions.ind <- ind_ranges %>%
    rowwise() %>%
    mutate(habitat_val = list(seq(min_env, max_env, length.out = 100))) %>%
    rename(ring = ID) %>%
    unnest(cols = c(habitat_val)) 
  
  # Get random effects
  rnd_eff.within <- get_rnd_effects(mod_within, habitat_var) %>%
    separate(ID, c("ring", "season"), remove = F) %>% mutate(season_type = "Within season")
  rnd_eff.between <- get_rnd_effects(mod_between, habitat_var) %>%
    mutate(ring = ID, season = NA, season_type = "Between seasons")
  
  rnd_eff.all <- rbind(rnd_eff.within, rnd_eff.between)
  
  predictions.ind <- merge(predictions.ind, rnd_eff.all, by = "ring", all.x = TRUE)
  predictions.ind$logit <- predictions.ind$intercept + predictions.ind$slope * predictions.ind$habitat_val
  predictions.ind$probability <- 1 / (1 + exp(-predictions.ind$logit))  
  
  ## Plot
  
  ### Set colours
  line_colour <- if (grepl("bathy", habitat_var)) {
         bathy_col   } else if (grepl("chl", habitat_var)) {
  chl_col  } else if (grepl("sst", habitat_var)) {
  sst_col } 
  
  pred_plot <- ggplot() +
      geom_line(data = predictions.ind, aes(x = habitat_val, y = probability, group = ID),
                colour = line_colour, alpha = 0.2) +
      facet_grid(~season_type) +
      theme_bw() +
      theme(legend.position = "none") +
      labs(y = "Probability of use")
  
  assign(paste0("waal_cro_", env_vars.pref[i], "_pred_plot"), pred_plot)
  
}

```

```{r waal_cro_preference_figures}

# [ 3 ] Final plot ----

## Conditional repeatability ----

waal_cro_habitat_condR_plot <- ggarrange(
  waal_cro_bathy_condR_plot +
    labs(x = "Bathymetry (m below sea level)") + 
    ylim(0, 1),
  waal_cro_chl_condR_plot +
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()),
  waal_cro_sst_condR_plot + 
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    ylim(0, 1) + 
    theme(axis.title.y = element_blank(),
          axis.text.y = element_blank()), 
  ncol = 3, 
  nrow = 1,
  align = "hv")

waal_cro_habitat_condR_plot <- annotate_figure(waal_cro_habitat_condR_plot, top = text_grob("WAAL Crozet", size = 14))

save(waal_cro_habitat_condR_plot, file = "Outputs/Plots/waal_cro_habitat_condR_plot.RData")


## Slopes -----

waal_cro_habitat_slopes_plot <- ggarrange(
  waal_cro_bathy_pred_plot + 
    labs(x = "Bathymetry (m below sea level)"), 
  waal_cro_chl_pred_plot +
    labs(x = expression("Chlorophyll concentration (" * mg ~ Chl ~ m^-3 * ")")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()),
  waal_cro_sst_pred_plot +
    labs(x = expression("Sea surface temperature (" * ~ degree ~ "C)")) + 
    theme(axis.title.y = element_blank(), 
          axis.text.y = element_blank()),
  ncol = 3,
  nrow = 1,
  align = "hv")

  
waal_cro_habitat_slopes_plot <- annotate_figure(waal_cro_habitat_slopes_plot,
                top = text_grob("WAAL Crozet", size = 14))

save(waal_cro_habitat_slopes_plot, file = "Outputs/Plots/waal_cro_habitat_slopes_plot.RData")

```

### Heritability

#### Habitat used

```{r waal_cro_heritability_used }

# Load & prep data ====
load("Outputs/Posteriors/waal_cro_habitat_used_ranef.RData")
waal_cro_pedigree.df <- read.csv("Data/Raw/waal_cro_pedigree.csv") %>% select(-Year)

data_prep <- prep_used_heritability_data(waal_cro_ranefs_used, waal_cro_pedigree.df)
waal_cro_used_h2.df <- data_prep$data
waal_cro_used_h2.Amat <- data_prep$Amat


# Fit models ====
models_list <- lapply(env_vars.used, function(e) {
  fit_habitat_heritability_model("estimate", e, waal_cro_used_h2.df, waal_cro_used_h2.Amat, priors_animal, "Outputs/Models/waal_cro_")
})
names(models_list) <- env_vars.used


# Extract h2 ==== 
waal_cro_h2_used.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chlA"   = "Chlorophyll",
      "SST"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_cro_h2_used.posteriors) <- NULL

save(waal_cro_h2_used.posteriors, file = "Outputs/Posteriors/waal_cro_habitat_used_h2_posteriors.RData")

```

#### Habitat preference

```{r waal_cro_heritability_pref }

# Extract slopes ====

load("Data/Processed/waal_cro_habitat_data.RData")
pnts_meta <- pnts_all %>% select(c(tripID, phase)) %>% distinct()

waal_cro_habitat_slopes <- extract_habitat_slopes(env_vars.pref, "waal_cro", pnts_meta)
save(waal_cro_habitat_slopes, file = "Outputs/Posteriors/waal_cro_habitat_slopes.RData")


# Estimate heritability ===

## Prep the data
waal_cro_pedigree.df <- read.csv("Data/Raw/waal_cro_pedigree.csv") %>% dplyr::select(-Year)

data_prep <- prep_slope_heritability_data(waal_cro_habitat_slopes, waal_cro_pedigree.df)
waal_cro_pref_h2.df <- data_prep$data
waal_cro_pref_h2.Amat <- data_prep$Amat


# Fit models ====

models_list <- lapply(env_vars.pref, function(e) {
  fit_habitat_heritability_model("slope", e, waal_cro_pref_h2.df, waal_cro_pref_h2.Amat, priors_animal, "Outputs/Models/waal_cro_")
})
names(models_list) <- env_vars.pref


# Extract h2 ==== 
waal_cro_h2_slope.posteriors <- purrr::map_df(models_list, ~ {
  data.frame(h2 = get_h2(.x))
}, .id = "env") %>%
  mutate(env = recode(env,
      "bathy" = "Bathymetry",
      "chl"   = "Chlorophyll",
      "sst"   = "SST")) %>%
  mutate(env = factor(env, levels = c("SST", "Chlorophyll", "Bathymetry"))) 

rownames(waal_cro_h2_slope.posteriors) <- NULL

save(waal_cro_h2_slope.posteriors, file = "Outputs/Posteriors/waal_cro_habitat_pref_h2_posteriors.RData")

ggplot(posteriors, aes(x = h2, y = env)) +
  geom_density_ridges(fill = "#1b9e77", alpha = 0.5, scale = 1) +
  labs(x = "Additive genetic heritability (h²)", y = "") +
  theme_bw()

```
