---
title: "Heritability results"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_purrr(v.1.2.1)_, _pedigree(v.1.4.2)_, _pedtricks(v.0.5.0)_, _MCMCglmm(v.2.36)_, _ape(v.5.8-1)_, _MasterBayes(v.2.59)_, _kinship2(v.1.9.6.2)_, _quadprog(v.1.5-8)_, _Matrix(v.1.5-4.1)_, _genetics(v.1.3.8.1.3)_, _mvtnorm(v.1.3-3)_, _MASS(v.7.3-60)_, _gtools(v.3.9.5)_, _gdata(v.3.0.1)_, _combinat(v.0.0-8)_, _coda(v.0.19-4.1)_, _ggridges(v.0.5.7)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Define & load the packages
packages <- c("dplyr", "magrittr", "ggplot2", "ggridges", "MasterBayes", "mvtnorm", "MASS", "MCMCglmm", "pedtricks", "pedigree", "purrr") 

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
#  if (any(installed_packages == FALSE)) {
#   install.packages(packages[!installed_packages])
# }

invisible(lapply(packages, library, character.only = TRUE))

## Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)

## Make sure using dplyr select
select <- dplyr::select

# Set colours
col_effort <- "#1b9e77"
col_bearing <- "#aaf0db"
col_slope <- "#0d4e3a"
col_used <- "#3cddac"

source("R/Functions/RPT_functions.R")

```

# Pedigree summary

```{r summarise_pedigrees}

pedigree_list <- list(
  bbal_birdis = "Data/Raw/bbal_birdis_pedigree.csv",
  waal_birdis = "Data/Raw/waal_birdis_pedigree.csv",
  waal_cro    = "Data/Raw/waal_cro_pedigree.csv" )

master_pedigree <- map_df(pedigree_list, function(path) {
  read.csv(path) %>%
    mutate(across(any_of(c("id", "dam", "sire")), as.character)) }, .id = "population") %>%
  select(-any_of("Year")) %>%
  mutate(across(c(dam, sire), ~na_if(., "0")))

spec_col <- c("bbal_birdis", "waal_birdis", "waal_cro")
summary.table <- vector(mode = "list", length = length(spec_col))

for (i in 1:length(spec_col)) {
  
  ped <- master_pedigree %>% filter(population == spec_col[i]) %>% select(-population)
  order_ped <- MasterBayes::orderPed(ped)
  depths = countGen(order_ped)
  
  df <- data.frame(
    N              = nrow(order_ped),
    founders       = sum(is.na(order_ped$sire) & is.na(order_ped$dam)),
    pct_paternal   = round(sum(!is.na(order_ped$sire)) / nrow(order_ped) * 100, 1),
    pct_maternal   = round(sum(!is.na(order_ped$dam)) / nrow(order_ped) * 100, 1),
    gen_depth_mean = round(mean(depths), 2),
    gen_depth_max  = max(depths),
    population     = spec_col[i])
  
  summary.table[[i]] <- df
  
}

table_s2 <- do.call("rbind", summary.table)
save(table_s2, file = "Outputs/Posteriors/table_s2_data.RData")

```


# Power analysis

```{r power_analysis}


## Fix some dam/sire encoding
# pedigree.df$sire[pedigree.df$sire == "0"] <- NA
# 
# pedigree.df$sire[pedigree.df$id == "1271009"] <- "1318880"
# pedigree.df$dam[pedigree.df$id == "1271009"] <- "1318939"
# 
# pedigree.df$sire[pedigree.df$id == "1434383"] <- "1139670"
# pedigree.df$dam[pedigree.df$id == "1434383"] <- "1146056"
# 
# pedigree.df$sire[pedigree.df$id == "4003739"] <- "5116152"
# pedigree.df$dam[pedigree.df$id == "4003739"] <- "5142862"
# 
# pedigree.df$sire[pedigree.df$id == "4003674"] <- "4000720"
# pedigree.df$dam[pedigree.df$id == "4003674"] <- "5110290"
# 
# pedigree.df$sire[pedigree.df$id == "5144330"] <- "5093486"
# pedigree.df$dam[pedigree.df$id == "5144330"] <- "5097831"


# CHANGE THIS BIT TO CYCLE THROUGH EACH COLONY #

# Load behaviour data
load("Data/Processed/behavioural_repeatability_data.RData") 
  
# Set parameters for simulated genetic analysis
VA <- seq(from=0.0, to =0.3, by=0.02)
nrep <- 200

spec_col <- c("bbal_birdis", "waal_birdis", "waal_cro") 

for (i in 1:length(spec_col)){
  
  species_i <- strsplit(spec_col[i], "_")[[1]][1]
  colony_i <- strsplit(spec_col[i], "_")[[1]][2]

  # Load data ====
  
  focal.dat <- rep_data %>% filter(species == species_i & colony == colony_i)
  pedigree.df <- read.csv(paste0("Data/Raw/", spec_col[i], "_pedigree.csv"))
  
  ## Check pedigree in correct format
  pedigree.df <- fix_ped(pedigree.df)
  pedigree.df <- MasterBayes::orderPed(pedigree.df)
  
  ## Prune to used dataset
  pedigree_pruned <- prune_pedigree(focal.dat, pedigree.df)
  
  #draw_ped(pedigree_pruned)
  
  # Simulate genetic analyis ====
  
  ## Make inverse genetic matrix
  ainv <- inverseA(pedigree_pruned)
  
  ## Create output wrapper
  all_va_results <- data.frame(VA = numeric(),
                               rep = integer(),
                               posterior = numeric(),
                               cilow = numeric(),
                               cihigh = numeric())
  
  ## Set progress bar
  n_total <- length(VA) * nrep
  pb <- txtProgressBar(min = 0, max = n_total, style = 3)
  counter <- 0
  
  for (j in 1:length(VA))  {
      
      print(paste0("Processing VA: ", VA[i], " (", j, " of ", length(VA), ")"))
    
      va_results <- expand.grid(VA=VA[j], rep=1:nrep, posterior=NA, cilow=NA, cihigh=NA)
  
      for(k in 1:nrep){
         
        xP <- phen_sim(pedigree_pruned, traits = 1, randomA = VA[j]*diag(1), randomE = (1-VA[j])*diag(1), verbose = F)[[1]]
        
        # wrap the MCMCglmm fit in tryCatch
        fit <- tryCatch({
          m0sim <- MCMCglmm(data = xP,
                            fixed = trait_1 ~ 1,
                            random =~ id,
                            ginverse = list(id = ainv$Ainv),
                            verbose = FALSE)
          
          h2 <- m0sim$VCV[,1] / (m0sim$VCV[,1] + m0sim$VCV[,2])
          
          c(median(h2), HPDinterval(h2)[1], HPDinterval(h2)[2])
          
        }, error = function(e) {
          # return NAs if it fails
          c(NA, NA, NA)
        })
        
        va_results[va_results$rep == k, c("posterior","cilow","cihigh")] <- fit
        
        ## Update progress bar
        counter <- counter + 1
        setTxtProgressBar(pb, counter)
      
      }
     
      all_va_results <- rbind(all_va_results, va_results)
    
    }

  save(all_va_results, file = paste0("Outputs/Posteriors/", spec_col[i], "_heritability_power_analysis.RData"))

}


# Plot results ----

for (i in 1:length(spec_col)) {
  
  load(paste0("Outputs/Posteriors/", spec_col[i], "_heritability_power_analysis.RData"))
  
  ggplot(all_va_results, aes(x=VA, y=posterior)) + 
    geom_point() 
  
  ggplot(all_va_results, aes(x=VA, y=cilow)) + 
    geom_point() 
  
  all_va_results %>% 
    group_by(VA) %>%
    summarise(coverage=mean(VA<=cihigh & VA>=cilow, na.rm = T), 
              power=mean(cilow>=quantile(all_va_results$posterior[all_va_results$VA==0], probs = 0.975), na.rm = T))
  
  # Another way of summarising power
  all_va_results %>% group_by(VA) %>%
    summarise(coverage=mean(VA<=cihigh & VA>=cilow, na.rm = T), power=mean(cilow>= all_va_results$posterior[all_va_results$VA==0], na.rm = T))
}


```

# Combine posteriors

```{r combine_posteriors}

# Foraging effort ====

posteriors_effort <- loadRData("Outputs/Posteriors/behaviour_heritability_posteriors.RData")

## Recode behaviour & spec col
posteriors_effort %<>%
  mutate(behaviour = recode(behaviour,
                            "duration.hrs" = "Trip duration",
                            "landings_per_day" = "Landings per day",
                            "transit_time.pct" = "Flight time",
                            "rest_time.pct" = "Rest time",
                            "transit_to_rest.ratio" = "Flight:Rest ratio"),
          spec_col = recode(spec_col,
                "bbal_birdis" = "BBAL Bird Island",
                "waal_birdis" = "WAAL Bird Island",
                "waal_cro" = "WAAL Crozet"),
         behaviour = factor(behaviour, levels = rev(c("Trip duration", "Rest time", "Landings per day", "Flight:Rest ratio", "Flight time"))))


# Foraging bearings ====

posteriors_bearings <- loadRData("Outputs/Posteriors/bearing_heritability_posteriors.RData")


# Habitat used ====

bbal_birdis.used <- loadRData("Outputs/Posteriors/bbal_birdis_habitat_used_h2_posteriors.RData") %>% mutate(spec_col = "BBAL Bird Island")
waal_birdis.used <- loadRData("Outputs/Posteriors/waal_birdis_habitat_used_h2_posteriors.RData") %>% mutate(spec_col = "WAAL Bird Island")
waal_cro.used <- loadRData("Outputs/Posteriors/waal_cro_habitat_used_h2_posteriors.RData") %>% mutate(spec_col = "WAAL Crozet")

posteriors_used <- rbind(bbal_birdis.used, waal_birdis.used, waal_cro.used)


# Habitat slopes =====

bbal_birdis.slopes <- loadRData("Outputs/Posteriors/bbal_birdis_habitat_pref_h2_posteriors.RData") %>% mutate(spec_col = "BBAL Bird Island")
waal_birdis.slopes <- loadRData("Outputs/Posteriors/waal_birdis_habitat_pref_h2_posteriors.RData") %>% mutate(spec_col = "WAAL Bird Island")
waal_cro.slopes <- loadRData("Outputs/Posteriors/waal_cro_habitat_pref_h2_posteriors.RData") %>% mutate(spec_col = "WAAL Crozet")

posteriors_slopes <- rbind(bbal_birdis.slopes, waal_birdis.slopes, waal_cro.slopes)




# Bind the behaviours into one plot ====

posteriors_effort %<>%
  select(c(spec_col, h2_post, behaviour)) %>%
  rename(h2 = h2_post) %>%
  mutate(group = "effort")

posteriors_bearings %<>%
  select(c(spec_col, h2)) %>%
  mutate(behaviour = "Foraging bearing",
         group = "bearings")

posteriors_slopes %<>%
  rename(behaviour = env) %>%
  mutate(behaviour = paste0(behaviour, " preference"),
         group = "slopes") %>%
  relocate(c(spec_col, h2), .before = behaviour) %>%
  mutate(behaviour = factor(
    behaviour,
    levels = rev(c("Bathymetry preference", "Chlorophyll preference", "SST preference"))))

posteriors_used %<>%
  rename(behaviour = env) %>%
  mutate(behaviour = paste0(behaviour, " used"),
         group = "used") %>%
  relocate(c(spec_col, h2), .before = behaviour) %>%
  mutate(behaviour = factor(
    behaviour,
    levels = rev(c("Bathymetry used", "Chlorophyll used", "SST used"))))

posteriors_all <- rbind(posteriors_effort, posteriors_bearings, posteriors_used, posteriors_slopes) 

# Change ordering for plot/table
posteriors_all %<>%
   mutate(behaviour = factor(behaviour, 
                             levels = c("Trip duration", "Rest time", "Landings per day", "Flight:Rest ratio", "Flight time",
                                        "Foraging bearing",
                                        "Bathymetry used", "Chlorophyll used", "SST used",
                                        "Bathymetry preference", "Chlorophyll preference", "SST preference"))) %>%
   arrange(spec_col, behaviour)

posteriors_all$behaviour <- factor(
  posteriors_all$behaviour,
  levels = rev(levels(posteriors_all$behaviour)))

save(posteriors_all, file = "Outputs/Posteriors/heritability_posteriors_all.RData")

```


