---
title: "Processing 1: Extract foraging trips"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

Purpose of script: Generalised script to extract foraging trips from GLS and GPS data

Inputs:

-   `Data/Raw/{species}_{colony}_gls.RData`

-   `Data/Raw/{species}{colony}_gps.RData`

Outputs:

-   `Data/Intermediate/{species}_{colony}gls_labelled.RData`

-   `Data/Intermediate/{species}{colony}_gps_labelled.RData`

# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit)

**locale:** *LC_COLLATE=English_United Kingdom.utf8*, *LC_CTYPE=English_United Kingdom.utf8*, *LC_MONETARY=English_United Kingdom.utf8*, *LC_NUMERIC=C* and *LC_TIME=English_United Kingdom.utf8*

**attached base packages:** *stats*, *graphics*, *grDevices*, *utils*, *datasets*, *methods* and *base*

**other attached packages:** *data.table(v.1.18.2.1)*, *ggplot2(v.4.0.2)*, *sf(v.1.0-20)*, *magrittr(v.2.0.3)* and *dplyr(v.1.2.0)*

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

source("R/0_setup.R")

# Define the packages
packages <- c("dplyr", "magrittr", "sf", "ggplot2", "data.table")

# Install packages not yet installed - change lib to library path
#installed_packages <- packages %in% rownames(installed.packages())

#if (any(installed_packages == FALSE)) {
#  install.packages(packages[!installed_packages], lib = "C:/Users/libraryPath")
#}

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

```

```{r set_population}

# Must be manually changed
spec_col <- "bbal_birdis" #"bbal_birdis" "bbal_ker" "waal_birdis" "waal_cro"

```

## Label GPS

```{r label_GPS}

# Load GPS
my_file <- list.files("Data/Raw/", pattern = paste0(spec_col, "_gps"))
my_gps <- loadRData(paste0("Data/Raw/", my_file)) 

my_gps %<>%
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(across(c(latitude, longitude), as.numeric)) %>% filter(!is.na(longitude) & !is.na(latitude)) %>%
  arrange(ring,datetime)


# Label GPS data with colony vs trip ----

my_gps$ringYr <- paste0(my_gps$ring, "_", my_gps$season)
my_rings <- unique(my_gps$ringYr)

gps_labelled.list <- list()

pb <- txtProgressBar(min = 0, max = length(my_rings), style = 3)

for (i in 1:length(my_rings)) {
  # Progress bar
  setTxtProgressBar(pb, i)
  
  dat_i <- subset(my_gps, ringYr == my_rings[i]) %>% distinct()
  
  dat_i %<>% 
    filter(!is.na(datetime)) %>%
    arrange(datetime) %>%
    select(c(ringYr, ring, season, datetime, longitude, latitude))
  
  ### Catch problem data
  if ( sum(dat_i$longitude) == 0 |
       nrow(dat_i) < 5 |
       is.na(dat_i$ring)[1] ) {
    next }
  
  ## Calculate speed ----
  mytrip.filtered.df <- dat_i
  
  mytrip.filtered.df$calc_speed <- calc_speed(mytrip.filtered.df)
  mytrip.filtered.df %<>% filter(calc_speed < 25)
  
  ## Calculate distance from colony ----
  mytrip.filtered.df$dist_col <- apply(mytrip.filtered.df, 1, function(row) {
    point_coords <- c(as.numeric(row["longitude"]), as.numeric(row["latitude"]))
    geosphere::distHaversine(point_coords, col_coords)
  })
  
  
  ### @CHECKSUMS@ ----
  # Ensure trips start and end within colony #
  
  ## Find minimum distance from colony
  min_col_dist <- median(head(sort(mytrip.filtered.df$dist_col), 10))
  
  if(min_col_dist > 10000) next 
  
  ### Interactively filter out bad trips (uncomment to run) ----
  # plot(mytrip.filtered.df$longitude, mytrip.filtered.df$latitude, type = "l",
  #      main = paste0(my_rings[i], "; ", i, " of ", length(my_rings)))
  # points(mytrip.filtered.df$longitude, mytrip.filtered.df$latitude, col = "red", pch = 20)
  # points(col_coords$longitude, col_coords$latitude, col = "green", pc = 15, cex = 2)

  # ## Ask for a decision
  # click <- locator(1)  # Wait for a single click
  # 
  # if (!is.null(click)) { 
  #   remove_ids <- c(remove_ids, my_rings[i])
  #   cat("Marked for removal:", my_rings[i], "\n")
  #   next
  # } 
  # 
  # # Output clean data
  # gps_cleaned.list[[i]] <- mytrip.filtered.df
  
  
  ### Determine whether or not at colony based on distance
  mytrip.filtered.df %<>%
    mutate(loc = ifelse(dist_col < 6000, "colony", "trip")) 
  
  mytrip.filtered.df <- mytrip.filtered.df[order(mytrip.filtered.df$datetime),]
 
  mytrip.filtered.df %<>% 
    mutate(bout = rep(1:length(rle(loc)$lengths), rle(loc)$lengths),
           boutID = paste(loc, ceiling(bout/2), sep = "_")) %>%
    group_by(boutID) %>%
    mutate(duration.mins = as.numeric(difftime(max(datetime), min(datetime), units = "mins")),
           duration.hours = duration.mins/60,
           duration.days = duration.hours/24) 
  
  ### Make sure each trip ends within 40km of colony (allows device to fail 30 mins before arrival assuming max speed 80km/h)
  mytrip.filtered.df %<>% group_by(boutID) %>% 
    mutate(ends_home = ifelse(dist_col[n()] > 40000, "N", "Y"))
  
  mytrip.filtered.df %<>% 
    mutate(ends_home = ifelse(loc == "colony", "Y", ends_home))
  
  mytrip.filtered.df %<>% filter(ends_home == "Y") %>% select(-ends_home)
  
  ## Save data to lists to output ----
  gps_labelled.list[[i]] <- mytrip.filtered.df

}

close(pb)

gps_labelled.df <- do.call("rbind", gps_labelled.list)
save(gps_labelled.df, file = paste0("Data/Intermediate/", spec_col, "_gps_labelled.RData"))


# SECOND FILTER: Manually check for errors in individual trip tracks (uncomment to run) ----

# load(paste0("Data/Intermediate/", spec_col, "_gps_labelled.RData"))
# 
# gps_labelled.df %<>% mutate(tripID = paste0(ringYr, "_", boutID))
# my_ids <- unique(gps_labelled.df$tripID)
# my_ids <- my_ids[!grepl("colony", my_ids)]
# 
# remove_trips <- c()
# 
# for (i in 1:length(my_ids)) {
# 
#   my_trip <- subset(gps_labelled.df, tripID == my_ids[i])
# 
#   # Interactively filter out bad trips #
#   plot(my_trip$longitude, my_trip$latitude, type = "l",
#        main = paste0(my_ids[i], "; ", i, " of ", length(my_ids)))
#   points(my_trip$longitude, my_trip$latitude, col = "red", pch = 20)
#   points(col_coords$longitude, col_coords$latitude, col = "green", pc = 15, cex = 2)
# 
#   ## Ask for a decision
#   click <- locator(1)  # Wait for a single click
# 
#   if (!is.null(click)) {
#     remove_trips <- c(remove_trips, my_ids[i])
#     cat("Marked for removal:", my_ids[i], "\n")
#     next
#   }
# 
# }
# 
# save(remove_trips, file = paste0("Data/Intermediate/", spec_col, "_trip_errors.RData"))
# 
# gps_labelled.df %<>% filter(!tripID %in% remove_trips)
# save(gps_labelled.df, file = paste0("Data/Intermediate/", spec_col, "_gps_labelled.RData"))

```

## Label GLS

```{r label_GLS}

# Load GLS data
my_file <- list.files("Data/Raw/", pattern = paste0(spec_col, "_gls"))
my_gls <- loadRData(paste0("Data/Raw/", my_file)) 

my_gls %<>%
  mutate(season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 )) %>%
  select(c(ring, season, datetime, immersion))

# Label GLS data with colony vs trip ----

## Get birdyears
my_gls$birdSeason <- paste(my_gls$ring, my_gls$season, sep = "_")
my_birds <- unique(my_gls$birdSeason)
gls_labelled.list <- vector(mode = "list", length = length(my_birds))

pb <- txtProgressBar(min = 0, max = length(my_birds), style = 3)

for (i in 1:length(my_birds)) {

  setTxtProgressBar(pb, i)

  # Isolate gls data
  gls_i <- subset(my_gls, birdSeason == my_birds[i])
  gls_i %<>% arrange(datetime) %>% distinct()

  ## Skip errors
  if (nrow(gls_i) == 0) next
  if (sum(gls_i$immersion) == 0) next

  # Label colony vs trip
  ## Look for dry periods > 8 hours
  gls_i %<>% arrange(datetime) %>%
    mutate(dryWet = ifelse(immersion <= 0.0*max(immersion, na.rm = T), "dry", "wet"),
           dur.mins = rep(rle(dryWet)$lengths, rle(dryWet)$lengths)*10,
           loc = ifelse(dryWet == "dry" & dur.mins > 480, "col", "trip"),
           tripID = rep(1:length(rle(loc)$values), rle(loc)$lengths))

  gls_labelled.list[[i]] <- gls_i

}

close(pb)

gls_labelled.df <- do.call("rbind", gls_labelled.list)

save(gls_labelled.df, file = paste0("Data/Intermediate/", spec_col, "_gls_labelled.RData"))

```
