---
title: "Processing 2: Extract at-sea activity"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

Purpose of script: Extract at-sea activity from GLS data.

Inputs:

-   `Data/Intermediate/{species}_{colony}_gls_labelled.RData`

-   `Data/Metadata/{BAS/Chize}_demo_complete.RData`

Outputs:

-   `Data/Intermediate/{species}_{colony}_individual_gls_trips.RData`


# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_data.table(v.1.18.2.1)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

source("R/0_setup.R")

# Define the packages
packages <- c("dplyr", "magrittr", "readxl", "ggplot2", "data.table", "tidyr")

# Install packages not yet installed - change lib to library path
#installed_packages <- packages %in% rownames(installed.packages())

#if (any(installed_packages == FALSE)) {
#  install.packages(packages[!installed_packages], lib = "C:/Users/libraryPath")
#}

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

```

## Cut GLS to breeding season

```{r cut_GLS_breeding}

spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")

for (i in 1:length(spec_col)) {
  
  load(paste0("Data/Intermediate/", spec_col[i], "_gls_labelled.RData"))
  
  gls_labelled.df %<>%
    mutate(ring = toupper(ring),
           ring = gsub(" ", "", ring),
           ringYr = paste(ring, season, sep = "_")) %>%
    filter(!is.na(ring))
  
  
  ### Load meta data ----
  
  meta_source <- ifelse(grepl("birdis", spec_col[i]), "BAS", "Chize")
  all_demo <- loadRData(paste0("Data/Metadata/", meta_source, "_demo_complete.RData"))
  
  all_demo %<>% mutate(species = tolower(species)) %>% filter(species == my_species) 
  
  ## Make dataframe long
  demoA <- all_demo %>% select(-c(bird2_sex, bird2_birthYr)) %>% 
    dplyr::rename(ring = bird1, sex = bird1_sex, birthYr = bird1_birthYr, partner = bird2) %>%
    filter(!is.na(ring)) 
  
  demoB <- all_demo %>% select(-c(bird1_sex, bird1_birthYr)) %>% 
    dplyr::rename(ring = bird2, sex = bird2_sex, birthYr = bird2_birthYr, partner = bird1) %>%
    filter(!is.na(ring)) 
  
  my_demo <- rbind(demoA, demoB)
  my_demo %<>% group_by(ring, season) %>% tidyr::fill(everything(), .direction = 'updown') %>% distinct()
  
  ## Manual breeding dates ----
  
  if (spec_col[i] == "waal_birdis") {
    
    median_lay <- "12-01" 
    median_hatch <- "03-10"
    
  } else if (spec_col[i] == "bba_birdis") {
    
    median_lay <- "10-13"
    median_hatch <- "01-05"
    
  } else if (spec_col[i] == "waal_cro") {
    
    median_lay <- "12-01"
    median_hatch <- "03-22"
    
  } else if (spec_col[i] == "bba_kerguelen") {
    
    median_lay <- "09-25"
    median_hatch <- "12-15"
  }
  
  ## Match in metadata and filter to breeding ----
  
  setDT(my_demo)
  my_demo %<>% select(ring, season, rs, lay_date, hatch_date, fail_date) %>%
    mutate(season = as.numeric(season)) %>% distinct()
  my_demo <- my_demo[, .SD[sample(.N, 1)], by = .(ring, season)]
  
  setDT(gls_labelled.df)
  
  # Merge using data.table
  gls_labelled.df <- merge(gls_labelled.df, my_demo, by = c("ring", "season"), all.x = T)
  
  ## Cut to breeding
  hatch_time = ifelse(my_species == "bba", 101, 121)
  brood_time = ifelse(my_species == "bba", 38, 48) # brooding ends 22/01
  
  gls_labelled.df %<>% rename(boutID = tripID)
  
  gls_labelled.df <- cut_to_breeding(gls_labelled.df, median_lay, median_hatch, hatch_time, brood_time)
  save(gls_labelled.df, file = paste0("Data/Intermediate/", spec_col[i], "_gls_labelled_subset.RData"))
  
}

```

## Extract behaviours

```{r extract_behaviours}

# Extract behaviours =====

spec_col <- c("bbal_birdis", "bbal_ker", "waal_birdis", "waal_cro")

for (i in 1:length(spec_col)) {
  
  message("Processing ", spec_col[i])
  
  load(paste0("Data/Intermediate/", spec_col[i], "_gls_labelled_subset.RData"))
  
  ## Deal with duplicates
  gls_labelled.df %<>%
    group_by(birdSeason) %>%
    mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"),
           datetime = lubridate::round_date(datetime, "10 minutes")) %>%
    distinct() %>%
    arrange(datetime)
  
  all_birds <- unique(gls_labelled.df$ringYr)
  all_trips.list <- vector(mode = "list", length = length(all_birds))
  
  species <- strsplit(spec_col[i], "_")[[1]][1]
  colony <- strsplit(spec_col[i], "_")[[1]][2]
  
  pb <- txtProgressBar(min = 0, max = length(all_birds), style = 3)
  
  for (j in 1:length(all_birds)) {
    
    setTxtProgressBar(pb, j)
    
    gls_j <- subset(gls_labelled.df, ringYr == all_birds[j]) %>%
      ungroup() %>%
      mutate(behav_state = ifelse(immersion <= 3, "transit", 
                                  ifelse(immersion >= 197, "rest", "aloft")),
             datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
      select(c(ring, season, datetime, ringYr, phase, loc, behav_state, 
               immersion, dryWet)) %>%
      relocate(ring, season, datetime, ringYr, phase, loc, behav_state, 
               immersion, dryWet)

    # Clean duplicates - sometimes have non-logging period recorded as co-deployed
    gls_j %<>%
      mutate(datetime = lubridate::round_date(datetime, "10 minutes")) %>%
      group_by(datetime) %>%
      arrange(desc(immersion)) %>%
      slice(1) %>%
      ungroup()
    
    trips_j <- gls_j
    trips_j$tripID = rep(1:length(rle(trips_j$loc)$values), rle(trips_j$loc)$lengths)
    
    trips_j %<>%
      group_by(ring, season, tripID) %>%
      mutate(start = datetime[1],
             end = datetime[n()],
             duration.mins = as.numeric(difftime(end, start, unit = "mins"))) %>%
      arrange(start) %>%
      filter(!is.na(datetime) & !is.na(end))
    
    ### Get rid of very short trips/col visits ----
    if (all(trips_j$duration.mins < 480)) next
    
    max_tries <- 5
    tries <- 0
    
    while(any(trips_j$duration.mins <= 480) & tries <= max_tries) {
      
      tries = tries + 1
      
      trips_j$loc <- ifelse(trips_j$duration.mins <= 480, 
                             ifelse(trips_j$loc == "trip", "col", "trip"),
                             trips_j$loc)
      trips_j$tripID <- rep(1:length(rle(trips_j$loc)$values), rle(trips_j$loc)$length)
      
      trips_j %<>%
        group_by(ring, season, tripID) %>%
        mutate(start = datetime[1],
               end = datetime[n()],
               duration.mins = as.numeric(difftime(end, start, unit = "mins"))) %>%
        arrange(start)
      
    }
    
    
    # Skip if only colony locations
    if(length(unique(trips_j$loc)) < 2) next
    
    ## Summarise behaviours ----
    trips_j %<>%
      group_by(tripID) %>%
      filter(loc == "trip") %>%
      mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S"),
             prev_state = lag(dryWet),
             bout_time = as.numeric(difftime(datetime, lag(datetime), units = "mins"))) %>%
      dplyr::summarise(species = species,
                       colony = colony,
                       ring = ring[1],
                       season = season[1],
                       phase = phase[1],
                       start_date = datetime[1],
                       end_date = max(datetime, na.rm = T),
                       loc = loc[1],
                       # Duration
                       duration.hrs = as.numeric(difftime(end_date, start_date, units = "hours") + 0.17),
                       duration.days = duration.hrs / 24,
                       # Transitions
                       transitions = sum(dryWet != prev_state, na.rm = TRUE),  # Count state transitions
                       transitions.var = var(transitions),
                       landing = sum(dryWet == "dry" & prev_state == "wet", na.rm = TRUE),
                       takeoff = sum(dryWet == "wet" & prev_state == "dry", na.rm = TRUE),
                       # Behaviours
                       transit_time.hrs = sum(bout_time[behav_state == "transit"], na.rm = T) / 60,
                       rest_time.hrs = sum(bout_time[behav_state == "rest"], na.rm = T) / 60,
                       transit_time.pct = transit_time.hrs/duration.hrs * 100,
                       rest_time.pct = rest_time.hrs/duration.hrs * 100) %>%
      ungroup() %>%
      filter(duration.hrs > 0) %>%
      filter(!(transit_time.pct > 100 | rest_time.pct > 100)) %>%
      mutate(landings_per_day = landing/duration.days) %>%
      select(-tripID) %>%
      filter(duration.days <= 30 & rest_time.pct > 0 & rest_time.pct < 100)
    
    # Add ring_season
    trips_j %<>%
      mutate(ring_season = paste(ring, season, sep = "_")) %>%
      relocate(ring_season, .before = colony)
    
    all_trips.list[[j]] <- trips_j
    
  }
  
  close(pb)
  trips_summary <- do.call("rbind", all_trips.list)
  
  # Do some filtering to ensure capture real foraging trips
  trips_summary %<>% filter(duration.days <= 30 &
                             rest_time.pct > 0 & 
                              rest_time.pct < 100,
                               landings_per_day > 0)
  
  # Remove unnecessary columns
  trips_summary %<>% 
    select(species, ring_season, colony, ring, season, phase, 
           start_date, duration.hrs, landings_per_day, transit_time.pct, rest_time.pct)
  
  # Output data
  save(trips_summary, file = paste0("Data/Intermediate/", spec_col[i], "_individual_gls_trips.RData"))
  
}



```

