---
title: "Processing 2: Extract at-sea activity"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 4
---

Purpose of script: Label GPS with behaviour, extract foraging components, sample available locations for habitat models

Inputs:

-   `Data/Intermediate/{species}_{colony}_gps_labelled.RData`

-   `Data/Metadata/breeding_dates.RData`

-   `Data/Metadata/ind_meta.RData`

Outputs:

-   `Data/Intermediate/{species}_{colony}_available_pnts_population_level.RData`


# Session info

**R version 4.3.1 (2023-06-16 ucrt)**

**Platform:** x86_64-w64-mingw32/x64 (64-bit) 

**locale:**
_LC_COLLATE=English_United Kingdom.utf8_, _LC_CTYPE=English_United Kingdom.utf8_, _LC_MONETARY=English_United Kingdom.utf8_, _LC_NUMERIC=C_ and _LC_TIME=English_United Kingdom.utf8_

**attached base packages:** 
_stats_, _graphics_, _grDevices_, _utils_, _datasets_, _methods_ and _base_

**other attached packages:** 
_adehabitatHR(v.0.4.22)_, _adehabitatLT(v.0.3.28)_, _CircStats(v.0.2-7)_, _boot(v.1.3-28.1)_, _MASS(v.7.3-60)_, _adehabitatMA(v.0.3.17)_, _ade4(v.1.7-24)_, _rnaturalearth(v.1.2.0)_, _marmap(v.1.0.12)_, _raster(v.3.6-32)_, _sp(v.2.2-1)_, _terra(v.1.8-29)_, _lubridate(v.1.9.4)_, _sf(v.1.0-20)_, _momentuHMM(v.1.5.8)_, _tidyr(v.1.3.2)_, _ggplot2(v.4.0.2)_, _magrittr(v.2.0.3)_ and _dplyr(v.1.2.0)_

# Set-up

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, eval = FALSE)

# Define the packages
packages <- c("dplyr", "magrittr", "ggplot2", "tidyr",
              "momentuHMM", "sf", "lubridate", "terra", "raster",
              "marmap", "rnaturalearth", "adehabitatHR")

# Install packages not yet installed - change lib to library path
# installed_packages <- packages %in% rownames(installed.packages())
# 
# if (any(installed_packages == FALSE)) {
#  install.packages(packages[!installed_packages])
# }

# Load packages
invisible(lapply(packages, library, character.only = TRUE))

source("R/Functions/RPT_functions.R")


# Suppress dplyr summarise warning
options(dplyr.summarise.inform = FALSE)
select <- dplyr::select

```

## Isolate foraging behaviour

### Prepare the data for a HMM

```{r prepare_hmm_data}

# Load the data ----

# Set manually
spec_col <- "bbal_birdis"  # "bbal_birdis" "bbal_ker" "waal_birdis", "waal_cro"

# Load the GPS data
my_file <- list.files("Data/intermediate/", pattern = paste0(spec_col, "_gps_labelled"))
my_gps <- loadRData(paste0("Data/intermediate/", my_file)) 

my_trip_meta <- my_gps %>% select(c(ring, season, datetime, dist_col, phase, boutID)) %>%
  group_by(ring, season) %>% mutate(max_range = max(dist_col, na.rm = T)) %>% ungroup()

my_gps %<>%
  mutate(ID = paste0(ring, "_", season, "_", boutID),
         season = ifelse(as.numeric(format(as.Date(datetime), "%m")) < 9,
                         as.numeric(format(as.Date(datetime), "%Y")),
                         as.numeric(format(as.Date(datetime), "%Y")) + 1 ))

### Prepare dataset ------

my_gps %<>% 
  # Check formatting
  mutate(datetime = as.POSIXct(datetime, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate(across(c(latitude, longitude), as.numeric)) %>% 
  filter(!is.na(longitude) & !is.na(latitude)) %>%
  # Filter for trips only
  filter(loc == "trip") %>%
  group_by(ID) %>%
  mutate(n_recs = n()) %>%
  # Remove short trips
  filter(n_recs > 3) %>% # loses 28 trips 
  select(-n_recs) %>%
  ungroup() 

my_gps.hmm <- my_gps %>%
  select(ID, datetime, longitude, latitude) %>%
  arrange(ID, datetime) %>%
  relocate(ID, longitude, latitude) %>% 
  data.frame() 

subset(my_gps.hmm, grepl("5072900", ID))

gps_hmm <- prepData(my_gps.hmm,
                    type = "LL", 
                    coordNames = c("longitude", "latitude")) 

# Remove erroneous step lengths that can't be larger than 40 m
hist(gps_hmm$step)
nrow(subset(gps_hmm, step > 40))/nrow(gps_hmm)
gps_hmm <- subset(gps_hmm, step < 40)

## Set 0s to very small numbers 
gps_hmm$step[gps_hmm$step == 0 | is.na(gps_hmm$step)] <- runif(sum(gps_hmm$step == 0 | is.na(gps_hmm$step))) / 10000
gps_hmm <- na.omit(gps_hmm)
gps_hmm %<>% filter(!is.na(angle) & !is.na(step))

save(gps_hmm, file = paste0("Data/Intermediate/", spec_col, "_HMMdata.RData"))

```

### Fit the HMM

NOTE: Initial value finding code taken from Clay et al. 2020.

Clay, T. A., Joo, R., Weimerskirch, H., Phillips, R. A., den Ouden, O., Basille, M., Clusella-Trullas, S., Assink, J. D., & Patrick, S. C. (2020). Sex-specific effects of wind on the flight decisions of a sexually dimorphic soaring bird. *Journal of Animal Ecology, 89*(8), 1811–1823. <https://doi.org/10.1111/1365-2656.13267>

```{r fit_the_HMM}

## Assign step lengths based on previously identified initial values
shape_0 <- c(12.42, 4.10, 0.33)
scale_0 <- c(3.62, 4.71, 0.17)
stepPar0 <- c(shape_0, scale_0)

## Assigning turning angles based on previously identified initial values
location_0 <- c(0.00302, 0.00343, 0.0291)
concentration_0 <- c(50.79, 1.27, 44.02)
anglePar0 <- c(location_0, concentration_0)

load(paste0("Data/Intermediate/", spec_col, "_HMMdata.RData"))

stateNames <- c("travel","search", "rest")

my_hmm <- fitHMM(
  data = gps_hmm,
  nbStates = 3,
  dist = list(step = "gamma", angle = "vm"),
  Par0 = list(step = stepPar0, angle = anglePar0),
  estAngleMean = list(angle = TRUE),
  stateNames = stateNames)

save(my_hmm, file = paste0("Outputs/Models/", spec_col, "_HMM.RData"))

## Plot pseudo-residuals
# plotPR(my_hmm)

```

```{r assign_behaviours}

load(paste0("Outputs/Models/", spec_col, "_HMM.RData"))

hmm_data_out <- my_hmm$data
hmm_data_out$State <- viterbi(my_hmm)

# Assess step/angle distributions
ggplot(aes(x = step, fill = as.factor(State)), data = hmm_data_out) + geom_histogram(alpha = 0.5)
ggplot(aes(x = angle, fill = as.factor(State)), data = hmm_data_out) + geom_histogram(alpha = 0.5)

# Label each state and check classification
hmm_data_out$State[hmm_data_out$State == 1] <- "Travel"
hmm_data_out$State[hmm_data_out$State == 2] <- "Search"
hmm_data_out$State[hmm_data_out$State == 3] <- "Rest"

# Check states
table(hmm_data_out$State)

# Calculate percentage time spent in each state 
hmm_data_out %>%
  group_by(State) %>%
  summarize(counts = n()) %>%
  mutate(per = counts / sum(counts) * 100) %>%
  collect()

# Merge with OG data
hmm_states <- hmm_data_out %>% select(c(ID, datetime, step, angle, State))

my_gps %<>% select(ring, season, datetime, longitude, latitude, dist_col, boutID, ID)
my_gps <- merge(my_gps, hmm_states, by = c("ID", "datetime"), all.x =T)
my_gps %<>% filter(!is.na(step) & !is.na(angle))

save(my_gps, file = paste0("Data/Intermediate/", my_species, "_", colony_exp, "_labelledHMM.RData"))

```

## Sample for available habitat

```{r sample_points}

# Load breeding dates
load("Data/Metadata/breeding_dates.RData")
load("Data/Metadata/ind_meta.RData")

spec_col <- c("bbal_ker", "bbal_birdis", "waal_birdis", "waal_cro")
ker_coords <- data.frame(longitude = 70.2333, latitude = -49.6833)
bi_coords <- data.frame(longitude = -38.05, latitude = -54.00)
cro_coords <- data.frame(longitude = 51.706972, latitude = -46.358639)

n.pnts = 6

for (i in 1:length(spec_col)) {

  set.seed(123) 
  
  print(paste0("Processing ", spec_col[i], "..."))
  
  load(paste0("Data/Intermediate/", spec_col[i], "_labelledHMM.RData"))
  
  my_species = strsplit(spec_col, "_")[[i]][1]
  my_colony = strsplit(spec_col, "_")[[i]][2]
  my_colony = ifelse(my_colony == "birdis", "birdisland", 
                     ifelse(my_colony == "ker", "kerguelen",
                            ifelse(my_colony == "cro", "crozet", my_colony)))
  
  ## Match in breeding phase
  my_breeding <- breed_dates %>% filter(species == my_species, colony == my_colony) %>%
         pivot_longer(cols = c(ring, partner), names_to = "role", values_to = "ring") %>%
         select(c(season, ring, hatch_date))
  
  ### Set hatch dates for when unknown
  if (my_colony == "birdis" & my_species == "waal") {
    median_hatch <- "03-10" } else if (my_colony == "birdis" & my_species == "bbal") { 
      median_hatch <- "01-05" } else if (my_colony == "crozet") {
        median_hatch <- "03-22" } else if (my_colony == "kerguelen") {
          median_hatch <- "12-15" }
  
  my_gps <- merge(my_gps, my_breeding, by = c("ring", "season"), all.x = T)
  my_gps %<>% mutate(date = as.Date(datetime),
                     hatch_date = ifelse(!is.na(hatch_date), 
                                         as.character(hatch_date),
                                         paste0(season, "-", median_hatch)),
                     hatch_date = as.Date(hatch_date),
                     phase = ifelse(date < hatch_date, "incubation", "brooding"),
                     year = year(datetime))

  ## Match in sex
  my_ind <- all_ind.meta %>% mutate(species = tolower(species),
                                    species = ifelse(species == "bba", "bbal", "waal")) %>% 
    filter(species == my_species) %>%
    select(ring, sex) %>% distinct()
  
  ### Remove sex if have > 2
  my_ind %<>% group_by(ring) %>% mutate(n_sexes = n_distinct(sex), sex = ifelse(n_sexes > 1, NA, sex)) %>% select(-n_sexes) %>% distinct()
  my_gps <- merge(my_gps, my_ind, by = "ring", all.x = T)
  
  ## Get colony coords
  if (grepl("ker", spec_col[i])) {
    col_coords <- ker_coords
  } else if (grepl("birdis", spec_col[i])) {
    col_coords <- bi_coords
  } else if (grepl("cro", spec_col[i])) {
    col_coords <- cro_coords }
  
  ## Calculate UD per phase * sex ----------------
  
  ## Set CRS
  if (my_colony == "kerguelen") {
    my_crs = 32742 # UTM zone 42S 
  } else if (my_colony == "birdisland") {
    my_crs = 32724 # UTM Zone 24S
  } else if (my_colony == "crozet") {
    my_crs = 32740 } # UTM Zone 40S
  
  # Make group ID to account for breeding phase (separate UD per phase)
  my_gps %<>% mutate(group_id = paste(phase, sex, sep = "_"))
  
  ud.df <- my_gps %>% filter(!is.na(sex)) %>% select(longitude, latitude, group_id) %>% 
    rename(x = longitude, y = latitude, ID = group_id) 
  coordinates(ud.df) <- ud.df[, c('x', 'y')]
  
  # proj4string(ud.df) <- CRS(SRS_string = paste0("EPSG:", my_crs))
  # ud.sf <- st_as_sf(ud.df, coords = c("x", "y"), crs = my_crs)
  # ud.df <- as(ud.sf, "Spatial")
  
  ## Split the UDs by group
  if(grepl("cro", spec_col[i])) {
    ud.est <- kernelUD(ud.df[, 3], h = "href", extent = 3, grid = 500)
  } else {  ud.est <- kernelUD(ud.df[, 3], h = "href") }
  
  ud.95 <- getverticeshr(ud.est, percent = 95)
  ud.list <- split(ud.95, ud.95$id)
  
  ## Make a land mask
  land <- ne_countries(scale = "medium", returnclass = "sf") %>% st_transform(my_crs)
  land <- st_make_valid(land)
  
  ### Buffer to remove coastline
  land_buffered <- st_buffer(land, dist = 20000)
  
  # List to store all sampled available points
  sampled_pts.list <- list()
  ind_trips <- unique(my_gps$ID) 
  
  # Loop through unique used points (could be per hour or other resolution)
  for (ind in 1:length(ind_trips)) {
    
    print(paste0("Processing ", ind, " of ", length(ind_trips)))
    
    gps_ind <- my_gps %>% filter(ID == ind_trips[ind]) %>% mutate(date_hourly = round(datetime, units = "hours"))
   
    # Get relevant UD
    grp_id <- unique(gps_ind$group_id)
    if(length(grp_id) > 1) { grp_id <- grep("incubation", grp_id, value = TRUE) }
    
    if (grp_id == "incubation_NA") {
      # Combine male and female UDs
      ud.grp <- st_union(
        st_geometry(st_as_sf(ud.list[["incubation_M"]])),
        st_geometry(st_as_sf(ud.list[["incubation_F"]])) ) %>%
        st_sf(geometry = .) %>%
        st_set_crs(4326) %>%
        st_transform(my_crs)
    } else if (grp_id == "brooding_NA") {
      # Combine male and female UDs
      ud.grp <- st_union(
        st_geometry(st_as_sf(ud.list[["brooding_M"]])),
        st_geometry(st_as_sf(ud.list[["brooding_F"]])) ) %>%
        st_sf(geometry = .) %>%
        st_set_crs(4326) %>%
        st_transform(my_crs)
    } else {
      # Standard single UD
      ud.grp <- st_as_sf(ud.list[[grp_id]]) %>%
        st_set_crs(4326) %>%
        st_transform(my_crs)
    }
    
    ud.grp <- st_transform(ud.grp, crs = my_crs)
    
    # Buffer and remove land
    ud.buffer <- st_buffer(ud.grp, dist = 15000)
    ud.buffer_noLand <- st_difference(ud.buffer, st_union(land_buffered)) %>%
      st_union() %>%
      st_make_valid()
    
    # Sample n.pnts random points
    sampled_pnts <- gps_ind %>% 
      group_by(date_hourly) %>% 
      summarise(random_pnts = list(st_sample(ud.buffer_noLand, n() * n.pnts)), .groups = "drop",
                tripID = boutID[1]) %>%
      unnest(random_pnts) %>%
      mutate(row_id = row_number())
    
    # Convert to lat/lon
    sampled_sf <- st_sf(geometry = sampled_pnts, crs = my_crs) %>%
      st_transform(4326)
    coords <- st_coordinates(sampled_sf)
    
    # Create data.frame
    avail_pnts <- sampled_sf %>%
      st_drop_geometry() %>%
      mutate(longitude = coords[,1],
             latitude = coords[,2],
             ring = rep(gps_ind$ring)[1], 
             season = rep(gps_ind$season)[1],
             date_hourly = sampled_pnts$date_hourly, 
             tripID = sampled_pnts$tripID,
             year = year(date_hourly),
             used = 0,
             phase = gps_ind$phase[1],
             sex = gps_ind$sex[1]) %>% 
      relocate(ring, season, year, tripID, date_hourly, longitude, latitude, used) %>%
      select(-row_id)
    
    sampled_pts.list[[ind]] <- avail_pnts
  }
 
  avail.pnts.df <- do.call("rbind", sampled_pts.list)
 
  # Bind in with used points
  used.pnts.df <- my_gps %>% 
                    mutate(date_hourly = round(datetime, units = "hours"), 
                           used = 1) %>% 
    rename(tripID = boutID) %>% 
    select(c(ring, season, year, tripID, date_hourly, longitude, latitude, used, phase, sex))
  
  all_pnts <- rbind(used.pnts.df, avail.pnts.df) %>% arrange(ring, date_hourly)
  
  # Calculate distance from the colony
  all_pnts$dist_col <- apply(all_pnts, 1, function(row) {
    point_coords <- c(as.numeric(row["longitude"]), as.numeric(row["latitude"]))
    geosphere::distHaversine(point_coords, col_coords)
  })
  all_pnts$dist_col <- all_pnts$dist_col/1000
  
  
  # Output
  save(all_pnts, file = paste0("Data/Intermediate/", spec_col[i], "_available_pnts_population_level.RData"))
  
  print(paste0(spec_col[i], " complete! ", 4-i, " more to go."))
  
}

```